### 2019-09-16 main.tex
Update on Overleaf.

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=2cm]{geometry}
\usepackage{natbib}

\title{An agent-based model for city networks based on interactions between firms}

\date{}

\begin{document}

\maketitle

\section{Introduction}

\citep{taylor2001specification} world city network driven by interactions between firms/behavior of firms

\citep{martinus2018global} combination of different types of proximity in inter-urban firm networks: economic, sociocultural, geopolitical - implies spatial and non-spatial

\citep{pan2017mapping} different types of linkages: e.g. services here

\section{Model}

\subsection{Rationale}

\begin{itemize}
    \item Single level model: scale of cities - firms induce the links but are not represented in the model
    \item Cities are defined by their profile regarding the proportion of firms in the different sectors
    \item Links between cities are created in an iterative way, with the combination of several drivers:
    \begin{itemize}
        \item geographical proximity (distance or effective accessibility depending on the scale)
        \item geopolitical proximity (belonging to the same ``administrative unit'' $\simeq$ country or single market $\rightarrow$ application to Brexit)
        \item city size (taking not population but an economic variable related to firms avoids adding the dimension of population, simplifies the model) $\rightarrow$ can be combined with geographical distance with a classical interaction model, see \cite{cottineau2015growing}, \cite{raimbault2018indirect}, \cite{favaro2011gibrat}
        \item economic proximity (e.g. cosine distance between sector proximity as done in \cite{2019arXiv190505106C}
        \item previous linkages (taking into account history of previous firm linkages)
    \end{itemize}
    \item These competing drivers are combined - how ? \textit{Options: (i) multiplicative function to get probabilities; (ii) generalization with a multi-attribute Cobb-Douglas; (iii) multi-objective optimization depending on sectors, using Pareto fronts ? - needs empirical evidence; (iv) copula-based combination as done by \cite{2019arXiv190505106C} - copula parameters are free or estimated from real data ?}
    \item The accumulated effects and therefore path-dependency are taken into account into this dynamical model
\end{itemize}

\subsection{Model description}

Let $1 \leq i \leq N$ cities defined at time $t$ by their economic structure such that $E_i(t)$ is the total economic volume (GDP) for city $i$ and $S_{ik}(t)$ is a matrix giving economic volume proportions within each economic sector $k$, assuming $K$ economic sectors.

% Q : evolution of volume / populations ? Populations not needed indeed.

The city network evolves at discrete time steps the following way:
\begin{enumerate}
    \item Economic structure of cities are evolved with a city interaction model which can be
    \begin{itemize}
        \item a direct interaction model, according to
        \begin{equation}
            \Delta E_i = g_0\cdot E_i + w_G \cdot \sum_j V_{ij}/<V_{ij}>
        \end{equation}
        \item the Marius model with economic exchanges
    \end{itemize}
    \textit{Baseline: no evolution - on short time spans may be relevant ?}
    \item A new linkage is established
    \begin{itemize}
        \item \textit{rq: as these are done sequentially, it is equivalent to have a fixed time step and a number of linkages per time step, or to do it link by link - in the data-driven version the actual values shall be parametrized, while in the synthetic version this has no importance}
        \item the link is drawn randomly with a probability for it to be $i \rightarrow j$ proportional to a function
    \begin{equation}
        p_{ij} = f(d_{ij},g_{ij},E_i,E_j,s(S_{ik},S_{jk}),w_{ij})
    \end{equation}
    with $d_{ij}$ the geographical distance, $g_{ij}$ the geopolitical/sociocultural distance, $g$ a similarity function between sector structure, $w_{ij}$ the relative size of already existing links.
    \item A simple multiplicative form between the different terms may be too restrictive as it assumes independence between components and an equal share between each. Copula capture a dependency structure but must be parametrized by estimation on real data (\textit{keep it as an option for the data-driven model}). An other option is to use a generalized Cobb-Douglas function \citep{vilcu2011geometric}, where exponents can be kept as free parameters. Furthermore, a natural formulation as a separable gravity model as for cities above is already a particular case of such a function, and taking an exponential gravity potential yields that the decay exponent corresponds to the Cobb-Douglas exponent. We therefore take
    \begin{equation}
     p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_F} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_T} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s(S_{ik},S_{jk})^{\gamma_S} \cdot \exp \left(- \gamma_D \cdot d_{ij}\right) \cdot \exp \left(- \gamma_G \cdot g_{ij}\right)
    \end{equation}

    \item Note that: (i) we take an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalize the exponents to 1 (which is different than renormalizing the probability), to ensure to include convex functions.
    \item We take for the sake of simplicity (\textit{should be amended/validated/tested with an asymmetric function in a further version}) we take a cosine similarity between the sector structures, given with such probability vectors (size effect is already taken into account) by
    \begin{equation}
        s\left(S_{ik},S_{jk}\right) = \sum_{k=1}^{K} S_{ik} S_{jk}
    \end{equation}

    \item The ``geopolitical/sociocultural'' distance remains abstract and should be parametrized (see setup below) and estimated in the data-driven setting.
    \end{itemize}
\end{enumerate}

\subsection{Model setup}

\subsubsection{Synthetic setup}

Let consider the simplest synthetic setup possible:
\begin{itemize}
    \item cities are distributed randomly in a uniform and isotropic space. The total width of the space $d_{max}$ fixed to 3000km to be approximatively the size of Europe. The number of cities is fixed as $N=700$ to similarly match the European urban system structure (GHS database \citep{Florczyk2019ghs}).
    \item city sizes are attributed with a scaling law (equivalent to have a Zipf law for population and a scaling law for economic size), following an exponent $\beta_0$ (fixed at $\beta_0 = 1.1$ in most experiments when not specified, approximate value for GDP from the GHS database, see \citep{raimbault:halshs-02284897})
    % from the GHS database : scaling GDP - pop ~ 1.1 for Europe, rank size 0.94 -> 1.034 : take 1.1 to simplify - can be paramterized ?
    \item country boundaries are constructed to obtain an approximately equal coverage of the space for a fixed number of countries, what remains consistent with non-correlated city sizes at the scale of the countries as our cities are randomly distributed in space \citep{simini2019testing} (\textit{Note: on this point this may be realistic if staying at a continental level, although we should include realistic size of countries; this paper is surely wrong when changing the dataset used and the scale of analysis, working at the scale of Mega-city regions e.g.}). Therefore, we do a basic spatial k-means clustering with $C = 20$ countries and attribute the country accordingly.
    \item firm sector structure: 
\end{itemize}

\subsection{Model indicators}

\begin{itemize}
    \item weighted degree distribution (\textit{if it is fat-tailed, log-log power law exponent and adjusted r-squared})
\end{itemize}

\subsection{Model parameters}

\section{Results}

\subsection{Implementation}

% Netlogo for vis + scala into spatialdata / urbangrowth ?

% On vine copulas :
%    https://link.springer.com/article/10.1023/A:1016725902970
%    https://www.groups.ma.tum.de/en/statistics/research/vine-copula-models/
%    R package : https://github.com/tnagler/VineCopula

\subsection{Synthetic city systems}

\subsubsection{Sensitivity analysis and grid exploration}

\subsubsection{PSE}

\subsubsection{A virtual case study: subsystem-xit}

\bibliographystyle{apalike}
\bibliography{biblio}

\end{document}

### 2019-09-25 main.tex
Update on Overleaf.

@@
 \citep{taylor2001specification} world city network driven by interactions between firms/behavior of firms
+\citep{martinus2018global} combination of different types of proximity in inter-urban firm networks: economic, sociocultural, geopolitical - implies spatial and non-spatial 
+\citep{pan2017mapping} different types of linkages: e.g. services here, but not only. All sectors should be taken into account first.
 
-\citep{martinus2018global} combination of different types of proximity in inter-urban firm networks: economic, sociocultural, geopolitical - implies spatial and non-spatial
+The originality of the present paper is to approach these issues within the system of cities framework \citep{berry1964cities}, where the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \citep{pumain1997pour}.
+We examine the interactions of european cities within firm linkages.
 
-\citep{pan2017mapping} different types of linkages: e.g. services here
+% We compare the model to real data on transnational firm networks defined by ownership links at city-level, by the means of the Bureau Van Dijk's \emph {FAME} database with a graph theory application.
 
@@
     \item The ``geopolitical/sociocultural'' distance remains abstract and should be parametrized (see setup below) and estimated in the data-driven setting.
+    \item \textcolor{red} TODO the path dependency $w_{ij}$ could (or should ?) be decomposed into in/out degree for origin and destination ? (then stronger self-reinforcment effect)
     \end{itemize}
@@
     \item firm sector structure: 
+    \begin{itemize}
+        \item random probabilities
+        \item uniform: then the term of proximity is cancelled, no interest
+        \item function of size: larger cities are more diverse, larger cities have more knowledge-based sectors. Test 1: log-normal with slightly increasing width and shifting mode.
+        % mode of log normal is exp(mu - sigma ^ 2) ; variance is [exp(sigma^2) - 1]*exp(2 mu + sigma ^ 2)
+        % when the pdf is 1 / (x sigma sqrt(2 pi)) * exp (- (ln x - mu) ^ 2 / 2 sigma^2 )
+        % -> for largest city stddev = sqrt(variance) = fields / 2 ; for smallest sigma = 1 / nfields
+        %  ==> solving for mu and sigma with linear mode and variance as a function of log(E_i)
+        % => sigma^2 is the unique positive root of f(X)=0 with f(X) = -3X - 2 ln(exp(X)-1) - ln(e_i)
+        % => mu = sigma^2 + ln(e_i)
+        %    ~ somehow dirty and not sure that most of the distrib for highest values is within [0,1]
     \end{itemize}
+\end{itemize}
+
+! links depend on the sectors: cobb douglas for sector proximity should depend on sector composition ? or several functions ?
 
@@
 
+\textit{find some indicators for polarization : when big cities take everything or when local clusters organize}
+
+Network indicators:
+
 \begin{itemize}
     \item weighted degree distribution (\textit{if it is fat-tailed, log-log power law exponent and adjusted r-squared})
+\end{itemize}
+
+Sector structure indicators:
+
+\begin{itemize}
+    \item sector diversity (Herfindhal ?) 
 \end{itemize}

### 2019-09-25 slides.tex
Update on Overleaf.

\input{header_slides.tex}

\begin{document}

\title{An agent-based model for city networks based on interactions between firms}
\author{J.~Raimbault$^{1,2,3\ast}$ and N. Zdanowska$^{1,3}$\\\medskip
$^{\ast}$\texttt{j.raimbault@ucl.ac.uk}
}

\institute{$^{1}$CASA, UCL\\
$^{2}$UPS CNRS 3611 Complex Systems Institute Paris\\
$^{3}$UMR CNRS 8504 G{\'e}ographie-cit{\'e}s
}

\date{CCS 2019\\
Satellite Simulation models\\
October 2nd 2019
}

\frame{\maketitle}

\section{Introduction}

\sframe{Urban firm networks}{

Cities as maximisation of socio-economic interactions

Network economies \cite{sassen1991global} \cite{castells1996information}

World city network driven by interactions between firms (Taylor, 2001; Martinus and Sigler, 2018)

Asymmetrical spatial interactions -- different relative position of cities

}

\sframe{An evolutionary approach to Urban Systems}{

Evolutionary Theory of Urban Systems \cite{pumain1997pour} \cite{pumain2006villes}

Adaptive cycles and diffusion on innovation (H{\"a}gerstrand, 1973)

Path dependence

Selection

Emergent structures

% discussion theorie Denise

}

\sframe{Firm linkages structuring urban systems}{

% examples de resultats pour l'Europe Centrale et Orientale

 \begin{center}
 \includegraphics[width=0.7\textwidth]{figures/Figure1.png}
 \end{center}

  \textit{Central and Eastern European cities within ownership links of firms in 2013.}

}

% si besoin je peux rajouter d'autres exemples sur le UK ou carrement des trucs de Celine sur les reseaux de firmes

\sframe{Urban Firm linkages and geo-economic processes}{

Metropolisation vs. regionalisation effects

Relationship local/global of multinational firm linkages (internationalisation)

Specialisation-driven factors for an economic network to emerge

Drivers of innovation

Macroeconomic exogenous chocs and resilience of urban systems

Can we access geographical and economic processes within urban networks of firms with a generative model?

}
\sframe{Towards a generative model of urban firm networks}{

% research question / objective
% du coup je te laisse formuler le broad research question, car je vois pas bien

}

\section{Model}

\sframe{Model rationale}{

}

\sframe{Formalization}{

Cities characterized by economic size $E_i$ (GDP) and economic structure $s_{ik}$

\bigskip

Starting from an initial network, at each time step:
\begin{enumerate}
    \item Evolve cities with an interaction model (\cite{raimbault2018indirect} or \cite{cottineau2015growing})
    \item Add a fixed number of links randomly, following a probability function of sizes, sector proximity, and geographical and socio-cultural proximities
\end{enumerate}

}

\sframe{Formalization}{

Probability for a new link follows a generalized Cobb-Douglas function

\begin{equation}
     p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_F} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_T} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s(S_{ik},S_{jk})^{\gamma_S} \cdot \exp \left(- \gamma_D \cdot d_{ij}\right) \cdot \exp \left(- \gamma_G \cdot g_{ij}\right)
\end{equation}

}

\sframe{Model indicators}{

Geographical indicators:

\begin{itemize}
    \item Internationalisation
    \item Metropolisation
    \item Regionalisation
    \item Specialisation
\end{itemize}

\bigskip

Network and flows indicators

}

\sframe{Simulation on synthetic systems of cities}{

% la ou avant il faut surement expliquer que c'est pas totalement synthetique, que tu pars de X villes europeennes de plus de 50 000 habitants (pour les plus nuls comme moi)
% si c'est completement synthetiques, l'histoire des villes europeenes c'est pour avoir les ordres de grandeur de la hierarchie et du nombre de villes, mais c'est pas parametrise sur des vraies donnees - mais oui ca ca sera explique dans comment construire le systeme de ville synthetique.

}

\section{Results}

\sframe{Implementation}{

% for now no dynamics of cities

}

\sframe{Simulation of urban networks}{

    % examples

    \centering

    \includegraphics[width=0.48\textwidth]{figures/ex_alleq-highgravity_seed-12102_t1500.png}\hspace{0.1cm}
    \includegraphics[width=0.48\textwidth]{figures/ex_alleq-lowgravity_seed-12102_t1500.png}

}

\sframe{Statistical convergence}{

Good convergence for most indicators: average sharpes ratios on a one-factor sampling above 5 (except for metropolization index $\simeq 1.5$)

% + histograms

}

\sframe{Interaction decay}{

    % One factor sampling results
    %  20190924_162740_ONEFACTOR_REPLICATIONS_SYNTHETIC_GRID

    \includegraphics[width=0.48\textwidth]{figures/internationalization-gravityDecay_errorbars.png}
    \includegraphics[width=0.48\textwidth]{figures/rhoDegreeSize-gravityDecay_errorbars.png}
}

\sframe{Sector proximity}{

    % One factor sampling results
    %  20190924_162740_ONEFACTOR_REPLICATIONS_SYNTHETIC_GRID

     \includegraphics[width=0.48\textwidth]{figures/internationalization-gammaSectors_errorbars.png}
    \includegraphics[width=0.48\textwidth]{figures/rhoDegreeSize-gammaSectors_errorbars.png}

}

\sframe{Grid exploration}{

}

% other experiments ? PSE on correlations ?

\section{Discussion}

% - future work etc (classical discussion)
% - on the role of model exploration and simulation to gain knowledge from the model

% comparing with real data?

\sframe{On the role of model exploration}{

}

\sframe{Discussion}{

}

\sframe{Conclusion}{

$\rightarrow$

\bigskip
\bigskip

\textbf{Open repository} at

\texttt{https://github.com/JusteRaimbault/ABMCitiesFirms}

\bigskip

\textbf{Acknowledgments}: thanks to the \textit{European Grid Infrastructure} for access to the infrastructure.

}

\sframe{Reserve slides}{

\centering

\Large

\textbf{Reserve Slides}

}

%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[allowframebreaks]
\frametitle{References}
\bibliographystyle{apalike}
\bibliography{biblio}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2019-09-25 slides.tex
ucl_template<3

@@
 
-\title{An agent-based model for city networks based on interactions between firms}
-\author{J.~Raimbault$^{1,2,3\ast}$ and N. Zdanowska$^{1,3}$\\\medskip
+\title[Agent-based modeling city networks]{An agent-based model for city networks based on interactions between firms}
+\author[Raimbault \and Zdanowska]{J.~Raimbault$^{1,2,3\ast}$ \and N. Zdanowska$^{1,3}$\\\medskip
 $^{\ast}$\texttt{j.raimbault@ucl.ac.uk}
@@
 
-\institute{$^{1}$CASA, UCL\\
+\institute[UCL]{$^{1}$CASA, UCL\\
 $^{2}$UPS CNRS 3611 Complex Systems Institute Paris\\
@@
 
-\date{CCS 2019\\
+\date[2nd October 2019]{CCS 2019\\
 Satellite Simulation models\\

### 2019-09-25 slides.tex
Update on Overleaf.

@@
 
-Cities as maximisation of socio-economic interactions
+General framework:
 
-Network economies \cite{sassen1991global} \cite{castells1996information}
+\begin{itemize}
+    \item Cities as maximisation of socio-economic interactions 
+    \item Network economies \cite{sassen1991global} \cite{castells1996information}
+    \item World city network driven by interactions between firms (Taylor, 2001; Martinus and Sigler, 2018)
+    \item Asymmetrical spatial interactions -- different relative position of cities
+\end{itemize}
 
-World city network driven by interactions between firms (Taylor, 2001; Martinus and Sigler, 2018)
-
-Asymmetrical spatial interactions -- different relative position of cities
+\bigskip
 
@@
 
-Evolutionary Theory of Urban Systems \cite{pumain1997pour} \cite{pumain2006villes}
+\begin{itemize}
+    \item Evolutionary Theory of Urban Systems \cite{pumain1997pour} \cite{pumain2006villes}
+    \item Adaptive cycles and diffusion on innovation (H{\"a}gerstrand, 1973) 
+    \item Path dependence \cite{martinsunley2006information}
+    \item Selection and emerging structures of systems
+\end{itemize}
 
-Adaptive cycles and diffusion on innovation (H{\"a}gerstrand, 1973)
-
-Path dependence
-
-Selection
-
-Emergent structures
-
+\bigskip
 % discussion theorie Denise
-
 }
@@
 
-Metropolisation vs. regionalisation effects
+Assumptions
 
-Relationship local/global of multinational firm linkages (internationalisation)
+\begin{itemize}
+    \item Metropolisation vs. regionalisation effects
+    \item Relationship local/global of multinational firm linkages 
+    \item Specialisation-driven factors and drivers of innovation
+    \item Macroeconomic exogenous chocs and resilience of urban systems 
+\end{itemize}
 
-Specialisation-driven factors for an economic network to emerge
-
-Drivers of innovation
-
-Macroeconomic exogenous chocs and resilience of urban systems
-
-Can we access geographical and economic processes within urban networks of firms with a generative model?
+How can we approach geographical and economic processes within urban networks of firms with a generative model?
 
@@
 
+Towards a generative model of urban firm networks
+
 % research question / objective
-% du coup je te laisse formuler le broad research question, car je vois pas bien
+%
 
@@
 % si c'est completement synthetiques, l'histoire des villes europeenes c'est pour avoir les ordres de grandeur de la hierarchie et du nombre de villes, mais c'est pas parametrise sur des vraies donnees - mais oui ca ca sera explique dans comment construire le systeme de ville synthetique.
+
+% ok desolee je n'avais pas bien compris
 
@@
 
-% comparing with real data?
+% further comparing with real data?
 

### 2019-09-26 slides.tex
.

--- 2019-09-25 slides.tex
@@
 
-World city network driven by interactions between firms (Taylor, 2001; Martinus and Sigler, 2018)
+World city network driven by interactions between firms \cite{taylor2001specification} \cite{martinus2018global}
 
@@
 Emergent structures
-
-% discussion theorie Denise
 
@@
 \begin{equation}
-     p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_F} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_T} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s(S_{ik},S_{jk})^{\gamma_S} \cdot \exp \left(- \gamma_D \cdot d_{ij}\right) \cdot \exp \left(- \gamma_G \cdot g_{ij}\right)
+     p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_F} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_T} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s\left(S_{ik},S_{jk}\right)^{\gamma_S} \cdot \exp \left(- \gamma_D \cdot d_{ij}\right) \cdot \exp \left(- \gamma_G \cdot g_{ij}\right)
 \end{equation}
@@
 
+\textit{Following \cite{raimbault2018space}, geosimulation models must be studied within synthetic controllable urban contexts (spatial sensitivity)}
+
+\bigskip
+
 }
@@
 % for now no dynamics of cities
+
+\begin{itemize}
+	\item Model implemented in NetLogo (good compromise interactivity / ergonomy), with fast data structures (matrix/table extensions)
+	\item Integrated into OpenMOLE \cite{reuillon2013openmole} for model exploration
+	\item Current implementation: only network dynamics 
+\end{itemize}
 
@@
 
+	\medskip
+
+	\textit{Networks at $t=1500$, for default parameters values and high gravity (left) and low gravity (right)}
+
 }
@@
 
-Good convergence for most indicators: average sharpes ratios on a one-factor sampling above 5 (except for metropolization index $\simeq 1.5$)
+$\rightarrow$ Good convergence for most indicators: average sharpes ratios on a one-factor sampling above 5 (except for metropolization index $\simeq 1.5$)
 
 % + histograms
+
+\bigskip
 
@@
 
+\textbf{Practical application}
+
+\textbf{Developments}
+
 }
@@
 \sframe{Conclusion}{
+
+$\rightarrow$ A simulation model to understand processes of network emergence
 
@@
 
-\textbf{Open repository} at
+\textbf{Open repository for model and results at}
 
 \texttt{https://github.com/JusteRaimbault/ABMCitiesFirms}
+
+\textbf{Simulation data at}
+
+\texttt{}% dataverse
 

### 2019-09-26 slides.tex

--- 2019-09-25 slides.tex
@@
     \item Network economies \cite{sassen1991global} \cite{castells1996information}
-    \item World city network driven by interactions between firms (Taylor, 2001; Martinus and Sigler, 2018)
+    \item World city network driven by interactions between firms \cite{taylor2001specification} \cite{martinus2018global}
     \item Asymmetrical spatial interactions -- different relative position of cities
@@
 
-\bigskip
-% discussion theorie Denise
 }

### 2019-09-26 slides.tex
Update on Overleaf.

@@
     \item Evolutionary Theory of Urban Systems \cite{pumain1997pour} \cite{pumain2006villes}
-    \item Adaptive cycles and diffusion on innovation (H{\"a}gerstrand, 1973) 
-    \item Path dependence \cite{martinsunley2006information}
+    \item Adaptive cycles and diffusion on innovation \cite{Hagerstrand1968Innovation}
+    \item Path dependence \cite{martin2006path}
     \item Selection and emerging structures of systems
@@
 
-  \textit{Central and Eastern European cities within ownership links of firms in 2013.}
+  \textit{Central and Eastern European cities within ownership links of firms in 2013}
 
@@
 
+\item Cities are defined by their profile regarding the proportion of firms in the different sectors. Links between cities are created in an iterative way, taking into account:
+    \begin{itemize}
+        \item geographical proximity (distance or effective accessibility)
+        \item geopolitical proximity (belonging to the same country or single market $\rightarrow$ application to Brexit)
+        \item city size (economic size as GDP) 
+        \item economic similarity (e.g. cosine distance between sector proximity as done in \cite{2019arXiv190505106C}
+        \item previous linkages 
+    \end{itemize}
+
+% j'ai mis ca plus pour moi,enfin pour moi c'est plus clair pour la presentation
+
 }
@@
     \item Evolve cities with an interaction model (\cite{raimbault2018indirect} or \cite{cottineau2015growing})
-    \item Add a fixed number of links randomly, following a probability function of sizes, sector proximity, and geographical and socio-cultural proximities
+    \item Add a fixed number of links randomly, following a probability function of sizes, sector proximity, and geographical and socio-cultural proximity
 \end{enumerate}

### 2019-09-26 slides.tex
Update on Overleaf.

@@
     \item Cities as maximisation of socio-economic interactions 
+    % que veux tu dire par maximisation ? je mettrais plutot "foster socio-eco"
     \item Network economies \cite{sassen1991global} \cite{castells1996information}
@@
     \item Adaptive cycles and diffusion on innovation \cite{Hagerstrand1968Innovation}
-    \item Path dependence \cite{martin2006path}
-    \item Selection and emerging structures of systems
+    \item Path dependence \cite{martin2006path} \cite{pumain2012urban}
+    \item Selection and emerging structures of systems % pas clair
+    \item Evolutionary models for urban systems dynamics: \cite{favaro2011gibrat} \cite{cottineau2015growing} \cite{schmitt2015half} \cite{raimbault2018indirect} \cite{raimbault2018calibration}
 \end{itemize}
@@
 % si besoin je peux rajouter d'autres exemples sur le UK ou carrement des trucs de Celine sur les reseaux de firmes
+% oui ca serait cool des trucs a toi sur le UK
 
@@
 
-How can we approach geographical and economic processes within urban networks of firms with a generative model?
+\bigskip
+
+How can we capture geographical and economic processes within urban networks of firms with a generative model?
 
@@
 
-\item Cities are defined by their profile regarding the proportion of firms in the different sectors. Links between cities are created in an iterative way, taking into account:
+\item Cities are defined by their GDP and their profile regarding the proportion of firms in the different sectors. Links between cities are created in an iterative way, taking into account:
+
+\medskip
+
     \begin{itemize}
@@
 % j'ai mis ca plus pour moi,enfin pour moi c'est plus clair pour la presentation
+%  => oui c'est exactement ce quil fallait mettre !
 
@@
 
-Cities characterized by economic size $E_i$ (GDP) and economic structure $s_{ik}$
+Cities characterized by economic size $E_i$ (GDP) and economic structure $S_{ik}$ (probability distribution of firms within $K$ sectors)
 
@@
 \begin{enumerate}
-    \item Evolve cities with an interaction model (\cite{raimbault2018indirect} or \cite{cottineau2015growing})
+    \item Evolve city sizes $E_i(t+1) = f(E_j(t))$ with an interaction model (\cite{raimbault2018indirect} or \cite{cottineau2015growing})
     \item Add a fixed number of links randomly, following a probability function of sizes, sector proximity, and geographical and socio-cultural proximity
@@
 
-\begin{equation}
-     p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_F} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_T} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s\left(S_{ik},S_{jk}\right)^{\gamma_S} \cdot \exp \left(- \gamma_D \cdot d_{ij}\right) \cdot \exp \left(- \gamma_G \cdot g_{ij}\right)
-\end{equation}
+\medskip
+
+\[
+     p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_F} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_T} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s\left(S_{ik},S_{jk}\right)^{\gamma_S} \cdot \exp \left(- \gamma_G \cdot d_{ij}\right) \cdot \exp \left(- \gamma_D \cdot g_{ij}\right)
+\]
+
+\medskip
+
+where $E=\sum_k E_k$, $W = \sum_{i,j} w_{ij}$, $s$ is cosine similarity, $d_{ij}$ euclidian distance, and $g_{ij}$ socio-cultural distance
+
+\medskip
+
+\textbf{Model parameters: } $\gamma_F, \gamma_T, \gamma_W, \gamma_S, \gamma_D = \frac{1}{d_G}, \gamma_G = \frac{1}{d_G}$
 
@@
 
-Geographical indicators:
+\textbf{Geographical indicators:}
 
 \begin{itemize}
-    \item Internationalisation
-    \item Metropolisation
-    \item Regionalisation
-    \item Specialisation
+    \item Internationalisation (modularity of countries in the network)
+    \item Metropolisation (correlation between weighted degree and city size)
+    \item Regionalisation (correlation between length and flow of links, stratified by size of extremities)
+    \item Specialisation (correlation between sector proximity and flow of links, stratified by size of extremities)
 \end{itemize}
@@
 
-Network and flows indicators
+\textbf{Network and flows indicators:}
+
+%  networkModularity, networkAvgCommunitySize, networkDegreeHierarchy, networkAvgDegree, networkDegreeEntropy, flowsHierarchy, flowsEntropy, rhoDegreeSize, rhoFlowDistancePos, rhoFlowDistance
+
+\begin{itemize}
+    \item Louvain modularity, community sizes
+    \item Degree and flows distribution (average, hierarchy, entropy)
+    \item Correlations (degree-size, flow-distance)
+\end{itemize}
 
@@
 % ok desolee je n'avais pas bien compris 
+% javais mal explique ^-^
 
-\textit{Following \cite{raimbault2018space}, geosimulation models must be studied within synthetic controllable urban contexts (spatial sensitivity)}
+\justify
+
+\textit{Following \cite{raimbault2018space}, geosimulation models must be studied within synthetic controllable urban contexts in order to (i) understand intrinsic behavior of the model and robust qualitative stylized facts; (ii) study the sensitivity to the spatial configuration}
 
 \bigskip
+
+Generation of a continent-scale urban system with stylized order of magnitude corresponding to Europe:
+
+\medskip
+
+\footnotesize
+
+\begin{enumerate}
+    \item Generate $N=700$ cities with size following a power law $E_i \eq E_0 \cdot i^{-\alpha}$ with $E_0 \eq 10^{11}$ and $\alpha=1.1$ (computed on Europe for GDP with cities larger than 50.000 inhabitants)
+    \item Distribute them randomly in space (\cite{simini2019testing} vs \cite{banos2011christaller})
+    \item Create countries with k-means clustering ($C \eq 30$)
+    \item 
+\end{enumerate}
 
@@
 
-\sframe{Implementation}{
-
-% for now no dynamics of cities
+\sframe{Implementation and experiments}{
 
@@
 	\item Model implemented in NetLogo (good compromise interactivity / ergonomy), with fast data structures (matrix/table extensions)
-	\item Integrated into OpenMOLE \cite{reuillon2013openmole} for model exploration
+	\item Integrated seamlessly into OpenMOLE \cite{reuillon2013openmole} for model exploration (NetLogoTask)
+\end{itemize}
+
+% for now no dynamics of cities
 	\item Current implementation: only network dynamics
-\end{itemize}
 
@@
 
-	\textit{Networks at $t=1500$, for default parameters values and high gravity (left) and low gravity (right)}
+	\textit{Networks at $t = 1500$, for default parameters values and high gravity (left) and low gravity (right)}
 
@@
 
+    \centering
+
+    \includegraphics[width=\textwidth]{figures/internationalization_countryGravityDecay200_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+}
+
+\sframe{Grid exploration}{
+
+    \centering
+
+    \includegraphics[width=\textwidth]{figures/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+}
+
+\sframe{Grid exploration}{
+
+    \centering
+
+    \includegraphics[width=\textwidth]{figures/rhoFlowDistance_countryGravityDecay2100_gammaDestination2_facetwrapgammaOrigin_colorgammaSectors.png}
 }

### 2019-09-26 slides.tex
Update on Overleaf.

@@
 
-\item Cities are defined by their GDP and their profile regarding the proportion of firms in the different sectors. Links between cities are created in an iterative way, taking into account:
+$\rightarrow$ Cities are defined by their GDP and their profile regarding the proportion of firms in the different sectors. Links between cities are created in an iterative way, taking into account:
 
@@
 
-\footnotesize
+\tiny
 
 \begin{enumerate}
-    \item Generate $N=700$ cities with size following a power law $E_i \eq E_0 \cdot i^{-\alpha}$ with $E_0 \eq 10^{11}$ and $\alpha=1.1$ (computed on Europe for GDP with cities larger than 50.000 inhabitants)
+    \item Generate $N=700$ cities with size following a power law $E_i \= E_0 \cdot i^{-\alpha}$ with $E_0 \= 10^{11}$ and $\alpha=1.1$ (computed on Europe for GDP with cities larger than 50.000 inhabitants)
     \item Distribute them randomly in space (\cite{simini2019testing} vs \cite{banos2011christaller})
     \item Create countries with k-means clustering ($C \eq 30$)
-    \item 
+    \item Distribute sectors such that (i) smaller cities are more specialized and (ii) larger cities are more knowledge-based, with a one dimensional axis to position sectors $1/K \ldots 1$ where the density $f\left(k\right)$ follows a log-normal with $\left(\mu,\sigma\right)$ such that $\sqrt{\Var f} = K/2$ for the largest, $\sqrt{\Var f} = 1/K$ for the smallest
 \end{enumerate}
@@
 % for now no dynamics of cities
-	\item Current implementation: only network dynamics
+
+\begin{itemize}
+	\item Current experiment: only network dynamics (short time scale)
+	\item One-factor sampling with 100 repetitions to assess statistical properties
+	\item Grid sampling with 20 repetitions for model behavior (2years 5month of computation)
+\end{itemize}
 

### 2019-09-30 slides.tex
Update on Overleaf.
.

@@
 \begin{enumerate}
-    \item Generate $N=700$ cities with size following a power law $E_i \= E_0 \cdot i^{-\alpha}$ with $E_0 \= 10^{11}$ and $\alpha=1.1$ (computed on Europe for GDP with cities larger than 50.000 inhabitants)
+    \item Generate $N \textrm{ = } 700$ cities with size following a power law $E_i \textrm{ = } E_0 \cdot i^{-\alpha}$ with $E_0 \textrm{ = } 10^{11}$ and $\alpha=1.1$ (computed on Europe for GDP with cities larger than 50.000 inhabitants)
     \item Distribute them randomly in space (\cite{simini2019testing} vs \cite{banos2011christaller})
@@
 
-\texttt{}% dataverse
+\texttt{https://doi.org/10.7910/DVN/UPX23S}
 

### 2019-10-02 slides.tex
slides_v1

--- 2019-09-26 slides.tex
@@
 \begin{itemize}
-    \item Cities as maximisation of socio-economic interactions 
-    % que veux tu dire par maximisation ? je mettrais plutot "foster socio-eco"
+    \item Cities as cross-overs of socio-economic interactions 
     \item Network economies \cite{sassen1991global} \cite{castells1996information}
@@
 
-% examples de resultats pour l'Europe Centrale et Orientale
-
  \begin{center}
@@
 
-% si besoin je peux rajouter d'autres exemples sur le UK ou carrement des trucs de Celine sur les reseaux de firmes
-% oui ca serait cool des trucs a toi sur le UK
+\sframe{Firm linkages structuring urban systems}{
 
-\sframe{Urban Firm linkages and geo-economic processes}{
+ \begin{center}
+ \includegraphics[width=0.7\textwidth]{figures/Figure2.jpg}
+ \end{center}
+
+  \textit{UK cities within transnational ownership links of firms in 2016}
+
+}
+
+% bon je ne sais pas si c'est utile mais au moins ca fait un exempple en plus (bon ok on ne voit rien)
+
+\sframe{Urban firm linkages and geo-economic processes}{
 
@@
 
-How can we capture geographical and economic processes within urban networks of firms with a generative model?
+$\rightarrow$ \textit{How can we capture geographical and economic processes within urban networks of firms with a generative model?}
 
-}
-\sframe{Towards a generative model of urban firm networks}{
-
-Towards a generative model of urban firm networks
-
-% research question / objective
-%
+\begin{itemize}
+	\item Indirect inference on processes
+	\item Includes path-dependency
+\end{itemize}
 
@@
 %  => oui c'est exactement ce quil fallait mettre !
-
+% question a la con, mais est-ce que ce n'est pas tautolgique de dire que les villes et les liens sont definis par le GDP, ou le lien est bien defini par le GDP des noeuds.
 }
@@
 
+% excuses-moi, mais je vais commencer a partir de la a te demander de m'expliquer, ca va te paraitre evident je suis desolee, mais c'est pour que je ne trompe pas en presentant
+% a quoi correspond le i et j? En gros c'est ville i, puis la ville j? Le k c'est bien le nombre de secteurs? 
+% -> oui exactement
+
 \bigskip
@@
 \begin{enumerate}
-    \item Evolve city sizes $E_i(t+1) = f(E_j(t))$ with an interaction model (\cite{raimbault2018indirect} or \cite{cottineau2015growing})
+    \item Evolve city sizes $E_i \left(t \textrm{ + } 1 \right) \textrm{ = } f\left(E_j\left(t\right)\right)$ with an interaction model (\cite{raimbault2018indirect} or \cite{cottineau2015growing})
+    % ici le = n'apparait pas dans la formule. le t c'est bien le temps?
+    %  => oui t est le temps ; les = et + et ( ) qui passent pas c'est la template ucl, j'arrive pas a fixer alors il faut passer par le reswitch en mode texte
     \item Add a fixed number of links randomly, following a probability function of sizes, sector proximity, and geographical and socio-cultural proximity
@@
 
-where $E=\sum_k E_k$, $W = \sum_{i,j} w_{ij}$, $s$ is cosine similarity, $d_{ij}$ euclidian distance, and $g_{ij}$ socio-cultural distance
+where $E \textrm{ = } \sum_k E_k$, $W \textrm{ = } \sum_{i,j} w_{ij}$, $s$ is a proximity measure given by cosine similarity, $d_{ij}$ euclidian distance, and $g_{ij}$ a socio-cultural distance
 
-\medskip
+\bigskip
+\bigskip
 
-\textbf{Model parameters: } $\gamma_F, \gamma_T, \gamma_W, \gamma_S, \gamma_D = \frac{1}{d_G}, \gamma_G = \frac{1}{d_G}$
+\textbf{Model parameters: } $\gamma_F, \gamma_T, \gamma_W, \gamma_S, \gamma_D \textrm{ = } \frac{1}{d_G}, \gamma_G \textrm{ = } \frac{1}{d_G}$
 
@@
 
-% la ou avant il faut surement expliquer que c'est pas totalement synthetique, que tu pars de X villes europeennes de plus de 50 000 habitants (pour les plus nuls comme moi)
-% si c'est completement synthetiques, l'histoire des villes europeenes c'est pour avoir les ordres de grandeur de la hierarchie et du nombre de villes, mais c'est pas parametrise sur des vraies donnees - mais oui ca ca sera explique dans comment construire le systeme de ville synthetique.
-
-% ok desolee je n'avais pas bien compris 
-% javais mal explique ^-^
-
 \justify
@@
 \begin{enumerate}
-    \item Generate $N=700$ cities with size following a power law $E_i \eq E_0 \cdot i^{-\alpha}$ with $E_0 \eq 10^{11}$ and $\alpha=1.1$ (computed on Europe for GDP with cities larger than 50.000 inhabitants)
+    \item Generate $N \textrm{ = } 700$ cities with size following a power law $E_i \textrm{ = } E_0 \cdot i^{-\alpha}$ with $E_0 \textrm{ = } 10^{11}$ and $\alpha \textrm{ = } 1.1$ (computed on Europe for GDP with cities larger than 50.000 inhabitants)
     \item Distribute them randomly in space (\cite{simini2019testing} vs \cite{banos2011christaller})
-    \item Create countries with k-means clustering ($C \eq 30$)
-    \item 
+    \item Create countries with k-means clustering ($C \textrm{ = } 30$)
+    \item Distribute sectors such that (i) smaller cities are more specialized and (ii) larger cities are more knowledge-based, with a one dimensional axis to position sectors $1/K \ldots 1$ where the density $f\left(k\right)$ follows a log-normal with $\left(\mu,\sigma\right)$ such that $\sqrt{\Var f} \textrm{ = } K/2$ for the largest, $\sqrt{\Var f} \textrm{ = } 1/K$ for the smallest
 \end{enumerate}
@@
 
+\textbf{Implementation}
+
 \begin{itemize}
@@
 
+\medskip
+
+\textbf{Experiments}
+
 % for now no dynamics of cities
-	\item Current implementation: only network dynamics
+
+\begin{itemize}
+	\item Current experiment: only network dynamics (short time scale)
+	\item One-factor sampling with 100 repetitions to assess statistical properties (good convergence, average sharpe ratios for indicators all larger than 5)
+	\item Grid sampling with 20 repetitions for model behavior %(2years 5months of computation)
+\end{itemize}
 
@@
 
-    \centering
-
+    \begin{center}
     \includegraphics[width=0.48\textwidth]{figures/ex_alleq-highgravity_seed-12102_t1500.png}\hspace{0.1cm}
     \includegraphics[width=0.48\textwidth]{figures/ex_alleq-lowgravity_seed-12102_t1500.png}
+	\end{center}
 
@@
 
-	\textit{Networks at $t = 1500$, for default parameters values and high gravity (left) and low gravity (right)}
+	\textit{Networks at $t \textrm{ = } 1500$, for default parameters values and high gravity (left) and low gravity (right)}
 
@@
 
-\sframe{Statistical convergence}{
+%\sframe{Statistical convergence}{
 
-$\rightarrow$ Good convergence for most indicators: average sharpes ratios on a one-factor sampling above 5 (except for metropolization index $\simeq 1.5$)
+%$\rightarrow$ Good convergence for most indicators: average sharpes ratios on a one-factor sampling above 5 (except for metropolization index $\simeq 1.5$)
 
@@
 
-\bigskip
+%\bigskip
+%}
 
-}
-
-\sframe{Interaction decay}{
+\sframe{Effect of interaction decay}{
 
@@
     \includegraphics[width=0.48\textwidth]{figures/rhoDegreeSize-gravityDecay_errorbars.png}
+
+    \medskip
+
+    \footnotesize
+
+    \textit{(Left) Internationalization index decreases exponentially with gravity decay; (Right) Correlation between city weighted degree and size.
+Both plots show a transition from a local to a global regime.}
+
 }
 
-\sframe{Sector proximity}{
+% alors la je ne comprend pas desolee l'interpretation de l'evolution des courbes, l'internationalisation (qui est defini par la modularite diminue avec le temp?
+% et le p(di,E) proba entre distance et poids economique? comment tu interpretes ca? il y a un seuil a la fin
+% du coup je ne comprend pas bien comment interpreter
+%  => c'est pas dans le temps, c'est en fonction du parametre de gravity ; internationalization comme c'est la modularite elle est minimisee plus c'est internataional
+
+\sframe{Effect of sector proximity}{
 
@@
 
+    \medskip
+    \footnotesize
+
+    \textit{(Left) Internationalization varies linearly with sector proximity $\gamma_S$; (Right) Correlation between degree and size exhibits a maximum, witnessing an intermediate regime where size is the most important}
+
 }
 
-\sframe{Grid exploration}{
+\sframe{Grid exploration: internationalization}{
 
-    \centering
+    \begin{center}
+    \includegraphics[width=\textwidth]{figures/internationalization_countryGravityDecay200_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+    \end{center}
 
-    \includegraphics[width=\textwidth]{figures/internationalization_countryGravityDecay200_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+    \medskip
+
+    \footnotesize
+
+    \textit{The transition as a function of interaction range depends on the influence of origin size $\gamma_F$; sector proximity $\gamma_S$ plays a role only for a large influence of the origin.}
+
 }
 
-\sframe{Grid exploration}{
+% alors tu as explique pendant la presentqtion pour Mike, mais je n'ai pas compris pourquoi il a trois graphiques 0 1 2 et pourquoi le 3 est si different:?
+% je pense qu'il faut reconmmer en un truc plus clair l'axe des ordonnees et les 0, 1, 2
+% dG c'est quoi, c'est toujours le pqs de temps? donc la meme chose que le gravity decay?
 
-    \centering
+\sframe{Grid exploration: community size}{
 
-    \includegraphics[width=\textwidth]{figures/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+    \begin{center}
+    	\includegraphics[height=0.6\textheight]{figures/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+    \end{center}
+
+    \vspace{-0.5cm}
+
+    \footnotesize
+
+\begin{itemize}
+	\item Maximal integration in term of community size is achieved at an intermediate value of $d_G$: emergence of a regional regime
+	\item Maximal size depends on the role of sectors $\gamma_S$, in a decreasing way when origin size is deactivated, and increasing way when $\gamma_F=1$
+	\item This regime disappear when origin size influence is too large
+\end{itemize}
+
 }
 
-\sframe{Grid exploration}{
+% alors ici il faut reconnomer l'axe des ordonnees ca peut etre community size?
+qui est defini comment d'ailleurs ici?
+C'est la ou tu as dit que le pic aussi rapide est interessant, donc en si peu de temps la taille du reseaux grossi c'est ca :'(?
+et du coup comment expliquer le 2 a droite?
 
-    \centering
+%\sframe{Grid exploration}{
+%    \centering
+%
+%    \includegraphics[width=\textwidth]{figures/rhoFlowDistance_countryGravityDecay2100_gammaDestination2_facetwrapgammaOrigin_colorgammaSectors.png}
+%}
 
-    \includegraphics[width=\textwidth]{figures/rhoFlowDistance_countryGravityDecay2100_gammaDestination2_facetwrapgammaOrigin_colorgammaSectors.png}
-}
+% idem la c'est pas facile a lire pour quelqu'un qui ne connait pas. c'est quoi rhoFlowDistance?
+% le gamma sector c'est le parametre de la variable secteurs donc s? 
+% en fait quand j'y pense pourquoi il yq u gradien du gamma et comment comprendre ces varations pour chqaue pqs de temps
 
@@
 
-\sframe{On the role of model exploration}{
+\sframe{Discussion}{
 
-}
-
-\sframe{Discussion}{
+\justify
+\footnotesize
 
@@
 
+$\rightarrow$ effect of exogenous shocks in the socio-economic structure (``subsystem-xit'')
+
+\medskip
+
 \textbf{Developments}
+
+$\rightarrow$ evolution of city sizes (co-evolution model)
+
+$\rightarrow$ parametrization/calibration on real data
+
+$\rightarrow$ role of path-dependency
+
+$\rightarrow$ towards a model with firm agents? (multi-scale ABM)
+
+\medskip
+
+\textbf{On the role of model exploration}
+
+$\rightarrow$ even with such a ``simple'' model (close to directly tractable stationary state), behavior is highly non-linear in many dimensions
+
+$\rightarrow$ model exploration allows to overcome hidden parameters (deactivated mechanisms or default parameter values)
+
+$\rightarrow$ exploration of intrinsic dynamics on synthetic data is a crucial step before an application on real data (disentangle effects of geography from model dynamics)
 
@@
 
-$\rightarrow$ A simulation model to understand processes of network emergence
+$\rightarrow$ A generative model to understand processes of economic network emergence
 
-$\rightarrow$
+\bigskip
+
+$\rightarrow$ Crucial role of model exploration to validate and extract knowledge from such a simulation model
 
@@
 
-\textbf{Simulation data at}
+\bigskip
 
-\texttt{}% dataverse
+\textbf{Simulation data at} \texttt{https://doi.org/10.7910/DVN/UPX23S}
 
@@
 \textbf{Acknowledgments}: thanks to the \textit{European Grid Infrastructure} for access to the infrastructure.
-
-}
-
-\sframe{Reserve slides}{
-
-\centering
-
-\Large
-
-\textbf{Reserve Slides}
 

### 2019-10-16 main.tex
Update on Overleaf.
.

@@
     \item The ``geopolitical/sociocultural'' distance remains abstract and should be parametrized (see setup below) and estimated in the data-driven setting.
-    \item \textcolor{red} TODO the path dependency $w_{ij}$ could (or should ?) be decomposed into in/out degree for origin and destination ? (then stronger self-reinforcment effect)
+    \item %\textcolor{red}
+    TODO the path dependency $w_{ij}$ could (or should ?) be decomposed into in/out degree for origin and destination ? (then stronger self-reinforcment effect)
     \end{itemize}
@@
 ! links depend on the sectors: cobb douglas for sector proximity should depend on sector composition ? or several functions ?
+
+\subsubsection{Real data setup}
+
+Data sources:
+\begin{itemize}
+	\item ORBIS database for company links
+	\item GHS UCDB for GDP of urban built-up areas
+	\item Functional urban areas (JRC-EC) for consistent ontology and spatial join between the different sources
+\end{itemize}
+
+If the links are ``complete'' within Europe, the model (i) can be initialized at $t_0$ with the network corresponding to the first date in the dataset; (ii) adjustment error is computed on the later dates.
+
+More probably, we may have to work on an extract of the ORBiS database including only links from and to UK-based companies
 

### 2019-10-16 main.tex
Update on Overleaf.

@@
 We examine the interactions of european cities within firm linkages.
+
+\cite{dai2016generative} simpler urban network generative model
 
@@
 \end{itemize}
+
+\subsection{Statistical analysis}
+
+% TODO stat models justifying processes included:
+% - poisson regression if we count a number of links?
+% - country couples fixed effect to estimate the country distance matrix (as a function of interaction range ?)
+% - multilevel modeling ?
+% - distribution and properties (spatial stats ?) of sector distributions and similarities
 
@@
 \begin{itemize}
-	\item ORBIS database for company links
+	\item AMADEUS database for company links
 	\item GHS UCDB for GDP of urban built-up areas
 	\item Functional urban areas (JRC-EC) for consistent ontology and spatial join between the different sources
+	\item Cities GDP can also be computed by aggregating company data: correlation between this approach and the GHS one must be checked
 \end{itemize}
@@
 
-More probably, we may have to work on an extract of the ORBiS database including only links from and to UK-based companies
+More probably, we may have to work on an extract of the AMADEUS database including only links from and to UK-based companies. This constraints imposes modifications on the level to remain consistent:
+\begin{itemize}
+    \item the adjustment error can only be partial and computed on the UK links only
+    \item an initialization method for the missing links must be implemented (e.g. through a statistical analysis? not far from the model then)
+    \item the model can be run without initial links, and an additional parameter sets the duration necessary to reach the time considered as the first date
+\end{itemize}
 
@@
 
+The model is implemented in NetLogo and integrated into the OpenMOLE platform for model exploration \cite{reuillon2013openmole}.
+
 \subsection{Synthetic city systems}
@@
 
+\subsection{Real application to Europe}
+
+% Q: do we do some "hybrid experiments"? - ex synthetic system of cities but real sector distribution
+
 \bibliographystyle{apalike}

### 2019-10-27 main.tex
Update on Overleaf.
.

@@
 
-\textit{find some indicators for polarization : when big cities take everything or when local clusters organize}
-
-Network indicators:
+\subsubsection{Geographical indicators:}
 
 \begin{itemize}
-    \item weighted degree distribution (\textit{if it is fat-tailed, log-log power law exponent and adjusted r-squared})
+    \item Internationalisation (modularity of countries in the network)
+    \item Metropolisation (correlation between weighted degree and city size)
+    \item Regionalisation (correlation between length and flow of links, stratified by size of extremities)
+    \item Specialisation (correlation between sector proximity and flow of links, stratified by size of extremities)
 \end{itemize}
 
-Sector structure indicators:
+\subsubsection{Network and flows indicators:}
+
+%  networkModularity, networkAvgCommunitySize, networkDegreeHierarchy, networkAvgDegree, networkDegreeEntropy, flowsHierarchy, flowsEntropy, rhoDegreeSize, rhoFlowDistancePos, rhoFlowDistance
 
 \begin{itemize}
-    \item sector diversity (Herfindhal ?) 
+    \item Louvain modularity, community sizes
+    \item Degree and flows distribution (average, hierarchy, entropy)
+    \item Correlations (degree-size, flow-distance)
 \end{itemize}
+
+% possible other indics:
+% weighted degree distribution (\textit{if it is fat-tailed, log-log power law exponent and adjusted r-squared})
+%Sector structure indicators: sector diversity (Herfindhal ?)
 
@@
 The model is implemented in NetLogo and integrated into the OpenMOLE platform for model exploration \cite{reuillon2013openmole}.
+NetLogo provides a good compromise between performance and interactive prototyping and exploration.
+%We use in particular data structures providing good performances for the type of operations needed (matrices operations for the construction of the utility function e.g.).
+
+%%%%%%%%%%%%%
+\begin{figure}
+    \includegraphics[width=0.48\textwidth]{figures/ex_alleq-highgravity_seed-12102_t1500.png}\hspace{0.1cm}
+    \includegraphics[width=0.48\textwidth]{figures/ex_alleq-lowgravity_seed-12102_t1500.png}	
+    \caption{Example of simulated networks at $t=1500$, with a high (resp. low) gravity decay parameter on the left (resp. right).\label{fig:example}}
+\end{figure}
+%%%%%%%%%%%%%
 
@@
 \subsubsection{Sensitivity analysis and grid exploration}
+
+\paragraph{One factor sampling}
+% 20190924_162740_ONEFACTOR_REPLICATIONS_SYNTHETIC_GRID
+
+\begin{figure}
+	\includegraphics[width=0.48\textwidth]{figures/internationalization-gravityDecay_errorbars.png}
+    \includegraphics[width=0.48\textwidth]{figures/rhoDegreeSize-gravityDecay_errorbars.png}\\
+
+\end{figure}
 

### 2019-11-12 main.tex

@@
 
+% possible target JOEG? https://academic.oup.com/joeg/
+
 \section{Introduction}
@@
 
-% We compare the model to real data on transnational firm networks defined by ownership links at city-level, by the means of the Bureau Van Dijk's \emph {FAME} database with a graph theory application.
+\cite{martinus2019brokerage} role of small countries
+
+\cite{dawley2019creating} regional policies for insertion in global network
+
+\cite{turkina2016structure} networks of industrial clusters
+
+\cite{gluckler2016relational} positioning within networks
+
+We compare the model to real data on transnational firm networks defined by ownership links at city-level, by the means of the Bureau Van Dijk's \emph {FAME} database with a graph theory application.
 
@@
 
+% stats to describe the structure of the database, nb of total links, nodes, missing values? + nb of firms per country,  total turnover generated per country etc..
+
 % TODO stat models justifying processes included:
@@
 % - distribution and properties (spatial stats ?) of sector distributions and similarities
+
+\subsection{Network analysis}
+
+% - network properties of European network (FUAs) in time -> centralities (! directed links), weighted in/out degree
+
+% - correlation with city size etc (check indics of the simulation model?)
+
+% - correlation FUAs gdp from ucdb / gdp by aggregating firms
 
@@
 
-More probably, we may have to work on an extract of the AMADEUS database including only links from and to UK-based companies. This constraints imposes modifications on the level to remain consistent:
-\begin{itemize}
-    \item the adjustment error can only be partial and computed on the UK links only
-    \item an initialization method for the missing links must be implemented (e.g. through a statistical analysis? not far from the model then)
-    \item the model can be run without initial links, and an additional parameter sets the duration necessary to reach the time considered as the first date
-\end{itemize}
+%%%
+% with new dataset we do not have a subsample of companies within a specific country anymore
+%More probably, we may have to work on an extract of the AMADEUS database including only links from and to UK-based companies. This constraints imposes modifications on the level to remain consistent:
+%\begin{itemize}
+%    \item the adjustment error can only be partial and computed on the UK links only
+%    \item an initialization method for the missing links must be implemented (e.g. through a statistical analysis? not far from the model then)
+%    \item the model can be run without initial links, and an additional parameter sets the duration necessary to reach the time considered as the first date
+%\end{itemize}
 
@@
 
+\begin{table}
+    \caption{List of parameters, associated processes, and ranges.}
+    \begin{center}
+    \begin{tabular}{|c|c|c|c|}
+    \hline
+    Parameter & Name & Process & Range \\\hline
+      $\gamma_F$ & Origin &  & $\left[0;10\right]$ \\
+    $\gamma_T$ & Destination & & $\left[0;10\right]$ \\
+    $\gamma_W$ & Previous links & Self-reinforcement & $\left[0;10\right]$  \\
+    $\gamma_D = \frac{1}{d_G}$ & Geographical distance & & $\left[0;10000\right]$ \\
+    $\gamma_G  = \frac{1}{d_G}$ & Socio-cultural distance & & $\left[0;10000\right]$ \\\hline
+    \end{tabular}
+    \end{center}
+\end{table}
+
 \section{Results}
@@
 %    R package : https://github.com/tnagler/VineCopula
+% Q: what would be the meaning on parametrizing interaction function with a real Vine copula? non-linear interactions? (on log, linear combination of factors)
 
@@
 \begin{figure}
-    \includegraphics[width=0.48\textwidth]{figures/ex_alleq-highgravity_seed-12102_t1500.png}\hspace{0.1cm}
+    \includegraphics[width=0.48\textwidth]{figures/ex_alleq-highgravity_seed-12102_t1500.png}
     \includegraphics[width=0.48\textwidth]{figures/ex_alleq-lowgravity_seed-12102_t1500.png}	
@@
 %%%%%%%%%%%%%
+
+\cite{railsback2017improving} netlogo perfs
 
@@
 	\includegraphics[width=0.48\textwidth]{figures/internationalization-gravityDecay_errorbars.png}
-    \includegraphics[width=0.48\textwidth]{figures/rhoDegreeSize-gravityDecay_errorbars.png}\\
+    \includegraphics[width=0.48\textwidth]{figures/rhoDegreeSize-gravityDecay_errorbars.png}
+    \caption{(Left) Internationalization index decreases exponentially with gravity decay; (Right) Correlation between city weighted degree and size.
+Both plots show a transition from a local to a global regime.}
+\end{figure}
 
+\begin{figure}
+    \includegraphics[width=0.48\textwidth]{figures/internationalization-gammaSectors_errorbars.png}
+    \includegraphics[width=0.48\textwidth]{figures/rhoDegreeSize-gammaSectors_errorbars.png}
+    \caption{(Left) Internationalization varies linearly with sector proximity $\gamma_S$; (Right) Correlation between degree and size exhibits a maximum, witnessing an intermediate regime where size is the most important}
+\end{figure}
+
+\paragraph{Grid exploration}
+
+\begin{figure}
+    \includegraphics[width=\textwidth]{figures/internationalization_countryGravityDecay200_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+    \caption{The transition as a function of interaction range depends on the influence of origin size $\gamma_F$; sector proximity $\gamma_S$ plays a role only for a large influence of the origin.}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[height=0.6\textheight]{figures/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+	\caption{Maximal integration in term of community size is achieved at an intermediate value of $d_G$: emergence of a regional regime; Maximal size depends on the role of sectors $\gamma_S$, in a decreasing way when origin size is deactivated, and increasing way when $\gamma_F=1$; This regime disappear when origin size influence is too large}
 \end{figure}
@@
 
+%%%%%%%%%%%%%%%
+\section{Discussion}
+
+% - evolution of city sizes (co-evolution model)
+
+% - role of path dependency
+
+% - towards a model with firm agents? (multi-scale ABM)
+
 \bibliographystyle{apalike}

### 2019-11-15 extended_abstract.tex
Update on Overleaf.

\documentclass[10pt,letterpaper]{article}
\usepackage[margin=1in]{geometry}
%\usepackage{natbib}
\usepackage{multicol}
\usepackage{graphicx}
\pagenumbering{gobble}
\usepackage{color}

\newcommand{\ea}[1]{{\color{blue}  #1}}

\title{\vspace{-2cm}A generative model for urban firm networks}
\author{J. Raimbault$^{1,2,3}$, N. Zdanowska$^{1,3}$ and E. Arcaute$^1$\\\medskip\small
$^{1}$ Centre for Advanced Spatial Analysis, UCL; $^{2}$ UPS CNRS 3611 ISC-PIF; $^{3}$ UMR CNRS 8504 G{\'e}ographie-cit{\'e}s
}
\date{}

\begin{document}

\maketitle

%\begin{abstract}
%    We introduce a generative network model for urban networks of interactions between firms. The growth of the network is simulated by adding new firm ownership links randomly, with probabilities determined by several factors, namely the economic size of origin and destination urban areas, industrial sector proximity, the strength of previous links, and geographical and socio-cultural distance. Simulation on synthetic systems of cities unveil phase transitions when changing interaction distance, and non-trivial patterns in model behavior. Future work include the calibration of the model on real company links for Europe.
%\end{abstract}

%\vspace{0.5cm}

The emergence of a global network of cities, defined by the interactions between firms, is a salient feature of recent globalization processes \cite{taylor2001specification}.  
\ea{Understanding the drivers for growth for such a network, can be used to foster innovation through} 
%Understanding drivers of this urban network growth have several concrete and theoretical implications, including 
policies for local economic clusters \cite{turkina2016structure}, \ea{development of strategic industries for the global insertion of regions}
%regional insertion into the global network 
\cite{dawley2019creating}, 
%strategic positioning of %
cities \cite{gluckler2016relational} and
countries \cite{martinus2019brokerage}.\ea{Linkages are characterized by interactions of different nature and determined by economic, socio-cultural, or geopolitical proximity \cite{martinus2018global}.}
%The role of different types of proximity, such as economic, socio-cultural, or geopolitical proximity, in the establishment of urban linkages have been shown empirically \cite{martinus2018global}.
%The type of linkage can furthermore be of different nature, such as firm linkages, services, economic exchanges \cite{pan2017mapping}.

This paper focuses on economic relations between urban areas defined by ownership links of firms. Since transnational firms structure is one of the determinant of the global economic space, it is one of the proxies to unveil geographical structures. \cite{2019arXiv191014652Z}. 
\ea{This work} introduces a generative network model to simulate the growth of such linkages at urban areas scale within the evolutionary systems of cities framework \cite{pumain2018evolutionary}. 
%Existing generative urban network models generally operate at a more abstract level with spatial interactions \cite{dai2016generative} or with other types of processes such as economic exchanges \cite{cottineau2015growing}. The model is at the level of urban areas.
We combine multiple factors which influence link formation, including economic intensity of origin and destination of urban areas, industrial sector proximity, the strength of previous links, and geographical and socio-cultural distance. The model allows thus to compare the effect of different factors on the final network structure, and to extrapolate their actual effect when calibrated on real data.

%\section{Generative network model}

Formally, the model grows links between cities which are characterized by their economic size $E_i$ (GDP) and their economic structure $S_{ik}$ in terms of activity sectors (probability distribution of firms within $K$ sectors). Starting with an existing network or no links, the model iteratively adds links, following a probability given by a generalized Cobb-Douglas function \cite{vilcu2011geometric} as 
\[
p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_F} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_T} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s\left(S_{ik},S_{jk}\right)^{\gamma_S} \cdot \exp \left(- \gamma_G \cdot d_{ij}\right) \cdot \exp \left(- \gamma_D \cdot g_{ij}\right)
\]
where $E  =  \sum_k E_k$, $W  = \sum_{i,j} w_{ij}$ weights of previous links, \ea{$s(S_{ik},S_{jk})$ is similarity function between activity sectors $S_{ik}$ and $S_{jk}$}, $d_{ij}$ euclidian distance, and $g_{ij}$ a socio-cultural distance. When a link already exists between two areas, its weight is incremented by one. The model is stopped either when a final time is reached or when a given number of links have been added. This formulation can be understood both as an economic utility, but also as a generalization of spatial interaction models. Model parameters are the Cobb-Douglas weights $\gamma_F,\gamma_T,\gamma_W,\gamma_s$ and \ea{distance}
%gravity 
decay parameters $\gamma_G,\gamma_D$. Note that the model includes path-dependency through the influence of previous links, and thus can not be reduced to a simple statistical model. The role of space furthermore introduces more complexity, yielding a generative model which has to be simulated.

We consider several macroscopic indicators to quantify the generated network. First, geographical structures are captured by (i) internationalisation (modularity of countries in the network); (ii) metropolisation (correlation between weighted degree and city size); (iii) regionalisation (correlation between length and flow of links, stratified by size of extremities); and (iv) Specialisation (correlation between sector proximity and flow of links, stratified by size of extremities). We also consider network indicators including Louvain modularity, community sizes, degree and link weight distributions (described by their average, hierarchy, and entropy), and correlations between degree and city size, and between link weight and distance.

%\section{Results}

The model is implemented in NetLogo and integrated into the OpenMOLE platform for model exploration to distribute the computation on high performance computing infrastructures. We first apply the model on synthetic systems of cities, at a continent scale with properties to resemble Europe: (i) 700 cities with an empirical power-law from the Global Human Settlements database for economic sizes; (ii) clustered into 30 countries; (iii) with synthetic distribution of sectors following log-normal distributions adjusted such that large cities have more high-value activities and are more diverse. \ea{The} network is generated from an empty initial network until \ea{reaching} $t_f=1500$. A first numerical experiment of one-factor sampling on all Cobb-Douglas parameters and 100 stochastic repetitions confirms the good statistical convergence (average Sharpe ratios for indicators all larger than 5). \ea{The} model behaviour is then studied with a grid sampling for parameters and 20 repetitions (\ea{the} computations run on the European Grid Infrastructure, \ea{and are} equivalent to 2.5 years of CPU time).

We show in Fig. 1 examples of generated networks, with very contrasted patterns obtained with different spatial interaction ranges. This behavior is confirmed when plotting the internationalization index against 
%gravity, 
\ea{the distance}, which shows an exponential decay. Similarly, the correlation between city weighted degree and city size exhibit a behaviour that can be interpreted as a phase transition between local and global networks. Other indicators exhibit non-trivial patterns. For example, when considering average community size of the final network, we obtain a maximal integration in term of communities at an intermediate value of the gravity decay, what can be interpreted as the emergence of a regional regime. 
%this next sentence is very difficult to understand. I'm not sure it's worth leavbing it in the abstract:
%The maximal average furthermore size depends on the role of sectors $\gamma_S$, in a decreasing manner when origin size is deactivated ($\gamma_F=0$), and increasing when it is activated ($\gamma_F=1$), witnessing a complex interaction between processes through this qualitative inversion.
\ea{The size of communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.}

\ea{As a matter of validation, the model is implemented for real data for Europe}
%An application on real data for Europe is currently being implemented, 
joining Functional Urban Areas with the Global Human Settlement Database for cities and their characteristics, with the AMADEUS database produced by the Bureau Van Dijk on ownership links between firms in Europe between 2010 and 2018 at postcode level in all types of sectors.

%\section{Future work}

\ea{Several scenarios are considered taking into consideration the current and future economic challenges, such as those given by Brexit.}
%Several model extensions and applications can be considered.

\begin{figure}
\vspace{-2cm}
\begin{center}
    \begin{minipage}[c]{0.25\textwidth}
        \includegraphics[width=\textwidth]{figures/ex_alleq-lowgravity_seed-12102_t1500.png}\\
        \includegraphics[width=\textwidth]{figures/ex_alleq-highgravity_seed-12102_t1500.png}
    \end{minipage}
    \begin{minipage}[c]{0.6\linewidth}
     \includegraphics[width=0.48\textwidth]{figures/internationalization-gravityDecay_errorbars.png}
    \includegraphics[width=0.48\textwidth]{figures/rhoDegreeSize-gravityDecay_errorbars.png}\\
    \includegraphics[width=\textwidth]{figures/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
    \end{minipage}
    \end{center}
    \caption{\footnotesize\textbf{(Left column)} Example of simulated networks for a synthetic system of cities, with low gravity range (top) and high gravity range (bottom); \textbf{(Top right)} Internationalization and correlation between size and degree as a function of \ea{distance for different decay parameters}; \textbf{(Bottom right)}}
\end{figure}

\footnotesize

\begin{multicols}{2}

\bibliographystyle{unsrt}
\bibliography{biblio}

\end{multicols}

\end{document}

### 2019-11-15 extended_abstract.tex
Update on Overleaf.

@@
 
-\title{\vspace{-2cm}A generative model for urban firm networks}
+\title{\vspace{-2.5cm}A generative model for urban firm networks}
 \author{J. Raimbault$^{1,2,3}$, N. Zdanowska$^{1,3}$ and E. Arcaute$^1$\\\medskip\small
@@
 
-%\vspace{0.5cm}
+\vspace{-0.5cm}
 
-The emergence of a global network of cities, defined by the interactions between firms, is a salient feature of recent globalization processes \cite{taylor2001specification}.  
-\ea{Understanding the drivers for growth for such a network, can be used to foster innovation through} 
-%Understanding drivers of this urban network growth have several concrete and theoretical implications, including 
-policies for local economic clusters \cite{turkina2016structure}, \ea{development of strategic industries for the global insertion of regions}
-%regional insertion into the global network 
-\cite{dawley2019creating}, 
-%strategic positioning of %
-cities \cite{gluckler2016relational} and
-countries \cite{martinus2019brokerage}.\ea{Linkages are characterized by interactions of different nature and determined by economic, socio-cultural, or geopolitical proximity \cite{martinus2018global}.}
-%The role of different types of proximity, such as economic, socio-cultural, or geopolitical proximity, in the establishment of urban linkages have been shown empirically \cite{martinus2018global}.
-%The type of linkage can furthermore be of different nature, such as firm linkages, services, economic exchanges \cite{pan2017mapping}.
+The emergence of a global network of cities, defined by interactions between firms, is a salient feature of recent globalization processes \cite{taylor2001specification}.
+Understanding the drivers for growth for such a network, can be used to foster innovation through policies for local economic clusters \cite{turkina2016structure}, development of strategic industries for global insertion of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
+Linkages are characterized by interactions of different nature and determined by economic, socio-cultural, or geopolitical proximity \cite{martinus2018global}.
 
 This paper focuses on economic relations between urban areas defined by ownership links of firms.
-Since transnational firms structure is one of the determinant of the global economic space, it is one of the proxies to unveil geographical structures.
-\cite{2019arXiv191014652Z}.
-\ea{This work} introduces a generative network model to simulate the growth of such linkages at urban areas scale within the evolutionary systems of cities framework \cite{pumain2018evolutionary}. 
-%Existing generative urban network models generally operate at a more abstract level with spatial interactions \cite{dai2016generative} or with other types of processes such as economic exchanges \cite{cottineau2015growing}.
-The model is at the level of urban areas.
-We combine multiple factors which influence link formation, including economic intensity of origin and destination of urban areas, industrial sector proximity, the strength of previous links, and geographical and socio-cultural distance.
+Since transnational firms structure is one of the determinant of the global economic space, it is one of the proxies to unveil geographical structures \cite{2019arXiv191014652Z}.
+This work introduces a generative network model to simulate the growth of such linkages at urban areas scale within the evolutionary systems of cities framework \cite{pumain2018evolutionary}.
+We combine multiple factors which influence link formation, as economic intensity of origin and destination of urban areas, industrial sector proximity, the strength of previous links, and geographical and socio-cultural distance.
 The model allows thus to compare the effect of different factors on the final network structure, and to extrapolate their actual effect when calibrated on real data.
 
-%\section{Generative network model}
-
-Formally, the model grows links between cities which are characterized by their economic size $E_i$ (GDP) and their economic structure $S_{ik}$ in terms of activity sectors (probability distribution of firms within $K$ sectors).
-Starting with an existing network or no links, the model iteratively adds links, following a probability given by a generalized Cobb-Douglas function \cite{vilcu2011geometric} as
+Formally, the model creates links between cities which are characterized by their economic size $E_i$ (GDP) and economic structure $S_{ik}$ in terms of activity sectors (probability distribution of firms within $K$ sectors).
+Starting with an existing network with no links, the model iteratively adds links, following a probability given by a generalized Cobb-Douglas function \cite{vilcu2011geometric} as
 \[
@@
 \]
-where $E  =  \sum_k E_k$, $W  = \sum_{i,j} w_{ij}$ weights of previous links, \ea{$s(S_{ik},S_{jk})$ is similarity function between activity sectors $S_{ik}$ and $S_{jk}$}, $d_{ij}$ euclidian distance, and $g_{ij}$ a socio-cultural distance.
-When a link already exists between two areas, its weight is incremented by one.
+where $E  =  \sum_k E_k$, $W  = \sum_{i,j} w_{ij}$ weights of previous links, $s(S_{ik},S_{jk})$ is similarity function between activity sectors $S_{ik}$ and $S_{jk}$, $d_{ij}$ euclidian distance, and $g_{ij}$ a socio-cultural distance.
+In the case of an already existing link between two areas, the weight of the latter is incremented by one.
 The model is stopped either when a final time is reached or when a given number of links have been added.
-This formulation can be understood both as an economic utility, but also as a generalization of spatial interaction models.
-Model parameters are the Cobb-Douglas weights $\gamma_F,\gamma_T,\gamma_W,\gamma_s$ and \ea{distance}
-%gravity 
-decay parameters $\gamma_G,\gamma_D$.
-Note that the model includes path-dependency through the influence of previous links, and thus can not be reduced to a simple statistical model.
-The role of space furthermore introduces more complexity, yielding a generative model which has to be simulated.
+This formulation can be considered both as an economic utility and a generalization of spatial interaction models.
+The model parameters are the Cobb-Douglas weights $\gamma_F,\gamma_T,\gamma_W,\gamma_s$ and distance decay parameters $\gamma_G,\gamma_D$.
+The model also includes path-dependency by integrating the influence of previous links.
+Therefore it can not be reduced to a simple statistical model.
+Finally, the role of space introduces even further complexity, yielding a generative model, which has to be simulated.
 
 We consider several macroscopic indicators to quantify the generated network.
-First, geographical structures are captured by (i) internationalisation (modularity of countries in the network); (ii) metropolisation (correlation between weighted degree and city size); (iii) regionalisation (correlation between length and flow of links, stratified by size of extremities); and (iv) Specialisation (correlation between sector proximity and flow of links, stratified by size of extremities).
-We also consider network indicators including Louvain modularity, community sizes, degree and link weight distributions (described by their average, hierarchy, and entropy), and correlations between degree and city size, and between link weight and distance.
+Firstly, geographical structures are captured by (i) internationalisation (modularity of countries in the network); (ii) metropolisation (correlation between weighted degree and city size); (iii) regionalisation (correlation between length and flow of links, stratified by size of extremities); and (iv) Specialisation (correlation between sector proximity and flow of links, stratified by size of extremities).
+We also consider several network indicators as Louvain modularity, community sizes, degree and link weight distributions (described by their average, hierarchy, and entropy), and correlations between degree and city size, and between link weight and distance.
 
-%\section{Results}
-
-The model is implemented in NetLogo and integrated into the OpenMOLE platform for model exploration to distribute the computation on high performance computing infrastructures.
-We first apply the model on synthetic systems of cities, at a continent scale with properties to resemble Europe: (i) 700 cities with an empirical power-law from the Global Human Settlements database for economic sizes; (ii) clustered into 30 countries; (iii) with synthetic distribution of sectors following log-normal distributions adjusted such that large cities have more high-value activities and are more diverse.
-\ea{The} network is generated from an empty initial network until \ea{reaching} $t_f=1500$.
-A first numerical experiment of one-factor sampling on all Cobb-Douglas parameters and 100 stochastic repetitions confirms the good statistical convergence (average Sharpe ratios for indicators all larger than 5).
-\ea{The} model behaviour is then studied with a grid sampling for parameters and 20 repetitions (\ea{the} computations run on the European Grid Infrastructure, \ea{and are} equivalent to 2.5 years of CPU time).
+The model is implemented in NetLogo and integrated into the OpenMOLE platform for model exploration to distribute the computation on high performance infrastructures.
+We build the model on a synthetic system of cities, at the scale of a continent with properties similar to Europe: (i) 700 cities with an empirical power-law from the Global Human Settlements database for economic sizes; (ii) clustered into 30 countries; (iii) with synthetic distribution of sectors following log-normal distributions adjusted in a way that large cities have more high-value activities and are more diverse.
+The network is generated from an empty initial network until reaching $t_f=1500$.
+A first numerical experiment of one-factor sampling on all Cobb-Douglas parameters and 100 stochastic repetitions confirms a good statistical convergence (average Sharpe ratios for indicators all larger than 5).
+The model behaviour is then studied with a grid sampling for parameters and 20 repetitions (the computations run on the European Grid Infrastructure, and are equivalent to 2.5 years of CPU time).
 
@@
 1 examples of generated networks, with very contrasted patterns obtained with different spatial interaction ranges.
-This behavior is confirmed when plotting the internationalization index against
-%gravity, 
-\ea{the distance}, which shows an exponential decay.
+This behavior is confirmed when plotting the internationalization index against the distance, which shows an exponential decay.
 Similarly, the correlation between city weighted degree and city size exhibit a behaviour that can be interpreted as a phase transition between local and global networks.
@@
 For example, when considering average community size of the final network, we obtain a maximal integration in term of communities at an intermediate value of the gravity decay, what can be interpreted as the emergence of a regional regime.
-%this next sentence is very difficult to understand. I'm not sure it's worth leavbing it in the abstract:
-%The maximal average furthermore size depends on the role of sectors $\gamma_S$, in a decreasing manner when origin size is deactivated ($\gamma_F=0$), and increasing when it is activated ($\gamma_F=1$), witnessing a complex interaction between processes through this qualitative inversion.
-\ea{The size of communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.}
+The size of communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.
+Altogether, these experiments reveal a complex interplay between processes and how the model produces diverse stylized facts.
 
-\ea{As a matter of validation, the model is implemented for real data for Europe}
-%An application on real data for Europe is currently being implemented, 
-joining Functional Urban Areas with the Global Human Settlement Database for cities and their characteristics, with the AMADEUS database produced by the Bureau Van Dijk on ownership links between firms in Europe between 2010 and 2018 at postcode level in all types of sectors.
+As a matter of validation, the model is implemented for real data for Europe joining Functional Urban Areas with the Global Human Settlement Database for cities and their characteristics, with the AMADEUS database produced by the Bureau Van Dijk on ownership links between firms in Europe between 2010 and 2018 at postcode level in all types of sectors.
+Calibration is done using a genetic NSGA2 algorithm, finding the parameter values giving network trajectories closest to reality.
+First results show in particular a role of the path-dependency parameter confirming the relevance of this complex generative model.
+Further work is needed to get more conclusions from this experiment but we already show the feasibility of model calibration on real data.
 
-%\section{Future work}
-
-\ea{Several scenarios are considered taking into consideration the current and future economic challenges, such as those given by Brexit.}
-%Several model extensions and applications can be considered.
+Several scenarios are considered taking into consideration the current and future economic challenges, such as those given by Brexit.
+This way, the model may be applied in future work to concrete policy questions.
 
@@
     \end{center}
-    \caption{\footnotesize\textbf{(Left column)} Example of simulated networks for a synthetic system of cities, with low gravity range (top) and high gravity range (bottom); \textbf{(Top right)} Internationalization and correlation between size and degree as a function of \ea{distance for different decay parameters}; \textbf{(Bottom right)}}
+    \vspace{-0.5cm}
+    \caption{\footnotesize\textbf{(Left column)} Example of simulated networks for a synthetic system of cities, with low gravity range (top) and high gravity range (bottom); \textbf{(Top right)} Internationalization and correlation between size and degree as a function of decay parameter; \textbf{(Bottom right)} Average network community size as a function of decay parameter, for varying influence of sectors ($\gamma_S$ in color) and of origin ($\gamma_F$ in columns).}
 \end{figure}

### 2019-11-15 extended_abstract.tex
Update on Overleaf.

@@
 First results show in particular a role of the path-dependency parameter confirming the relevance of this complex generative model.
-Further work is needed to get more conclusions from this experiment but we already show the feasibility of model calibration on real data.
-
+Further work is needed to get more interpretations from this experiment, but our study demonstrates the feasibility of model calibration on real data.
 Several scenarios are considered taking into consideration the current and future economic challenges, such as those given by Brexit.
-This way, the model may be applied in future work to concrete policy questions.
+The model may be applied in future work to assess policy issues and recommendations.
 

### 2019-11-15 extended_abstract.tex
Update on Overleaf.
.

@@
 
-\title{\vspace{-2.5cm}A generative model for urban firm networks}
+\title{\vspace{-2.7cm}A generative model for urban firm networks}
 \author{J. Raimbault$^{1,2,3}$, N. Zdanowska$^{1,3}$ and E. Arcaute$^1$\\\medskip\small
@@
 Simulation on synthetic systems of cities unveil phase transitions when changing interaction distance, and non-trivial patterns in model behavior.
-Future work include the calibration of the model on real company links for Europe.
+We also give first results of  calibrating the model on real company links for Europe using a genetic algorithm.
 %\end{abstract}
+
+% Keywords
+% Urban Firm Networks
+% Generative Network Model
+% Model Exploration
 
@@
 The size of communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.
-Altogether, these experiments reveal a complex interplay between processes and how the model produces diverse stylized facts.
+All these experiments reveal a complex interplay between processes and how the model produces diverse stylized facts.
 

### 2020-03-04 main.tex
Update on Overleaf.

--- 2019-10-16 main.tex
@@
-\documentclass{article}
+\documentclass[11pt]{article}
 \usepackage[utf8]{inputenc}
-\usepackage[margin=2cm]{geometry}
+\usepackage[margin=1in]{geometry}
 \usepackage{natbib}
+\usepackage{graphicx}
+\usepackage{mdframed}
 
@@
 
+\author{J. Raimbault$^{1,2,3}$, N. Zdanowska$^{1,3}$ and E. Arcaute$^1$\\\medskip\small
+$^{1}$ Centre for Advanced Spatial Analysis, UCL; $^{2}$ UPS CNRS 3611 ISC-PIF; $^{3}$ UMR CNRS 8504 G{\'e}ographie-cit{\'e}s
+}
+
 \maketitle
+% NetSci format: A4, 11pt, 1in margin, max 12p + up to 3p references
+
+\begin{abstract}
+    This paper introduces a generative model for urban networks defined by interactions between firms.
+The aim is to test several geographical properties emerging in such a network as metropolisation, internationalisation or regionalisation.
+The growth of the network is simulated by adding randomly new links defined by ownership between firms, with probabilities of occurring depending on the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past as well as the geographical and socio-cultural distance.
+The simulation on a synthetic system of cities unveil phase transitions when changing interaction distance, and non-trivial patterns in the model behavior.
+Calibration of the model on real data on company ownership linkages for Europe in 2018 is also undertaken.
+Future work will consist in observing thanks to this model the impact of economic crisis on such networks.
+
+\end{abstract}
 
@@
 
-\citep{taylor2001specification} world city network driven by interactions between firms/behavior of firms
-\citep{martinus2018global} combination of different types of proximity in inter-urban firm networks: economic, sociocultural, geopolitical - implies spatial and non-spatial 
-\citep{pan2017mapping} different types of linkages: e.g. services here, but not only. All sectors should be taken into account first.
+The intensification of globalisation of the economy since the late 20th century resulted in the rise of the network society \citep{castells2000networksociety} and the emergence of a world-city system \citep{taylor2001specification}, characterized by highly interconnected urban centres at the global level \citep{sassen1991}.
+These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity implying spatial and non spatial patterns \citep{martinus2018global}.
+Interactions of economic nature between transnational firms, are together with trade among the most representative of the current economic trends \citep{taylor2001specification}.
+Understanding the drivers of growth and geographical properties of such a urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \citep{turkina2016structure}.
+Straightening these city-interactions can be crucial for revitalising certain regions \cite{Clarke2018}, by developing strategic industries for global insertion of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
+The functioning of such a city-network and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future exogenous shocks such as an economic crisis, a single market collapse or a post-Brexit scenario for the European Union.
 
-The originality of the present paper is to approach these issues within the system of cities framework \citep{berry1964cities}, where the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \citep{pumain1997pour}.
-We examine the interactions of european cities within firm linkages.
+The use of network models to understand processes underlying the emergence and dynamics of such urban networks has already been widely tackled in the literature.
+\cite{dai2016simulating} propose a generative urban network model at the regional scale, combining geographical factors with network topological properties.
+Air transportation networks can also be understood as urban networks, and \cite{zanin2013modelling} reviews several models for such networks.
+Trade network have been extensively studied from the viewpoint of complex networks, but are generally driven at the country-level \cite{garlaschelli2005structure}.
+Input-output models, mostly used in Regional science to represent flows between geographical areas, are considered some cases a type of urban generative model \cite{jin1993generation}.
+In that stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for abstract flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2020indirect}.
 
-\cite{dai2016generative} simpler urban network generative model
+%\citep{martinus2018global} combination of different types of proximity in inter-urban firm networks: economic, sociocultural, geopolitical - implies spatial and non-spatial 
+%\citep{pan2017mapping} different types of linkages: e.g. services here, but not only. All sectors should be taken into account first.
+%\cite{martinus2019brokerage} role of small countries
+%\cite{dawley2019creating} regional policies for insertion in global network
+%\cite{gluckler2016relational} positioning within networks
 
-% We compare the model to real data on transnational firm networks defined by ownership links at city-level, by the means of the Bureau Van Dijk's \emph {FAME} database with a graph theory application.
+The originality of the present paper is to approach these issues within the system of cities framework \citep{berry1964cities}, where the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}.
+We examine the interactions of European cities within firm linkages defined by corporate ownership links.
+Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geographical structures \cite{2019arXiv191014652Z}.
+More precisely, our contribution relies on the following points: (i) we propose an empirical analysis of the firm ownership urban network for the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database; (ii) we introduce a generative network model to simulate the growth of such linkages at urban areas scale, which combines multiple factors influencing link formation, as economic intensity of origin and destination of urban areas, industrial sector proximity, the strength of previous links, and geographical and socio-cultural distance; (iii) the model allows us to compare the effect of different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; and (iv) we calibrate the model on the empirical network.
+
+The rest of this paper is organized as follows: we first proceed to an empirical analysis of the network studied.
+We then describe the generative model for urban networks, and its sensitivity analysis and exploration on synthetic data.
+We then parametrize and calibrate the model on real European firm ownership data.
+We finally discuss implications of our results and possible extensions.
+
+\section{Empirical network analysis}
+
+% stats to describe the structure of the database, nb of total links, nodes, missing values? + nb of firms per country,  total turnover generated per country etc..
+
+Companies from the AMADEUS database are georeferenced using the Geonames database (using postcode or address depending on the information available).
+We then join this data with the GHS-FUA dataset for functional urban areas \cite{Florczyk2019ghs}, which are the relevant objects for applying our model.
+Starting from a firm-level dataset of 1,562,862 nodes and  1,866,936 links, we obtain a directed network of 573 urban areas and 9323 ownership links.
+Weight of links are obtained by computing the owned share of turnover at destination, i.e.
+$w_{ij} = \sum_{k \in i,l \in j} s_{kl} \cdot T_l$ where $s_{kl}$ of the ownership share of company $l$ by company $k$ and $T_l$ the turnover of company $l$.
+Node attributes include an economic size of urban areas, that we compute as the sum of turnovers of companies within the area.
+This quantity is highly correlated with area-level GDP included in the GHS database (Spearman correlation $\rho = 0.71$) and also with population ($\rho = 0.60$), so we expect to capture size effects while keeping the consistence of a single main data source.
+An other attribute of nodes is the industrial sector composition, expressed as a proportion of turnover in the area associated to a given industrial sector.
+Detailed sectors of firms are available up to the fourth digit for the NACE classification in the raw data.
+We consider the first level classification (21 categories) and compute sector proportions accordingly.
+
+The urban network has heavy tail weighted degree and edge weight distributions, as shown in Fig.~\ref{fig:nwdist}.
+We fit power law and log-normal distributions, including minimal value cutoff following \cite{clauset2009power}, using the \texttt{powerlaw} R package \cite{powerlawpackage}.
+For both, log-normal distributions appear to be a better fit ($\mu=18.8$, $\sigma=2.3$ for weighted degree; $\mu=13.8$, $\sigma=2.8$ for edge weight), and include a much broader range of the empirical distributions with lower estimated cut-offs.
+These heavy tail properties suggest the relevance of a generative model including self-reinforcing processes which are known to produce such distributions.
+
+%%%%%%%%%%%%%
+\begin{figure}
+    \centering
+    \includegraphics[width=0.49\textwidth]{figures/edgeweight.png}
+    \includegraphics[width=0.49\textwidth]{figures/degreeDistr.png}
+    \caption{Distributions of network properties.
+\textit{(Left)} Cumulated distribution for edge weight, with power law fit (red) and log-normal fit (green); \textit{(Right)} Cumulated distribution for weighted degree.}
+    \label{fig:nwdist}
+\end{figure}
+%%%%%%%%%%%%%
+
+Regarding the community structure of the network, different modularity maximization algorithms (run on the undirected corresponding network) give a same undirected modularity of 0.38, and a directed modularity \citep{nicosia2009extending} of 0.34 for the greedy algorithm of \citep{clauset2004finding} and 0.36 for the Louvain algorithm \citep{blondel2008fast}, corresponding to 36 communities (resp.
+15).
+In comparison, with 29 countries, the classification of urban areas by country has a modularity of 0.32.
+This means that the structure of flows has a certain regional integration.
+Indeed, a null model assigning random labels with 30 communities, bootstrap 1000 times, yields an average modularity of $0.049 \pm 0.002$, confirming the statistical significance of these communities.
+
+To have a first insight into factors playing a role in the formation of links, we proceed to a statistical analysis which is similar to unconstrained spatial interaction models \cite{wilson1975some}.
+% poisson \cite{flowerdew1988fitting}
 
@@
 
-\begin{itemize}
-    \item Single level model: scale of cities - firms induce the links but are not represented in the model
-    \item Cities are defined by their profile regarding the proportion of firms in the different sectors
-    \item Links between cities are created in an iterative way, with the combination of several drivers:
-    \begin{itemize}
-        \item geographical proximity (distance or effective accessibility depending on the scale)
-        \item geopolitical proximity (belonging to the same ``administrative unit'' $\simeq$ country or single market $\rightarrow$ application to Brexit)
-        \item city size (taking not population but an economic variable related to firms avoids adding the dimension of population, simplifies the model) $\rightarrow$ can be combined with geographical distance with a classical interaction model, see \cite{cottineau2015growing}, \cite{raimbault2018indirect}, \cite{favaro2011gibrat}
-        \item economic proximity (e.g. cosine distance between sector proximity as done in \cite{2019arXiv190505106C}
-        \item previous linkages (taking into account history of previous firm linkages)
-    \end{itemize}
-    \item These competing drivers are combined - how ?
-\textit{Options: (i) multiplicative function to get probabilities; (ii) generalization with a multi-attribute Cobb-Douglas; (iii) multi-objective optimization depending on sectors, using Pareto fronts ?
-- needs empirical evidence; (iv) copula-based combination as done by \cite{2019arXiv190505106C} - copula parameters are free or estimated from real data ?}
-    \item The accumulated effects and therefore path-dependency are taken into account into this dynamical model
-\end{itemize}
-
-\subsection{Statistical analysis}
-
-% TODO stat models justifying processes included:
-% - poisson regression if we count a number of links?
-% - country couples fixed effect to estimate the country distance matrix (as a function of interaction range ?)
-% - multilevel modeling ?
-% - distribution and properties (spatial stats ?) of sector distributions and similarities
+The model is expected to capture a single macroscopic urban scale, i.e.
+the level of the urban system where basic entities are cities.
+Links are induced by underlying firms but these are not explicit.
+Cities are defined by an economic size, but also by an industrial sector profile.
+As \citep{martinus2018global} point out, a combination of several distances may play a role in establishing linkages.
+Geographical distance and industrial proximity and complementarity will typically be in concurrence \cite{cottineau2020nested}.
+Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalized to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets despite the European common market (confirmed relevant by the fixed effect regression described above); (iii) city economic size, which has furthermore scaling laws properties; (iv) industrial proximity in terms of firm sector composition; (v) previous linkages (influence of history).
+This last point is crucial to capture accumulated effects and include path-dependency, which fits well the use of a simulation model instead of a more simple statistical model.
 
@@
 Let $1 \leq i \leq N$ cities defined at time $t$ by their economic structure such that $E_i(t)$ is the total economic volume (GDP) for city $i$ and $S_{ik}(t)$ is a matrix giving economic volume proportions within each economic sector $k$, assuming $K$ economic sectors.
+Formally, the model creates links between cities which are characterized by their economic size $E_i$ (GDP) and economic structure $S_{ik}$ in terms of activity sectors (probability distribution of firms within $K$ sectors).
+Starting with an existing network with no links, the model iteratively adds links, following a probability given by a generalized Cobb-Douglas function \cite{vilcu2011geometric} as
 
-% Q : evolution of volume / populations ? Populations not needed indeed.
-
-The city network evolves at discrete time steps the following way:
-\begin{enumerate}
-    \item Economic structure of cities are evolved with a city interaction model which can be
-    \begin{itemize}
-        \item a direct interaction model, according to
 \begin{equation}
-            \Delta E_i = g_0\cdot E_i + w_G \cdot \sum_j V_{ij}/<V_{ij}>
-        \end{equation}
-        \item the Marius model with economic exchanges
-    \end{itemize}
-    \textit{Baseline: no evolution - on short time spans may be relevant ?}
-    \item A new linkage is established
-    \begin{itemize}
-        \item \textit{rq: as these are done sequentially, it is equivalent to have a fixed time step and a number of linkages per time step, or to do it link by link - in the data-driven version the actual values shall be parametrized, while in the synthetic version this has no importance}
-        \item the link is drawn randomly with a probability for it to be $i \rightarrow j$ proportional to a function
-    \begin{equation}
-        p_{ij} = f(d_{ij},g_{ij},E_i,E_j,s(S_{ik},S_{jk}),w_{ij})
-    \end{equation}
-    with $d_{ij}$ the geographical distance, $g_{ij}$ the geopolitical/sociocultural distance, $g$ a similarity function between sector structure, $w_{ij}$ the relative size of already existing links.
-    \item A simple multiplicative form between the different terms may be too restrictive as it assumes independence between components and an equal share between each.
-Copula capture a dependency structure but must be parametrized by estimation on real data (\textit{keep it as an option for the data-driven model}).
-An other option is to use a generalized Cobb-Douglas function \citep{vilcu2011geometric}, where exponents can be kept as free parameters.
-Furthermore, a natural formulation as a separable gravity model as for cities above is already a particular case of such a function, and taking an exponential gravity potential yields that the decay exponent corresponds to the Cobb-Douglas exponent.
-We therefore take
-    \begin{equation}
-     p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_F} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_T} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s(S_{ik},S_{jk})^{\gamma_S} \cdot \exp \left(- \gamma_D \cdot d_{ij}\right) \cdot \exp \left(- \gamma_G \cdot g_{ij}\right)
+p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_F} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_T} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s\left(S_{ik},S_{jk}\right)^{\gamma_S} \cdot \exp \left(- \gamma_G \cdot d_{ij}\right) \cdot \exp \left(- \gamma_D \cdot g_{ij}\right)
 \end{equation}
 
-    \item Note that: (i) we take an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalize the exponents to 1 (which is different than renormalizing the probability), to ensure to include convex functions.
-    \item We take for the sake of simplicity (\textit{should be amended/validated/tested with an asymmetric function in a further version}) we take a cosine similarity between the sector structures, given with such probability vectors (size effect is already taken into account) by
+where $E  =  \sum_k E_k$, $W  = \sum_{i,j} w_{ij}$ weights of previous links, $s(S_{ik},S_{jk})$ is similarity function between activity sectors $S_{ik}$ and $S_{jk}$, $d_{ij}$ euclidian distance, and $g_{ij}$ a socio-cultural distance.
+In the case of an already existing link between two areas, the weight of the latter is incremented by one.
+The model is stopped either when a final time is reached or when a given number of links have been added.
+This formulation can be considered both as an economic utility and a generalization of spatial interaction models.
+
+We take a cosine similarity between the sector structures, given with such probability vectors (size effect is already taken into account) by
 \begin{equation}
@@
 
-    \item The ``geopolitical/sociocultural'' distance remains abstract and should be parametrized (see setup below) and estimated in the data-driven setting.
-    \item %\textcolor{red}
-    TODO the path dependency $w_{ij}$ could (or should ?) be decomposed into in/out degree for origin and destination ? (then stronger self-reinforcment effect)
-    \end{itemize}
-\end{enumerate}
+Note that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalize the exponents to 1 (which is different than renormalizing the probability), to ensure to include convex functions.
+The ``geopolitical/sociocultural'' distance remains abstract and should be parametrized (see setup below) and estimated in the data-driven setting.
+
+We assume no evolution of economic size in this simple version of the model, which means that adjusted parameter are valid on a restricted time period
 
@@
 
-\subsubsection{Synthetic setup}
+\paragraph{Synthetic setup}
 
-Let consider the simplest synthetic setup possible:
+We build the model on a synthetic system of cities, at the scale of a continent with properties similar to Europe: (i) 700 cities with an empirical power-law from the Global Human Settlements database for economic sizes; (ii) clustered into 30 countries; (iii) with synthetic distribution of sectors following log-normal distributions adjusted in a way that large cities have more high-value activities and are more diverse.
+The network is generated from an empty initial network until reaching $t_f=1500$.
+
+A simple synthetic setup is considered.
 \begin{itemize}
@@
 
-! links depend on the sectors: cobb douglas for sector proximity should depend on sector composition ? or several functions ?
+\subsection{Model parameters and indicators}
 
-\subsubsection{Real data setup}
+The model parameters are the Cobb-Douglas weights $\gamma_F,\gamma_T,\gamma_W,\gamma_s$ and distance decay parameters $\gamma_G,\gamma_D$.
+The model also includes path-dependency by integrating the influence of previous links.
+Therefore it can not be reduced to a simple statistical model.
+Finally, the role of space introduces even further complexity, yielding a generative model, which has to be simulated.
 
-Data sources:
-\begin{itemize}
-	\item AMADEUS database for company links
-	\item GHS UCDB for GDP of urban built-up areas
-	\item Functional urban areas (JRC-EC) for consistent ontology and spatial join between the different sources
-	\item Cities GDP can also be computed by aggregating company data: correlation between this approach and the GHS one must be checked
-\end{itemize}
+We consider several macroscopic indicators to quantify the generated network.
+Firstly, geographical structures are captured by (i) internationalisation (modularity of countries in the network); (ii) metropolisation (correlation between weighted degree and city size).
+Two other thematic indicators, which are regionalisation (correlation between length and flow of links, stratified by size of extremities), and specialisation (correlation between sector proximity and flow of links, stratified by size of extremities), are not included in this study for the sake of simplicity as each has a high dimension.
+We also consider several network indicators as Louvain modularity, sizes of maximal modularity communities, degree and link weight distributions (described by their average, hierarchy, and entropy), and correlations between degree and city size, and between link weight and distance.
 
-If the links are ``complete'' within Europe, the model (i) can be initialized at $t_0$ with the network corresponding to the first date in the dataset; (ii) adjustment error is computed on the later dates.
+\section{Model exploration and calibration}
 
-More probably, we may have to work on an extract of the AMADEUS database including only links from and to UK-based companies. This constraints imposes modifications on the level to remain consistent:
-\begin{itemize}
-    \item the adjustment error can only be partial and computed on the UK links only
-    \item an initialization method for the missing links must be implemented (e.g. through a statistical analysis? not far from the model then)
-    \item the model can be run without initial links, and an additional parameter sets the duration necessary to reach the time considered as the first date
-\end{itemize}
+The model is implemented in NetLogo, which provides a good compromise between performance and interactive prototyping and exploration \cite{railsback2017improving}.
+Numerical experiments are done by integrating the model into the OpenMOLE software for model exploration \cite{reuillon2013openmole}, which provides sensitivity analysis and calibration methods and a transparent access to high-performance computing infrastructures.
 
-\subsection{Model indicators}
+Model source code, aggregated network data (raw initial firm network can not be made available for data ownership reasons), and results are available on the open git repository of the project at \url{https://github.com/JusteRaimbault/ABMCitiesFirms}.
+Large simulation result files are available on the dataverse repository at \url{https://doi.org/10.7910/DVN/UPX23S}.
 
-\textit{find some indicators for polarization : when big cities take everything or when local clusters organize}
+We show in Fig.~\ref{fig:example} examples of generated networks, with very contrasted patterns obtained with different spatial interaction ranges.
 
-Network indicators:
-
-\begin{itemize}
-    \item weighted degree distribution (\textit{if it is fat-tailed, log-log power law exponent and adjusted r-squared})
-\end{itemize}
-
-Sector structure indicators:
-
-\begin{itemize}
-    \item sector diversity (Herfindhal ?) 
-\end{itemize}
-
-\subsection{Model parameters}
-
-\section{Results}
-
-\subsection{Implementation}
-
-% Netlogo for vis + scala into spatialdata / urbangrowth ?
-
-% On vine copulas :
-%    https://link.springer.com/article/10.1023/A:1016725902970
-%    https://www.groups.ma.tum.de/en/statistics/research/vine-copula-models/
-%    R package : https://github.com/tnagler/VineCopula
-
-The model is implemented in NetLogo and integrated into the OpenMOLE platform for model exploration \cite{reuillon2013openmole}.
+%%%%%%%%%%%%%
+\begin{figure}
+    \includegraphics[width=0.48\textwidth]{figures/ex_alleq-highgravity_seed-12102_t1500.png}
+    \includegraphics[width=0.48\textwidth]{figures/ex_alleq-lowgravity_seed-12102_t1500.png}\\
+    \begin{mdframed}
+    \includegraphics[width=0.48\textwidth]{figures/ex_real_highgravity_t1500.png}
+    \end{mdframed}
+    \begin{mdframed}
+    \includegraphics[width=0.48\textwidth]{figures/ex_real_lowgravity_t1500.png}
+    \end{mdframed}
+    \caption{Example of simulated networks at $t=1500$, for a synthetic system of cities (Top row) and the European urban network (Bottom row).
+, with a high (resp.
+low) gravity decay parameter for the left column (resp.
+right).\label{fig:example}}
+\end{figure}
+%%%%%%%%%%%%%
 
@@
 
-\subsubsection{Sensitivity analysis and grid exploration}
+\paragraph{Sensitivity analysis}
 
-\subsubsection{PSE}
+\paragraph{One factor sampling}
+% 20190924_162740_ONEFACTOR_REPLICATIONS_SYNTHETIC_GRID
 
-\subsubsection{A virtual case study: subsystem-xit}
+A first numerical experiment of one-factor sampling on all Cobb-Douglas parameters and 100 stochastic repetitions confirms a good statistical convergence (average Sharpe ratios for indicators all larger than 5).
+Behavior of some indicators are shown in Fig.~\ref{fig:onefactor}.
 
-\subsection{Real application to Europe}
+\begin{figure}
+	\includegraphics[width=0.48\textwidth,height=0.25\textheight]{figures/internationalization-gravityDecay_errorbars.png}
+    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figures/rhoDegreeSize-gravityDecay_errorbars.png}\\
+    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figures/internationalization-gammaSectors_errorbars.png}
+    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figures/rhoDegreeSize-gammaSectors_errorbars.png}
+    \caption{(Top Left) Internationalization index decreases exponentially with gravity decay; (Top Right) Correlation between city weighted degree and size.
+Both plots show a transition from a local to a global regime.
+(Bottom Left) Internationalization varies linearly with sector proximity $\gamma_S$; (Bottom Right) Correlation between degree and size exhibits a maximum, witnessing an intermediate regime where size is the most important.
+\label{fig:onefactor}}
+\end{figure}
 
-% Q: do we do some "hybrid experiments"? - ex synthetic system of cities but real sector distribution
+\paragraph{Grid exploration}
 
-\bibliographystyle{apalike}
+The model behaviour is then studied with a grid sampling for parameters and 20 repetitions (the computations run on the European Grid Infrastructure, and are equivalent to 2.5 years of CPU time).
+
+The influence of gravity decay parameters is confirmed when plotting the internationalization index against the distance, which shows an exponential decay.
+Similarly, the correlation between city weighted degree and city size exhibit a behaviour that can be interpreted as a phase transition between local and global networks.
+Other indicators exhibit non-trivial patterns.
+For example, when considering average community size of the final network, we obtain a maximal integration in term of communities at an intermediate value of the gravity decay, what can be interpreted as the emergence of a regional regime.
+The size of communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.
+All these experiments reveal a complex interplay between processes and how the model produces diverse stylized facts.
+
+\begin{figure}
+    \includegraphics[width=\textwidth]{figures/internationalization_countryGravityDecay200_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+    \caption{The transition as a function of interaction range depends on the influence of origin size $\gamma_F$; sector proximity $\gamma_S$ plays a role only for a large influence of the origin.}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[width=\textwidth]{figures/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+	\caption{Maximal integration in term of community size is achieved at an intermediate value of $d_G$: emergence of a regional regime; Maximal size depends on the role of sectors $\gamma_S$, in a decreasing way when origin size is deactivated, and increasing way when $\gamma_F=1$; This regime disappear when origin size influence is too large}
+\end{figure}
+
+% Saltelli exploration of the model
+% -> confirm relative importance of params and search for inhomogeneity/interaction effects
+
+\subsection{Model calibration on the European urban network}
+
+As a matter of validation, the model is implemented for real data for Europe joining Functional Urban Areas with the Global Human Settlement Database for cities and their characteristics, with the AMADEUS database produced by the Bureau Van Dijk on ownership links between firms in Europe in 2018 at postcode level in all types of sectors.
+
+\subsubsection{Model calibration}
+
+Calibration is done using a genetic NSGA2 algorithm, finding the parameter values giving network trajectories closest to reality.
+First results show in particular a role of the path-dependency parameter confirming the relevance of this complex generative model.
+Further work is needed to get more interpretations from this experiment, but our study demonstrates the feasibility of model calibration on real data.
+Several scenarios are considered taking into consideration the current and future economic challenges, such as those given by Brexit.
+The model may be applied in future work to assess policy issues and recommendations.
+
+%%%%%%%%%%%%%%%
+\section{Discussion}
+
+This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated thanks to OpenMole and calibrated on real data on ownership linkages of firms for Europe.
+The simulation on a synthetic system of cities unveil phase transitions when changing interaction distance, and non-trivial patterns in the model behavior.
+This kind of model can have practical application in the future in terms of testing the effect of exogenous shocks in the socio-economic structure as the upcoming effect of Brexit on the European market.
+
+Even in such a simple model (close to directly tractable stationary state) the behaviour is highly non-linear in many dimensions.
+Our study shows how crucial model exploration is to overcome hidden parameters (deactivated mechanisms or default parameter values).
+This paper emphasis on the fact that exploration of intrinsic dynamics on synthetic data is of great importance should be driven before an application on real data so as to put aside the geographical effects from model dynamics.
+
+Several possible extensions of the current work can be undertaken in the future as a co-evolution model with an evolution of city sizes or a model with firm agents leading to a multi-scale agent-based model.
+A more precise study of the role of path dependency can also be undertaken in the creation of the links.
+Other formulations of the model might be taken into consideration as other formulation of the combination of factors or multi-objective optimisation depending on sectors using Pareto fronts.
+
+% - other formulation of the combination of factors
+%multi-objective optimization depending on sectors, using Pareto fronts ?
+- needs empirical evidence; (iv) copula-based combination as done by \cite{2019arXiv190505106C} - copula parameters are free or estimated from real data
+% The function used can be understood as a separable spatial interaction model. As a possible extension, Copula capture a dependency structure but must be parametrised by estimation on real data.
+
+% - evolution of city sizes (co-evolution model)
+% - role of path dependency
+% - towards a model with firm agents? (multi-scale ABM)
+
+% je suis nulle avec overleaf, mais est-ce possible de mettre la biblio dans un ordre alphabétique? J'imagine qu'il y a un moyen pour le faire d'une maniere non manuelle.
+% -> switch de unsrt a apalike
+
+% \bibliographystyle{apalike}
+\bibliographystyle{unsrt}
 \bibliography{biblio}

### 2020-03-04 main.tex

@@
 Air transportation networks can also be understood as urban networks, and \cite{zanin2013modelling} reviews several models for such networks.
-Trade network have been extensively studied from the viewpoint of complex networks, but are generally driven at the country-level \cite{garlaschelli2005structure}.
-Input-output models, mostly used in Regional science to represent flows between geographical areas, are considered some cases a type of urban generative model \cite{jin1993generation}.
-In that stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for abstract flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2020indirect}.
+Trade network have been extensively studied from the viewpoint of complex networks, but are generally driven at the country-level \cite{garlaschelli2005structure} because of scarce data at city-level.
+Input-output models, mostly used in Regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}.
+In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for abstract flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2020indirect}.
 
@@
 
-% je suis nulle avec overleaf, mais est-ce possible de mettre la biblio dans un ordre alphabétique? J'imagine qu'il y a un moyen pour le faire d'une maniere non manuelle.
-% -> switch de unsrt a apalike
-
-% \bibliographystyle{apalike}
-\bibliographystyle{unsrt}
+\bibliographystyle{apalike}
+%\bibliographystyle{unsrt}
 \bibliography{biblio}

### 2020-03-04 readme_amadeus.tex

\documentclass[10pt]{article}
\usepackage[margin=2cm]{geometry}

\begin{document}

\title{Amadeus Van Dijk database}

\maketitle

\section*{Synthesis}

\subsection*{Table nodes}

\subsection*{Table links}

\subsection*{Counts}

\begin{itemize}
    \item 3,053,540 firms
    \item 2,715,188 with coordinates
    \item 2,653,434 within Europe, 2,653,034 within EU (due to the way Amadeus is constructed)
    \item 1,920,909 within FUAs
    \item 1,495,651 within FUAs with a link
    \item 753,224 within FUAs, linked and with turnover
\end{itemize}

\begin{itemize}
    \item 1,866,936 links
    \item 794,782 with an origin European country, 1,691,007 with a destination European country, 761,173 with both % most links are intercontinental ! we get with the amadeus either links from EU, or to EU -> international study interesting ? 
    \item 616,031 with an origin FUA, 1,254,374 with a destination FUAs
    \item 555,196 between FUAs % FIXME proper join
    \item 161,303 between FUAs, with ownership \% information and destination turnover
    \item when aggregating at FUA (without years), 4,978 links (on 451,584 possible, very sparse)
\end{itemize}

\subsection*{FUA Network}

Quite modular (0.35); fixed effect by country significant

\end{document}

### 2020-03-04 main.tex
Update on Overleaf.

@@
     This paper introduces a generative model for urban networks defined by interactions between firms.
-The aim is to test several geographical properties emerging in such a network as metropolisation, internationalisation or regionalisation.
+The aim is to test several geographical properties emerging in such a network as metropolisation, internationalisation and regionalisation and test the effect of specialisation on these phenomena.
 The growth of the network is simulated by adding randomly new links defined by ownership between firms, with probabilities of occurring depending on the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past as well as the geographical and socio-cultural distance.
@@
 Air transportation networks can also be understood as urban networks, and \cite{zanin2013modelling} reviews several models for such networks.
-Trade network have been extensively studied from the viewpoint of complex networks, but are generally driven at the country-level \cite{garlaschelli2005structure} because of scarce data at city-level.
+Trade network have been extensively studied from complex networks perspective, but are generally driven at the country-level \cite{garlaschelli2005structure} because of lack of data at city-level.
 Input-output models, mostly used in Regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}.
@@
 Links are induced by underlying firms but these are not explicit.
-Cities are defined by an economic size, but also by an industrial sector profile.
+Cities are defined by their economic size, but also by an industrial sector profile.
 As \citep{martinus2018global} point out, a combination of several distances may play a role in establishing linkages.
-Geographical distance and industrial proximity and complementarity will typically be in concurrence \cite{cottineau2020nested}.
+Geographical distance, industrial proximity and complementarity will typically be in concurrence \cite{cottineau2020nested}.
 Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalized to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets despite the European common market (confirmed relevant by the fixed effect regression described above); (iii) city economic size, which has furthermore scaling laws properties; (iv) industrial proximity in terms of firm sector composition; (v) previous linkages (influence of history).

### 2020-03-04 main.tex
.

--- 2020-03-04 main.tex
@@
 \usepackage[margin=1in]{geometry}
-\usepackage{natbib}
-\usepackage{graphicx}
-\usepackage{mdframed}
+\usepackage{natbib,graphicx,mdframed,hyperref,threeparttable}
 
@@
 
-\author{J. Raimbault$^{1,2,3}$, N. Zdanowska$^{1,3}$ and E. Arcaute$^1$\\\medskip\small
-$^{1}$ Centre for Advanced Spatial Analysis, UCL; $^{2}$ UPS CNRS 3611 ISC-PIF; $^{3}$ UMR CNRS 8504 G{\'e}ographie-cit{\'e}s
+\author{J. Raimbault$^{1,2,3,\ast}$, N. Zdanowska$^{1,3}$ and E. Arcaute$^1$\\\medskip\small
+$^{1}$ Centre for Advanced Spatial Analysis, UCL; $^{2}$ UPS CNRS 3611 ISC-PIF; $^{3}$ UMR CNRS 8504 G{\'e}ographie-cit{\'e}s\\
+$^{\ast}$ \texttt{juste.raimbault@polytechnique.edu}
 }
@@
 Weight of links are obtained by computing the owned share of turnover at destination, i.e.
-$w_{ij} = \sum_{k \in i,l \in j} s_{kl} \cdot T_l$ where $s_{kl}$ of the ownership share of company $l$ by company $k$ and $T_l$ the turnover of company $l$.
+$w_{ij} = \sum_{k \in i,l \in j} p_{kl} \cdot T_l$ where $p_{kl}$ of the ownership share of company $l$ by company $k$ and $T_l$ the turnover of company $l$.
 Node attributes include an economic size of urban areas, that we compute as the sum of turnovers of companies within the area.
@@
 We consider the first level classification (21 categories) and compute sector proportions accordingly.
+This allows us to define an industrial proximity matrix between urban areas, taking a cosine similarity between vectors as $s_{ij} = \sum_{k=1}^{K} S_{ik}\cdot S_{jk}$ where $S_{ik}$ are sector proportions.
 
@@
 To have a first insight into factors playing a role in the formation of links, we proceed to a statistical analysis which is similar to unconstrained spatial interaction models \cite{wilson1975some}.
+We test five different statistical models, including progressively the processes candidate to be included in the network generative model.
+The most general model is
+\[
+\log(w_{ij}) = \log(d_{ij}) + \log(T_i) + \log(T_j) + \log(s_{ij}) + \alpha_{c_i,c_j} + \varepsilon_{ij}
+\]
+Estimation results are given in Table~\ref{tab:reg}.
 % poisson \cite{flowerdew1988fitting}
+
+%%%%%%%%%%%%%
+\begin{table}[h]
+\caption{Ordinary Least Square estimations of statistical models explaining $\log(w_{ij})$ (standard errors in parenthesis).
+***, ** and * respectively indicate 0.01, 0.05 and 0.1 levels of significance.
+For the fixed effect by origin-destination pairs of countries, we give the proportion of significant coefficients (where $p<0.1$).
+Models (1) to (4) are ordinary least square models, while model (5) is a Poisson generalized linear model with logarithmic link.\label{tab:reg}}
+\medskip
+\begin{tabular}{|l|c|c|c|c|c|}
+\hline
+Model  & (1) & (2) & (3) & (4) & (5) \\ 
+\hline
+$\log(d_{ij})$ &      -0.06** (0.03) &   -0.11** (0.05)  & -0.41*** (0.02)  & -0.35*** (0.04)  &  -0.26*** ($5\cdot 10^{-6}$) \\
+$\log(T_i)$ &   &   & 0.56*** (0.01) &  0.56*** (0.01) & 0.79*** ($2\cdot 10^{-6}$) \\
+$\log(T_j)$ &     &   & 0.39*** (0.01) &  0.39*** (0.01) & 0.66***  ($1.5\cdot 10^{-6}$) \\
+$\log(s_{ij})$ &     &   &  &  0.19*** (0.07) & 0.53*** ($9\cdot 10^{-6}$)  \\
+Countries &    &  28.3\% &   &  26.5\% & 100\% \\
+\hline
+$R^2$ &       0.00059   &  0.17 & 0.21 &  0.31  &  0.56 \\
+AIC &        44304   &  43578  &  42131  & 41917 &  $4.41\cdot 10^{11}$ \\
+\hline
+\end{tabular}
+\end{table}
+%%%%%%%%%%%%%
 
@@
 
-We take a cosine similarity between the sector structures, given with such probability vectors (size effect is already taken into account) by
-\begin{equation}
-    s\left(S_{ik},S_{jk}\right) = \sum_{k=1}^{K} S_{ik} S_{jk}
-\end{equation}
+%We take a cosine similarity between the sector structures, given with such probability vectors (size effect is already taken into account) by
 
@@
 
-\bibliographystyle{apalike}
-%\bibliographystyle{unsrt}
+%\bibliographystyle{apalike}
+\bibliographystyle{unsrt}
 \bibliography{biblio}

### 2020-03-04 main.tex
stats

@@
 We test five different statistical models, including progressively the processes candidate to be included in the network generative model.
-The most general model is
+The most general statistical model is
 \[
@@
 \]
+where $\alpha_{c_i,c_j}$ is a fixed effect term for the couple of countries $c_i,c_j$.
+We use a multiplicative form by taking logarithms as it is usually done for spatial interaction models.
 Estimation results are given in Table~\ref{tab:reg}.
-% poisson \cite{flowerdew1988fitting}
+We also fit a Poisson generalized linear model with a logarithmic link on the rounded weights \cite{flowerdew1988fitting}.
+We find that distance alone (model 1) has a significant negative effect but with an explained variance close to zero.
+The second model adds fixed effects by pair of countries, for which around a third are significant, and with a better $R^2$ of 0.17.
+Adding origin and destination economic size (turnovers $T_i$ and $T$, model 3) slightly improve the explanative power, but the best model among the linear ones is the model 4 with all factors and fixed effects, both in terms of explained variance ($R^2 = 0.31$) and Akaike information criterion (what means that there is no overfitting).
+The Poisson model with all variables is much better regarding pseudo-$R^2$ and provides much more significant estimates (all fixed effects significant and much smaller standard deviations).
+Across all models, we find consistent qualitative stylized facts: (i) distance has a negative influence; (ii) economic sizes have a positive influence, with the size of the origin being more important (what is consistent with ownership links, strongly driven by decision at the headquarters); (iii) industrial proximity has a positive influence, i.e.
+similarity is more important than complementarity for this dimension; (iv) an important part of country fixed-effects are significant; and (v) the order of magnitude of the influence of each factor is roughly the same, in the sense that no process totally dominates the others.
+This suggests the relevance of including all these different dimensions in the network generative model.
 
@@
 For the fixed effect by origin-destination pairs of countries, we give the proportion of significant coefficients (where $p<0.1$).
-Models (1) to (4) are ordinary least square models, while model (5) is a Poisson generalized linear model with logarithmic link.\label{tab:reg}}
+Models (1) to (4) are ordinary least square models, while model (5) is a Poisson generalized linear model with logarithmic link.
+For this model, AIC is not given as not comparable, and the pseudo-$R^2$ is given.\label{tab:reg}}
 \medskip
@@
 $R^2$ &       0.00059   &  0.17 & 0.21 &  0.31  &  0.56 \\
-AIC &        44304   &  43578  &  42131  & 41917 &  $4.41\cdot 10^{11}$ \\
+AIC &        44304   &  43578  &  42131  & 41917 &   \\
 \hline

### 2020-03-05 main.tex
Update on Overleaf.

@@
 
-The intensification of globalisation of the economy since the late 20th century resulted in the rise of the network society \citep{castells2000networksociety} and the emergence of a world-city system \citep{taylor2001specification}, characterized by highly interconnected urban centres at the global level \citep{sassen1991}.
-These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity implying spatial and non spatial patterns \citep{martinus2018global}.
-Interactions of economic nature between transnational firms, are together with trade among the most representative of the current economic trends \citep{taylor2001specification}.
-Understanding the drivers of growth and geographical properties of such a urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \citep{turkina2016structure}.
+The intensification of globalisation of the economy since the late 20th century resulted in the rise of the network society \cite{castells2000networksociety} and the emergence of a world-city system \cite{taylor2001specification}, characterized by highly interconnected urban centres at the global level \cite{sassen1991}.
+These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity implying spatial and non spatial patterns \cite{martinus2018global}.
+Interactions of economic nature between transnational firms, are together with trade among the most representative of the current economic trends \cite{taylor2001specification}.
+Understanding the drivers of growth and geographical properties of such a urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}.
 Straightening these city-interactions can be crucial for revitalising certain regions \cite{Clarke2018}, by developing strategic industries for global insertion of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
@@
 Air transportation networks can also be understood as urban networks, and \cite{zanin2013modelling} reviews several models for such networks.
-Trade network have been extensively studied from complex networks perspective, but are generally driven at the country-level \cite{garlaschelli2005structure} because of lack of data at city-level.
+Trade network have been extensively studied from complex networks perspective, but are generally driven at the country-level \cite{garlaschelli2005structure} because of lack of city-level data.
 Input-output models, mostly used in Regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}.
@@
 
-The originality of the present paper is to approach these issues within the system of cities framework \citep{berry1964cities}, where the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}.
+In this paper we choose to focus on urban networks of firms to question how can we understand and capture geographical and economic processes as agglomeration economies \cite{Glaeser2001} characteristic to spatial economy \cite{Fujitaetal2001} within a generative model.
+The originality of the present paper is to approach these issues within the system of cities framework \cite{berry1964cities}, where the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}.
 We examine the interactions of European cities within firm linkages defined by corporate ownership links.
 Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geographical structures \cite{2019arXiv191014652Z}.
+
 More precisely, our contribution relies on the following points: (i) we propose an empirical analysis of the firm ownership urban network for the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database; (ii) we introduce a generative network model to simulate the growth of such linkages at urban areas scale, which combines multiple factors influencing link formation, as economic intensity of origin and destination of urban areas, industrial sector proximity, the strength of previous links, and geographical and socio-cultural distance; (iii) the model allows us to compare the effect of different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; and (iv) we calibrate the model on the empirical network.
@@
 
-Regarding the community structure of the network, different modularity maximization algorithms (run on the undirected corresponding network) give a same undirected modularity of 0.38, and a directed modularity \citep{nicosia2009extending} of 0.34 for the greedy algorithm of \citep{clauset2004finding} and 0.36 for the Louvain algorithm \citep{blondel2008fast}, corresponding to 36 communities (resp.
+Regarding the community structure of the network, different modularity maximization algorithms (run on the undirected corresponding network) give a same undirected modularity of 0.38, and a directed modularity \cite{nicosia2009extending} of 0.34 for the greedy algorithm of \cite{clauset2004finding} and 0.36 for the Louvain algorithm \cite{blondel2008fast}, corresponding to 36 communities (resp.
 15).
@@
 Cities are defined by their economic size, but also by an industrial sector profile.
-As \citep{martinus2018global} point out, a combination of several distances may play a role in establishing linkages.
+As \cite{martinus2018global} point out, a combination of several distances may play a role in establishing linkages.
 Geographical distance, industrial proximity and complementarity will typically be in concurrence \cite{cottineau2020nested}.
@@
 The total width of the space $d_{max}$ fixed to 3000km to be approximatively the size of Europe.
-The number of cities is fixed as $N=700$ to similarly match the European urban system structure (GHS database \citep{Florczyk2019ghs}).
-    \item city sizes are attributed with a scaling law (equivalent to have a Zipf law for population and a scaling law for economic size), following an exponent $\beta_0$ (fixed at $\beta_0 = 1.1$ in most experiments when not specified, approximate value for GDP from the GHS database, see \citep{raimbault:halshs-02284897})
+The number of cities is fixed as $N=700$ to similarly match the European urban system structure (GHS database \cite{Florczyk2019ghs}).
+    \item city sizes are attributed with a scaling law (equivalent to have a Zipf law for population and a scaling law for economic size), following an exponent $\beta_0$ (fixed at $\beta_0 = 1.1$ in most experiments when not specified, approximate value for GDP from the GHS database, see \cite{raimbault:halshs-02284897})
     % from the GHS database : scaling GDP - pop ~ 1.1 for Europe, rank size 0.94 -> 1.034 : take 1.1 to simplify - can be paramterized ?
-    \item country boundaries are constructed to obtain an approximately equal coverage of the space for a fixed number of countries, what remains consistent with non-correlated city sizes at the scale of the countries as our cities are randomly distributed in space \citep{simini2019testing} (\textit{Note: on this point this may be realistic if staying at a continental level, although we should include realistic size of countries; this paper is surely wrong when changing the dataset used and the scale of analysis, working at the scale of Mega-city regions e.g.}).
+    \item country boundaries are constructed to obtain an approximately equal coverage of the space for a fixed number of countries, what remains consistent with non-correlated city sizes at the scale of the countries as our cities are randomly distributed in space \cite{simini2019testing} (\textit{Note: on this point this may be realistic if staying at a continental level, although we should include realistic size of countries; this paper is surely wrong when changing the dataset used and the scale of analysis, working at the scale of Mega-city regions e.g.}).
 Therefore, we do a basic spatial k-means clustering with $C = 20$ countries and attribute the country accordingly.
@@
 
-This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated thanks to OpenMole and calibrated on real data on ownership linkages of firms for Europe.
+This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMole platform and calibrated on real data on ownership linkages of firms for Europe.
 The simulation on a synthetic system of cities unveil phase transitions when changing interaction distance, and non-trivial patterns in the model behavior.
@@
 Several possible extensions of the current work can be undertaken in the future as a co-evolution model with an evolution of city sizes or a model with firm agents leading to a multi-scale agent-based model.
-A more precise study of the role of path dependency can also be undertaken in the creation of the links.
+A more precise study of the role of path dependency can also be undertaken in order to understand the creation of the future links.
 Other formulations of the model might be taken into consideration as other formulation of the combination of factors or multi-objective optimisation depending on sectors using Pareto fronts.
@@
 
-%\bibliographystyle{apalike}
-\bibliographystyle{unsrt}
+\bibliographystyle{apalike}
+%\bibliographystyle{unsrt}
 \bibliography{biblio}

### 2020-03-05 main.tex
.

--- 2020-03-04 main.tex
@@
 We use a multiplicative form by taking logarithms as it is usually done for spatial interaction models.
+Furthermore, origin and/or destination constrained models are not relevant in our case as links include not only turnover but also a share of ownership.
 Estimation results are given in Table~\ref{tab:reg}.
@@
 similarity is more important than complementarity for this dimension; (iv) an important part of country fixed-effects are significant; and (v) the order of magnitude of the influence of each factor is roughly the same, in the sense that no process totally dominates the others.
-This suggests the relevance of including all these different dimensions in the network generative model.
+This suggests the relevance of including all these different dimensions in the generative network model.
 
@@
 Models (1) to (4) are ordinary least square models, while model (5) is a Poisson generalized linear model with logarithmic link.
-For this model, AIC is not given as not comparable, and the pseudo-$R^2$ is given.\label{tab:reg}}
+We compare model performance using $R^2$, Mean Square Error, and Akaike Information Criterion.
+For the Poisson model, AIC is not given as not comparable, and the pseudo-$R^2$ is given.\label{tab:reg}}
 \medskip
@@
 $R^2$ &       0.00059   &  0.17 & 0.21 &  0.31  &  0.56 \\
+MSE & 7.75 & 6.84 & 6.10 & 5.33 & 8.72 \\
 AIC &        44304   &  43578  &  42131  & 41917 &   \\
@@
 
-where $E  =  \sum_k E_k$, $W  = \sum_{i,j} w_{ij}$ weights of previous links, $s(S_{ik},S_{jk})$ is similarity function between activity sectors $S_{ik}$ and $S_{jk}$, $d_{ij}$ euclidian distance, and $g_{ij}$ a socio-cultural distance.
-In the case of an already existing link between two areas, the weight of the latter is incremented by one.
-The model is stopped either when a final time is reached or when a given number of links have been added.
-This formulation can be considered both as an economic utility and a generalization of spatial interaction models.
-
-%We take a cosine similarity between the sector structures, given with such probability vectors (size effect is already taken into account) by
+where $E  =  \sum_k E_k$, $W  = \sum_{i,j} w_{ij}$ weights of previous links, $s(S_{ik},S_{jk})$ is a similarity function between activity sectors $S_{ik}$ and $S_{jk}$ taken as a cosine similarity, $d_{ij}$ euclidian distance, and $g_{ij}$ a socio-cultural distance.
+In the case of an already existing link between two areas, the weight of the latter is incremented by a unit volume $w_0$.
+The model is stopped either when a final time is reached or when the cumulated volume of links reaches a target quantity.
+This formulation can be considered both as an economic utility model and a generalization of spatial interaction models.
 
@@
 The ``geopolitical/sociocultural'' distance remains abstract and should be parametrized (see setup below) and estimated in the data-driven setting.
-
-We assume no evolution of economic size in this simple version of the model, which means that adjusted parameter are valid on a restricted time period
+We assume no evolution of economic size in this simple version of the model, which means that adjusted parameter are valid on a restricted time period.
 
@@
 
+Besides model dynamics, its initialization (or parametrization on real data) regarding the spatial distribution, sizes, countries and industrial sector compositions, is also important to understand model behavior.
+\cite{raimbault2019space} furthermore showed that the initial spatial configuration can have significant impacts on model outcomes.
+We therefore consider two versions: one with a synthetic system of cities which can be stochastically repeated, and one other based on real data.
+
 \paragraph{Synthetic setup}
 
-We build the model on a synthetic system of cities, at the scale of a continent with properties similar to Europe: (i) 700 cities with an empirical power-law from the Global Human Settlements database for economic sizes; (ii) clustered into 30 countries; (iii) with synthetic distribution of sectors following log-normal distributions adjusted in a way that large cities have more high-value activities and are more diverse.
-The network is generated from an empty initial network until reaching $t_f=1500$.
+The synthetic system of cities is generated to be at the scale of a continent with properties similar to Europe: (i) $N=700$ cities with an empirical power-law from the Global Human Settlements database for economic sizes ($\alpha = 1.1$), distributed randomly into a uniform and isotropic square space of width $d_{max}=3000km$; (ii) clustered into $C = 20$ countries using a k-means algorithm (this remains consistent with non-correlated city sizes at the scale of the countries as our cities are randomly distributed in space \citep{simini2019testing}); (iii) with synthetic distribution of sectors following log-normal distributions adjusted in a way that large cities have more high-value activities and are more diverse.
+More precisely for this last point, assuming sectors are distributed continuously along a one-dimensional axis, cities are assumed to have a sector distribution with a log-normal density function.
+The largest city is constrained to have a standard deviation and a mode of $1/2$ and the smallest $1/K$, with a linear interpolation between the two for other cities.
+Parameters of the log-normal law are obtained by solving the corresponding two equations system for each city, and each distribution is discretized and renormalized to be a vector of probabilities.
+In synthetic experiments, the network is generated from an empty initial network until reaching $t_f=1500$, with a unit link volume of $w_0 = 1$.
 
-A simple synthetic setup is considered.
-\begin{itemize}
-    \item cities are distributed randomly in a uniform and isotropic space.
-The total width of the space $d_{max}$ fixed to 3000km to be approximatively the size of Europe.
-The number of cities is fixed as $N=700$ to similarly match the European urban system structure (GHS database \citep{Florczyk2019ghs}).
-    \item city sizes are attributed with a scaling law (equivalent to have a Zipf law for population and a scaling law for economic size), following an exponent $\beta_0$ (fixed at $\beta_0 = 1.1$ in most experiments when not specified, approximate value for GDP from the GHS database, see \citep{raimbault:halshs-02284897})
-    % from the GHS database : scaling GDP - pop ~ 1.1 for Europe, rank size 0.94 -> 1.034 : take 1.1 to simplify - can be paramterized ?
-    \item country boundaries are constructed to obtain an approximately equal coverage of the space for a fixed number of countries, what remains consistent with non-correlated city sizes at the scale of the countries as our cities are randomly distributed in space \citep{simini2019testing} (\textit{Note: on this point this may be realistic if staying at a continental level, although we should include realistic size of countries; this paper is surely wrong when changing the dataset used and the scale of analysis, working at the scale of Mega-city regions e.g.}).
-Therefore, we do a basic spatial k-means clustering with $C = 20$ countries and attribute the country accordingly.
-    \item firm sector structure: 
-    \begin{itemize}
-        \item random probabilities
-        \item uniform: then the term of proximity is cancelled, no interest
-        \item function of size: larger cities are more diverse, larger cities have more knowledge-based sectors. Test 1: log-normal with slightly increasing width and shifting mode.
-        % mode of log normal is exp(mu - sigma ^ 2) ; variance is [exp(sigma^2) - 1]*exp(2 mu + sigma ^ 2)
-        % when the pdf is 1 / (x sigma sqrt(2 pi)) * exp (- (ln x - mu) ^ 2 / 2 sigma^2 )
-        % -> for largest city stddev = sqrt(variance) = fields / 2 ; for smallest sigma = 1 / nfields
-        %  ==> solving for mu and sigma with linear mode and variance as a function of log(E_i)
-        % => sigma^2 is the unique positive root of f(X)=0 with f(X) = -3X - 2 ln(exp(X)-1) - ln(e_i)
-        % => mu = sigma^2 + ln(e_i)
-        %    ~ somehow dirty and not sure that most of the distrib for highest values is within [0,1]
-    \end{itemize}
-\end{itemize}
+\paragraph{Real setup}
+
+The real setup is done using the empirical network analyzed above.
+Urban areas are distributed according to their positions in the real geographical space, and economic size and sector composition attributes are used.
+The socio-economic distance is constructed using information from the statistical modeling, by considering fixed-effect coefficients estimated for each pair of countries.
+Initial network is empty, and the real links are taken as targets for the generated network.
+We constrain the total volume of flows by setting $w_0 = \sum_{ij} w_{ij} / t_{f}$ when the number of time steps $t_f$ is given as a parameter.
 
@@
 Numerical experiments are done by integrating the model into the OpenMOLE software for model exploration \cite{reuillon2013openmole}, which provides sensitivity analysis and calibration methods and a transparent access to high-performance computing infrastructures.
-
 Model source code, aggregated network data (raw initial firm network can not be made available for data ownership reasons), and results are available on the open git repository of the project at \url{https://github.com/JusteRaimbault/ABMCitiesFirms}.
 Large simulation result files are available on the dataverse repository at \url{https://doi.org/10.7910/DVN/UPX23S}.
-
 We show in Fig.~\ref{fig:example} examples of generated networks, with very contrasted patterns obtained with different spatial interaction ranges.
+Low values of the gravity decay parameter yield local networks around main cities, while long interaction ranges yield more integrated networks.
+We also show networks simulated in the real setting.
 
@@
     \includegraphics[width=0.48\textwidth]{figures/ex_alleq-lowgravity_seed-12102_t1500.png}\\
+    \begin{minipage}{0.48\textwidth}
     \begin{mdframed}
-    \includegraphics[width=0.48\textwidth]{figures/ex_real_highgravity_t1500.png}
+    \includegraphics[width=\textwidth]{figures/ex_real_highgravity_t1500.png}
     \end{mdframed}
+    \end{minipage}
+    \begin{minipage}{0.48\textwidth}
     \begin{mdframed}
-    \includegraphics[width=0.48\textwidth]{figures/ex_real_lowgravity_t1500.png}
+    \includegraphics[width=\textwidth]{figures/ex_real_lowgravity_t1500.png}
     \end{mdframed}
+    \end{minipage}
     \caption{Example of simulated networks at $t=1500$, for a synthetic system of cities (Top row) and the European urban network (Bottom row).

### 2020-03-05 main.tex
saltelli

@@
 \usepackage[margin=1in]{geometry}
-\usepackage{natbib,graphicx,mdframed,hyperref,threeparttable}
+\usepackage{natbib,graphicx,mdframed,hyperref,threeparttable,amssymb}
 
@@
 
+We first study extensively the behavior of the model on synthetic systems of cities, to isolate its intrinsic behavior independently of a contingent geographical situation.
+
 \paragraph{Sensitivity analysis}
 
+A global sensitivity analysis first unveils an aggregated influence of parameters on model indicators.
+We apply Saltelli's Global Sensitivity method introduced by \cite{saltelli2008global}.
+This produces values for each parameter and each output of a first order sensitivity index, defined as $Var\left[\mathbb{E}_{\mathbf{X}_{\sim i}}\left(Y_j | X_i\right)\right] / Var\left[Y\right]$, for parameter $X_i$ and output $Y_j$, $\mathbf{X}_{\sim i}$ being all other parameters.
+This captures the total influence of $X_i$ all other things being equal.
+The total order index is given by $\mathbb{E}_{\mathbf{X}_{\sim i}} \left[Var(Y_j | \mathbf{X}_{\sim i})\right] / Var\left[Y_j\right]$ and captures non-homogeneity in behavior and interaction between parameters.
+Indices were estimated with a Sobol sequence sampling of 20,000 model runs \cite{saltelli2010variance}.
+We give in Table~\ref{tab:saltelli} estimated results.
+We find that indicators related to modularity (internationalization and optimal modularity) are mostly influenced by geographical parameters.
+On the contrary, indicators of network structure are driven by origin and destination sizes.
+Sectors and link history have a second order by not negligible role.
+Altogether, the analysis witnesses of strong interaction effects between parameters, since for several indicators the total order index is one order of magnitude larger than the first order index.
+This means that the generative model captures some complex behavior of the system.
+The very different important parameters depending on outputs studied confirms the relevance of a more exhaustive grid exploration.
+
+%%%%%%%%%%%%%
+\begin{table}
+\caption{Saltelli sensitivity indices, for indicators in rows and parameters in columns. We give for each pair the first order index (F) and the total order index (T).\label{tab:saltelli}}
+\hspace{-1cm}\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|c|c|}
+\hline
+ & \multicolumn{2}{|c|}{$\gamma_G$} & \multicolumn{2}{|c|}{$\gamma_D$} & \multicolumn{2}{|c|}{$\gamma_S$} & \multicolumn{2}{|c|}{$\gamma_W$} & \multicolumn{2}{|c|}{$\gamma_F$} & \multicolumn{2}{|c|}{$\gamma_T$} \\
+ & F & T & F & T & F & T & F & T & F & T & F & T \\
+ \hline
+Internationalization & 0.2 & 0.3 & 0.7 & 0.7 & 0.001 & 0.009 & $4\cdot 10^{-4}$ & 0.007 & 0.03 & 0.04 & 0.02 & 0.04 \\
+Metropolization & 0.02 & 0.1 & 0.02 & 0.2 & 0.002 & 0.1 & 0.001 & 0.09 & 0.2 & 0.6 & 0.3 & 0.6 \\
+Modularity & 0.3 & 0.4 & 0.6 & 0.6 & 0.004 & 0.02 & $3\cdot 10^{-4}$ & 0.01 & 0.005 & 0.03 & 0.002 & 0.03 \\
+Avg. com. size & 0.008 & 0.09 & 0.01 & 0.1 & 0.002 & 0.07 & 0.003 & 0.04 & 0.3 & 0.6 & 0.4 & 0.6 \\
+Degree entropy & 0.006 & 0.02 & 0.003 & 0.02 & 0.006 & 0.03 & 0.008 & 0.02 & 0.5 & 0.5 & 0.5 & 0.5 \\
+Weight entropy & 0.04 & 0.1 & 0.03 & 0.1 & 0.008 & 0.08 & 0.01 & 0.07 & 0.4 & 0.5 & 0.4 & 0.5 \\\hline
+\end{tabular}
+\end{table}
+%%%%%%%%%%%%%
+
 \paragraph{One factor sampling}
-% 20190924_162740_ONEFACTOR_REPLICATIONS_SYNTHETIC_GRID
 
 A first numerical experiment of one-factor sampling on all Cobb-Douglas parameters and 100 stochastic repetitions confirms a good statistical convergence (average Sharpe ratios for indicators all larger than 5).
+We use thus 20 repetitions in following larger experiments ().
 Behavior of some indicators are shown in Fig.~\ref{fig:onefactor}.

### 2020-03-05 main.tex
Update on Overleaf.

@@
 \begin{abstract}
-    This paper introduces a generative model for urban networks defined by interactions between firms.
-The aim is to test several geographical properties emerging in such a network as metropolisation, internationalisation and regionalisation and test the effect of specialisation on these phenomena.
+    This paper introduces a generative model for urban networks defined by interactions between firms to test several geographical properties emerging in such a network as metropolisation or internationalisation.
 The growth of the network is simulated by adding randomly new links defined by ownership between firms, with probabilities of occurring depending on the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past as well as the geographical and socio-cultural distance.
@@
 Understanding the drivers of growth and geographical properties of such a urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}.
-Straightening these city-interactions can be crucial for revitalising certain regions \cite{Clarke2018}, by developing strategic industries for global insertion of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
+Straightening these city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}, by developing strategic industries for global insertion of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
 The functioning of such a city-network and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future exogenous shocks such as an economic crisis, a single market collapse or a post-Brexit scenario for the European Union.
@@
 Trade network have been extensively studied from complex networks perspective, but are generally driven at the country-level \cite{garlaschelli2005structure} because of lack of city-level data.
-Input-output models, mostly used in Regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}.
+Input-output models, mostly used in regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}.
 In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for abstract flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2020indirect}.
@@
 Companies from the AMADEUS database are georeferenced using the Geonames database (using postcode or address depending on the information available).
-We then join this data with the GHS-FUA dataset for functional urban areas \cite{Florczyk2019ghs}, which are the relevant objects for applying our model.
+We then join this data with the GHS-FUA dataset for Functional Urban Areas \cite{Florczyk2019ghs}, which are the most relevant objects for applying our model.
 Starting from a firm-level dataset of 1,562,862 nodes and  1,866,936 links, we obtain a directed network of 573 urban areas and 9323 ownership links.
@@
 An other attribute of nodes is the industrial sector composition, expressed as a proportion of turnover in the area associated to a given industrial sector.
-Detailed sectors of firms are available up to the fourth digit for the NACE classification in the raw data.
+Detailed sectors of firms are available up to the four digits of the NACE classification in the raw data.
 We consider the first level classification (21 categories) and compute sector proportions accordingly.
@@
 In comparison, with 29 countries, the classification of urban areas by country has a modularity of 0.32.
-This means that the structure of flows has a certain regional integration.
+This means that the structure of flows present certain regional patterns.
 Indeed, a null model assigning random labels with 30 communities, bootstrap 1000 times, yields an average modularity of $0.049 \pm 0.002$, confirming the statistical significance of these communities.
 
-To have a first insight into factors playing a role in the formation of links, we proceed to a statistical analysis which is similar to unconstrained spatial interaction models \cite{wilson1975some}.
+To have a first insight into factors playing a role in the formation of links, we proceed to a statistical analysis similar to unconstrained spatial interaction models \cite{wilson1975some}.
 We test five different statistical models, including progressively the processes candidate to be included in the network generative model.
@@
 The second model adds fixed effects by pair of countries, for which around a third are significant, and with a better $R^2$ of 0.17.
-Adding origin and destination economic size (turnovers $T_i$ and $T$, model 3) slightly improve the explanative power, but the best model among the linear ones is the model 4 with all factors and fixed effects, both in terms of explained variance ($R^2 = 0.31$) and Akaike information criterion (what means that there is no overfitting).
+Adding origin and destination economic size (turnovers $T_i$ and $T$, model 3) slightly improve the explanative power, but the best model among the linear ones is the model 4 with all factors and fixed effects, both in terms of explained variance ($R^2 = 0.31$) and Akaike information criterion.
+The la(what means that there is no overfitting).
 The Poisson model with all variables is much better regarding pseudo-$R^2$ and provides much more significant estimates (all fixed effects significant and much smaller standard deviations).

### 2020-03-05 main.tex
Update on Overleaf.

@@
 
-Regarding the community structure of the network, different modularity maximization algorithms (run on the undirected corresponding network) give a same undirected modularity of 0.38, and a directed modularity \cite{nicosia2009extending} of 0.34 for the greedy algorithm of \cite{clauset2004finding} and 0.36 for the Louvain algorithm \cite{blondel2008fast}, corresponding to 36 communities (resp.
+Regarding the community structure of the network, different modularity maximisation algorithms (run on the undirected corresponding network) give a same undirected modularity of 0.38, and a directed modularity \cite{nicosia2009extending} of 0.34 for the greedy algorithm of \cite{clauset2004finding} and 0.36 for the Louvain algorithm \cite{blondel2008fast}, corresponding to 36 communities (resp.
 15).
@@
 The second model adds fixed effects by pair of countries, for which around a third are significant, and with a better $R^2$ of 0.17.
-Adding origin and destination economic size (turnovers $T_i$ and $T$, model 3) slightly improve the explanative power, but the best model among the linear ones is the model 4 with all factors and fixed effects, both in terms of explained variance ($R^2 = 0.31$) and Akaike information criterion.
-The la(what means that there is no overfitting).
+Adding origin and destination economic size (turnovers $T_i$ and $T$, model 3) slightly improves the explanative power, but the best model among the linear ones is the model 4 with all factors and fixed effects, both in terms of explained variance ($R^2 = 0.31$) and Akaike information criterion.
+The latter indicates the absence of overfitting.
 The Poisson model with all variables is much better regarding pseudo-$R^2$ and provides much more significant estimates (all fixed effects significant and much smaller standard deviations).
@@
 We compare model performance using $R^2$, Mean Square Error, and Akaike Information Criterion.
-For the Poisson model, AIC is not given as not comparable, and the pseudo-$R^2$ is given.\label{tab:reg}}
+For the Poisson model, AIC is not given as it is not comparable, and the pseudo-$R^2$ is given either.\label{tab:reg}}
 \medskip
@@
 As \cite{martinus2018global} point out, a combination of several distances may play a role in establishing linkages.
-Geographical distance, industrial proximity and complementarity will typically be in concurrence \cite{cottineau2020nested}.
+Geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}.
 Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalized to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets despite the European common market (confirmed relevant by the fixed effect regression described above); (iii) city economic size, which has furthermore scaling laws properties; (iv) industrial proximity in terms of firm sector composition; (v) previous linkages (influence of history).
@@
 The model is stopped either when a final time is reached or when the cumulated volume of links reaches a target quantity.
-This formulation can be considered both as an economic utility model and a generalization of spatial interaction models.
+This formulation can be considered both as an economic utility model and a generalisation of spatial interaction models.
 
-Note that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalize the exponents to 1 (which is different than renormalizing the probability), to ensure to include convex functions.
-The ``geopolitical/sociocultural'' distance remains abstract and should be parametrized (see setup below) and estimated in the data-driven setting.
+Note that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalize the exponents to 1 (which is different than renormalising the probability), to ensure to include convex functions.
+The ``geopolitical/sociocultural'' distance remains abstract and should be parametrised (see setup below) and estimated in the data-driven setting.
 We assume no evolution of economic size in this simple version of the model, which means that adjusted parameter are valid on a restricted time period.
@@
 
-Besides model dynamics, its initialization (or parametrization on real data) regarding the spatial distribution, sizes, countries and industrial sector compositions, is also important to understand model behavior.
-\cite{raimbault2019space} furthermore showed that the initial spatial configuration can have significant impacts on model outcomes.
+Besides model dynamics, its initialisation (or parametrisation on real data) regarding the spatial distribution, sizes, countries and industrial sector compositions, is also important to understand model behavior.
+In addition  \cite{raimbault2019space} showed that the initial spatial configuration can have significant impacts on model outcomes.
 We therefore consider two versions: one with a synthetic system of cities which can be stochastically repeated, and one other based on real data.
@@
 The largest city is constrained to have a standard deviation and a mode of $1/2$ and the smallest $1/K$, with a linear interpolation between the two for other cities.
-Parameters of the log-normal law are obtained by solving the corresponding two equations system for each city, and each distribution is discretized and renormalized to be a vector of probabilities.
+Parameters of the log-normal law are obtained by solving the corresponding two equations system for each city, and each distribution is discretised and renormalised to be a vector of probabilities.
 In synthetic experiments, the network is generated from an empty initial network until reaching $t_f=1500$, with a unit link volume of $w_0 = 1$.
@@
 The real setup is done using the empirical network analyzed above.
-Urban areas are distributed according to their positions in the real geographical space, and economic size and sector composition attributes are used.
+Urban areas are distributed according to their positions in the actual geographical space, and economic size and sector composition attributes are used.
 The socio-economic distance is constructed using information from the statistical modeling, by considering fixed-effect coefficients estimated for each pair of countries.
@@
 Numerical experiments are done by integrating the model into the OpenMOLE software for model exploration \cite{reuillon2013openmole}, which provides sensitivity analysis and calibration methods and a transparent access to high-performance computing infrastructures.
-Model source code, aggregated network data (raw initial firm network can not be made available for data ownership reasons), and results are available on the open git repository of the project at \url{https://github.com/JusteRaimbault/ABMCitiesFirms}.
+Model source code, aggregated network data (raw initial firm network can not be made available for data ownership reasons), and results are available on the open Git repository of the project at \url{https://github.com/JusteRaimbault/ABMCitiesFirms}.
 Large simulation result files are available on the dataverse repository at \url{https://doi.org/10.7910/DVN/UPX23S}.
@@
 We give in Table~\ref{tab:saltelli} estimated results.
-We find that indicators related to modularity (internationalization and optimal modularity) are mostly influenced by geographical parameters.
+We find that indicators related to modularity (internationalisation and optimal modularity) are mostly influenced by geographical parameters.
 On the contrary, indicators of network structure are driven by origin and destination sizes.
@@
  \hline
-Internationalization & 0.2 & 0.3 & 0.7 & 0.7 & 0.001 & 0.009 & $4\cdot 10^{-4}$ & 0.007 & 0.03 & 0.04 & 0.02 & 0.04 \\
-Metropolization & 0.02 & 0.1 & 0.02 & 0.2 & 0.002 & 0.1 & 0.001 & 0.09 & 0.2 & 0.6 & 0.3 & 0.6 \\
+Internationalisation & 0.2 & 0.3 & 0.7 & 0.7 & 0.001 & 0.009 & $4\cdot 10^{-4}$ & 0.007 & 0.03 & 0.04 & 0.02 & 0.04 \\
+Metropolisation & 0.02 & 0.1 & 0.02 & 0.2 & 0.002 & 0.1 & 0.001 & 0.09 & 0.2 & 0.6 & 0.3 & 0.6 \\
 Modularity & 0.3 & 0.4 & 0.6 & 0.6 & 0.004 & 0.02 & $3\cdot 10^{-4}$ & 0.01 & 0.005 & 0.03 & 0.002 & 0.03 \\
@@
     \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figures/rhoDegreeSize-gammaSectors_errorbars.png}
-    \caption{(Top Left) Internationalization index decreases exponentially with gravity decay; (Top Right) Correlation between city weighted degree and size.
+    \caption{(Top Left) Internationalisation index decreases exponentially with gravity decay; (Top Right) Correlation between city weighted degree and size.
 Both plots show a transition from a local to a global regime.
-(Bottom Left) Internationalization varies linearly with sector proximity $\gamma_S$; (Bottom Right) Correlation between degree and size exhibits a maximum, witnessing an intermediate regime where size is the most important.
+(Bottom Left) Internationalisation varies linearly with sector proximity $\gamma_S$; (Bottom Right) Correlation between degree and size exhibits a maximum, witnessing an intermediate regime where size is the most important.
 \label{fig:onefactor}}
@@
 
-The influence of gravity decay parameters is confirmed when plotting the internationalization index against the distance, which shows an exponential decay.
+The influence of gravity decay parameters is confirmed when plotting the internationalisation index against the distance, which shows an exponential decay.
 Similarly, the correlation between city weighted degree and city size exhibit a behaviour that can be interpreted as a phase transition between local and global networks.
@@
 The size of communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.
-All these experiments reveal a complex interplay between processes and how the model produces diverse stylized facts.
+All these experiments reveal a complex interplay between processes and how the model produces diverse stylised facts.
 
@@
 
-As a matter of validation, the model is implemented for real data for Europe joining Functional Urban Areas with the Global Human Settlement Database for cities and their characteristics, with the AMADEUS database produced by the Bureau Van Dijk on ownership links between firms in Europe in 2018 at postcode level in all types of sectors.
+As a matter of validation, the model is implemented for real data for Europe joining Functional Urban Areas with the Global Human Settlement Database for cities and their characteristics, with the AMADEUS database produced by the Bureau Van Dijk on ownership links between firms in Europe in 2018 available at postcode level in all types of sectors given by NACE codes.
 
@@
 % - other formulation of the combination of factors
-%multi-objective optimization depending on sectors, using Pareto fronts ?
+%multi-objective optimisation depending on sectors, using Pareto fronts ?
 - needs empirical evidence; (iv) copula-based combination as done by \cite{2019arXiv190505106C} - copula parameters are free or estimated from real data

### 2020-03-05 main.tex
.

--- 2020-03-05 main.tex
@@
 A first numerical experiment of one-factor sampling on all Cobb-Douglas parameters and 100 stochastic repetitions confirms a good statistical convergence (average Sharpe ratios for indicators all larger than 5).
-We use thus 20 repetitions in following larger experiments ().
+We use thus 20 repetitions in following larger experiments (in the case of a normal distribution, a 95\% confidence interval on the average is of size $2\sigma \cdot 1.96 / \sqrt{n}$, what gives $n=16$ repetitions for a CI of width of the standard deviation).
 Behavior of some indicators are shown in Fig.~\ref{fig:onefactor}.
+For example, internationalization and the correlation between economic size and weighted degree both show a phase transition as a function of $\gamma_G$, from a local to a global regime.
+The sector proximity parameter $\gamma_S$ influences the internationalization linearly, but induces a maximum for the correlation between size and degree, which can be interpreted as a setting where size and the total volume in a city are the most correlated.
 
+%%%%%%%%%%%%%
 \begin{figure}
-	\includegraphics[width=0.48\textwidth,height=0.25\textheight]{figures/internationalization-gravityDecay_errorbars.png}
-    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figures/rhoDegreeSize-gravityDecay_errorbars.png}\\
-    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figures/internationalization-gammaSectors_errorbars.png}
-    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figures/rhoDegreeSize-gammaSectors_errorbars.png}
+	\includegraphics[width=0.48\textwidth,height=0.2\textheight]{figures/internationalization-gravityDecay_errorbars.png}
+    \includegraphics[width=0.48\textwidth,height=0.2\textheight]{figures/rhoDegreeSize-gravityDecay_errorbars.png}\\
+    \includegraphics[width=0.48\textwidth,height=0.2\textheight]{figures/internationalization-gammaSectors_errorbars.png}
+    \includegraphics[width=0.48\textwidth,height=0.2\textheight]{figures/rhoDegreeSize-gammaSectors_errorbars.png}
     \caption{(Top Left) Internationalization index decreases exponentially with gravity decay; (Top Right) Correlation between city weighted degree and size.
@@
 \end{figure}
+%%%%%%%%%%%%%
 
@@
 The model behaviour is then studied with a grid sampling for parameters and 20 repetitions (the computations run on the European Grid Infrastructure, and are equivalent to 2.5 years of CPU time).
+Distance decay and sector  parameters are varied with 10 steps, other parameters with three.
+The influence of gravity decay parameters is confirmed when plotting the internationalization index against the distance, which exhibits an exponential decay as shown in Fig.~\ref{fig:internat}.
+It interacts with the role of sectors and the size of the origin, with a more significant effect of sector proximity when size has the largest influence (third panel).
 
-The influence of gravity decay parameters is confirmed when plotting the internationalization index against the distance, which shows an exponential decay.
-Similarly, the correlation between city weighted degree and city size exhibit a behaviour that can be interpreted as a phase transition between local and global networks.
-Other indicators exhibit non-trivial patterns.
-For example, when considering average community size of the final network, we obtain a maximal integration in term of communities at an intermediate value of the gravity decay, what can be interpreted as the emergence of a regional regime.
-The size of communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.
-All these experiments reveal a complex interplay between processes and how the model produces diverse stylized facts.
-
+%%%%%%%%%%%%%%
 \begin{figure}
     \includegraphics[width=\textwidth]{figures/internationalization_countryGravityDecay200_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
-    \caption{The transition as a function of interaction range depends on the influence of origin size $\gamma_F$; sector proximity $\gamma_S$ plays a role only for a large influence of the origin.}
+    \caption{The transition as a function of interaction range depends on the influence of origin size $\gamma_F$; sector proximity $\gamma_S$ plays a role only for a large influence of the origin.\label{fig:internat}}
 \end{figure}
+%%%%%%%%%%%%%%
 
+Other indicators exhibit non-trivial patterns.
+For example, when considering average community size of the final network as shown in Fig.~\ref{fig:comsize}, we obtain a maximal integration in term of communities at an intermediate value of the gravity decay, what can be interpreted as the emergence of a regional regime.
+The size of communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.
+In particular, we observe a qualitative inversion of the role of $\gamma_S$ when introducing an effect of the origin (switching $\gamma_F$ from 0 to 1, first panel to second panel).
+The maximal community size disappears when $\gamma_F = 2$, implying a regime with small local communities where large cities mostly connect with there hinterland.
+These experiments reveal a complex interplay between processes and how the model produces diverse stylized facts.
+
+%%%%%%%%%%%%%%
 \begin{figure}
     \includegraphics[width=\textwidth]{figures/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
-	\caption{Maximal integration in term of community size is achieved at an intermediate value of $d_G$: emergence of a regional regime; Maximal size depends on the role of sectors $\gamma_S$, in a decreasing way when origin size is deactivated, and increasing way when $\gamma_F=1$; This regime disappear when origin size influence is too large}
+	\caption{Maximal integration in term of community size is achieved at an intermediate value of $d_G$: emergence of a regional regime; Maximal size depends on the role of sectors $\gamma_S$, in a decreasing way when origin size is deactivated, and increasing way when $\gamma_F=1$; This regime disappear when origin size influence is too large.\label{fig:comsize}}
 \end{figure}
-
-% Saltelli exploration of the model
-% -> confirm relative importance of params and search for inhomogeneity/interaction effects
+%%%%%%%%%%%%%%
 
@@
 
-As a matter of validation, the model is implemented for real data for Europe joining Functional Urban Areas with the Global Human Settlement Database for cities and their characteristics, with the AMADEUS database produced by the Bureau Van Dijk on ownership links between firms in Europe in 2018 at postcode level in all types of sectors.
+We then apply the model to a real system of cities by calibrating it on the European ownership network.
+Model setup is done following the real setup described above.
+The number of time steps $t_f$ is left as an additional parameter, which in a sense determines the granularity of the cumulative network generation process.
+The objective functions for calibration are the average mean squared error on logarithms of weights given by $\varepsilon_L = \frac{1}{N^2} \sum_{i,j} \left(\log w_{ij} - \log \hat{w}_{ij} \right)^2$ if $\hat{w}_{ij}$ are the simulated weights, and the logarithm of the mean squared error $\varepsilon_M = \log\left(\frac{1}{N^2} \sum_{i,j} \left(w_{ij} - \hat{w}_{ij}\right)^2 \right)$.
+\cite{raimbault2018indirect} showed that these two are complementary to calibrate urban systems models.
+Calibration is done using a genetic NSGA2 algorithm \cite{deb2002fast}, which is particularly suited for such a bi-objective calibration of a stochastic simulation model.
+We use a population of 200 and run the algorithm for 20,000 generations, on 1000 islands in parallel.
 
-\subsubsection{Model calibration}
+%%%%%%%%%%%%%%
+\begin{figure}
+	\begin{center}
+    \includegraphics[width=0.5\textwidth]{figures/pareto_colorgammaLinks.png}
+    \end{center}
+	\caption{Pareto front obtained for the bi-objective model calibration.\label{fig:calibration}}
+\end{figure}
+%%%%%%%%%%%%%%
 
-Calibration is done using a genetic NSGA2 algorithm, finding the parameter values giving network trajectories closest to reality.
-First results show in particular a role of the path-dependency parameter confirming the relevance of this complex generative model.
-Further work is needed to get more interpretations from this experiment, but our study demonstrates the feasibility of model calibration on real data.
-Several scenarios are considered taking into consideration the current and future economic challenges, such as those given by Brexit.
-The model may be applied in future work to assess policy issues and recommendations.
+We show in Fig.~\ref{fig:calibration} the calibration result, as a Pareto front of compromise points between the two objectives.
+These correspond to parameter values giving network trajectories closest to reality.
+The values for $\varepsilon_L$ are always lower than for statistical models, our model outperforming these in this predictive sense.
+% add discussion param values
+This furthermore shows in particular a role of the path-dependency parameter, which calibrated value is quite high (8 in average on the Pareto front), confirming the relevance of this complex generative model including path-dependency.
+These results altogether demonstrate the feasibility of model calibration on real data.
 
@@
 
+% Several scenarios are considered taking into consideration the current and future economic challenges, such as those given by Brexit.
+The model may be applied in future work to assess policy issues and recommendations.
+
 % - other formulation of the combination of factors

### 2020-03-05 main.tex

--- 2020-03-05 main.tex
@@
 A first numerical experiment of one-factor sampling on all Cobb-Douglas parameters and 100 stochastic repetitions confirms a good statistical convergence (average Sharpe ratios for indicators all larger than 5).
-We use thus 20 repetitions in following larger experiments ().
+We use thus 20 repetitions in following larger experiments (in the case of a normal distribution, a 95\% confidence interval on the average is of size $2\sigma \cdot 1.96 / \sqrt{n}$, what gives $n=16$ repetitions for a CI of width of the standard deviation).
 Behavior of some indicators are shown in Fig.~\ref{fig:onefactor}.
+For example, internationalization and the correlation between economic size and weighted degree both show a phase transition as a function of $\gamma_G$, from a local to a global regime.
+The sector proximity parameter $\gamma_S$ influences the internationalization linearly, but induces a maximum for the correlation between size and degree, which can be interpreted as a setting where size and the total volume in a city are the most correlated.
 
+%%%%%%%%%%%%%
 \begin{figure}
@@
 The model behaviour is then studied with a grid sampling for parameters and 20 repetitions (the computations run on the European Grid Infrastructure, and are equivalent to 2.5 years of CPU time).
+Distance decay and sector  parameters are varied with 10 steps, other parameters with three.
+The influence of gravity decay parameters is confirmed when plotting the internationalization index against the distance, which exhibits an exponential decay as shown in Fig.~\ref{fig:internat}.
+It interacts with the role of sectors and the size of the origin, with a more significant effect of sector proximity when size has the largest influence (third panel).
 
-The influence of gravity decay parameters is confirmed when plotting the internationalisation index against the distance, which shows an exponential decay.
-Similarly, the correlation between city weighted degree and city size exhibit a behaviour that can be interpreted as a phase transition between local and global networks.
-Other indicators exhibit non-trivial patterns.
-For example, when considering average community size of the final network, we obtain a maximal integration in term of communities at an intermediate value of the gravity decay, what can be interpreted as the emergence of a regional regime.
-The size of communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.
-All these experiments reveal a complex interplay between processes and how the model produces diverse stylised facts.
-
+%%%%%%%%%%%%%%
 \begin{figure}
     \includegraphics[width=\textwidth]{figures/internationalization_countryGravityDecay200_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
-    \caption{The transition as a function of interaction range depends on the influence of origin size $\gamma_F$; sector proximity $\gamma_S$ plays a role only for a large influence of the origin.}
+    \caption{The transition as a function of interaction range depends on the influence of origin size $\gamma_F$; sector proximity $\gamma_S$ plays a role only for a large influence of the origin.\label{fig:internat}}
 \end{figure}
+%%%%%%%%%%%%%%
 
+Other indicators exhibit non-trivial patterns.
+For example, when considering average community size of the final network as shown in Fig.~\ref{fig:comsize}, we obtain a maximal integration in term of communities at an intermediate value of the gravity decay, what can be interpreted as the emergence of a regional regime.
+The size of communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.
+In particular, we observe a qualitative inversion of the role of $\gamma_S$ when introducing an effect of the origin (switching $\gamma_F$ from 0 to 1, first panel to second panel).
+The maximal community size disappears when $\gamma_F = 2$, implying a regime with small local communities where large cities mostly connect with there hinterland.
+These experiments reveal a complex interplay between processes and how the model produces diverse stylised facts.
+
+%%%%%%%%%%%%%%
 \begin{figure}
     \includegraphics[width=\textwidth]{figures/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
-	\caption{Maximal integration in term of community size is achieved at an intermediate value of $d_G$: emergence of a regional regime; Maximal size depends on the role of sectors $\gamma_S$, in a decreasing way when origin size is deactivated, and increasing way when $\gamma_F=1$; This regime disappear when origin size influence is too large}
+	\caption{Maximal integration in term of community size is achieved at an intermediate value of $d_G$: emergence of a regional regime; Maximal size depends on the role of sectors $\gamma_S$, in a decreasing way when origin size is deactivated, and increasing way when $\gamma_F=1$; This regime disappear when origin size influence is too large.\label{fig:comsize}}
 \end{figure}
-
-% Saltelli exploration of the model
-% -> confirm relative importance of params and search for inhomogeneity/interaction effects
+%%%%%%%%%%%%%%
 
@@
 
-As a matter of validation, the model is implemented for real data for Europe joining Functional Urban Areas with the Global Human Settlement Database for cities and their characteristics, with the AMADEUS database produced by the Bureau Van Dijk on ownership links between firms in Europe in 2018 available at postcode level in all types of sectors given by NACE codes.
+We then apply the model to a real system of cities by calibrating it on the European ownership network.
+Model setup is done following the real setup described above.
+The number of time steps $t_f$ is left as an additional parameter, which in a sense determines the granularity of the cumulative network generation process.
+The objective functions for calibration are the average mean squared error on logarithms of weights given by $\varepsilon_L = \frac{1}{N^2} \sum_{i,j} \left(\log w_{ij} - \log \hat{w}_{ij} \right)^2$ if $\hat{w}_{ij}$ are the simulated weights, and the logarithm of the mean squared error $\varepsilon_M = \log\left(\frac{1}{N^2} \sum_{i,j} \left(w_{ij} - \hat{w}_{ij}\right)^2 \right)$.
+\cite{raimbault2018indirect} showed that these two are complementary to calibrate urban systems models.
+Calibration is done using a genetic NSGA2 algorithm \cite{deb2002fast}, which is particularly suited for such a bi-objective calibration of a stochastic simulation model.
+We use a population of 200 and run the algorithm for 20,000 generations, on 1000 islands in parallel.
 
-\subsubsection{Model calibration}
+%%%%%%%%%%%%%%
+\begin{figure}
+	\begin{center}
+    \includegraphics[width=0.5\textwidth]{figures/pareto_colorgammaLinks.png}
+    \end{center}
+	\caption{Pareto front obtained for the bi-objective model calibration.\label{fig:calibration}}
+\end{figure}
+%%%%%%%%%%%%%%
 
-Calibration is done using a genetic NSGA2 algorithm, finding the parameter values giving network trajectories closest to reality.
-First results show in particular a role of the path-dependency parameter confirming the relevance of this complex generative model.
-Further work is needed to get more interpretations from this experiment, but our study demonstrates the feasibility of model calibration on real data.
-Several scenarios are considered taking into consideration the current and future economic challenges, such as those given by Brexit.
-The model may be applied in future work to assess policy issues and recommendations.
+We show in Fig.~\ref{fig:calibration} the calibration result, as a Pareto front of compromise points between the two objectives.
+These correspond to parameter values giving network trajectories closest to reality.
+The values for $\varepsilon_L$ are always lower than for statistical models, our model outperforming these in this predictive sense.
+% add discussion param values
+This furthermore shows in particular a role of the path-dependency parameter, which calibrated value is quite high (8 in average on the Pareto front), confirming the relevance of this complex generative model including path-dependency.
+These results altogether demonstrate the feasibility of model calibration on real data.
 

### 2020-04-22 main.tex
Update on Overleaf.
.

@@
 
-\title{An agent-based model for city networks based on interactions between firms}
+\title{\vspace{-2cm}A generative model for urban firm networks}
 
@@
 
-\author{J. Raimbault$^{1,2,3,\ast}$, N. Zdanowska$^{1,3}$ and E. Arcaute$^1$\\\medskip\small
-$^{1}$ Centre for Advanced Spatial Analysis, UCL; $^{2}$ UPS CNRS 3611 ISC-PIF; $^{3}$ UMR CNRS 8504 G{\'e}ographie-cit{\'e}s\\
+\author{J. Raimbault$^{1,2,3,\ast}$, N. Zdanowska$^{1,3}$\smallskip\\
+$^{1}$ Centre for Advanced Spatial Analysis, UCL\\
+$^{2}$ UPS CNRS 3611 ISC-PIF\\
+$^{3}$ UMR CNRS 8504 G{\'e}ographie-cit{\'e}s\\
 $^{\ast}$ \texttt{juste.raimbault@polytechnique.edu}
@@
 \begin{abstract}
-    This paper introduces a generative model for urban networks defined by interactions between firms to test several geographical properties emerging in such a network as metropolisation or internationalisation.
-The growth of the network is simulated by adding randomly new links defined by ownership between firms, with probabilities of occurring depending on the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past as well as the geographical and socio-cultural distance.
-The simulation on a synthetic system of cities unveil phase transitions when changing interaction distance, and non-trivial patterns in the model behavior.
-Calibration of the model on real data on company ownership linkages for Europe in 2018 is also undertaken.
-Future work will consist in observing thanks to this model the impact of economic crisis on such networks.
+The emergence of entangled urban networks is a salient feature of globalisation processes.
+Understanding the drivers behind the growth of such networks, and in particular urban firm networks, is crucial for the economic resilience of urban systems.
+We introduce in this paper a generative network model for firm networks at the urban area level, which includes several complementary processes which are the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past as well as the geographical and socio-cultural distance.
+An empirical network analysis on European firm ownership data confirms a role for each factor.
+We then simulate network growth for synthetic systems of cities, unveiling stylized facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range.
+We finally calibrate the model on the European network, outperforming statistical models and showing a strong role of path-dependency.
+Potential applications of the model include the study of mitigation policies for exogenous shocks such as economic crisis.
 
@@
 Input-output models, mostly used in regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}.
-In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for abstract flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2020indirect}.
-
-%\citep{martinus2018global} combination of different types of proximity in inter-urban firm networks: economic, sociocultural, geopolitical - implies spatial and non-spatial 
-%\citep{pan2017mapping} different types of linkages: e.g. services here, but not only. All sectors should be taken into account first.
-%\cite{martinus2019brokerage} role of small countries
-%\cite{dawley2019creating} regional policies for insertion in global network
-%\cite{gluckler2016relational} positioning within networks
+In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for abstract flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2018indirect}.
 
 In this paper we choose to focus on urban networks of firms to question how can we understand and capture geographical and economic processes as agglomeration economies \cite{Glaeser2001} characteristic to spatial economy \cite{Fujitaetal2001} within a generative model.
-The originality of the present paper is to approach these issues within the system of cities framework \cite{berry1964cities}, where the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}.
+The specific question we tackle is the complementarity of different processes to generate such urban networks.
+One originality of the present paper is to approach these issues within the system of cities framework \cite{berry1964cities}, where the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}.
 We examine the interactions of European cities within firm linkages defined by corporate ownership links.
@@
 \cite{raimbault2018indirect} showed that these two are complementary to calibrate urban systems models.
-Calibration is done using a genetic NSGA2 algorithm \cite{deb2002fast}, which is particularly suited for such a bi-objective calibration of a stochastic simulation model.
-We use a population of 200 and run the algorithm for 20,000 generations, on 1000 islands in parallel.
+Calibration is done using a genetic NSGA2 algorithm \cite{deb2002fast}, which is particularly suited for such a bi-objective calibration of a simulation model.
+Stochasticity is taken into account by adding the estimation confidence as an additional objective, and we filter the final population to have at least 20 samples.
+We use a population of 200 individuals and run the algorithm for 20,000 generations, on 1000 islands in parallel.
 
@@
 	\begin{center}
-    \includegraphics[width=0.5\textwidth]{figures/pareto_colorgammaLinks.png}
+    \includegraphics[width=0.5\textwidth]{figures/pareto_colorgammaOrigin.png}
     \end{center}
-	\caption{Pareto front obtained for the bi-objective model calibration.\label{fig:calibration}}
+	\caption{Pareto front obtained for the bi-objective model calibration for $\varepsilon_L$ and $\varepsilon_M$ objectives.
+$\varepsilon_L$ can be compared to the MSE of statistical models fitted above, and the generative model outperform these for several parametrizations (lowest MSE 5.33 for statistical models).
+Point size gives the number of stochastic samples and color the value of $\gamma_F$.\label{fig:calibration}}
 \end{figure}
@@
 We show in Fig.~\ref{fig:calibration} the calibration result, as a Pareto front of compromise points between the two objectives.
-These correspond to parameter values giving network trajectories closest to reality.
-The values for $\varepsilon_L$ are always lower than for statistical models, our model outperforming these in this predictive sense.
-% add discussion param values
-This furthermore shows in particular a role of the path-dependency parameter, which calibrated value is quite high (8 in average on the Pareto front), confirming the relevance of this complex generative model including path-dependency.
-These results altogether demonstrate the feasibility of model calibration on real data.
+As the mean squared error on logarithm is the same than the one for statistical models, it can be directly compared.
+The other objective would correspond to statistical models directly on variables: therefore, our model covers in a sense a set of intermediate models which make a compromise between the two objectives, and is therefore much more flexible.
+The values for $\varepsilon_L$ are for a large proportion of points lower than for statistical models, our model outperforming these in this predictive sense.
+As shown in Fig.~\ref{fig:calibration}, increasing values for $\gamma_F$ lead to better performance regarding $\varepsilon_L$: increasing the importance of the origin will provide a better fit on all order of magnitudes of link weights (while $\varepsilon_M$ favors mostly the largest links given the fat-tail distribution).
+
+When considering optimal points with $\varepsilon_L < 5.0$, the adjusted parameters are: long range gravity interactions and socio-economic interactions given by $d_G = 1/\gamma_G = 5205 \pm 1775$ and $d_D = 1/ \gamma_D = 109 \pm 29$ (the maximal $g_{ij}$ is 11.8); a stronger influence of destination than origin as $\gamma_F = 1.45 \pm 0.09$ and $\gamma_T = 2.0 \pm 0.06$; a comparable influence of sectors as $\gamma_S = 1.75 \pm 0.17$; a fine time granularity as $t_f = 4960 \pm 109$; and a very strong influence of previous links as $\gamma_W = 7.99 \pm 3.77$.
+This last point furthermore shows in particular a role of the path-dependency parameter, confirming the relevance of this complex generative model including path-dependency.
+The interaction of this process with others implies an inversion of the relative importance of origin and destination (on the contrary to statistical models).
+The fact that calibrated interaction distances are large is consistent with long range economic interactions within Europe.
+Besides these stylized facts, these calibration results altogether demonstrate the feasibility of model application to real data.
 
@@
 This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMole platform and calibrated on real data on ownership linkages of firms for Europe.
-The simulation on a synthetic system of cities unveil phase transitions when changing interaction distance, and non-trivial patterns in the model behavior.
-This kind of model can have practical application in the future in terms of testing the effect of exogenous shocks in the socio-economic structure as the upcoming effect of Brexit on the European market.
+The simulation on a synthetic system of cities provides an broad knowledge on model behavior in its parameter space, and its calibration suggests its potential application to policy issues.
+This kind of model can indeed have practical applications in the future in terms of testing the effect of exogenous shocks in the socio-economic structure as the upcoming effect of Brexit on the European market.
 
@@
 
-% Several scenarios are considered taking into consideration the current and future economic challenges, such as those given by Brexit.
-The model may be applied in future work to assess policy issues and recommendations.
+\newpage
 
-% - other formulation of the combination of factors
-%multi-objective optimisation depending on sectors, using Pareto fronts ?
-- needs empirical evidence; (iv) copula-based combination as done by \cite{2019arXiv190505106C} - copula parameters are free or estimated from real data
-% The function used can be understood as a separable spatial interaction model. As a possible extension, Copula capture a dependency structure but must be parametrised by estimation on real data.
-
-% - evolution of city sizes (co-evolution model)
-% - role of path dependency
-% - towards a model with firm agents? (multi-scale ABM)
-
-\bibliographystyle{apalike}
-%\bibliographystyle{unsrt}
+%\bibliographystyle{apalike}
+\bibliographystyle{unsrt}
 \bibliography{biblio}

### 2020-08-31 main.tex
Update on Overleaf.

@@
-\documentclass[11pt]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[margin=1in]{geometry}
-\usepackage{natbib,graphicx,mdframed,hyperref,threeparttable,amssymb}
+\documentclass[10pt,letterpaper]{article}
+\input{header.tex}
 
-\title{\vspace{-2cm}A generative model for urban firm networks}
+\usepackage{ragged2e}
+\begin{document}
+\justify
+\vspace*{0.2in}
 
-\date{}
+% Title must be 250 characters or less.
+\begin{flushleft}
+{\Large
+\textbf\newline{A generative model for urban firm networks} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+}
+% alternative titles?
+% Modeling the growth of urban firm networks
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+\\
+Juste Raimbault\textsuperscript{1,2,3*},
+Natalia Zdanowska\textsuperscript{1,3},
+Elsa Arcaute\textsuperscript{3}
+\\
+\bigskip
+\textbf{1} Centre for Advanced Spatial Analysis, UCL, London, United Kingdom
+\\
+\textbf{2} UPS CNRS 3611 ISC-PIF, Paris, France
+\\
+\textbf{3} UMR CNRS 8504 G{\'e}ographie-cit{\'e}s, Paris, France
+\\
+\bigskip
 
-\begin{document}
+% Use the asterisk to denote corresponding authorship and provide email address in note below.
+* juste.raimbault@polytechnique.edu
 
-\author{J. Raimbault$^{1,2,3,\ast}$, N. Zdanowska$^{1,3}$\smallskip\\
-$^{1}$ Centre for Advanced Spatial Analysis, UCL\\
-$^{2}$ UPS CNRS 3611 ISC-PIF\\
-$^{3}$ UMR CNRS 8504 G{\'e}ographie-cit{\'e}s\\
-$^{\ast}$ \texttt{juste.raimbault@polytechnique.edu}
-}
+\end{flushleft}
 
-\maketitle
-% NetSci format: A4, 11pt, 1in margin, max 12p + up to 3p references
+% Plos one requirements https://journals.plos.org/plosone/s/submission-guidelines
+% Deadline for the Cities as Complex Systems Call for Papers: 4th September https://collections.plos.org/s/cities
+% Manuscripts can be any length. There are no restrictions on word count, number of figures, or amount of supporting information. 
+% PLOS uses “Vancouver” references style
+% We need to reorganise the paper to have more or less the following sections (which can be reordered or renamed):
+% 1. Title page: List title, authors, and affiliations as first page of manuscript 2. Abstract 3. Introduction 4. Materials and Methods 5. Results 6. Discussion 7. Conclusions (optional)
 
-\begin{abstract}
-The emergence of entangled urban networks is a salient feature of globalisation processes.
-Understanding the drivers behind the growth of such networks, and in particular urban firm networks, is crucial for the economic resilience of urban systems.
-We introduce in this paper a generative network model for firm networks at the urban area level, which includes several complementary processes which are the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past as well as the geographical and socio-cultural distance.
-An empirical network analysis on European firm ownership data confirms a role for each factor.
+\section*{Abstract}
+The emergence of interconnected urban networks is a crucial feature of globalisation processes.
+Understanding the drivers behind the growth of such networks -- in particular urban firm networks --, is essential for the economic resilience of urban systems.
+We introduce in this paper a generative network model for firm networks at urban area level including several complementary processes: the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distance.
+An empirical network analysis on European firm ownership data confirms the relevance of each of these factors.
 We then simulate network growth for synthetic systems of cities, unveiling stylized facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range.
 We finally calibrate the model on the European network, outperforming statistical models and showing a strong role of path-dependency.
-Potential applications of the model include the study of mitigation policies for exogenous shocks such as economic crisis.
+Potential applications of the model include the study of mitigation policies to deal with exogenous shocks such as economic crisis or potential lockdowns of countries.
 
-\end{abstract}
+\linenumbers
 
-\section{Introduction}
+\section*{Introduction}
 
-The intensification of globalisation of the economy since the late 20th century resulted in the rise of the network society \cite{castells2000networksociety} and the emergence of a world-city system \cite{taylor2001specification}, characterized by highly interconnected urban centres at the global level \cite{sassen1991}.
-These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity implying spatial and non spatial patterns \cite{martinus2018global}.
-Interactions of economic nature between transnational firms, are together with trade among the most representative of the current economic trends \cite{taylor2001specification}.
-Understanding the drivers of growth and geographical properties of such a urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}.
-Straightening these city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}, by developing strategic industries for global insertion of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
-The functioning of such a city-network and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future exogenous shocks such as an economic crisis, a single market collapse or a post-Brexit scenario for the European Union.
+\cite{derudder2018central}
+\cite{storme2019introducing} cities firms interactions
 
-The use of network models to understand processes underlying the emergence and dynamics of such urban networks has already been widely tackled in the literature.
-\cite{dai2016simulating} propose a generative urban network model at the regional scale, combining geographical factors with network topological properties.
-Air transportation networks can also be understood as urban networks, and \cite{zanin2013modelling} reviews several models for such networks.
-Trade network have been extensively studied from complex networks perspective, but are generally driven at the country-level \cite{garlaschelli2005structure} because of lack of city-level data.
+Networks constitute the new social morphology of our societies, since the intensification of globalisation of the economy in the late 20th century \cite{castells2000networksociety}.
+A world-city system has since then emerged \cite{taylor2001specification}, characterized by highly interconnected urban centres at the global level \cite{sassen1991}.
+These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity implying spatial and non-spatial patterns \cite{martinus2018global}.
+Interactions of economic nature between transnational firms, are together with international trade patterns among the most representative of the current economic trends \cite{taylor2001specification}.
+Understanding the drivers of growth and geographical properties of such an urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}.
+Straightening these city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}, by developing strategic industries for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
+The functioning of such a city-network and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.
+
+The use of network models to understand processes underlying the emergence and dynamics of such networks has already been tackled in the literature, but still not extensively at city level.
+Generative urban network models combining geographical factors with network topological properties exist at the regional scale {\cite{dai2016simulating}}.
+%Air transportation networks can also be understood as urban networks.
+\cite{zanin2013modelling} reviews several models for such networks.
+Trade networks have been extensively studied from the complex networks perspective, but these are generally driven at the country-level \cite{garlaschelli2005structure} because of a lack of city-level data.
 Input-output models, mostly used in regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}.
-In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for abstract flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2018indirect}.
+In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for different types of flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2018indirect}.
 
-In this paper we choose to focus on urban networks of firms to question how can we understand and capture geographical and economic processes as agglomeration economies \cite{Glaeser2001} characteristic to spatial economy \cite{Fujitaetal2001} within a generative model.
+In this paper we choose to focus on urban networks of firms to question and capture geographical and economic processes as internationalisation, metropolisation, regionalisation and specialisation of cities within a generative model.
 The specific question we tackle is the complementarity of different processes to generate such urban networks.
-One originality of the present paper is to approach these issues within the system of cities framework \cite{berry1964cities}, where the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}.
+The originality of the present paper is to examine these issues within the system of cities framework \cite{berry1964cities}.
+In this approach the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}.
 We examine the interactions of European cities within firm linkages defined by corporate ownership links.
-Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geographical structures \cite{2019arXiv191014652Z}.
+Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2019arXiv191014652Z}.
 
-More precisely, our contribution relies on the following points: (i) we propose an empirical analysis of the firm ownership urban network for the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database; (ii) we introduce a generative network model to simulate the growth of such linkages at urban areas scale, which combines multiple factors influencing link formation, as economic intensity of origin and destination of urban areas, industrial sector proximity, the strength of previous links, and geographical and socio-cultural distance; (iii) the model allows us to compare the effect of different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; and (iv) we calibrate the model on the empirical network.
+More precisely, our contribution consists in the following: (i) we propose an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database \cite{AMADEUS2018}; (ii) we introduce a generative network model to simulate the growth of such linkages at urban areas scale, which combines multiple factors influencing link formation, as economic intensity of origin and destination of urban areas, industrial sector proximity, the strength of previous links, and geographical and socio-cultural distance; (iii) the model allows us to compare the effect of different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; and (iv) we calibrate the model on the empirical network.
 
-The rest of this paper is organized as follows: we first proceed to an empirical analysis of the network studied.
-We then describe the generative model for urban networks, and its sensitivity analysis and exploration on synthetic data.
-We then parametrize and calibrate the model on real European firm ownership data.
+The rest of this paper is organized as follows: We first describe the generative model for urban networks and proceed to an empirical analysis of the network used as a case study.
+We then do a sensitivity analysis and exploration of the model on synthetic data.
+The model is calibrated on real European firm ownership data.
 We finally discuss implications of our results and possible extensions.
 
-\section{Empirical network analysis}
+\section*{Materials and Methods}
+
+\subsection*{Generative urban network model}
+
+\subsubsection*{Rationale}
+
+The model is expected to capture a single macroscopic urban scale, i.e.
+the level of the urban system where basic entities are cities.
+Links are induced by underlying firms but these are not explicit.
+Cities are defined by their economic size, but also by an industrial sector profile.
+As the authors of \cite{martinus2018global} point out, a combination of several distances may play a role in establishing linkages.
+Geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}.
+Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalized to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (confirmed relevant by the fixed effect regression described above); (iii) city economic size, which has furthermore scaling laws properties; (iv) industrial proximity in terms of firm sector composition; (v) previous linkages (influence of already existing links in the past).
+This last point is crucial to capture accumulated effects and include path-dependency, which fits well the use of a simulation model instead of a more simple statistical model.
+
+\subsubsection*{Model description}
+
+Let $1 \leq i \leq N$ cities defined at time $t$ by their economic structure such that $E_i(t)$ is the total economic volume (GDP) for city $i$ and $S_{ik}(t)$ is a matrix giving economic volume proportions within each economic sector $k$, assuming $K$ economic sectors.
+Formally, the model creates links between cities which are characterized by their economic size $E_i$ (GDP) and economic structure $S_{ik}$ in terms of activity sectors (probability distribution of firms within $K$ sectors).
+Starting with an existing network with no links, the model iteratively adds links, following a probability given by a generalized Cobb-Douglas function \cite{vilcu2011geometric} as
+
+\begin{equation}
+p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_O} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_D} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s\left(S_{ik},S_{jk}\right)^{\gamma_S} \cdot \exp \left(- d_{ij} / d_0\right) \cdot \exp \left(- c_{ij} / c_0\right)
+\end{equation}
+
+where $E  =  \sum_k E_k$, $W  = \sum_{i,j} w_{ij}$ weights of previous links, $s(S_{ik},S_{jk})$ is a similarity function between activity sectors $S_{ik}$ and $S_{jk}$ taken as a cosine similarity, $d_{ij}$ euclidian distance, and $g_{ij}$ a socio-cultural distance.
+In the case of an already existing link between two areas, the weight of the latter is incremented by a unit volume $w_0$.
+The model is stopped either when a final time is reached or when the cumulated volume of links reaches a target quantity.
+This formulation can be considered both as an economic utility model and a generalisation of spatial interaction models.
+
+Note that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalise the exponents to 1 (which is different than renormalising the probability), to ensure to include convex functions.
+The ``geopolitical/sociocultural'' distance remains abstract and should be parametrised (see setup below) and estimated in the data-driven setting.
+We assume no evolution of economic size in this simple version of the model, which means that adjusted parameter are valid on a restricted time period.
+
+\subsubsection*{Model setup}
+
+Besides model dynamics, its initialisation (or parametrisation on real data) regarding the spatial distribution, sizes, countries and industrial sector compositions, is also important to understand model behavior.
+In addition  \cite{raimbault2019space} showed that the initial spatial configuration can have significant impacts on model outcomes.
+We therefore consider two versions: one with a synthetic system of cities which can be stochastically repeated, and one other based on real data.
+
+\paragraph{Synthetic setup}
+
+The synthetic system of cities is generated to be at the scale of a continent with properties similar to Europe: (i) $N=700$ cities with an empirical power-law from the Global Human Settlements database for economic sizes ($\alpha = 1.1$), distributed randomly into a uniform and isotropic square space of width $d_{max}=3000km$;
+(ii) clustered into $C = 20$ countries using a k-means algorithm (this remains consistent with non-correlated city sizes at the scale of the countries as our cities are randomly distributed in space \cite{simini2019testing});
+(iii) with synthetic distribution of sectors following log-normal distributions adjusted in a way that large cities have more high-value activities and are more diverse.
+
+More precisely for this last point, assuming sectors are distributed continuously along a one-dimensional axis, cities are assumed to have a sector distribution with a log-normal density function.
+The largest city is constrained to have a standard deviation and a mode of $1/2$ and the smallest $1/K$, with a linear interpolation between the two for other cities.
+Parameters of the log-normal law are obtained by solving the corresponding two equations system for each city, and each distribution is discretised and renormalised to be a vector of probabilities.
+See \ref{S2_Text} for all details of this synthetic setup.
+
+In synthetic experiments, the network is generated from an empty initial network until reaching $t_f=1500$, with a unit link volume of $w_0 = 1$.
+%\todo{explore effect of $t_f$ and/or $w_0$; be closer to real setup?} % -> more interesting to explore effect of urban hierarchy
+
+\paragraph{Real setup}
+
+The real setup is done using the empirical network analyzed above.
+Urban areas are distributed according to their positions in the actual geographical space, and economic size and sector composition attributes are used.
+The socio-economic distance is constructed using information from the statistical modeling, by considering fixed-effect coefficients estimated for each pair of countries.
+See \ref{S2_Text} for more details on the construction of the socio-economic distance matrix.
+
+Initial network is empty, and the real links are taken as targets for the generated network.
+We constrain the total volume of flows by setting $w_0 = \sum_{ij} w_{ij}^{(obs)} / t_{f}$ when the number of time steps $t_f$ is given as a parameter.
+
+\subsection*{Model parameters and indicators}
+
+The model parameters are the Cobb-Douglas weights $\gamma_F,\gamma_T,\gamma_W,\gamma_s$ and distance decay parameters $\gamma_G,\gamma_D$.
+As the model also includes path-dependency by integrating the influence of previous links, it can not be reduced to a simple statistical model.
+Finally, the role of space introduces even further complexity, yielding a generative model, which has to be simulated.
+
+We consider two main macroscopic indicators to quantify the generated network.
+Firstly geographical structures are captured by internationalisation (modularity of countries in the network).
+The situation of a minimised modularity corresponds to dense connections between cities of different countries of Europe and sparse connections between cities of the same European country.
+In practice, modularity is computed by
+\begin{equation}
+
+\end{equation}
+% put the actual modularity
+% add some reference values - some kind of null model
+
+Secondly we consider the metropolisation indicator as the correlation between the weighted degree and the GDP of the city.
+With this indicator we capture the effect of large cities or metropolises of regional importance attracting the biggest amount of links and involved in high interactions with other cities.
+It is computed with a Pearson correlation estimator as
+\begin{equation}
+    M = \hat{\rho} \left[d_i , E_i \right]
+\end{equation}
+
+%I think we need to keep this paragraph but shorten it, because to my understanding network average community size is tested in Figure 5, but not the Louvain modularity for ex.
+%We also consider several network indicators as Louvain modularity, sizes of maximal modularity communities, degree and link weight distributions (described by their average, hierarchy, and entropy), and correlations between degree and city size, and between link weight and distance.
+
+We take a geographical approach to study the behavior of the generative model. Other related approaches include indicators of economic complexity \cite{hidalgo2007product}. % develop
+These are less relevant in our case as the sector structure of cities does not evolve in the model.
+
+\subsection*{Empirical data}
 
@@
 
-Companies from the AMADEUS database are georeferenced using the Geonames database (using postcode or address depending on the information available).
-We then join this data with the GHS-FUA dataset for Functional Urban Areas \cite{Florczyk2019ghs}, which are the most relevant objects for applying our model.
+Companies from \emph {AMADEUS} are georeferenced using the Geonames database (using postcode or address depending on the information available).
+We then join this data with the Global Human Settlement dataset GHS-FUA for Functional Urban Areas \cite{Florczyk2019ghs} -- the most relevant objects for applying our model.
 Starting from a firm-level dataset of 1,562,862 nodes and 1,866,936 links, we obtain a directed network of 573 urban areas and 9323 ownership links.
-Weight of links are obtained by computing the owned share of turnover at destination, i.e.
+The weight of links is obtained by computing the owned share of turnover at destination, i.e.
 $w_{ij} = \sum_{k \in i,l \in j} p_{kl} \cdot T_l$ where $p_{kl}$ of the ownership share of company $l$ by company $k$ and $T_l$ the turnover of company $l$.
 Node attributes include an economic size of urban areas, that we compute as the sum of turnovers of companies within the area.
-This quantity is highly correlated with area-level GDP included in the GHS database (Spearman correlation $\rho = 0.71$) and also with population ($\rho = 0.60$), so we expect to capture size effects while keeping the consistence of a single main data source.
-An other attribute of nodes is the industrial sector composition, expressed as a proportion of turnover in the area associated to a given industrial sector.
-Detailed sectors of firms are available up to the four digits of the NACE classification in the raw data.
+This quantity is highly correlated with the GDP of the area-level included in the GHS database (Spearman correlation $\rho = 0.71$) and also with population ($\rho = 0.60$).
+We then expect to capture size effects while keeping the consistence of a single main data source.
+An other attribute of the nodes is the industrial sector composition, expressed as a proportion of turnover in the area associated to a given industrial sector.
+Detailed sectors of firms are available up to the four digits of the NACE classification in the raw data \cite{EUROSTAT2008}.
 We consider the first level classification (21 categories) and compute sector proportions accordingly.
@@
 For both, log-normal distributions appear to be a better fit ($\mu=18.8$, $\sigma=2.3$ for weighted degree; $\mu=13.8$, $\sigma=2.8$ for edge weight), and include a much broader range of the empirical distributions with lower estimated cut-offs.
-These heavy tail properties suggest the relevance of a generative model including self-reinforcing processes which are known to produce such distributions.
+These heavy tail properties suggest the relevance of a generative model including self-reinforcing processes, which are known to produce such distributions.
 
@@
 
-Regarding the community structure of the network, different modularity maximisation algorithms (run on the undirected corresponding network) give a same undirected modularity of 0.38, and a directed modularity \cite{nicosia2009extending} of 0.34 for the greedy algorithm of \cite{clauset2004finding} and 0.36 for the Louvain algorithm \cite{blondel2008fast}, corresponding to 36 communities (resp.
+Regarding the community structure of the network, different modularity maximisation algorithms (run on the undirected corresponding network) give the same undirected modularity of 0.38, and a directed modularity \cite{nicosia2009extending} of 0.34 for the greedy algorithm of \cite{clauset2004finding} and 0.36 for the Louvain algorithm \cite{blondel2008fast}, corresponding to 36 communities (resp.
 15).
@@
 Indeed, a null model assigning random labels with 30 communities, bootstrap 1000 times, yields an average modularity of $0.049 \pm 0.002$, confirming the statistical significance of these communities.
+We show in Fig.~\ref{fig:map_communities} a map of the communities obtained.
 
-To have a first insight into factors playing a role in the formation of links, we proceed to a statistical analysis similar to unconstrained spatial interaction models \cite{wilson1975some}.
-We test five different statistical models, including progressively the processes candidate to be included in the network generative model.
-The most general statistical model is
-\[
+%%%%%%%%%%%%%%
+\begin{figure}
+    \centering
+    \includegraphics{}
+    \caption{{\bf Map of network communities obtained by modularity maximisation.}}
+    \label{fig:map_communities}
+\end{figure}
+%%%%%%%%%%%%%%
+
+\subsubsection*{Statistical analysis}
+
+In order to understand which are the most determining geo-economic factors influencing creation of links between cities, we run a statistical analysis as applied in the case of unconstrained spatial interaction models \cite{wilson1975some}.
+We test five different statistical models, including progressively the processes to be included in the network generative model.
+The most general statistical model is:
+
+\begin{equation}
 \log(w_{ij}) = \log(d_{ij}) + \log(T_i) + \log(T_j) + \log(s_{ij}) + \alpha_{c_i,c_j} + \varepsilon_{ij}
-\]
-where $\alpha_{c_i,c_j}$ is a fixed effect term for the couple of countries $c_i,c_j$.
-We use a multiplicative form by taking logarithms as it is usually done for spatial interaction models.
-Furthermore, origin and/or destination constrained models are not relevant in our case as links include not only turnover but also a share of ownership.
+\end{equation}
+
+where $\alpha_{c_i,c_j}$ is a fixed effect term for the coupling of countries $c_i,c_j$.
+A multiplicative form is applied by taking logarithms as usually done for spatial interaction models.
+Furthermore, origin and/or destination constrained models are not relevant in our case, as links include information about the turnover and a share of ownership at the same time.
 Estimation results are given in Table~\ref{tab:reg}.
 We also fit a Poisson generalized linear model with a logarithmic link on the rounded weights \cite{flowerdew1988fitting}.
-We find that distance alone (model 1) has a significant negative effect but with an explained variance close to zero.
-The second model adds fixed effects by pair of countries, for which around a third are significant, and with a better $R^2$ of 0.17.
-Adding origin and destination economic size (turnovers $T_i$ and $T$, model 3) slightly improves the explanative power, but the best model among the linear ones is the model 4 with all factors and fixed effects, both in terms of explained variance ($R^2 = 0.31$) and Akaike information criterion.
-The latter indicates the absence of overfitting.
-The Poisson model with all variables is much better regarding pseudo-$R^2$ and provides much more significant estimates (all fixed effects significant and much smaller standard deviations).
-Across all models, we find consistent qualitative stylized facts: (i) distance has a negative influence; (ii) economic sizes have a positive influence, with the size of the origin being more important (what is consistent with ownership links, strongly driven by decision at the headquarters); (iii) industrial proximity has a positive influence, i.e.
-similarity is more important than complementarity for this dimension; (iv) an important part of country fixed-effects are significant; and (v) the order of magnitude of the influence of each factor is roughly the same, in the sense that no process totally dominates the others.
-This suggests the relevance of including all these different dimensions in the generative network model.
+The findings show that when distance only is taken into account (model 1), it has a significant negative effect on link creation, but with an explained variance close to zero.
+Model 2 adds fixed effects by pair of countries, for which around a third are significant, and with a better $R^2$ of 0.17.
+Adding the economic size at origin and destination (turnovers $T_i$ and $T$, model 3) slightly improves the explanative power (model 3).
+The best model among the linear ones is the model 4, with all factors and fixed effects, both in terms of explained variance ($R^2 = 0.31$) and Akaike information criterion.
+Model 4 also indicates the absence of overfitting.
+Model 5, which is a Poisson model with all variables is much a better fit regarding pseudo-$R^2$ and provides much more significant estimates (all fixed effects significant and much smaller standard deviations).
+Across all models, consistent qualitative stylized facts regarding link creation between cities are revealed: (i) distance has a negative influence; (ii) economic sizes have a positive effect, with the size of the origin city being more important (consistent in the ownership links context, as strongly driven by decision at the headquarter/owner level); (iii) industrial proximity has a positive influence, i.e.
+similarity is more important than complementarity for this dimension; (iv) an large part of country fixed-effects are statistically significant; and (v) the order of magnitude of the influence of each factor is roughly the same, which means that no process totally dominates the others.
+This confirms the relevance of including all these different dimensions in the generative network model, since they have a comparable explanatory role in statistical models.
 
 %%%%%%%%%%%%%
-\begin{table}[h]
-\caption{Ordinary Least Square estimations of statistical models explaining $\log(w_{ij})$ (standard errors in parenthesis).
+\begin{table}[!ht]
+\begin{adjustwidth}{-2.25in}{0in}
+\caption{{\bf Statistical models.} Ordinary Least Square estimations of statistical models explaining $\log(w_{ij})$ (standard errors in parenthesis).
 ***, ** and * respectively indicate 0.01, 0.05 and 0.1 levels of significance.
-For the fixed effect by origin-destination pairs of countries, we give the proportion of significant coefficients (where $p<0.1$).
+For the fixed effect by origin-destination pairs of countries, the proportion of significant coefficients is attributed (where $p<0.1$).
 Models (1) to (4) are ordinary least square models, while model (5) is a Poisson generalized linear model with logarithmic link.
 We compare model performance using $R^2$, Mean Square Error, and Akaike Information Criterion.
-For the Poisson model, AIC is not given as it is not comparable, and the pseudo-$R^2$ is given either.\label{tab:reg}}
+For the Poisson model (model 5), AIC is not given as it is not comparable, and the pseudo-$R^2$ is given either.
+\label{tab:reg}}
 \medskip
@@
 \end{tabular}
+\end{adjustwidth}
 \end{table}
@@
 
-\section{Model}
+%\todo{which model is what is not clear between the text and the table / econometric conventions}
 
-\subsection{Rationale}
+%\section*{Results}
 
-The model is expected to capture a single macroscopic urban scale, i.e.
-the level of the urban system where basic entities are cities.
-Links are induced by underlying firms but these are not explicit.
-Cities are defined by their economic size, but also by an industrial sector profile.
-As \cite{martinus2018global} point out, a combination of several distances may play a role in establishing linkages.
-Geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}.
-Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalized to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets despite the European common market (confirmed relevant by the fixed effect regression described above); (iii) city economic size, which has furthermore scaling laws properties; (iv) industrial proximity in terms of firm sector composition; (v) previous linkages (influence of history).
-This last point is crucial to capture accumulated effects and include path-dependency, which fits well the use of a simulation model instead of a more simple statistical model.
-
-\subsection{Model description}
-
-Let $1 \leq i \leq N$ cities defined at time $t$ by their economic structure such that $E_i(t)$ is the total economic volume (GDP) for city $i$ and $S_{ik}(t)$ is a matrix giving economic volume proportions within each economic sector $k$, assuming $K$ economic sectors.
-Formally, the model creates links between cities which are characterized by their economic size $E_i$ (GDP) and economic structure $S_{ik}$ in terms of activity sectors (probability distribution of firms within $K$ sectors).
-Starting with an existing network with no links, the model iteratively adds links, following a probability given by a generalized Cobb-Douglas function \cite{vilcu2011geometric} as
-
-\begin{equation}
-p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_F} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_T} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s\left(S_{ik},S_{jk}\right)^{\gamma_S} \cdot \exp \left(- \gamma_G \cdot d_{ij}\right) \cdot \exp \left(- \gamma_D \cdot g_{ij}\right)
-\end{equation}
-
-where $E  =  \sum_k E_k$, $W  = \sum_{i,j} w_{ij}$ weights of previous links, $s(S_{ik},S_{jk})$ is a similarity function between activity sectors $S_{ik}$ and $S_{jk}$ taken as a cosine similarity, $d_{ij}$ euclidian distance, and $g_{ij}$ a socio-cultural distance.
-In the case of an already existing link between two areas, the weight of the latter is incremented by a unit volume $w_0$.
-The model is stopped either when a final time is reached or when the cumulated volume of links reaches a target quantity.
-This formulation can be considered both as an economic utility model and a generalisation of spatial interaction models.
-
-Note that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalize the exponents to 1 (which is different than renormalising the probability), to ensure to include convex functions.
-The ``geopolitical/sociocultural'' distance remains abstract and should be parametrised (see setup below) and estimated in the data-driven setting.
-We assume no evolution of economic size in this simple version of the model, which means that adjusted parameter are valid on a restricted time period.
-
-\subsection{Model setup}
-
-Besides model dynamics, its initialisation (or parametrisation on real data) regarding the spatial distribution, sizes, countries and industrial sector compositions, is also important to understand model behavior.
-In addition  \cite{raimbault2019space} showed that the initial spatial configuration can have significant impacts on model outcomes.
-We therefore consider two versions: one with a synthetic system of cities which can be stochastically repeated, and one other based on real data.
-
-\paragraph{Synthetic setup}
-
-The synthetic system of cities is generated to be at the scale of a continent with properties similar to Europe: (i) $N=700$ cities with an empirical power-law from the Global Human Settlements database for economic sizes ($\alpha = 1.1$), distributed randomly into a uniform and isotropic square space of width $d_{max}=3000km$; (ii) clustered into $C = 20$ countries using a k-means algorithm (this remains consistent with non-correlated city sizes at the scale of the countries as our cities are randomly distributed in space \citep{simini2019testing}); (iii) with synthetic distribution of sectors following log-normal distributions adjusted in a way that large cities have more high-value activities and are more diverse.
-More precisely for this last point, assuming sectors are distributed continuously along a one-dimensional axis, cities are assumed to have a sector distribution with a log-normal density function.
-The largest city is constrained to have a standard deviation and a mode of $1/2$ and the smallest $1/K$, with a linear interpolation between the two for other cities.
-Parameters of the log-normal law are obtained by solving the corresponding two equations system for each city, and each distribution is discretised and renormalised to be a vector of probabilities.
-In synthetic experiments, the network is generated from an empty initial network until reaching $t_f=1500$, with a unit link volume of $w_0 = 1$.
-
-\paragraph{Real setup}
-
-The real setup is done using the empirical network analyzed above.
-Urban areas are distributed according to their positions in the actual geographical space, and economic size and sector composition attributes are used.
-The socio-economic distance is constructed using information from the statistical modeling, by considering fixed-effect coefficients estimated for each pair of countries.
-Initial network is empty, and the real links are taken as targets for the generated network.
-We constrain the total volume of flows by setting $w_0 = \sum_{ij} w_{ij} / t_{f}$ when the number of time steps $t_f$ is given as a parameter.
-
-\subsection{Model parameters and indicators}
-
-The model parameters are the Cobb-Douglas weights $\gamma_F,\gamma_T,\gamma_W,\gamma_s$ and distance decay parameters $\gamma_G,\gamma_D$.
-The model also includes path-dependency by integrating the influence of previous links.
-Therefore it can not be reduced to a simple statistical model.
-Finally, the role of space introduces even further complexity, yielding a generative model, which has to be simulated.
-
-We consider several macroscopic indicators to quantify the generated network.
-Firstly, geographical structures are captured by (i) internationalisation (modularity of countries in the network); (ii) metropolisation (correlation between weighted degree and city size).
-Two other thematic indicators, which are regionalisation (correlation between length and flow of links, stratified by size of extremities), and specialisation (correlation between sector proximity and flow of links, stratified by size of extremities), are not included in this study for the sake of simplicity as each has a high dimension.
-We also consider several network indicators as Louvain modularity, sizes of maximal modularity communities, degree and link weight distributions (described by their average, hierarchy, and entropy), and correlations between degree and city size, and between link weight and distance.
-
-\section{Model exploration and calibration}
+\section*{Model exploration and calibration}
 
@@
 
-\subsection{Synthetic city systems}
+\subsection*{Synthetic city systems}
 
@@
 Indices were estimated with a Sobol sequence sampling of 20,000 model runs \cite{saltelli2010variance}.
-We give in Table~\ref{tab:saltelli} estimated results.
-We find that indicators related to modularity (internationalisation and optimal modularity) are mostly influenced by geographical parameters.
+We give in \ref{S4_Text} the full estimated results.
+To summarize, we find that indicators related to modularity (internationalisation and optimal modularity) are mostly influenced by geographical parameters.
 On the contrary, indicators of network structure are driven by origin and destination sizes.
-Sectors and link history have a second order by not negligible role.
+The influence of sectors and historical links has a secondary but not negligible role.
 Altogether, the analysis witnesses of strong interaction effects between parameters, since for several indicators the total order index is one order of magnitude larger than the first order index.
 This means that the generative model captures some complex behavior of the system.
-The very different important parameters depending on outputs studied confirms the relevance of a more exhaustive grid exploration.
-
-%%%%%%%%%%%%%
-\begin{table}
-\caption{Saltelli sensitivity indices, for indicators in rows and parameters in columns. We give for each pair the first order index (F) and the total order index (T).\label{tab:saltelli}}
-\hspace{-1cm}\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|c|c|}
-\hline
- & \multicolumn{2}{|c|}{$\gamma_G$} & \multicolumn{2}{|c|}{$\gamma_D$} & \multicolumn{2}{|c|}{$\gamma_S$} & \multicolumn{2}{|c|}{$\gamma_W$} & \multicolumn{2}{|c|}{$\gamma_F$} & \multicolumn{2}{|c|}{$\gamma_T$} \\
- & F & T & F & T & F & T & F & T & F & T & F & T \\
- \hline
-Internationalisation & 0.2 & 0.3 & 0.7 & 0.7 & 0.001 & 0.009 & $4\cdot 10^{-4}$ & 0.007 & 0.03 & 0.04 & 0.02 & 0.04 \\
-Metropolisation & 0.02 & 0.1 & 0.02 & 0.2 & 0.002 & 0.1 & 0.001 & 0.09 & 0.2 & 0.6 & 0.3 & 0.6 \\
-Modularity & 0.3 & 0.4 & 0.6 & 0.6 & 0.004 & 0.02 & $3\cdot 10^{-4}$ & 0.01 & 0.005 & 0.03 & 0.002 & 0.03 \\
-Avg. com. size & 0.008 & 0.09 & 0.01 & 0.1 & 0.002 & 0.07 & 0.003 & 0.04 & 0.3 & 0.6 & 0.4 & 0.6 \\
-Degree entropy & 0.006 & 0.02 & 0.003 & 0.02 & 0.006 & 0.03 & 0.008 & 0.02 & 0.5 & 0.5 & 0.5 & 0.5 \\
-Weight entropy & 0.04 & 0.1 & 0.03 & 0.1 & 0.008 & 0.08 & 0.01 & 0.07 & 0.4 & 0.5 & 0.4 & 0.5 \\\hline
-\end{tabular}
-\end{table}
-%%%%%%%%%%%%%
+% The very different important parameters depending on outputs studied confirms the relevance of a more exhaustive grid exploration.
 
@@
 Other indicators exhibit non-trivial patterns.
-For example, when considering average community size of the final network as shown in Fig.~\ref{fig:comsize}, we obtain a maximal integration in term of communities at an intermediate value of the gravity decay, what can be interpreted as the emergence of a regional regime.
+For example, when considering average community size of the final network as shown in Fig.~\ref{fig:comsize}, we obtain a maximal integration in terms of communities at an intermediate value of the gravity decay, what can be interpreted as the emergence of a regional regime.
 The size of communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.
@@
     \includegraphics[width=\textwidth]{figures/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
-	\caption{Maximal integration in term of community size is achieved at an intermediate value of $d_G$: emergence of a regional regime; Maximal size depends on the role of sectors $\gamma_S$, in a decreasing way when origin size is deactivated, and increasing way when $\gamma_F=1$; This regime disappear when origin size influence is too large.\label{fig:comsize}}
+	\caption{Maximal integration in terms of community size is achieved at an intermediate value of $d_G$: emergence of a regional regime; Maximal size depends on the role of sectors $\gamma_S$, in a decreasing way when origin size is deactivated, and increasing way when $\gamma_F=1$; This regime disappear when origin size influence is too large.\label{fig:comsize}}
 \end{figure}
@@
 
-\subsection{Model calibration on the European urban network}
+\paragraph{Impact of urban hierarchy}
+
+We run a specific experiment to study the impact of urban hierarchy on firms dynamics.
+
+\subsection*{Model calibration on the European urban network}
 
@@
 	\begin{center}
-    \includegraphics[width=0.5\textwidth]{figures/pareto_colorgammaOrigin.png}
+    \includegraphics[width=0.75\textwidth]{figures/pareto_colorgammaOrigin.png}
     \end{center}
@@
 
+\subsection*{Application: impact of economic shock}
+
+Based on calibrated parameter values, we design scenarios to evaluate the impact of economic shocks.
+These can be the consequence of diverse factors, including socio-economic, geopolitical factors (example of Brexit), and natural disasters such as COVID-19.
+We simulate a full border closure by setting $\gamma_D = 0$, and intermediate scenarios.
+Setting also $\gamma_G$ to 0 allows simulating inter-country lock-downs.
+% run simulations
+% note that time scales are in fact smaller - or assume that crisis keeping border closed is rather long
+
 %%%%%%%%%%%%%%%
-\section{Discussion}
+\section*{Discussion}
 
-This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMole platform and calibrated on real data on ownership linkages of firms for Europe.
+This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMole platform and calibrated on real data on ownership linkages of firms for Europe from Amadeus dataset.
 The simulation on a synthetic system of cities provides an broad knowledge on model behavior in its parameter space, and its calibration suggests its potential application to policy issues.
-This kind of model can indeed have practical applications in the future in terms of testing the effect of exogenous shocks in the socio-economic structure as the upcoming effect of Brexit on the European market.
+This kind of model can therefore have practical applications in the future in terms of testing the effect of exogenous shocks in the socio-economic structure: as the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to COVID19 on the single market and on the strongly foreign dependent economic sectors (see Appendix).
 
@@
 Our study shows how crucial model exploration is to overcome hidden parameters (deactivated mechanisms or default parameter values).
-This paper emphasis on the fact that exploration of intrinsic dynamics on synthetic data is of great importance should be driven before an application on real data so as to put aside the geographical effects from model dynamics.
+This paper emphasis on the fact that exploration of intrinsic dynamics on synthetic data is of great importance and should be driven before an application on real data so as to put aside the geographical effects from model dynamics.
 
@@
 
+\section*{Supporting information}
+
+\paragraph*{S1 Text.}
+\label{S1_Text}
+{\bf UK's economic sectors dependent on international links.}
+
+\paragraph*{S2 Text.}
+\label{S2_Text}
+{\bf Socio-economic distance matrix.}
+
+\paragraph*{S3 Text.}
+\label{S3_Text}
+{\bf Synthetic model setup.}
+
+\paragraph*{S4 Text.}
+\label{S4_Text}
+{\bf Sensitivity analysis.}
+
+\section*{Acknowledgments}
+
+Results obtained in this paper were computed on the vo.complex-system.eu virtual organization of the European Grid Infrastructure (http://www.egi.eu).
+We thank the European Grid Infrastructure and its supporting National Grid Initiatives (France-Grilles in particular) for providing the technical support and infrastructure.
+
+\bibliographystyle{plos2015}
+\bibliography{biblio}
+
+%%%%%%%%%%%%%
+\newpage
+\section*{S1 Text: UK's economic sectors dependent on international links}
+
+Based on \emph {AMADEUS} dataset, the most vulnerable economic sectors in the United Kingdom have been highlighted.
+The later are the ones due to be mostly affected by eventual future lock-downs of the international economy (international ownership links).
+
+When considering the UK sectors generating the highest turnover in 2018 mostly dependent on foreign capital they own (out-going links from the UK) these are extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods.
+In terms of in-going links to the UK, activities of head offices and holding companies, sale of cars and manufacture of motor vehicles are sectors mostly dependent on foreign ownership coming from abroad.
+
+\FloatBarrier
+
+\begin{table}[!ht]
+\begin{adjustwidth}{-2.25in}{0in}
+\caption{Top 10 UK industries depending on foreign ownership expressed in total turnover (out-going links from the UK)\label{tab:uk_outgoing}}
+\centering
+\medskip
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+Classification  & NACE Industries & Total turnover \\ 
+\hline
+1 &     Extraction of crude petroleum &   96,458,258,642  \\
+\hline
+2 & Manufacture of tobacco products &  5,477,272,464  \\
+\hline
+3 & Manufacture of basic pharmaceutical products &  4,825,881,930 \\
+\hline
+4 &  Retail sale in non-specialised stores with food, beverages or tobacco predominating  &   3,681,373,374  \\
+\hline
+5 &  Wireless telecommunications activities &   3,379,297,218  \\
+\hline
+6 & Event catering activities & 2,871,593,754  \\
+\hline
+7 &  Other mining and quarrying n.e.c. &  1,590,548,873  \\
+\hline
+8 & Business and other management consultancy activities &  1,394,373,014 \\
+\hline
+9 &  Activities of head offices &  1,322,692,365 \\
+\hline
+10 & Manufacture of air and spacecraft and related machinery &  1,292,065,497 \\
+\hline
+\end{tabular}
+\end{adjustwidth}
+\end{table}
+
+\begin{table}[!ht]
+\begin{adjustwidth}{-2.25in}{0in}
+\caption{Top 10 UK industries depending on foreign ownership expressed in total turnover (in-going links to the UK)\label{tab:uk_ingoing}}
+\centering
+\medskip
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+Classification  & NACE Industries & Total turnover \\ 
+\hline
+1 &      Activities of head offices &  358,099,870  \\
+\hline
+2 & Activities of holding companies  & 259,008,123  \\
+\hline
+3 & Other business support service activities n.e.c. & 183,098,170 \\
+\hline
+4 &  Security and commodity contracts brokerage  &  72,482,167 \\
+\hline
+5 &  Manufacture of motor vehicles  &   65,162,757 \\
+\hline
+6 &  Sale of cars and light motor vehicles  &    61,019,026  \\
+\hline
+7 &  Other financial service activities, except insurance and pension funding & 47,003,937 \\
+\hline
+8 &  Other personal service activities n.e.c.  & 37,926,668  \\
+\hline
+9 &  Wholesale of solid, liquid and gaseous fuels and related products & 36,693,362 \\
+\hline
+10 &  Other information technology and computer service activities  &  35,668,779 \\
+\hline
+\end{tabular}
+\end{adjustwidth}
+\end{table}
+
+\newpage
+When expressed in total number of links, activities mostly dependent on out-going links from the UK are activities of holding companies, of head offices and fund management activities.
+In terms of in-going links these are business and consultancy activities and renting and operating of own and leased real estate.
+
+\begin{table}[!ht]
+\begin{adjustwidth}{-2.25in}{0in}
+\caption{Top 10 UK industries depending on foreign ownership expressed in total links (out-going links from the UK)\label{tab:uk_outgoinglinks}}
+\centering
+\medskip
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+Classification  & NACE Industries & Total links \\ 
+\hline
+1 & Activities of holding companies &   686  \\
+\hline
+2 & Other business support service activities n.e.c.  &  567  \\
+\hline
+3 & Activities of head offices &  503 \\
+\hline
+4 &  Fund management activities &   407  \\
+\hline
+5 &  Business and other management consultancy activities &  398  \\
+\hline
+6 &  Extraction of crude petroleum &  366  \\
+\hline
+7 &  Other activities auxiliary to financial services, except insurance and pension funding & 240 \\
+\hline
+8 &  Development of building projects  & 221 \\
+\hline
+9 &  Manufacture of basic pharmaceutical products &  212 \\
+\hline
+10 &  Other information technology and computer service activities  &  181 \\
+\hline
+\end{tabular}
+\end{adjustwidth}
+\end{table}
+
+\begin{table}[!ht]
+\begin{adjustwidth}{-2.25in}{0in}
+\caption{Top 10 UK industries depending on foreign ownership expressed in total links (in-going links to the UK)\label{tab:uk_ingoinglinks}}
+\centering
+\medskip
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+Classification  & NACE Industries & Total links \\ 
+\hline
+1 &      Other business support service activities n.e.c. &   16,140  \\
+\hline
+2 & Renting and operating of own or leased real estate  &  6,320  \\
+\hline
+3 & Business and other management consultancy activities &  6,094  \\
+\hline
+4 &  Activities of head offices &   5,628  \\
+\hline
+5 &  Computer consultancy activities &   5,047  \\
+\hline
+6 &  Activities of holding companies &  4,728  \\
+\hline
+7 &  Other personal service activities n.e.c. &  3,996 \\
+\hline
+8 &  Development of building projects &  3,946 \\
+\hline
+9 &  Other information technology and computer service activities &  3,757 \\
+\hline
+10 &  Other financial service activities, except insurance and pension funding  &  3,321 \\
+\hline
+\end{tabular}
+\end{adjustwidth}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%
 \newpage
 
-%\bibliographystyle{apalike}
-\bibliographystyle{unsrt}
-\bibliography{biblio}
+\section*{S2 Text: Socio-economic distance matrix}
+
+% \todo{more detail / descriptive stats of the matrix / function of the distance between countries}
+
+%%%%%%%%%%%%%%%%%%
+\newpage
+
+\section*{S3 Text: Synthetic model setup}
+
+The rationale for the synthetic sector setup is the following:
+\begin{itemize}
+    \item 
+\end{itemize}
+
+%solve for mu,sigma of the log-normal as a function of gdp
+%       we aim at having a log-normal mostly in [0,1] such that
+%       (mode,variance) =  1 / K for the smallest log(E_i) and = 1/2 for the largest log(E_i)
+%        writing
+%        e_i = (log(E_i) - log(E_imin))/(log(E_imax) - log(E_imin)) * (1/2 - 1/K) + 1/K
+%        this yields
+%         (1)  mu - sigma^2 = log(e_i)
+%         (2) -3 sigma^2 - 2 log(exp(sigma^2) - 1) = log(e_i)
+%        => sigma^2 is the unique positive root of f(X)=0 with f(X) = -3X - 2 ln(exp(X) - 1) - ln(e_i)
+
+%%%%%%%%%%%%%%%%%%
+\newpage
+
+\section*{S4 Text: Sensitivity analysis}
+
+%%%%%%%%%%%%%
+\begin{table}
+\begin{adjustwidth}{-2.25in}{0in}
+\caption{Saltelli sensitivity indices, for indicators in rows and parameters in columns. We give for each pair the first order index (F) and the total order index (T).\label{tab:saltelli}}
+\hspace{-1cm}\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|c|c|}
+\hline
+ & \multicolumn{2}{|c|}{$\gamma_G$} & \multicolumn{2}{|c|}{$\gamma_D$} & \multicolumn{2}{|c|}{$\gamma_S$} & \multicolumn{2}{|c|}{$\gamma_W$} & \multicolumn{2}{|c|}{$\gamma_F$} & \multicolumn{2}{|c|}{$\gamma_T$} \\
+ & F & T & F & T & F & T & F & T & F & T & F & T \\
+ \hline
+Internationalisation & 0.2 & 0.3 & 0.7 & 0.7 & 0.001 & 0.009 & $4\cdot 10^{-4}$ & 0.007 & 0.03 & 0.04 & 0.02 & 0.04 \\
+Metropolisation & 0.02 & 0.1 & 0.02 & 0.2 & 0.002 & 0.1 & 0.001 & 0.09 & 0.2 & 0.6 & 0.3 & 0.6 \\
+Modularity & 0.3 & 0.4 & 0.6 & 0.6 & 0.004 & 0.02 & $3\cdot 10^{-4}$ & 0.01 & 0.005 & 0.03 & 0.002 & 0.03 \\
+Avg. com. size & 0.008 & 0.09 & 0.01 & 0.1 & 0.002 & 0.07 & 0.003 & 0.04 & 0.3 & 0.6 & 0.4 & 0.6 \\
+Degree entropy & 0.006 & 0.02 & 0.003 & 0.02 & 0.006 & 0.03 & 0.008 & 0.02 & 0.5 & 0.5 & 0.5 & 0.5 \\
+Weight entropy & 0.04 & 0.1 & 0.03 & 0.1 & 0.008 & 0.08 & 0.01 & 0.07 & 0.4 & 0.5 & 0.4 & 0.5 \\\hline
+\end{tabular}
+\end{adjustwidth}
+\end{table}
+%%%%%%%%%%%%%
 

### 2020-08-31 main.tex
Update on Overleaf.

@@
 In synthetic experiments, the network is generated from an empty initial network until reaching $t_f=1500$, with a unit link volume of $w_0 = 1$.
-%\todo{explore effect of $t_f$ and/or $w_0$; be closer to real setup?} % -> more interesting to explore effect of urban hierarchy
 
@@
 
-The model parameters are the Cobb-Douglas weights $\gamma_F,\gamma_T,\gamma_W,\gamma_s$ and distance decay parameters $\gamma_G,\gamma_D$.
+The model parameters are the Cobb-Douglas weights $\gamma_O,\gamma_D,\gamma_W,\gamma_S$ and distance decay parameters $d_0,c_0$.
 As the model also includes path-dependency by integrating the influence of previous links, it can not be reduced to a simple statistical model.
@@
 Model 2 adds fixed effects by pair of countries, for which around a third are significant, and with a better $R^2$ of 0.17.
-Adding the economic size at origin and destination (turnovers $T_i$ and $T$, model 3) slightly improves the explanative power (model 3).
+Adding the economic size at origin and destination (turnovers $T_i$ and $T$, model 3) slightly improves the explanatory power (model 3).
 The best model among the linear ones is the model 4, with all factors and fixed effects, both in terms of explained variance ($R^2 = 0.31$) and Akaike information criterion.

### 2020-09-01 main.tex
Update on Overleaf.

@@
 \begin{figure}
-    \centering
-    \includegraphics{}
-    \caption{{\bf Map of network communities obtained by modularity maximisation.}}
+    \begin{center}
+    \includegraphics[width=\linewidth]{figures/map_communities_louvain.png}
+    \end{center}
+    \caption{{\bf Map of network communities obtained by modularity maximisation.} Area of circles gives the population with the Functional Urban Area, and color the community it belongs to.
+Urban Areas in communities of size smaller than 3 were not colored for readability.}
     \label{fig:map_communities}
@@
 %%%%%%%%%%%%%
-
-%\todo{which model is what is not clear between the text and the table / econometric conventions}
-
-%\section*{Results}
 
@@
 We run a specific experiment to study the impact of urban hierarchy on firms dynamics.
+Rank-size law in cities is a central stylized fact to urban theories, despite the discussions on how to estimate the power-law exponent, its actual values \cite{cottineau2017metazipf} and how its depends on the definition of cities considered \cite{corral2020truncated}.
+In that particular context, understanding how this parameter can impact urban dynamics within artificial systems of cities in which it can be modified to extreme values not observed in real systems, should inform on underlying processes.
+% too abstract: clarify why hypothetical systems of cities bring info on real ones
+We change the initial hierarchy $\alpha$ from 0.2 to 2.0 with a step of 0.1, and also vary interaction range $d_0$.
 
@@
 
-When considering optimal points with $\varepsilon_L < 5.0$, the adjusted parameters are: long range gravity interactions and socio-economic interactions given by $d_G = 1/\gamma_G = 5205 \pm 1775$ and $d_D = 1/ \gamma_D = 109 \pm 29$ (the maximal $g_{ij}$ is 11.8); a stronger influence of destination than origin as $\gamma_F = 1.45 \pm 0.09$ and $\gamma_T = 2.0 \pm 0.06$; a comparable influence of sectors as $\gamma_S = 1.75 \pm 0.17$; a fine time granularity as $t_f = 4960 \pm 109$; and a very strong influence of previous links as $\gamma_W = 7.99 \pm 3.77$.
+When considering optimal points with $\varepsilon_L < 5.0$, the adjusted parameters are: long range gravity interactions and socio-economic interactions given by $d_0 = 5205 \pm 1775$ and $c_0 = 109 \pm 29$ (the maximal $c_{ij}$ is 11.8); a stronger influence of destination than origin as $\gamma_O = 1.45 \pm 0.09$ and $\gamma_D = 2.0 \pm 0.06$; a comparable influence of sectors as $\gamma_S = 1.75 \pm 0.17$; a fine time granularity as $t_f = 4960 \pm 109$; and a very strong influence of previous links as $\gamma_W = 7.99 \pm 3.77$.
 This last point furthermore shows in particular a role of the path-dependency parameter, confirming the relevance of this complex generative model including path-dependency.

### 2020-09-01 communications/abstract_spatialdatascience.tex
reorg_repo

--- 2019-11-15 extended_abstract.tex
@@
 
-%\begin{abstract}
-%    We introduce a generative network model for urban networks of interactions between firms.
-The growth of the network is simulated by adding new firm ownership links randomly, with probabilities determined by several factors, namely the economic size of origin and destination urban areas, industrial sector proximity, the strength of previous links, and geographical and socio-cultural distance.
-Simulation on synthetic systems of cities unveil phase transitions when changing interaction distance, and non-trivial patterns in model behavior.
-Future work include the calibration of the model on real company links for Europe.
-%\end{abstract}
-
 \vspace{-0.5cm}
 
-The emergence of a global network of cities, defined by interactions between firms, is a salient feature of recent globalization processes \cite{taylor2001specification}.
-Understanding the drivers for growth for such a network, can be used to foster innovation through policies for local economic clusters \cite{turkina2016structure}, development of strategic industries for global insertion of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
-Linkages are characterized by interactions of different nature and determined by economic, socio-cultural, or geopolitical proximity \cite{martinus2018global}.
+The emergence of a global network of cities, defined by interactions between firms, is at the heart of recent globalization processes \cite{taylor2001specification}.
+Understanding growth mechanisms of such a network is crucial to foster innovation, shape local economy policies \cite{turkina2016structure}, and global industrial strategies for regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
+Urban networks are characterized by interactions of different nature and determined by economic, socio-cultural, or geopolitical proximity \cite{martinus2018global}.
+This paper focuses on economic relations between urban areas defined by ownership links of firms.
+Since transnational firms structure is one of the drivers of the global economic space, it is one of the proxies to unveil geographical structures \cite{2019arXiv191014652Z}.
 
-This paper focuses on economic relations between urban areas defined by ownership links of firms.
-Since transnational firms structure is one of the determinant of the global economic space, it is one of the proxies to unveil geographical structures \cite{2019arXiv191014652Z}.
-This work introduces a generative network model to simulate the growth of such linkages at urban areas scale within the evolutionary systems of cities framework \cite{pumain2018evolutionary}.
-We combine multiple factors which influence link formation, as economic intensity of origin and destination of urban areas, industrial sector proximity, the strength of previous links, and geographical and socio-cultural distance.
+This work introduces a generative network model to simulate the growth of such links at urban areas scale within the evolutionary systems of cities framework \cite{pumain2018evolutionary}.
+We combine multiple factors which influence link formation, as economic intensity of the origin and destination of urban areas, industrial sector proximity between cities, the strength of links from the past, and geographical and socio-cultural distance.
 The model allows thus to compare the effect of different factors on the final network structure, and to extrapolate their actual effect when calibrated on real data.
-
 Formally, the model creates links between cities which are characterized by their economic size $E_i$ (GDP) and economic structure $S_{ik}$ in terms of activity sectors (probability distribution of firms within $K$ sectors).
@@
 where $E  =  \sum_k E_k$, $W  = \sum_{i,j} w_{ij}$ weights of previous links, $s(S_{ik},S_{jk})$ is similarity function between activity sectors $S_{ik}$ and $S_{jk}$, $d_{ij}$ euclidian distance, and $g_{ij}$ a socio-cultural distance.
-In the case of an already existing link between two areas, the weight of the latter is incremented by one.
-The model is stopped either when a final time is reached or when a given number of links have been added.
-This formulation can be considered both as an economic utility and a generalization of spatial interaction models.
-The model parameters are the Cobb-Douglas weights $\gamma_F,\gamma_T,\gamma_W,\gamma_s$ and distance decay parameters $\gamma_G,\gamma_D$.
-The model also includes path-dependency by integrating the influence of previous links.
-Therefore it can not be reduced to a simple statistical model.
-Finally, the role of space introduces even further complexity, yielding a generative model, which has to be simulated.
 
@@
 We build the model on a synthetic system of cities, at the scale of a continent with properties similar to Europe: (i) 700 cities with an empirical power-law from the Global Human Settlements database for economic sizes; (ii) clustered into 30 countries; (iii) with synthetic distribution of sectors following log-normal distributions adjusted in a way that large cities have more high-value activities and are more diverse.
-The network is generated from an empty initial network until reaching $t_f=1500$.
-A first numerical experiment of one-factor sampling on all Cobb-Douglas parameters and 100 stochastic repetitions confirms a good statistical convergence (average Sharpe ratios for indicators all larger than 5).
-The model behaviour is then studied with a grid sampling for parameters and 20 repetitions (the computations run on the European Grid Infrastructure, and are equivalent to 2.5 years of CPU time).
 
-We show in Fig.
-1 examples of generated networks, with very contrasted patterns obtained with different spatial interaction ranges.
-This behavior is confirmed when plotting the internationalization index against the distance, which shows an exponential decay.
-Similarly, the correlation between city weighted degree and city size exhibit a behaviour that can be interpreted as a phase transition between local and global networks.
-Other indicators exhibit non-trivial patterns.
-For example, when considering average community size of the final network, we obtain a maximal integration in term of communities at an intermediate value of the gravity decay, what can be interpreted as the emergence of a regional regime.
-The size of communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.
-Altogether, these experiments reveal a complex interplay between processes and how the model produces diverse stylized facts.
-
-As a matter of validation, the model is implemented for real data for Europe joining Functional Urban Areas with the Global Human Settlement Database for cities and their characteristics, with the AMADEUS database produced by the Bureau Van Dijk on ownership links between firms in Europe between 2010 and 2018 at postcode level in all types of sectors.
+As a matter of validation, the model is compared with real data for Europe : Functional Urban Areas from the Global Human Settlement Database for cities and their characteristics, and the AMADEUS database produced by the Bureau Van Dijk on ownership links between firms in Europe between 2010 and 2018 at postcode level in all types of sectors.
 Calibration is done using a genetic NSGA2 algorithm, finding the parameter values giving network trajectories closest to reality.
 First results show in particular a role of the path-dependency parameter confirming the relevance of this complex generative model.
-Further work is needed to get more interpretations from this experiment, but our study demonstrates the feasibility of model calibration on real data.
+Further work is needed, although our study demonstrates the feasibility of model calibration on real data.
 Several scenarios are considered taking into consideration the current and future economic challenges, such as those given by Brexit.
@@
 
-\begin{figure}
-\vspace{-2cm}
-\begin{center}
-    \begin{minipage}[c]{0.25\textwidth}
-        \includegraphics[width=\textwidth]{figures/ex_alleq-lowgravity_seed-12102_t1500.png}\\
-        \includegraphics[width=\textwidth]{figures/ex_alleq-highgravity_seed-12102_t1500.png}
-    \end{minipage}
-    \begin{minipage}[c]{0.6\linewidth}
-     \includegraphics[width=0.48\textwidth]{figures/internationalization-gravityDecay_errorbars.png}
-    \includegraphics[width=0.48\textwidth]{figures/rhoDegreeSize-gravityDecay_errorbars.png}\\
-    \includegraphics[width=\textwidth]{figures/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
-    \end{minipage}
-    \end{center}
-    \vspace{-0.5cm}
-    \caption{\footnotesize\textbf{(Left column)} Example of simulated networks for a synthetic system of cities, with low gravity range (top) and high gravity range (bottom); \textbf{(Top right)} Internationalization and correlation between size and degree as a function of decay parameter; \textbf{(Bottom right)} Average network community size as a function of decay parameter, for varying influence of sectors ($\gamma_S$ in color) and of origin ($\gamma_F$ in columns).}
-\end{figure}
+%\begin{figure}
+%\vspace{-2cm}
+%\begin{center}
+%    \begin{minipage}[c]{0.25\textwidth}
+%        \includegraphics[width=\textwidth]{figures/ex_alleq-lowgravity_seed-12102_t1500.png}\\
+%        \includegraphics[width=\textwidth]{figures/ex_alleq-highgravity_seed-12102_t1500.png}
+%    \end{minipage}
+%    \begin{minipage}[c]{0.6\linewidth}
+%     \includegraphics[width=0.48\textwidth]{figures/internationalization-gravityDecay_errorbars.pn%g}
+%    \includegraphics[width=0.48\textwidth]{figures/rhoDegreeSize-gravityDecay_errorbars.png}\\
+%    \includegraphics[width=\textwidth]{figures/networkAvgCommunitySize_countryGravityDecay2100_gam%maDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+%    \end{minipage}
+%    \end{center}
+%    \vspace{-0.5cm}
+%    \caption{\footnotesize\textbf{(Left column)} Example of simulated networks for a synthetic %system of cities, with low gravity range (top) and high gravity range (bottom); \textbf{(Top %right)} Internationalization and correlation between size and degree as a function of decay %parameter; \textbf{(Bottom right)} Average network community size as a function of decay %parameter, for varying influence of sectors ($\gamma_S$ in color) and of origin ($\gamma_F$ in %columns).}
+%\end{figure}
 

### 2020-09-01 figures.sh

@@ -1,0 +1,18 @@
+
+JPGQUALITY=50
+PDFRESOLUTION=200
+WIDTH=1000
+HORIZONTALPADDING=10
+VERTICALPADDING=10
+
+# Fig 1
+montage figuresraw/edgeweight.png figuresraw/degreeDistr.png -tile 2x1 -geometry +"$HORIZONTALPADDING"+0 -resize "$WIDTH"x figures/Fig1.png
+
+# Fig 2
+cp figuresraw/map_communities_louvain.png figures/Fig2.png
+
+# Fig 3
+montage figuresraw/ex_alleq-highgravity_seed-12102_t1500.png figuresraw/ex_alleq-lowgravity_seed-12102_t1500.png figuresraw/ex_real_highgravity_t1500.png figuresraw/ex_real_lowgravity_t1500.png -tile 2x2 -geometry +"$HORIZONTALPADDING"+"$VERTICALPADDING" -resize "$WIDTH"x figures/Fig3.png
+
+# Fig 4
+montage figuresraw/internationalisation-gravityDecay_errorbars.png figuresraw/metropolisation-gravityDecay_errorbars.png figuresraw/internationalisation-gammaSectors_errorbars.png figuresraw/metropolisation-gammaSectors_errorbars.png -tile 2x2 -geometry +"$HORIZONTALPADDING"+"$VERTICALPADDING" -resize "$WIDTH"x figures/Fig4.png

### 2020-09-01 figures.sh
figs

@@ -16,3 +16,9 @@ montage figuresraw/ex_alleq-highgravity_seed-12102_t1500.png figuresraw/ex_alleq
 
 # Fig 4
 montage figuresraw/internationalisation-gravityDecay_errorbars.png figuresraw/metropolisation-gravityDecay_errorbars.png figuresraw/internationalisation-gammaSectors_errorbars.png figuresraw/metropolisation-gammaSectors_errorbars.png -tile 2x2 -geometry +"$HORIZONTALPADDING"+"$VERTICALPADDING" -resize "$WIDTH"x figures/Fig4.png
+
+# Fig 5
+cp figuresraw/internationalization_countryGravityDecay200_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png figures/Fig5.png
+
+# Fig 6
+cp figuresraw/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png figures/Fig6.png

### 2020-09-01 main.tex

@@
 The situation of a minimised modularity corresponds to dense connections between cities of different countries of Europe and sparse connections between cities of the same European country.
-In practice, modularity is computed by
+In practice, directed modularity is computed by
 \begin{equation}
@@
 
-We take a geographical approach to study the behavior of the generative model. Other related approaches include indicators of economic complexity \cite{hidalgo2007product}. % develop
+We take a geographical approach to study the behavior of the generative model.
+Other related approaches include indicators of economic complexity, estimating for example the diversity of products produced by countries and how they depend on each other \cite{hidalgo2007product}.
 These are less relevant in our case as the sector structure of cities does not evolve in the model.
@@
 \subsection*{Empirical data}
-
-% stats to describe the structure of the database, nb of total links, nodes, missing values? + nb of firms per country,  total turnover generated per country etc..
 
@@
 \begin{figure}
-    \centering
-    \includegraphics[width=0.49\textwidth]{figures/edgeweight.png}
-    \includegraphics[width=0.49\textwidth]{figures/degreeDistr.png}
+    \begin{center}
+        \includegraphics[width=\linewidth]{figures/Fig1.png}
+    \end{center}
     \caption{Distributions of network properties.
@@
 Indeed, a null model assigning random labels with 30 communities, bootstrap 1000 times, yields an average modularity of $0.049 \pm 0.002$, confirming the statistical significance of these communities.
-We show in Fig.~\ref{fig:map_communities} a map of the communities obtained.
+We show in Fig.~\ref{fig:fig2} a map of the communities obtained.
 
@@
     \begin{center}
-    \includegraphics[width=\linewidth]{figures/map_communities_louvain.png}
+    \includegraphics[width=\linewidth]{figures/Fig2.png}
     \end{center}
@@
 Urban Areas in communities of size smaller than 3 were not colored for readability.}
-    \label{fig:map_communities}
+    \label{fig:fig2}
 \end{figure}
@@
 Large simulation result files are available on the dataverse repository at \url{https://doi.org/10.7910/DVN/UPX23S}.
-We show in Fig.~\ref{fig:example} examples of generated networks, with very contrasted patterns obtained with different spatial interaction ranges.
+We show in Fig.~\ref{fig:fig3} examples of generated networks, with very contrasted patterns obtained with different spatial interaction ranges.
 Low values of the gravity decay parameter yield local networks around main cities, while long interaction ranges yield more integrated networks.
@@
 \begin{figure}
-    \includegraphics[width=0.48\textwidth]{figures/ex_alleq-highgravity_seed-12102_t1500.png}
-    \includegraphics[width=0.48\textwidth]{figures/ex_alleq-lowgravity_seed-12102_t1500.png}\\
-    \begin{minipage}{0.48\textwidth}
-    \begin{mdframed}
-    \includegraphics[width=\textwidth]{figures/ex_real_highgravity_t1500.png}
-    \end{mdframed}
-    \end{minipage}
-    \begin{minipage}{0.48\textwidth}
-    \begin{mdframed}
-    \includegraphics[width=\textwidth]{figures/ex_real_lowgravity_t1500.png}
-    \end{mdframed}
-    \end{minipage}
+    \begin{center}
+        \includegraphics[width=\linewidth]{figures/Fig3.png}
+    \end{center}
     \caption{Example of simulated networks at $t=1500$, for a synthetic system of cities (Top row) and the European urban network (Bottom row).
@@
 low) gravity decay parameter for the left column (resp.
-right).\label{fig:example}}
+right).\label{fig:fig3}}
 \end{figure}
@@
 Behavior of some indicators are shown in Fig.~\ref{fig:onefactor}.
-For example, internationalization and the correlation between economic size and weighted degree both show a phase transition as a function of $\gamma_G$, from a local to a global regime.
+For example, internationalisation and metropolisation both show a transition as a function of $\gamma_G$, from a local to a global regime.
 The sector proximity parameter $\gamma_S$ influences the internationalization linearly, but induces a maximum for the correlation between size and degree, which can be interpreted as a setting where size and the total volume in a city are the most correlated.
@@
 \begin{figure}
-	\includegraphics[width=0.48\textwidth,height=0.25\textheight]{figures/internationalization-gravityDecay_errorbars.png}
-    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figures/rhoDegreeSize-gravityDecay_errorbars.png}\\
-    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figures/internationalization-gammaSectors_errorbars.png}
-    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figures/rhoDegreeSize-gammaSectors_errorbars.png}
+	\begin{center}
+	    \includegraphics[width=\linewidth]{figures/Fig4.png}
+	\end{center}
     \caption{(Top Left) Internationalisation index decreases exponentially with gravity decay; (Top Right) Correlation between city weighted degree and size.
@@
 (Bottom Left) Internationalisation varies linearly with sector proximity $\gamma_S$; (Bottom Right) Correlation between degree and size exhibits a maximum, witnessing an intermediate regime where size is the most important.
-\label{fig:onefactor}}
+\label{fig:fig4}}
 \end{figure}
@@
 \begin{figure}
-    \includegraphics[width=\textwidth]{figures/internationalization_countryGravityDecay200_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
-    \caption{The transition as a function of interaction range depends on the influence of origin size $\gamma_F$; sector proximity $\gamma_S$ plays a role only for a large influence of the origin.\label{fig:internat}}
+    \includegraphics[width=\linewidth]{figures/Fig5.png}
+    \caption{The transition as a function of interaction range depends on the influence of origin size $\gamma_O$; sector proximity $\gamma_S$ plays a role only for a large influence of the origin.\label{fig:fig5}}
 \end{figure}

### 2020-09-01 main.tex
Update on Overleaf.

@@
 
-When considering optimal points with $\varepsilon_L < 5.0$, the adjusted parameters are: long range gravity interactions and socio-economic interactions given by $d_0 = 5205 \pm 1775$ and $c_0 = 109 \pm 29$ (the maximal $c_{ij}$ is 11.8); a stronger influence of destination than origin as $\gamma_O = 1.45 \pm 0.09$ and $\gamma_D = 2.0 \pm 0.06$; a comparable influence of sectors as $\gamma_S = 1.75 \pm 0.17$; a fine time granularity as $t_f = 4960 \pm 109$; and a very strong influence of previous links as $\gamma_W = 7.99 \pm 3.77$.
+When considering optimal points with $\varepsilon_L < 5.0$, the adjusted parameters are: long range gravity interactions and socio-economic interactions given by $d_0 = 5972 \pm 2148$ and $c_0 = 113 \pm 33$ (the maximal $c_{ij}$ is 11.8); a stronger influence of destination than origin as $\gamma_O = 1.46 \pm 0.10$ and $\gamma_D = 2.0 \pm 0.05$; a comparable influence of sectors as $\gamma_S = 1.76 \pm 0.21$; a fine time granularity as $t_f = 4951 \pm 124$; and a very strong influence of previous links as $\gamma_W = 6.43 \pm 4.11$.
 This last point furthermore shows in particular a role of the path-dependency parameter, confirming the relevance of this complex generative model including path-dependency.

### 2020-09-01 figures.sh
fig7

@@ -22,3 +22,6 @@ cp figuresraw/internationalization_countryGravityDecay200_gammaDestination0_face
 
 # Fig 6
 cp figuresraw/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png figures/Fig6.png
+
+# Fig 7
+cp figuresraw/pareto_colorgammaOrigin.png figures/Fig7.png

### 2020-09-01 S2_Text.tex
Update on Overleaf.

--- 2020-09-01 main.tex
@@
-\documentclass[10pt,letterpaper]{article}
-\input{header.tex}
+\input{si.tex}
+\begin{document}
+\vspace*{0.35in}
+\justify
 
-\usepackage{ragged2e}
-\begin{document}
-\justify
-\vspace*{0.2in}
+\section*{S2 Text: Model setup and sensitivity analysis}
 
-% Title must be 250 characters or less.
-\begin{flushleft}
-{\Large
-\textbf\newline{A generative model for urban firm networks} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
-}
-% alternative titles?
-% Modeling the growth of urban firm networks
-\newline
-% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
-\\
-Juste Raimbault\textsuperscript{1,2,3*},
-Natalia Zdanowska\textsuperscript{1,3},
-Elsa Arcaute\textsuperscript{3}
-\\
-\bigskip
-\textbf{1} Centre for Advanced Spatial Analysis, UCL, London, United Kingdom
-\\
-\textbf{2} UPS CNRS 3611 ISC-PIF, Paris, France
-\\
-\textbf{3} UMR CNRS 8504 G{\'e}ographie-cit{\'e}s, Paris, France
-\\
-\bigskip
-
-% Use the asterisk to denote corresponding authorship and provide email address in note below.
-* juste.raimbault@polytechnique.edu
-
-\end{flushleft}
-
-% Plos one requirements https://journals.plos.org/plosone/s/submission-guidelines
-% Deadline for the Cities as Complex Systems Call for Papers: 4th September https://collections.plos.org/s/cities
-% Manuscripts can be any length. There are no restrictions on word count, number of figures, or amount of supporting information. 
-% PLOS uses “Vancouver” references style
-% We need to reorganise the paper to have more or less the following sections (which can be reordered or renamed):
-% 1. Title page: List title, authors, and affiliations as first page of manuscript 2. Abstract 3. Introduction 4. Materials and Methods 5. Results 6. Discussion 7. Conclusions (optional)
-
-\section*{Abstract}
-The emergence of interconnected urban networks is a crucial feature of globalisation processes.
-Understanding the drivers behind the growth of such networks -- in particular urban firm networks --, is essential for the economic resilience of urban systems.
-We introduce in this paper a generative network model for firm networks at urban area level including several complementary processes: the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distance.
-An empirical network analysis on European firm ownership data confirms the relevance of each of these factors.
-We then simulate network growth for synthetic systems of cities, unveiling stylized facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range.
-We finally calibrate the model on the European network, outperforming statistical models and showing a strong role of path-dependency.
-Potential applications of the model include the study of mitigation policies to deal with exogenous shocks such as economic crisis or potential lockdowns of countries.
-
-\linenumbers
-
-\section*{Introduction}
-
-\cite{derudder2018central}
-\cite{storme2019introducing} cities firms interactions
-
-Networks constitute the new social morphology of our societies, since the intensification of globalisation of the economy in the late 20th century \cite{castells2000networksociety}.
-A world-city system has since then emerged \cite{taylor2001specification}, characterized by highly interconnected urban centres at the global level \cite{sassen1991}.
-These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity implying spatial and non-spatial patterns \cite{martinus2018global}.
-Interactions of economic nature between transnational firms, are together with international trade patterns among the most representative of the current economic trends \cite{taylor2001specification}.
-Understanding the drivers of growth and geographical properties of such an urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}.
-Straightening these city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}, by developing strategic industries for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
-The functioning of such a city-network and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.
-
-The use of network models to understand processes underlying the emergence and dynamics of such networks has already been tackled in the literature, but still not extensively at city level.
-Generative urban network models combining geographical factors with network topological properties exist at the regional scale {\cite{dai2016simulating}}.
-%Air transportation networks can also be understood as urban networks.
-\cite{zanin2013modelling} reviews several models for such networks.
-Trade networks have been extensively studied from the complex networks perspective, but these are generally driven at the country-level \cite{garlaschelli2005structure} because of a lack of city-level data.
-Input-output models, mostly used in regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}.
-In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for different types of flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2018indirect}.
-
-In this paper we choose to focus on urban networks of firms to question and capture geographical and economic processes as internationalisation, metropolisation, regionalisation and specialisation of cities within a generative model.
-The specific question we tackle is the complementarity of different processes to generate such urban networks.
-The originality of the present paper is to examine these issues within the system of cities framework \cite{berry1964cities}.
-In this approach the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}.
-We examine the interactions of European cities within firm linkages defined by corporate ownership links.
-Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2019arXiv191014652Z}.
-
-More precisely, our contribution consists in the following: (i) we propose an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database \cite{AMADEUS2018}; (ii) we introduce a generative network model to simulate the growth of such linkages at urban areas scale, which combines multiple factors influencing link formation, as economic intensity of origin and destination of urban areas, industrial sector proximity, the strength of previous links, and geographical and socio-cultural distance; (iii) the model allows us to compare the effect of different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; and (iv) we calibrate the model on the empirical network.
-
-The rest of this paper is organized as follows: We first describe the generative model for urban networks and proceed to an empirical analysis of the network used as a case study.
-We then do a sensitivity analysis and exploration of the model on synthetic data.
-The model is calibrated on real European firm ownership data.
-We finally discuss implications of our results and possible extensions.
-
-\section*{Materials and Methods}
-
-\subsection*{Generative urban network model}
-
-\subsubsection*{Rationale}
-
-The model is expected to capture a single macroscopic urban scale, i.e.
-the level of the urban system where basic entities are cities.
-Links are induced by underlying firms but these are not explicit.
-Cities are defined by their economic size, but also by an industrial sector profile.
-As the authors of \cite{martinus2018global} point out, a combination of several distances may play a role in establishing linkages.
-Geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}.
-Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalized to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (confirmed relevant by the fixed effect regression described above); (iii) city economic size, which has furthermore scaling laws properties; (iv) industrial proximity in terms of firm sector composition; (v) previous linkages (influence of already existing links in the past).
-This last point is crucial to capture accumulated effects and include path-dependency, which fits well the use of a simulation model instead of a more simple statistical model.
-
-\subsubsection*{Model description}
-
-Let $1 \leq i \leq N$ cities defined at time $t$ by their economic structure such that $E_i(t)$ is the total economic volume (GDP) for city $i$ and $S_{ik}(t)$ is a matrix giving economic volume proportions within each economic sector $k$, assuming $K$ economic sectors.
-Formally, the model creates links between cities which are characterized by their economic size $E_i$ (GDP) and economic structure $S_{ik}$ in terms of activity sectors (probability distribution of firms within $K$ sectors).
-Starting with an existing network with no links, the model iteratively adds links, following a probability given by a generalized Cobb-Douglas function \cite{vilcu2011geometric} as
-
-\begin{equation}
-p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_O} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_D} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s\left(S_{ik},S_{jk}\right)^{\gamma_S} \cdot \exp \left(- d_{ij} / d_0\right) \cdot \exp \left(- c_{ij} / c_0\right)
-\end{equation}
-
-where $E  =  \sum_k E_k$, $W  = \sum_{i,j} w_{ij}$ weights of previous links, $s(S_{ik},S_{jk})$ is a similarity function between activity sectors $S_{ik}$ and $S_{jk}$ taken as a cosine similarity, $d_{ij}$ euclidian distance, and $g_{ij}$ a socio-cultural distance.
-In the case of an already existing link between two areas, the weight of the latter is incremented by a unit volume $w_0$.
-The model is stopped either when a final time is reached or when the cumulated volume of links reaches a target quantity.
-This formulation can be considered both as an economic utility model and a generalisation of spatial interaction models.
-
-Note that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalise the exponents to 1 (which is different than renormalising the probability), to ensure to include convex functions.
-The ``geopolitical/sociocultural'' distance remains abstract and should be parametrised (see setup below) and estimated in the data-driven setting.
-We assume no evolution of economic size in this simple version of the model, which means that adjusted parameter are valid on a restricted time period.
-
-\subsubsection*{Model setup}
-
-Besides model dynamics, its initialisation (or parametrisation on real data) regarding the spatial distribution, sizes, countries and industrial sector compositions, is also important to understand model behavior.
-In addition  \cite{raimbault2019space} showed that the initial spatial configuration can have significant impacts on model outcomes.
-We therefore consider two versions: one with a synthetic system of cities which can be stochastically repeated, and one other based on real data.
-
-\paragraph{Synthetic setup}
-
-The synthetic system of cities is generated to be at the scale of a continent with properties similar to Europe: (i) $N=700$ cities with an empirical power-law from the Global Human Settlements database for economic sizes ($\alpha = 1.1$), distributed randomly into a uniform and isotropic square space of width $d_{max}=3000km$;
-(ii) clustered into $C = 20$ countries using a k-means algorithm (this remains consistent with non-correlated city sizes at the scale of the countries as our cities are randomly distributed in space \cite{simini2019testing});
-(iii) with synthetic distribution of sectors following log-normal distributions adjusted in a way that large cities have more high-value activities and are more diverse.
-
-More precisely for this last point, assuming sectors are distributed continuously along a one-dimensional axis, cities are assumed to have a sector distribution with a log-normal density function.
-The largest city is constrained to have a standard deviation and a mode of $1/2$ and the smallest $1/K$, with a linear interpolation between the two for other cities.
-Parameters of the log-normal law are obtained by solving the corresponding two equations system for each city, and each distribution is discretised and renormalised to be a vector of probabilities.
-See \ref{S2_Text} for all details of this synthetic setup.
-
-In synthetic experiments, the network is generated from an empty initial network until reaching $t_f=1500$, with a unit link volume of $w_0 = 1$.
-
-\paragraph{Real setup}
-
-The real setup is done using the empirical network analyzed above.
-Urban areas are distributed according to their positions in the actual geographical space, and economic size and sector composition attributes are used.
-The socio-economic distance is constructed using information from the statistical modeling, by considering fixed-effect coefficients estimated for each pair of countries.
-See \ref{S2_Text} for more details on the construction of the socio-economic distance matrix.
-
-Initial network is empty, and the real links are taken as targets for the generated network.
-We constrain the total volume of flows by setting $w_0 = \sum_{ij} w_{ij}^{(obs)} / t_{f}$ when the number of time steps $t_f$ is given as a parameter.
-
-\subsection*{Model parameters and indicators}
-
-The model parameters are the Cobb-Douglas weights $\gamma_O,\gamma_D,\gamma_W,\gamma_S$ and distance decay parameters $d_0,c_0$.
-As the model also includes path-dependency by integrating the influence of previous links, it can not be reduced to a simple statistical model.
-Finally, the role of space introduces even further complexity, yielding a generative model, which has to be simulated.
-
-We consider two main macroscopic indicators to quantify the generated network.
-Firstly geographical structures are captured by internationalisation (modularity of countries in the network).
-The situation of a minimised modularity corresponds to dense connections between cities of different countries of Europe and sparse connections between cities of the same European country.
-In practice, directed modularity is computed by
-\begin{equation}
-
-\end{equation}
-% put the actual modularity
-% add some reference values - some kind of null model
-
-Secondly we consider the metropolisation indicator as the correlation between the weighted degree and the GDP of the city.
-With this indicator we capture the effect of large cities or metropolises of regional importance attracting the biggest amount of links and involved in high interactions with other cities.
-It is computed with a Pearson correlation estimator as
-\begin{equation}
-    M = \hat{\rho} \left[d_i , E_i \right]
-\end{equation}
-
-%I think we need to keep this paragraph but shorten it, because to my understanding network average community size is tested in Figure 5, but not the Louvain modularity for ex.
-%We also consider several network indicators as Louvain modularity, sizes of maximal modularity communities, degree and link weight distributions (described by their average, hierarchy, and entropy), and correlations between degree and city size, and between link weight and distance.
-
-We take a geographical approach to study the behavior of the generative model.
-Other related approaches include indicators of economic complexity, estimating for example the diversity of products produced by countries and how they depend on each other \cite{hidalgo2007product}.
-These are less relevant in our case as the sector structure of cities does not evolve in the model.
-
-\subsection*{Empirical data}
-
-Companies from \emph {AMADEUS} are georeferenced using the Geonames database (using postcode or address depending on the information available).
-We then join this data with the Global Human Settlement dataset GHS-FUA for Functional Urban Areas \cite{Florczyk2019ghs} -- the most relevant objects for applying our model.
-Starting from a firm-level dataset of 1,562,862 nodes and 1,866,936 links, we obtain a directed network of 573 urban areas and 9323 ownership links.
-The weight of links is obtained by computing the owned share of turnover at destination, i.e.
-$w_{ij} = \sum_{k \in i,l \in j} p_{kl} \cdot T_l$ where $p_{kl}$ of the ownership share of company $l$ by company $k$ and $T_l$ the turnover of company $l$.
-Node attributes include an economic size of urban areas, that we compute as the sum of turnovers of companies within the area.
-This quantity is highly correlated with the GDP of the area-level included in the GHS database (Spearman correlation $\rho = 0.71$) and also with population ($\rho = 0.60$).
-We then expect to capture size effects while keeping the consistence of a single main data source.
-An other attribute of the nodes is the industrial sector composition, expressed as a proportion of turnover in the area associated to a given industrial sector.
-Detailed sectors of firms are available up to the four digits of the NACE classification in the raw data \cite{EUROSTAT2008}.
-We consider the first level classification (21 categories) and compute sector proportions accordingly.
-This allows us to define an industrial proximity matrix between urban areas, taking a cosine similarity between vectors as $s_{ij} = \sum_{k=1}^{K} S_{ik}\cdot S_{jk}$ where $S_{ik}$ are sector proportions.
-
-The urban network has heavy tail weighted degree and edge weight distributions, as shown in Fig.~\ref{fig:nwdist}.
-We fit power law and log-normal distributions, including minimal value cutoff following \cite{clauset2009power}, using the \texttt{powerlaw} R package \cite{powerlawpackage}.
-For both, log-normal distributions appear to be a better fit ($\mu=18.8$, $\sigma=2.3$ for weighted degree; $\mu=13.8$, $\sigma=2.8$ for edge weight), and include a much broader range of the empirical distributions with lower estimated cut-offs.
-These heavy tail properties suggest the relevance of a generative model including self-reinforcing processes, which are known to produce such distributions.
-
-%%%%%%%%%%%%%
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=\linewidth]{figures/Fig1.png}
-    \end{center}
-    \caption{Distributions of network properties.
-\textit{(Left)} Cumulated distribution for edge weight, with power law fit (red) and log-normal fit (green); \textit{(Right)} Cumulated distribution for weighted degree.}
-    \label{fig:nwdist}
-\end{figure}
-%%%%%%%%%%%%%
-
-Regarding the community structure of the network, different modularity maximisation algorithms (run on the undirected corresponding network) give the same undirected modularity of 0.38, and a directed modularity \cite{nicosia2009extending} of 0.34 for the greedy algorithm of \cite{clauset2004finding} and 0.36 for the Louvain algorithm \cite{blondel2008fast}, corresponding to 36 communities (resp.
-15).
-In comparison, with 29 countries, the classification of urban areas by country has a modularity of 0.32.
-This means that the structure of flows present certain regional patterns.
-Indeed, a null model assigning random labels with 30 communities, bootstrap 1000 times, yields an average modularity of $0.049 \pm 0.002$, confirming the statistical significance of these communities.
-We show in Fig.~\ref{fig:fig2} a map of the communities obtained.
-
-%%%%%%%%%%%%%%
-\begin{figure}
-    \begin{center}
-    \includegraphics[width=\linewidth]{figures/Fig2.png}
-    \end{center}
-    \caption{{\bf Map of network communities obtained by modularity maximisation.} Area of circles gives the population with the Functional Urban Area, and color the community it belongs to.
-Urban Areas in communities of size smaller than 3 were not colored for readability.}
-    \label{fig:fig2}
-\end{figure}
-%%%%%%%%%%%%%%
-
-\subsubsection*{Statistical analysis}
-
-In order to understand which are the most determining geo-economic factors influencing creation of links between cities, we run a statistical analysis as applied in the case of unconstrained spatial interaction models \cite{wilson1975some}.
-We test five different statistical models, including progressively the processes to be included in the network generative model.
-The most general statistical model is:
-
-\begin{equation}
-\log(w_{ij}) = \log(d_{ij}) + \log(T_i) + \log(T_j) + \log(s_{ij}) + \alpha_{c_i,c_j} + \varepsilon_{ij}
-\end{equation}
-
-where $\alpha_{c_i,c_j}$ is a fixed effect term for the coupling of countries $c_i,c_j$.
-A multiplicative form is applied by taking logarithms as usually done for spatial interaction models.
-Furthermore, origin and/or destination constrained models are not relevant in our case, as links include information about the turnover and a share of ownership at the same time.
-Estimation results are given in Table~\ref{tab:reg}.
-We also fit a Poisson generalized linear model with a logarithmic link on the rounded weights \cite{flowerdew1988fitting}.
-The findings show that when distance only is taken into account (model 1), it has a significant negative effect on link creation, but with an explained variance close to zero.
-Model 2 adds fixed effects by pair of countries, for which around a third are significant, and with a better $R^2$ of 0.17.
-Adding the economic size at origin and destination (turnovers $T_i$ and $T$, model 3) slightly improves the explanatory power (model 3).
-The best model among the linear ones is the model 4, with all factors and fixed effects, both in terms of explained variance ($R^2 = 0.31$) and Akaike information criterion.
-Model 4 also indicates the absence of overfitting.
-Model 5, which is a Poisson model with all variables is much a better fit regarding pseudo-$R^2$ and provides much more significant estimates (all fixed effects significant and much smaller standard deviations).
-Across all models, consistent qualitative stylized facts regarding link creation between cities are revealed: (i) distance has a negative influence; (ii) economic sizes have a positive effect, with the size of the origin city being more important (consistent in the ownership links context, as strongly driven by decision at the headquarter/owner level); (iii) industrial proximity has a positive influence, i.e.
-similarity is more important than complementarity for this dimension; (iv) an large part of country fixed-effects are statistically significant; and (v) the order of magnitude of the influence of each factor is roughly the same, which means that no process totally dominates the others.
-This confirms the relevance of including all these different dimensions in the generative network model, since they have a comparable explanatory role in statistical models.
-
-%%%%%%%%%%%%%
-\begin{table}[!ht]
-\begin{adjustwidth}{-2.25in}{0in}
-\caption{{\bf Statistical models.} Ordinary Least Square estimations of statistical models explaining $\log(w_{ij})$ (standard errors in parenthesis).
-***, ** and * respectively indicate 0.01, 0.05 and 0.1 levels of significance.
-For the fixed effect by origin-destination pairs of countries, the proportion of significant coefficients is attributed (where $p<0.1$).
-Models (1) to (4) are ordinary least square models, while model (5) is a Poisson generalized linear model with logarithmic link.
-We compare model performance using $R^2$, Mean Square Error, and Akaike Information Criterion.
-For the Poisson model (model 5), AIC is not given as it is not comparable, and the pseudo-$R^2$ is given either.
-\label{tab:reg}}
-\medskip
-\begin{tabular}{|l|c|c|c|c|c|}
-\hline
-Model  & (1) & (2) & (3) & (4) & (5) \\ 
-\hline
-$\log(d_{ij})$ &      -0.06** (0.03) &   -0.11** (0.05)  & -0.41*** (0.02)  & -0.35*** (0.04)  &  -0.26*** ($5\cdot 10^{-6}$) \\
-$\log(T_i)$ &   &   & 0.56*** (0.01) &  0.56*** (0.01) & 0.79*** ($2\cdot 10^{-6}$) \\
-$\log(T_j)$ &     &   & 0.39*** (0.01) &  0.39*** (0.01) & 0.66***  ($1.5\cdot 10^{-6}$) \\
-$\log(s_{ij})$ &     &   &  &  0.19*** (0.07) & 0.53*** ($9\cdot 10^{-6}$)  \\
-Countries &    &  28.3\% &   &  26.5\% & 100\% \\
-\hline
-$R^2$ &       0.00059   &  0.17 & 0.21 &  0.31  &  0.56 \\
-MSE & 7.75 & 6.84 & 6.10 & 5.33 & 8.72 \\
-AIC &        44304   &  43578  &  42131  & 41917 &   \\
-\hline
-\end{tabular}
-\end{adjustwidth}
-\end{table}
-%%%%%%%%%%%%%
-
-\section*{Model exploration and calibration}
-
-The model is implemented in NetLogo, which provides a good compromise between performance and interactive prototyping and exploration \cite{railsback2017improving}.
-Numerical experiments are done by integrating the model into the OpenMOLE software for model exploration \cite{reuillon2013openmole}, which provides sensitivity analysis and calibration methods and a transparent access to high-performance computing infrastructures.
-Model source code, aggregated network data (raw initial firm network can not be made available for data ownership reasons), and results are available on the open Git repository of the project at \url{https://github.com/JusteRaimbault/ABMCitiesFirms}.
-Large simulation result files are available on the dataverse repository at \url{https://doi.org/10.7910/DVN/UPX23S}.
-We show in Fig.~\ref{fig:fig3} examples of generated networks, with very contrasted patterns obtained with different spatial interaction ranges.
-Low values of the gravity decay parameter yield local networks around main cities, while long interaction ranges yield more integrated networks.
-We also show networks simulated in the real setting.
-
-%%%%%%%%%%%%%
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=\linewidth]{figures/Fig3.png}
-    \end{center}
-    \caption{Example of simulated networks at $t=1500$, for a synthetic system of cities (Top row) and the European urban network (Bottom row).
-, with a high (resp.
-low) gravity decay parameter for the left column (resp.
-right).\label{fig:fig3}}
-\end{figure}
-%%%%%%%%%%%%%
-
-\subsection*{Synthetic city systems}
-
-We first study extensively the behavior of the model on synthetic systems of cities, to isolate its intrinsic behavior independently of a contingent geographical situation.
-
-\paragraph{Sensitivity analysis}
-
-A global sensitivity analysis first unveils an aggregated influence of parameters on model indicators.
-We apply Saltelli's Global Sensitivity method introduced by \cite{saltelli2008global}.
-This produces values for each parameter and each output of a first order sensitivity index, defined as $Var\left[\mathbb{E}_{\mathbf{X}_{\sim i}}\left(Y_j | X_i\right)\right] / Var\left[Y\right]$, for parameter $X_i$ and output $Y_j$, $\mathbf{X}_{\sim i}$ being all other parameters.
-This captures the total influence of $X_i$ all other things being equal.
-The total order index is given by $\mathbb{E}_{\mathbf{X}_{\sim i}} \left[Var(Y_j | \mathbf{X}_{\sim i})\right] / Var\left[Y_j\right]$ and captures non-homogeneity in behavior and interaction between parameters.
-Indices were estimated with a Sobol sequence sampling of 20,000 model runs \cite{saltelli2010variance}.
-We give in \ref{S4_Text} the full estimated results.
-To summarize, we find that indicators related to modularity (internationalisation and optimal modularity) are mostly influenced by geographical parameters.
-On the contrary, indicators of network structure are driven by origin and destination sizes.
-The influence of sectors and historical links has a secondary but not negligible role.
-Altogether, the analysis witnesses of strong interaction effects between parameters, since for several indicators the total order index is one order of magnitude larger than the first order index.
-This means that the generative model captures some complex behavior of the system.
-% The very different important parameters depending on outputs studied confirms the relevance of a more exhaustive grid exploration.
-
-\paragraph{One factor sampling}
-
-A first numerical experiment of one-factor sampling on all Cobb-Douglas parameters and 100 stochastic repetitions confirms a good statistical convergence (average Sharpe ratios for indicators all larger than 5).
-We use thus 20 repetitions in following larger experiments (in the case of a normal distribution, a 95\% confidence interval on the average is of size $2\sigma \cdot 1.96 / \sqrt{n}$, what gives $n=16$ repetitions for a CI of width of the standard deviation).
-Behavior of some indicators are shown in Fig.~\ref{fig:onefactor}.
-For example, internationalisation and metropolisation both show a transition as a function of $\gamma_G$, from a local to a global regime.
-The sector proximity parameter $\gamma_S$ influences the internationalization linearly, but induces a maximum for the correlation between size and degree, which can be interpreted as a setting where size and the total volume in a city are the most correlated.
-
-%%%%%%%%%%%%%
-\begin{figure}
-	\begin{center}
-	    \includegraphics[width=\linewidth]{figures/Fig4.png}
-	\end{center}
-    \caption{(Top Left) Internationalisation index decreases exponentially with gravity decay; (Top Right) Correlation between city weighted degree and size.
-Both plots show a transition from a local to a global regime.
-(Bottom Left) Internationalisation varies linearly with sector proximity $\gamma_S$; (Bottom Right) Correlation between degree and size exhibits a maximum, witnessing an intermediate regime where size is the most important.
-\label{fig:fig4}}
-\end{figure}
-%%%%%%%%%%%%%
-
-\paragraph{Grid exploration}
-
-The model behaviour is then studied with a grid sampling for parameters and 20 repetitions (the computations run on the European Grid Infrastructure, and are equivalent to 2.5 years of CPU time).
-Distance decay and sector  parameters are varied with 10 steps, other parameters with three.
-The influence of gravity decay parameters is confirmed when plotting the internationalization index against the distance, which exhibits an exponential decay as shown in Fig.~\ref{fig:internat}.
-It interacts with the role of sectors and the size of the origin, with a more significant effect of sector proximity when size has the largest influence (third panel).
-
-%%%%%%%%%%%%%%
-\begin{figure}
-    \includegraphics[width=\linewidth]{figures/Fig5.png}
-    \caption{The transition as a function of interaction range depends on the influence of origin size $\gamma_O$; sector proximity $\gamma_S$ plays a role only for a large influence of the origin.\label{fig:fig5}}
-\end{figure}
-%%%%%%%%%%%%%%
-
-Other indicators exhibit non-trivial patterns.
-For example, when considering average community size of the final network as shown in Fig.~\ref{fig:comsize}, we obtain a maximal integration in terms of communities at an intermediate value of the gravity decay, what can be interpreted as the emergence of a regional regime.
-The size of communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.
-In particular, we observe a qualitative inversion of the role of $\gamma_S$ when introducing an effect of the origin (switching $\gamma_F$ from 0 to 1, first panel to second panel).
-The maximal community size disappears when $\gamma_F = 2$, implying a regime with small local communities where large cities mostly connect with there hinterland.
-These experiments reveal a complex interplay between processes and how the model produces diverse stylised facts.
-
-%%%%%%%%%%%%%%
-\begin{figure}
-    \includegraphics[width=\textwidth]{figures/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
-	\caption{Maximal integration in terms of community size is achieved at an intermediate value of $d_G$: emergence of a regional regime; Maximal size depends on the role of sectors $\gamma_S$, in a decreasing way when origin size is deactivated, and increasing way when $\gamma_F=1$; This regime disappear when origin size influence is too large.\label{fig:comsize}}
-\end{figure}
-%%%%%%%%%%%%%%
-
-\paragraph{Impact of urban hierarchy}
-
-We run a specific experiment to study the impact of urban hierarchy on firms dynamics.
-Rank-size law in cities is a central stylized fact to urban theories, despite the discussions on how to estimate the power-law exponent, its actual values \cite{cottineau2017metazipf} and how its depends on the definition of cities considered \cite{corral2020truncated}.
-In that particular context, understanding how this parameter can impact urban dynamics within artificial systems of cities in which it can be modified to extreme values not observed in real systems, should inform on underlying processes.
-% too abstract: clarify why hypothetical systems of cities bring info on real ones
-We change the initial hierarchy $\alpha$ from 0.2 to 2.0 with a step of 0.1, and also vary interaction range $d_0$.
-
-\subsection*{Model calibration on the European urban network}
-
-We then apply the model to a real system of cities by calibrating it on the European ownership network.
-Model setup is done following the real setup described above.
-The number of time steps $t_f$ is left as an additional parameter, which in a sense determines the granularity of the cumulative network generation process.
-The objective functions for calibration are the average mean squared error on logarithms of weights given by $\varepsilon_L = \frac{1}{N^2} \sum_{i,j} \left(\log w_{ij} - \log \hat{w}_{ij} \right)^2$ if $\hat{w}_{ij}$ are the simulated weights, and the logarithm of the mean squared error $\varepsilon_M = \log\left(\frac{1}{N^2} \sum_{i,j} \left(w_{ij} - \hat{w}_{ij}\right)^2 \right)$.
-\cite{raimbault2018indirect} showed that these two are complementary to calibrate urban systems models.
-Calibration is done using a genetic NSGA2 algorithm \cite{deb2002fast}, which is particularly suited for such a bi-objective calibration of a simulation model.
-Stochasticity is taken into account by adding the estimation confidence as an additional objective, and we filter the final population to have at least 20 samples.
-We use a population of 200 individuals and run the algorithm for 20,000 generations, on 1000 islands in parallel.
-
-%%%%%%%%%%%%%%
-\begin{figure}
-	\begin{center}
-    \includegraphics[width=0.75\textwidth]{figures/pareto_colorgammaOrigin.png}
-    \end{center}
-	\caption{Pareto front obtained for the bi-objective model calibration for $\varepsilon_L$ and $\varepsilon_M$ objectives.
-$\varepsilon_L$ can be compared to the MSE of statistical models fitted above, and the generative model outperform these for several parametrizations (lowest MSE 5.33 for statistical models).
-Point size gives the number of stochastic samples and color the value of $\gamma_F$.\label{fig:calibration}}
-\end{figure}
-%%%%%%%%%%%%%%
-
-We show in Fig.~\ref{fig:calibration} the calibration result, as a Pareto front of compromise points between the two objectives.
-As the mean squared error on logarithm is the same than the one for statistical models, it can be directly compared.
-The other objective would correspond to statistical models directly on variables: therefore, our model covers in a sense a set of intermediate models which make a compromise between the two objectives, and is therefore much more flexible.
-The values for $\varepsilon_L$ are for a large proportion of points lower than for statistical models, our model outperforming these in this predictive sense.
-As shown in Fig.~\ref{fig:calibration}, increasing values for $\gamma_F$ lead to better performance regarding $\varepsilon_L$: increasing the importance of the origin will provide a better fit on all order of magnitudes of link weights (while $\varepsilon_M$ favors mostly the largest links given the fat-tail distribution).
-
-When considering optimal points with $\varepsilon_L < 5.0$, the adjusted parameters are: long range gravity interactions and socio-economic interactions given by $d_0 = 5972 \pm 2148$ and $c_0 = 113 \pm 33$ (the maximal $c_{ij}$ is 11.8); a stronger influence of destination than origin as $\gamma_O = 1.46 \pm 0.10$ and $\gamma_D = 2.0 \pm 0.05$; a comparable influence of sectors as $\gamma_S = 1.76 \pm 0.21$; a fine time granularity as $t_f = 4951 \pm 124$; and a very strong influence of previous links as $\gamma_W = 6.43 \pm 4.11$.
-This last point furthermore shows in particular a role of the path-dependency parameter, confirming the relevance of this complex generative model including path-dependency.
-The interaction of this process with others implies an inversion of the relative importance of origin and destination (on the contrary to statistical models).
-The fact that calibrated interaction distances are large is consistent with long range economic interactions within Europe.
-Besides these stylized facts, these calibration results altogether demonstrate the feasibility of model application to real data.
-
-\subsection*{Application: impact of economic shock}
-
-Based on calibrated parameter values, we design scenarios to evaluate the impact of economic shocks.
-These can be the consequence of diverse factors, including socio-economic, geopolitical factors (example of Brexit), and natural disasters such as COVID-19.
-We simulate a full border closure by setting $\gamma_D = 0$, and intermediate scenarios.
-Setting also $\gamma_G$ to 0 allows simulating inter-country lock-downs.
-% run simulations
-% note that time scales are in fact smaller - or assume that crisis keeping border closed is rather long
-
-%%%%%%%%%%%%%%%
-\section*{Discussion}
-
-This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMole platform and calibrated on real data on ownership linkages of firms for Europe from Amadeus dataset.
-The simulation on a synthetic system of cities provides an broad knowledge on model behavior in its parameter space, and its calibration suggests its potential application to policy issues.
-This kind of model can therefore have practical applications in the future in terms of testing the effect of exogenous shocks in the socio-economic structure: as the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to COVID19 on the single market and on the strongly foreign dependent economic sectors (see Appendix).
-
-Even in such a simple model (close to directly tractable stationary state) the behaviour is highly non-linear in many dimensions.
-Our study shows how crucial model exploration is to overcome hidden parameters (deactivated mechanisms or default parameter values).
-This paper emphasis on the fact that exploration of intrinsic dynamics on synthetic data is of great importance and should be driven before an application on real data so as to put aside the geographical effects from model dynamics.
-
-Several possible extensions of the current work can be undertaken in the future as a co-evolution model with an evolution of city sizes or a model with firm agents leading to a multi-scale agent-based model.
-A more precise study of the role of path dependency can also be undertaken in order to understand the creation of the future links.
-Other formulations of the model might be taken into consideration as other formulation of the combination of factors or multi-objective optimisation depending on sectors using Pareto fronts.
-
-\section*{Supporting information}
-
-\paragraph*{S1 Text.}
-\label{S1_Text}
-{\bf UK's economic sectors dependent on international links.}
-
-\paragraph*{S2 Text.}
-\label{S2_Text}
-{\bf Socio-economic distance matrix.}
-
-\paragraph*{S3 Text.}
-\label{S3_Text}
-{\bf Synthetic model setup.}
-
-\paragraph*{S4 Text.}
-\label{S4_Text}
-{\bf Sensitivity analysis.}
-
-\section*{Acknowledgments}
-
-Results obtained in this paper were computed on the vo.complex-system.eu virtual organization of the European Grid Infrastructure (http://www.egi.eu).
-We thank the European Grid Infrastructure and its supporting National Grid Initiatives (France-Grilles in particular) for providing the technical support and infrastructure.
-
-\bibliographystyle{plos2015}
-\bibliography{biblio}
-
-%%%%%%%%%%%%%
-\newpage
-\section*{S1 Text: UK's economic sectors dependent on international links}
-
-Based on \emph {AMADEUS} dataset, the most vulnerable economic sectors in the United Kingdom have been highlighted.
-The later are the ones due to be mostly affected by eventual future lock-downs of the international economy (international ownership links).
-
-When considering the UK sectors generating the highest turnover in 2018 mostly dependent on foreign capital they own (out-going links from the UK) these are extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods.
-In terms of in-going links to the UK, activities of head offices and holding companies, sale of cars and manufacture of motor vehicles are sectors mostly dependent on foreign ownership coming from abroad.
-
-\FloatBarrier
-
-\begin{table}[!ht]
-\begin{adjustwidth}{-2.25in}{0in}
-\caption{Top 10 UK industries depending on foreign ownership expressed in total turnover (out-going links from the UK)\label{tab:uk_outgoing}}
-\centering
-\medskip
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Classification  & NACE Industries & Total turnover \\ 
-\hline
-1 &     Extraction of crude petroleum &   96,458,258,642  \\
-\hline
-2 & Manufacture of tobacco products &  5,477,272,464  \\
-\hline
-3 & Manufacture of basic pharmaceutical products &  4,825,881,930 \\
-\hline
-4 &  Retail sale in non-specialised stores with food, beverages or tobacco predominating  &   3,681,373,374  \\
-\hline
-5 &  Wireless telecommunications activities &   3,379,297,218  \\
-\hline
-6 & Event catering activities & 2,871,593,754  \\
-\hline
-7 &  Other mining and quarrying n.e.c. &  1,590,548,873  \\
-\hline
-8 & Business and other management consultancy activities &  1,394,373,014 \\
-\hline
-9 &  Activities of head offices &  1,322,692,365 \\
-\hline
-10 & Manufacture of air and spacecraft and related machinery &  1,292,065,497 \\
-\hline
-\end{tabular}
-\end{adjustwidth}
-\end{table}
-
-\begin{table}[!ht]
-\begin{adjustwidth}{-2.25in}{0in}
-\caption{Top 10 UK industries depending on foreign ownership expressed in total turnover (in-going links to the UK)\label{tab:uk_ingoing}}
-\centering
-\medskip
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Classification  & NACE Industries & Total turnover \\ 
-\hline
-1 &      Activities of head offices &  358,099,870  \\
-\hline
-2 & Activities of holding companies  & 259,008,123  \\
-\hline
-3 & Other business support service activities n.e.c. & 183,098,170 \\
-\hline
-4 &  Security and commodity contracts brokerage  &  72,482,167 \\
-\hline
-5 &  Manufacture of motor vehicles  &   65,162,757 \\
-\hline
-6 &  Sale of cars and light motor vehicles  &    61,019,026  \\
-\hline
-7 &  Other financial service activities, except insurance and pension funding & 47,003,937 \\
-\hline
-8 &  Other personal service activities n.e.c.  & 37,926,668  \\
-\hline
-9 &  Wholesale of solid, liquid and gaseous fuels and related products & 36,693,362 \\
-\hline
-10 &  Other information technology and computer service activities  &  35,668,779 \\
-\hline
-\end{tabular}
-\end{adjustwidth}
-\end{table}
-
-\newpage
-When expressed in total number of links, activities mostly dependent on out-going links from the UK are activities of holding companies, of head offices and fund management activities.
-In terms of in-going links these are business and consultancy activities and renting and operating of own and leased real estate.
-
-\begin{table}[!ht]
-\begin{adjustwidth}{-2.25in}{0in}
-\caption{Top 10 UK industries depending on foreign ownership expressed in total links (out-going links from the UK)\label{tab:uk_outgoinglinks}}
-\centering
-\medskip
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Classification  & NACE Industries & Total links \\ 
-\hline
-1 & Activities of holding companies &   686  \\
-\hline
-2 & Other business support service activities n.e.c.  &  567  \\
-\hline
-3 & Activities of head offices &  503 \\
-\hline
-4 &  Fund management activities &   407  \\
-\hline
-5 &  Business and other management consultancy activities &  398  \\
-\hline
-6 &  Extraction of crude petroleum &  366  \\
-\hline
-7 &  Other activities auxiliary to financial services, except insurance and pension funding & 240 \\
-\hline
-8 &  Development of building projects  & 221 \\
-\hline
-9 &  Manufacture of basic pharmaceutical products &  212 \\
-\hline
-10 &  Other information technology and computer service activities  &  181 \\
-\hline
-\end{tabular}
-\end{adjustwidth}
-\end{table}
-
-\begin{table}[!ht]
-\begin{adjustwidth}{-2.25in}{0in}
-\caption{Top 10 UK industries depending on foreign ownership expressed in total links (in-going links to the UK)\label{tab:uk_ingoinglinks}}
-\centering
-\medskip
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Classification  & NACE Industries & Total links \\ 
-\hline
-1 &      Other business support service activities n.e.c. &   16,140  \\
-\hline
-2 & Renting and operating of own or leased real estate  &  6,320  \\
-\hline
-3 & Business and other management consultancy activities &  6,094  \\
-\hline
-4 &  Activities of head offices &   5,628  \\
-\hline
-5 &  Computer consultancy activities &   5,047  \\
-\hline
-6 &  Activities of holding companies &  4,728  \\
-\hline
-7 &  Other personal service activities n.e.c. &  3,996 \\
-\hline
-8 &  Development of building projects &  3,946 \\
-\hline
-9 &  Other information technology and computer service activities &  3,757 \\
-\hline
-10 &  Other financial service activities, except insurance and pension funding  &  3,321 \\
-\hline
-\end{tabular}
-\end{adjustwidth}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%
-\newpage
-
-\section*{S2 Text: Socio-economic distance matrix}
+\subsection*{Socio-economic distance matrix}
 
@@
 
-%%%%%%%%%%%%%%%%%%
-\newpage
-
-\section*{S3 Text: Synthetic model setup}
+\subsection*{Synthetic model setup}
 
@@
 \begin{itemize}
-    \item 
+    \item Sectors distributions follow log-normal densities with most mass in $\left[0;1\right]$
 \end{itemize}
@@
 
-%%%%%%%%%%%%%%%%%%
-\newpage
-
-\section*{S4 Text: Sensitivity analysis}
+\section*{Sensitivity analysis}
 
@@
 \hline
- & \multicolumn{2}{|c|}{$\gamma_G$} & \multicolumn{2}{|c|}{$\gamma_D$} & \multicolumn{2}{|c|}{$\gamma_S$} & \multicolumn{2}{|c|}{$\gamma_W$} & \multicolumn{2}{|c|}{$\gamma_F$} & \multicolumn{2}{|c|}{$\gamma_T$} \\
+ & \multicolumn{2}{|c|}{$\gamma_G$} & \multicolumn{2}{|c|}{$\gamma_D$} & \multicolumn{2}{|c|}{$\gamma_S$} & \multicolumn{2}{|c|}{$\gamma_W$} & \multicolumn{2}{|c|}{$\gamma_O$} & \multicolumn{2}{|c|}{$\gamma_D$} \\
  & F & T & F & T & F & T & F & T & F & T & F & T \\

### 2020-09-01 figures.sh
.

@@ -24,4 +24,8 @@ cp figuresraw/internationalization_countryGravityDecay200_gammaDestination0_face
 cp figuresraw/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png figures/Fig6.png
 
 # Fig 7
-cp figuresraw/pareto_colorgammaOrigin.png figures/Fig7.png
+
+# Fig 8
+cp figuresraw/pareto_colorgammaOrigin.png figures/Fig8.png
+
+# Fig 9

### 2020-09-01 S1_Text.tex

--- 2020-09-01 main.tex
@@
-\documentclass[10pt,letterpaper]{article}
-\input{header.tex}
+\input{si.tex}
+\begin{document}
+\vspace*{0.35in}
+\justify
 
-\usepackage{ragged2e}
-\begin{document}
-\justify
-\vspace*{0.2in}
-
-% Title must be 250 characters or less.
-\begin{flushleft}
-{\Large
-\textbf\newline{A generative model for urban firm networks} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
-}
-% alternative titles?
-% Modeling the growth of urban firm networks
-\newline
-% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
-\\
-Juste Raimbault\textsuperscript{1,2,3*},
-Natalia Zdanowska\textsuperscript{1,3},
-Elsa Arcaute\textsuperscript{3}
-\\
-\bigskip
-\textbf{1} Centre for Advanced Spatial Analysis, UCL, London, United Kingdom
-\\
-\textbf{2} UPS CNRS 3611 ISC-PIF, Paris, France
-\\
-\textbf{3} UMR CNRS 8504 G{\'e}ographie-cit{\'e}s, Paris, France
-\\
-\bigskip
-
-% Use the asterisk to denote corresponding authorship and provide email address in note below.
-* juste.raimbault@polytechnique.edu
-
-\end{flushleft}
-
-% Plos one requirements https://journals.plos.org/plosone/s/submission-guidelines
-% Deadline for the Cities as Complex Systems Call for Papers: 4th September https://collections.plos.org/s/cities
-% Manuscripts can be any length. There are no restrictions on word count, number of figures, or amount of supporting information. 
-% PLOS uses “Vancouver” references style
-% We need to reorganise the paper to have more or less the following sections (which can be reordered or renamed):
-% 1. Title page: List title, authors, and affiliations as first page of manuscript 2. Abstract 3. Introduction 4. Materials and Methods 5. Results 6. Discussion 7. Conclusions (optional)
-
-\section*{Abstract}
-The emergence of interconnected urban networks is a crucial feature of globalisation processes.
-Understanding the drivers behind the growth of such networks -- in particular urban firm networks --, is essential for the economic resilience of urban systems.
-We introduce in this paper a generative network model for firm networks at urban area level including several complementary processes: the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distance.
-An empirical network analysis on European firm ownership data confirms the relevance of each of these factors.
-We then simulate network growth for synthetic systems of cities, unveiling stylized facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range.
-We finally calibrate the model on the European network, outperforming statistical models and showing a strong role of path-dependency.
-Potential applications of the model include the study of mitigation policies to deal with exogenous shocks such as economic crisis or potential lockdowns of countries.
-
-\linenumbers
-
-\section*{Introduction}
-
-\cite{derudder2018central}
-\cite{storme2019introducing} cities firms interactions
-
-Networks constitute the new social morphology of our societies, since the intensification of globalisation of the economy in the late 20th century \cite{castells2000networksociety}.
-A world-city system has since then emerged \cite{taylor2001specification}, characterized by highly interconnected urban centres at the global level \cite{sassen1991}.
-These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity implying spatial and non-spatial patterns \cite{martinus2018global}.
-Interactions of economic nature between transnational firms, are together with international trade patterns among the most representative of the current economic trends \cite{taylor2001specification}.
-Understanding the drivers of growth and geographical properties of such an urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}.
-Straightening these city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}, by developing strategic industries for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
-The functioning of such a city-network and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.
-
-The use of network models to understand processes underlying the emergence and dynamics of such networks has already been tackled in the literature, but still not extensively at city level.
-Generative urban network models combining geographical factors with network topological properties exist at the regional scale {\cite{dai2016simulating}}.
-%Air transportation networks can also be understood as urban networks.
-\cite{zanin2013modelling} reviews several models for such networks.
-Trade networks have been extensively studied from the complex networks perspective, but these are generally driven at the country-level \cite{garlaschelli2005structure} because of a lack of city-level data.
-Input-output models, mostly used in regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}.
-In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for different types of flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2018indirect}.
-
-In this paper we choose to focus on urban networks of firms to question and capture geographical and economic processes as internationalisation, metropolisation, regionalisation and specialisation of cities within a generative model.
-The specific question we tackle is the complementarity of different processes to generate such urban networks.
-The originality of the present paper is to examine these issues within the system of cities framework \cite{berry1964cities}.
-In this approach the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}.
-We examine the interactions of European cities within firm linkages defined by corporate ownership links.
-Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2019arXiv191014652Z}.
-
-More precisely, our contribution consists in the following: (i) we propose an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database \cite{AMADEUS2018}; (ii) we introduce a generative network model to simulate the growth of such linkages at urban areas scale, which combines multiple factors influencing link formation, as economic intensity of origin and destination of urban areas, industrial sector proximity, the strength of previous links, and geographical and socio-cultural distance; (iii) the model allows us to compare the effect of different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; and (iv) we calibrate the model on the empirical network.
-
-The rest of this paper is organized as follows: We first describe the generative model for urban networks and proceed to an empirical analysis of the network used as a case study.
-We then do a sensitivity analysis and exploration of the model on synthetic data.
-The model is calibrated on real European firm ownership data.
-We finally discuss implications of our results and possible extensions.
-
-\section*{Materials and Methods}
-
-\subsection*{Generative urban network model}
-
-\subsubsection*{Rationale}
-
-The model is expected to capture a single macroscopic urban scale, i.e.
-the level of the urban system where basic entities are cities.
-Links are induced by underlying firms but these are not explicit.
-Cities are defined by their economic size, but also by an industrial sector profile.
-As the authors of \cite{martinus2018global} point out, a combination of several distances may play a role in establishing linkages.
-Geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}.
-Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalized to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (confirmed relevant by the fixed effect regression described above); (iii) city economic size, which has furthermore scaling laws properties; (iv) industrial proximity in terms of firm sector composition; (v) previous linkages (influence of already existing links in the past).
-This last point is crucial to capture accumulated effects and include path-dependency, which fits well the use of a simulation model instead of a more simple statistical model.
-
-\subsubsection*{Model description}
-
-Let $1 \leq i \leq N$ cities defined at time $t$ by their economic structure such that $E_i(t)$ is the total economic volume (GDP) for city $i$ and $S_{ik}(t)$ is a matrix giving economic volume proportions within each economic sector $k$, assuming $K$ economic sectors.
-Formally, the model creates links between cities which are characterized by their economic size $E_i$ (GDP) and economic structure $S_{ik}$ in terms of activity sectors (probability distribution of firms within $K$ sectors).
-Starting with an existing network with no links, the model iteratively adds links, following a probability given by a generalized Cobb-Douglas function \cite{vilcu2011geometric} as
-
-\begin{equation}
-p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_O} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_D} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s\left(S_{ik},S_{jk}\right)^{\gamma_S} \cdot \exp \left(- d_{ij} / d_0\right) \cdot \exp \left(- c_{ij} / c_0\right)
-\end{equation}
-
-where $E  =  \sum_k E_k$, $W  = \sum_{i,j} w_{ij}$ weights of previous links, $s(S_{ik},S_{jk})$ is a similarity function between activity sectors $S_{ik}$ and $S_{jk}$ taken as a cosine similarity, $d_{ij}$ euclidian distance, and $g_{ij}$ a socio-cultural distance.
-In the case of an already existing link between two areas, the weight of the latter is incremented by a unit volume $w_0$.
-The model is stopped either when a final time is reached or when the cumulated volume of links reaches a target quantity.
-This formulation can be considered both as an economic utility model and a generalisation of spatial interaction models.
-
-Note that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalise the exponents to 1 (which is different than renormalising the probability), to ensure to include convex functions.
-The ``geopolitical/sociocultural'' distance remains abstract and should be parametrised (see setup below) and estimated in the data-driven setting.
-We assume no evolution of economic size in this simple version of the model, which means that adjusted parameter are valid on a restricted time period.
-
-\subsubsection*{Model setup}
-
-Besides model dynamics, its initialisation (or parametrisation on real data) regarding the spatial distribution, sizes, countries and industrial sector compositions, is also important to understand model behavior.
-In addition  \cite{raimbault2019space} showed that the initial spatial configuration can have significant impacts on model outcomes.
-We therefore consider two versions: one with a synthetic system of cities which can be stochastically repeated, and one other based on real data.
-
-\paragraph{Synthetic setup}
-
-The synthetic system of cities is generated to be at the scale of a continent with properties similar to Europe: (i) $N=700$ cities with an empirical power-law from the Global Human Settlements database for economic sizes ($\alpha = 1.1$), distributed randomly into a uniform and isotropic square space of width $d_{max}=3000km$;
-(ii) clustered into $C = 20$ countries using a k-means algorithm (this remains consistent with non-correlated city sizes at the scale of the countries as our cities are randomly distributed in space \cite{simini2019testing});
-(iii) with synthetic distribution of sectors following log-normal distributions adjusted in a way that large cities have more high-value activities and are more diverse.
-
-More precisely for this last point, assuming sectors are distributed continuously along a one-dimensional axis, cities are assumed to have a sector distribution with a log-normal density function.
-The largest city is constrained to have a standard deviation and a mode of $1/2$ and the smallest $1/K$, with a linear interpolation between the two for other cities.
-Parameters of the log-normal law are obtained by solving the corresponding two equations system for each city, and each distribution is discretised and renormalised to be a vector of probabilities.
-See \ref{S2_Text} for all details of this synthetic setup.
-
-In synthetic experiments, the network is generated from an empty initial network until reaching $t_f=1500$, with a unit link volume of $w_0 = 1$.
-
-\paragraph{Real setup}
-
-The real setup is done using the empirical network analyzed above.
-Urban areas are distributed according to their positions in the actual geographical space, and economic size and sector composition attributes are used.
-The socio-economic distance is constructed using information from the statistical modeling, by considering fixed-effect coefficients estimated for each pair of countries.
-See \ref{S2_Text} for more details on the construction of the socio-economic distance matrix.
-
-Initial network is empty, and the real links are taken as targets for the generated network.
-We constrain the total volume of flows by setting $w_0 = \sum_{ij} w_{ij}^{(obs)} / t_{f}$ when the number of time steps $t_f$ is given as a parameter.
-
-\subsection*{Model parameters and indicators}
-
-The model parameters are the Cobb-Douglas weights $\gamma_O,\gamma_D,\gamma_W,\gamma_S$ and distance decay parameters $d_0,c_0$.
-As the model also includes path-dependency by integrating the influence of previous links, it can not be reduced to a simple statistical model.
-Finally, the role of space introduces even further complexity, yielding a generative model, which has to be simulated.
-
-We consider two main macroscopic indicators to quantify the generated network.
-Firstly geographical structures are captured by internationalisation (modularity of countries in the network).
-The situation of a minimised modularity corresponds to dense connections between cities of different countries of Europe and sparse connections between cities of the same European country.
-In practice, directed modularity is computed by
-\begin{equation}
-
-\end{equation}
-% put the actual modularity
-% add some reference values - some kind of null model
-
-Secondly we consider the metropolisation indicator as the correlation between the weighted degree and the GDP of the city.
-With this indicator we capture the effect of large cities or metropolises of regional importance attracting the biggest amount of links and involved in high interactions with other cities.
-It is computed with a Pearson correlation estimator as
-\begin{equation}
-    M = \hat{\rho} \left[d_i , E_i \right]
-\end{equation}
-
-%I think we need to keep this paragraph but shorten it, because to my understanding network average community size is tested in Figure 5, but not the Louvain modularity for ex.
-%We also consider several network indicators as Louvain modularity, sizes of maximal modularity communities, degree and link weight distributions (described by their average, hierarchy, and entropy), and correlations between degree and city size, and between link weight and distance.
-
-We take a geographical approach to study the behavior of the generative model.
-Other related approaches include indicators of economic complexity, estimating for example the diversity of products produced by countries and how they depend on each other \cite{hidalgo2007product}.
-These are less relevant in our case as the sector structure of cities does not evolve in the model.
-
-\subsection*{Empirical data}
-
-Companies from \emph {AMADEUS} are georeferenced using the Geonames database (using postcode or address depending on the information available).
-We then join this data with the Global Human Settlement dataset GHS-FUA for Functional Urban Areas \cite{Florczyk2019ghs} -- the most relevant objects for applying our model.
-Starting from a firm-level dataset of 1,562,862 nodes and 1,866,936 links, we obtain a directed network of 573 urban areas and 9323 ownership links.
-The weight of links is obtained by computing the owned share of turnover at destination, i.e.
-$w_{ij} = \sum_{k \in i,l \in j} p_{kl} \cdot T_l$ where $p_{kl}$ of the ownership share of company $l$ by company $k$ and $T_l$ the turnover of company $l$.
-Node attributes include an economic size of urban areas, that we compute as the sum of turnovers of companies within the area.
-This quantity is highly correlated with the GDP of the area-level included in the GHS database (Spearman correlation $\rho = 0.71$) and also with population ($\rho = 0.60$).
-We then expect to capture size effects while keeping the consistence of a single main data source.
-An other attribute of the nodes is the industrial sector composition, expressed as a proportion of turnover in the area associated to a given industrial sector.
-Detailed sectors of firms are available up to the four digits of the NACE classification in the raw data \cite{EUROSTAT2008}.
-We consider the first level classification (21 categories) and compute sector proportions accordingly.
-This allows us to define an industrial proximity matrix between urban areas, taking a cosine similarity between vectors as $s_{ij} = \sum_{k=1}^{K} S_{ik}\cdot S_{jk}$ where $S_{ik}$ are sector proportions.
-
-The urban network has heavy tail weighted degree and edge weight distributions, as shown in Fig.~\ref{fig:nwdist}.
-We fit power law and log-normal distributions, including minimal value cutoff following \cite{clauset2009power}, using the \texttt{powerlaw} R package \cite{powerlawpackage}.
-For both, log-normal distributions appear to be a better fit ($\mu=18.8$, $\sigma=2.3$ for weighted degree; $\mu=13.8$, $\sigma=2.8$ for edge weight), and include a much broader range of the empirical distributions with lower estimated cut-offs.
-These heavy tail properties suggest the relevance of a generative model including self-reinforcing processes, which are known to produce such distributions.
-
-%%%%%%%%%%%%%
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=\linewidth]{figures/Fig1.png}
-    \end{center}
-    \caption{Distributions of network properties.
-\textit{(Left)} Cumulated distribution for edge weight, with power law fit (red) and log-normal fit (green); \textit{(Right)} Cumulated distribution for weighted degree.}
-    \label{fig:nwdist}
-\end{figure}
-%%%%%%%%%%%%%
-
-Regarding the community structure of the network, different modularity maximisation algorithms (run on the undirected corresponding network) give the same undirected modularity of 0.38, and a directed modularity \cite{nicosia2009extending} of 0.34 for the greedy algorithm of \cite{clauset2004finding} and 0.36 for the Louvain algorithm \cite{blondel2008fast}, corresponding to 36 communities (resp.
-15).
-In comparison, with 29 countries, the classification of urban areas by country has a modularity of 0.32.
-This means that the structure of flows present certain regional patterns.
-Indeed, a null model assigning random labels with 30 communities, bootstrap 1000 times, yields an average modularity of $0.049 \pm 0.002$, confirming the statistical significance of these communities.
-We show in Fig.~\ref{fig:fig2} a map of the communities obtained.
-
-%%%%%%%%%%%%%%
-\begin{figure}
-    \begin{center}
-    \includegraphics[width=\linewidth]{figures/Fig2.png}
-    \end{center}
-    \caption{{\bf Map of network communities obtained by modularity maximisation.} Area of circles gives the population with the Functional Urban Area, and color the community it belongs to.
-Urban Areas in communities of size smaller than 3 were not colored for readability.}
-    \label{fig:fig2}
-\end{figure}
-%%%%%%%%%%%%%%
-
-\subsubsection*{Statistical analysis}
-
-In order to understand which are the most determining geo-economic factors influencing creation of links between cities, we run a statistical analysis as applied in the case of unconstrained spatial interaction models \cite{wilson1975some}.
-We test five different statistical models, including progressively the processes to be included in the network generative model.
-The most general statistical model is:
-
-\begin{equation}
-\log(w_{ij}) = \log(d_{ij}) + \log(T_i) + \log(T_j) + \log(s_{ij}) + \alpha_{c_i,c_j} + \varepsilon_{ij}
-\end{equation}
-
-where $\alpha_{c_i,c_j}$ is a fixed effect term for the coupling of countries $c_i,c_j$.
-A multiplicative form is applied by taking logarithms as usually done for spatial interaction models.
-Furthermore, origin and/or destination constrained models are not relevant in our case, as links include information about the turnover and a share of ownership at the same time.
-Estimation results are given in Table~\ref{tab:reg}.
-We also fit a Poisson generalized linear model with a logarithmic link on the rounded weights \cite{flowerdew1988fitting}.
-The findings show that when distance only is taken into account (model 1), it has a significant negative effect on link creation, but with an explained variance close to zero.
-Model 2 adds fixed effects by pair of countries, for which around a third are significant, and with a better $R^2$ of 0.17.
-Adding the economic size at origin and destination (turnovers $T_i$ and $T$, model 3) slightly improves the explanatory power (model 3).
-The best model among the linear ones is the model 4, with all factors and fixed effects, both in terms of explained variance ($R^2 = 0.31$) and Akaike information criterion.
-Model 4 also indicates the absence of overfitting.
-Model 5, which is a Poisson model with all variables is much a better fit regarding pseudo-$R^2$ and provides much more significant estimates (all fixed effects significant and much smaller standard deviations).
-Across all models, consistent qualitative stylized facts regarding link creation between cities are revealed: (i) distance has a negative influence; (ii) economic sizes have a positive effect, with the size of the origin city being more important (consistent in the ownership links context, as strongly driven by decision at the headquarter/owner level); (iii) industrial proximity has a positive influence, i.e.
-similarity is more important than complementarity for this dimension; (iv) an large part of country fixed-effects are statistically significant; and (v) the order of magnitude of the influence of each factor is roughly the same, which means that no process totally dominates the others.
-This confirms the relevance of including all these different dimensions in the generative network model, since they have a comparable explanatory role in statistical models.
-
-%%%%%%%%%%%%%
-\begin{table}[!ht]
-\begin{adjustwidth}{-2.25in}{0in}
-\caption{{\bf Statistical models.} Ordinary Least Square estimations of statistical models explaining $\log(w_{ij})$ (standard errors in parenthesis).
-***, ** and * respectively indicate 0.01, 0.05 and 0.1 levels of significance.
-For the fixed effect by origin-destination pairs of countries, the proportion of significant coefficients is attributed (where $p<0.1$).
-Models (1) to (4) are ordinary least square models, while model (5) is a Poisson generalized linear model with logarithmic link.
-We compare model performance using $R^2$, Mean Square Error, and Akaike Information Criterion.
-For the Poisson model (model 5), AIC is not given as it is not comparable, and the pseudo-$R^2$ is given either.
-\label{tab:reg}}
-\medskip
-\begin{tabular}{|l|c|c|c|c|c|}
-\hline
-Model  & (1) & (2) & (3) & (4) & (5) \\ 
-\hline
-$\log(d_{ij})$ &      -0.06** (0.03) &   -0.11** (0.05)  & -0.41*** (0.02)  & -0.35*** (0.04)  &  -0.26*** ($5\cdot 10^{-6}$) \\
-$\log(T_i)$ &   &   & 0.56*** (0.01) &  0.56*** (0.01) & 0.79*** ($2\cdot 10^{-6}$) \\
-$\log(T_j)$ &     &   & 0.39*** (0.01) &  0.39*** (0.01) & 0.66***  ($1.5\cdot 10^{-6}$) \\
-$\log(s_{ij})$ &     &   &  &  0.19*** (0.07) & 0.53*** ($9\cdot 10^{-6}$)  \\
-Countries &    &  28.3\% &   &  26.5\% & 100\% \\
-\hline
-$R^2$ &       0.00059   &  0.17 & 0.21 &  0.31  &  0.56 \\
-MSE & 7.75 & 6.84 & 6.10 & 5.33 & 8.72 \\
-AIC &        44304   &  43578  &  42131  & 41917 &   \\
-\hline
-\end{tabular}
-\end{adjustwidth}
-\end{table}
-%%%%%%%%%%%%%
-
-\section*{Model exploration and calibration}
-
-The model is implemented in NetLogo, which provides a good compromise between performance and interactive prototyping and exploration \cite{railsback2017improving}.
-Numerical experiments are done by integrating the model into the OpenMOLE software for model exploration \cite{reuillon2013openmole}, which provides sensitivity analysis and calibration methods and a transparent access to high-performance computing infrastructures.
-Model source code, aggregated network data (raw initial firm network can not be made available for data ownership reasons), and results are available on the open Git repository of the project at \url{https://github.com/JusteRaimbault/ABMCitiesFirms}.
-Large simulation result files are available on the dataverse repository at \url{https://doi.org/10.7910/DVN/UPX23S}.
-We show in Fig.~\ref{fig:fig3} examples of generated networks, with very contrasted patterns obtained with different spatial interaction ranges.
-Low values of the gravity decay parameter yield local networks around main cities, while long interaction ranges yield more integrated networks.
-We also show networks simulated in the real setting.
-
-%%%%%%%%%%%%%
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=\linewidth]{figures/Fig3.png}
-    \end{center}
-    \caption{Example of simulated networks at $t=1500$, for a synthetic system of cities (Top row) and the European urban network (Bottom row).
-, with a high (resp.
-low) gravity decay parameter for the left column (resp.
-right).\label{fig:fig3}}
-\end{figure}
-%%%%%%%%%%%%%
-
-\subsection*{Synthetic city systems}
-
-We first study extensively the behavior of the model on synthetic systems of cities, to isolate its intrinsic behavior independently of a contingent geographical situation.
-
-\paragraph{Sensitivity analysis}
-
-A global sensitivity analysis first unveils an aggregated influence of parameters on model indicators.
-We apply Saltelli's Global Sensitivity method introduced by \cite{saltelli2008global}.
-This produces values for each parameter and each output of a first order sensitivity index, defined as $Var\left[\mathbb{E}_{\mathbf{X}_{\sim i}}\left(Y_j | X_i\right)\right] / Var\left[Y\right]$, for parameter $X_i$ and output $Y_j$, $\mathbf{X}_{\sim i}$ being all other parameters.
-This captures the total influence of $X_i$ all other things being equal.
-The total order index is given by $\mathbb{E}_{\mathbf{X}_{\sim i}} \left[Var(Y_j | \mathbf{X}_{\sim i})\right] / Var\left[Y_j\right]$ and captures non-homogeneity in behavior and interaction between parameters.
-Indices were estimated with a Sobol sequence sampling of 20,000 model runs \cite{saltelli2010variance}.
-We give in \ref{S4_Text} the full estimated results.
-To summarize, we find that indicators related to modularity (internationalisation and optimal modularity) are mostly influenced by geographical parameters.
-On the contrary, indicators of network structure are driven by origin and destination sizes.
-The influence of sectors and historical links has a secondary but not negligible role.
-Altogether, the analysis witnesses of strong interaction effects between parameters, since for several indicators the total order index is one order of magnitude larger than the first order index.
-This means that the generative model captures some complex behavior of the system.
-% The very different important parameters depending on outputs studied confirms the relevance of a more exhaustive grid exploration.
-
-\paragraph{One factor sampling}
-
-A first numerical experiment of one-factor sampling on all Cobb-Douglas parameters and 100 stochastic repetitions confirms a good statistical convergence (average Sharpe ratios for indicators all larger than 5).
-We use thus 20 repetitions in following larger experiments (in the case of a normal distribution, a 95\% confidence interval on the average is of size $2\sigma \cdot 1.96 / \sqrt{n}$, what gives $n=16$ repetitions for a CI of width of the standard deviation).
-Behavior of some indicators are shown in Fig.~\ref{fig:onefactor}.
-For example, internationalisation and metropolisation both show a transition as a function of $\gamma_G$, from a local to a global regime.
-The sector proximity parameter $\gamma_S$ influences the internationalization linearly, but induces a maximum for the correlation between size and degree, which can be interpreted as a setting where size and the total volume in a city are the most correlated.
-
-%%%%%%%%%%%%%
-\begin{figure}
-	\begin{center}
-	    \includegraphics[width=\linewidth]{figures/Fig4.png}
-	\end{center}
-    \caption{(Top Left) Internationalisation index decreases exponentially with gravity decay; (Top Right) Correlation between city weighted degree and size.
-Both plots show a transition from a local to a global regime.
-(Bottom Left) Internationalisation varies linearly with sector proximity $\gamma_S$; (Bottom Right) Correlation between degree and size exhibits a maximum, witnessing an intermediate regime where size is the most important.
-\label{fig:fig4}}
-\end{figure}
-%%%%%%%%%%%%%
-
-\paragraph{Grid exploration}
-
-The model behaviour is then studied with a grid sampling for parameters and 20 repetitions (the computations run on the European Grid Infrastructure, and are equivalent to 2.5 years of CPU time).
-Distance decay and sector  parameters are varied with 10 steps, other parameters with three.
-The influence of gravity decay parameters is confirmed when plotting the internationalization index against the distance, which exhibits an exponential decay as shown in Fig.~\ref{fig:internat}.
-It interacts with the role of sectors and the size of the origin, with a more significant effect of sector proximity when size has the largest influence (third panel).
-
-%%%%%%%%%%%%%%
-\begin{figure}
-    \includegraphics[width=\linewidth]{figures/Fig5.png}
-    \caption{The transition as a function of interaction range depends on the influence of origin size $\gamma_O$; sector proximity $\gamma_S$ plays a role only for a large influence of the origin.\label{fig:fig5}}
-\end{figure}
-%%%%%%%%%%%%%%
-
-Other indicators exhibit non-trivial patterns.
-For example, when considering average community size of the final network as shown in Fig.~\ref{fig:comsize}, we obtain a maximal integration in terms of communities at an intermediate value of the gravity decay, what can be interpreted as the emergence of a regional regime.
-The size of communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.
-In particular, we observe a qualitative inversion of the role of $\gamma_S$ when introducing an effect of the origin (switching $\gamma_F$ from 0 to 1, first panel to second panel).
-The maximal community size disappears when $\gamma_F = 2$, implying a regime with small local communities where large cities mostly connect with there hinterland.
-These experiments reveal a complex interplay between processes and how the model produces diverse stylised facts.
-
-%%%%%%%%%%%%%%
-\begin{figure}
-    \includegraphics[width=\textwidth]{figures/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
-	\caption{Maximal integration in terms of community size is achieved at an intermediate value of $d_G$: emergence of a regional regime; Maximal size depends on the role of sectors $\gamma_S$, in a decreasing way when origin size is deactivated, and increasing way when $\gamma_F=1$; This regime disappear when origin size influence is too large.\label{fig:comsize}}
-\end{figure}
-%%%%%%%%%%%%%%
-
-\paragraph{Impact of urban hierarchy}
-
-We run a specific experiment to study the impact of urban hierarchy on firms dynamics.
-Rank-size law in cities is a central stylized fact to urban theories, despite the discussions on how to estimate the power-law exponent, its actual values \cite{cottineau2017metazipf} and how its depends on the definition of cities considered \cite{corral2020truncated}.
-In that particular context, understanding how this parameter can impact urban dynamics within artificial systems of cities in which it can be modified to extreme values not observed in real systems, should inform on underlying processes.
-% too abstract: clarify why hypothetical systems of cities bring info on real ones
-We change the initial hierarchy $\alpha$ from 0.2 to 2.0 with a step of 0.1, and also vary interaction range $d_0$.
-
-\subsection*{Model calibration on the European urban network}
-
-We then apply the model to a real system of cities by calibrating it on the European ownership network.
-Model setup is done following the real setup described above.
-The number of time steps $t_f$ is left as an additional parameter, which in a sense determines the granularity of the cumulative network generation process.
-The objective functions for calibration are the average mean squared error on logarithms of weights given by $\varepsilon_L = \frac{1}{N^2} \sum_{i,j} \left(\log w_{ij} - \log \hat{w}_{ij} \right)^2$ if $\hat{w}_{ij}$ are the simulated weights, and the logarithm of the mean squared error $\varepsilon_M = \log\left(\frac{1}{N^2} \sum_{i,j} \left(w_{ij} - \hat{w}_{ij}\right)^2 \right)$.
-\cite{raimbault2018indirect} showed that these two are complementary to calibrate urban systems models.
-Calibration is done using a genetic NSGA2 algorithm \cite{deb2002fast}, which is particularly suited for such a bi-objective calibration of a simulation model.
-Stochasticity is taken into account by adding the estimation confidence as an additional objective, and we filter the final population to have at least 20 samples.
-We use a population of 200 individuals and run the algorithm for 20,000 generations, on 1000 islands in parallel.
-
-%%%%%%%%%%%%%%
-\begin{figure}
-	\begin{center}
-    \includegraphics[width=0.75\textwidth]{figures/pareto_colorgammaOrigin.png}
-    \end{center}
-	\caption{Pareto front obtained for the bi-objective model calibration for $\varepsilon_L$ and $\varepsilon_M$ objectives.
-$\varepsilon_L$ can be compared to the MSE of statistical models fitted above, and the generative model outperform these for several parametrizations (lowest MSE 5.33 for statistical models).
-Point size gives the number of stochastic samples and color the value of $\gamma_F$.\label{fig:calibration}}
-\end{figure}
-%%%%%%%%%%%%%%
-
-We show in Fig.~\ref{fig:calibration} the calibration result, as a Pareto front of compromise points between the two objectives.
-As the mean squared error on logarithm is the same than the one for statistical models, it can be directly compared.
-The other objective would correspond to statistical models directly on variables: therefore, our model covers in a sense a set of intermediate models which make a compromise between the two objectives, and is therefore much more flexible.
-The values for $\varepsilon_L$ are for a large proportion of points lower than for statistical models, our model outperforming these in this predictive sense.
-As shown in Fig.~\ref{fig:calibration}, increasing values for $\gamma_F$ lead to better performance regarding $\varepsilon_L$: increasing the importance of the origin will provide a better fit on all order of magnitudes of link weights (while $\varepsilon_M$ favors mostly the largest links given the fat-tail distribution).
-
-When considering optimal points with $\varepsilon_L < 5.0$, the adjusted parameters are: long range gravity interactions and socio-economic interactions given by $d_0 = 5972 \pm 2148$ and $c_0 = 113 \pm 33$ (the maximal $c_{ij}$ is 11.8); a stronger influence of destination than origin as $\gamma_O = 1.46 \pm 0.10$ and $\gamma_D = 2.0 \pm 0.05$; a comparable influence of sectors as $\gamma_S = 1.76 \pm 0.21$; a fine time granularity as $t_f = 4951 \pm 124$; and a very strong influence of previous links as $\gamma_W = 6.43 \pm 4.11$.
-This last point furthermore shows in particular a role of the path-dependency parameter, confirming the relevance of this complex generative model including path-dependency.
-The interaction of this process with others implies an inversion of the relative importance of origin and destination (on the contrary to statistical models).
-The fact that calibrated interaction distances are large is consistent with long range economic interactions within Europe.
-Besides these stylized facts, these calibration results altogether demonstrate the feasibility of model application to real data.
-
-\subsection*{Application: impact of economic shock}
-
-Based on calibrated parameter values, we design scenarios to evaluate the impact of economic shocks.
-These can be the consequence of diverse factors, including socio-economic, geopolitical factors (example of Brexit), and natural disasters such as COVID-19.
-We simulate a full border closure by setting $\gamma_D = 0$, and intermediate scenarios.
-Setting also $\gamma_G$ to 0 allows simulating inter-country lock-downs.
-% run simulations
-% note that time scales are in fact smaller - or assume that crisis keeping border closed is rather long
-
-%%%%%%%%%%%%%%%
-\section*{Discussion}
-
-This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMole platform and calibrated on real data on ownership linkages of firms for Europe from Amadeus dataset.
-The simulation on a synthetic system of cities provides an broad knowledge on model behavior in its parameter space, and its calibration suggests its potential application to policy issues.
-This kind of model can therefore have practical applications in the future in terms of testing the effect of exogenous shocks in the socio-economic structure: as the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to COVID19 on the single market and on the strongly foreign dependent economic sectors (see Appendix).
-
-Even in such a simple model (close to directly tractable stationary state) the behaviour is highly non-linear in many dimensions.
-Our study shows how crucial model exploration is to overcome hidden parameters (deactivated mechanisms or default parameter values).
-This paper emphasis on the fact that exploration of intrinsic dynamics on synthetic data is of great importance and should be driven before an application on real data so as to put aside the geographical effects from model dynamics.
-
-Several possible extensions of the current work can be undertaken in the future as a co-evolution model with an evolution of city sizes or a model with firm agents leading to a multi-scale agent-based model.
-A more precise study of the role of path dependency can also be undertaken in order to understand the creation of the future links.
-Other formulations of the model might be taken into consideration as other formulation of the combination of factors or multi-objective optimisation depending on sectors using Pareto fronts.
-
-\section*{Supporting information}
-
-\paragraph*{S1 Text.}
-\label{S1_Text}
-{\bf UK's economic sectors dependent on international links.}
-
-\paragraph*{S2 Text.}
-\label{S2_Text}
-{\bf Socio-economic distance matrix.}
-
-\paragraph*{S3 Text.}
-\label{S3_Text}
-{\bf Synthetic model setup.}
-
-\paragraph*{S4 Text.}
-\label{S4_Text}
-{\bf Sensitivity analysis.}
-
-\section*{Acknowledgments}
-
-Results obtained in this paper were computed on the vo.complex-system.eu virtual organization of the European Grid Infrastructure (http://www.egi.eu).
-We thank the European Grid Infrastructure and its supporting National Grid Initiatives (France-Grilles in particular) for providing the technical support and infrastructure.
-
-\bibliographystyle{plos2015}
-\bibliography{biblio}
-
-%%%%%%%%%%%%%
-\newpage
 \section*{S1 Text: UK's economic sectors dependent on international links}
@@
 
-%%%%%%%%%%%%%%%%%%
-\newpage
-
-\section*{S2 Text: Socio-economic distance matrix}
-
-% \todo{more detail / descriptive stats of the matrix / function of the distance between countries}
-
-%%%%%%%%%%%%%%%%%%
-\newpage
-
-\section*{S3 Text: Synthetic model setup}
-
-The rationale for the synthetic sector setup is the following:
-\begin{itemize}
-    \item 
-\end{itemize}
-
-%solve for mu,sigma of the log-normal as a function of gdp
-%       we aim at having a log-normal mostly in [0,1] such that
-%       (mode,variance) =  1 / K for the smallest log(E_i) and = 1/2 for the largest log(E_i)
-%        writing
-%        e_i = (log(E_i) - log(E_imin))/(log(E_imax) - log(E_imin)) * (1/2 - 1/K) + 1/K
-%        this yields
-%         (1)  mu - sigma^2 = log(e_i)
-%         (2) -3 sigma^2 - 2 log(exp(sigma^2) - 1) = log(e_i)
-%        => sigma^2 is the unique positive root of f(X)=0 with f(X) = -3X - 2 ln(exp(X) - 1) - ln(e_i)
-
-%%%%%%%%%%%%%%%%%%
-\newpage
-
-\section*{S4 Text: Sensitivity analysis}
-
-%%%%%%%%%%%%%
-\begin{table}
-\begin{adjustwidth}{-2.25in}{0in}
-\caption{Saltelli sensitivity indices, for indicators in rows and parameters in columns. We give for each pair the first order index (F) and the total order index (T).\label{tab:saltelli}}
-\hspace{-1cm}\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|c|c|}
-\hline
- & \multicolumn{2}{|c|}{$\gamma_G$} & \multicolumn{2}{|c|}{$\gamma_D$} & \multicolumn{2}{|c|}{$\gamma_S$} & \multicolumn{2}{|c|}{$\gamma_W$} & \multicolumn{2}{|c|}{$\gamma_F$} & \multicolumn{2}{|c|}{$\gamma_T$} \\
- & F & T & F & T & F & T & F & T & F & T & F & T \\
- \hline
-Internationalisation & 0.2 & 0.3 & 0.7 & 0.7 & 0.001 & 0.009 & $4\cdot 10^{-4}$ & 0.007 & 0.03 & 0.04 & 0.02 & 0.04 \\
-Metropolisation & 0.02 & 0.1 & 0.02 & 0.2 & 0.002 & 0.1 & 0.001 & 0.09 & 0.2 & 0.6 & 0.3 & 0.6 \\
-Modularity & 0.3 & 0.4 & 0.6 & 0.6 & 0.004 & 0.02 & $3\cdot 10^{-4}$ & 0.01 & 0.005 & 0.03 & 0.002 & 0.03 \\
-Avg. com. size & 0.008 & 0.09 & 0.01 & 0.1 & 0.002 & 0.07 & 0.003 & 0.04 & 0.3 & 0.6 & 0.4 & 0.6 \\
-Degree entropy & 0.006 & 0.02 & 0.003 & 0.02 & 0.006 & 0.03 & 0.008 & 0.02 & 0.5 & 0.5 & 0.5 & 0.5 \\
-Weight entropy & 0.04 & 0.1 & 0.03 & 0.1 & 0.008 & 0.08 & 0.01 & 0.07 & 0.4 & 0.5 & 0.4 & 0.5 \\\hline
-\end{tabular}
-\end{adjustwidth}
-\end{table}
-%%%%%%%%%%%%%
-
 \end{document}

### 2020-09-01 main.tex

@@
 
+% comment upon Netherlands Poland cluster justifying the ownership from Netherlands of polish firms (Zdanowska, 2018+ Zdanowska 2019) + Germany Czech Republic Hungary Romania Bulgaria corridor (firms+ trade road to Turkey)
+
 %%%%%%%%%%%%%%
@@
 The size of communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.
-In particular, we observe a qualitative inversion of the role of $\gamma_S$ when introducing an effect of the origin (switching $\gamma_F$ from 0 to 1, first panel to second panel).
-The maximal community size disappears when $\gamma_F = 2$, implying a regime with small local communities where large cities mostly connect with there hinterland.
+In particular, we observe a qualitative inversion of the role of $\gamma_S$ when introducing an effect of the origin (switching $\gamma_O$ from 0 to 1, first panel to second panel).
+The maximal community size disappears when $\gamma_O = 2$, implying a regime with small local communities where large cities mostly connect with there hinterland.
 These experiments reveal a complex interplay between processes and how the model produces diverse stylised facts.
@@
 \begin{figure}
-    \includegraphics[width=\textwidth]{figures/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
-	\caption{Maximal integration in terms of community size is achieved at an intermediate value of $d_G$: emergence of a regional regime; Maximal size depends on the role of sectors $\gamma_S$, in a decreasing way when origin size is deactivated, and increasing way when $\gamma_F=1$; This regime disappear when origin size influence is too large.\label{fig:comsize}}
+    \includegraphics[width=\textwidth]{figures/Fig6.png}
+	\caption{Maximal integration in terms of community size is achieved at an intermediate value of $d_G$: emergence of a regional regime; Maximal size depends on the role of sectors $\gamma_S$, in a decreasing way when origin size is deactivated, and increasing way when $\gamma_F=1$; This regime disappear when origin size influence is too large.\label{fig:fig6}}
 \end{figure}
@@
 
+%%%%%%%%%%%%%
+\begin{figure}
+    \begin{center}
+        %\includegraphics[width=\linewidth]{figures/Fig7.png}
+    \end{center}
+    \caption{\textbf{Impact of urban hierarchy.}\label{fig:fig7}}
+\end{figure}
+%%%%%%%%%%%%%
+
 \subsection*{Model calibration on the European urban network}
@@
 Stochasticity is taken into account by adding the estimation confidence as an additional objective, and we filter the final population to have at least 20 samples.
-We use a population of 200 individuals and run the algorithm for 20,000 generations, on 1000 islands in parallel.
+We use a population of 200 individuals and run the algorithm for 50,000 generations, on 1000 islands in parallel.
 
@@
 	\begin{center}
-    \includegraphics[width=0.75\textwidth]{figures/pareto_colorgammaOrigin.png}
+    \includegraphics[width=0.75\linewidth]{figures/Fig8.png}
     \end{center}
-	\caption{Pareto front obtained for the bi-objective model calibration for $\varepsilon_L$ and $\varepsilon_M$ objectives.
+	\caption{\textbf{Model calibration results.} Pareto front obtained for the bi-objective model calibration for $\varepsilon_L$ and $\varepsilon_M$ objectives.
 $\varepsilon_L$ can be compared to the MSE of statistical models fitted above, and the generative model outperform these for several parametrizations (lowest MSE 5.33 for statistical models).
-Point size gives the number of stochastic samples and color the value of $\gamma_F$.\label{fig:calibration}}
+Point size gives the number of stochastic samples and color the value of $\gamma_O$.\label{fig:fig8}}
 \end{figure}
@@
 
-We show in Fig.~\ref{fig:calibration} the calibration result, as a Pareto front of compromise points between the two objectives.
+We show in Fig.~\ref{fig:fig8} the calibration result, as a Pareto front of compromise points between the two objectives.
 As the mean squared error on logarithm is the same than the one for statistical models, it can be directly compared.
@@
 The values for $\varepsilon_L$ are for a large proportion of points lower than for statistical models, our model outperforming these in this predictive sense.
-As shown in Fig.~\ref{fig:calibration}, increasing values for $\gamma_F$ lead to better performance regarding $\varepsilon_L$: increasing the importance of the origin will provide a better fit on all order of magnitudes of link weights (while $\varepsilon_M$ favors mostly the largest links given the fat-tail distribution).
+As shown in Fig.~\ref{fig:fig7}, increasing values for $\gamma_O$ lead to better performance regarding $\varepsilon_L$: increasing the importance of the origin will provide a better fit on all order of magnitudes of link weights (while $\varepsilon_M$ favors mostly the largest links given the fat-tail distribution).
 
@@
 The simulation on a synthetic system of cities provides an broad knowledge on model behavior in its parameter space, and its calibration suggests its potential application to policy issues.
-This kind of model can therefore have practical applications in the future in terms of testing the effect of exogenous shocks in the socio-economic structure: as the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to COVID19 on the single market and on the strongly foreign dependent economic sectors (see Appendix).
+This kind of model can therefore have practical applications in the future in terms of testing the effect of exogenous shocks in the socio-economic structure as the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to COVID19 on the single market and on the strongly foreign dependent economic sectors.
+More precisely the UK sectors generating the highest turnover in 2018 mostly dependent on foreign capital they own (out-going links from the UK) are extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods.
+In terms of in-going links to the UK, activities of head offices and holding companies, sale of cars and manufacture of motor vehicles are sectors mostly dependent on foreign ownership coming from abroad.
 
@@
 \label{S2_Text}
-{\bf Socio-economic distance matrix.}
+{\bf Model setup and sensitivity analysis.}
 
-\paragraph*{S3 Text.}
-\label{S3_Text}
-{\bf Synthetic model setup.}
-
-\paragraph*{S4 Text.}
-\label{S4_Text}
-{\bf Sensitivity analysis.}
+Socio-economic distance matrix, Synthetic model setup and Sensitivity analysis.
 
@@
 
-%%%%%%%%%%%%%
-\newpage
-\section*{S1 Text: UK's economic sectors dependent on international links}
-
-Based on \emph {AMADEUS} dataset, the most vulnerable economic sectors in the United Kingdom have been highlighted.
-The later are the ones due to be mostly affected by eventual future lock-downs of the international economy (international ownership links).
-
-When considering the UK sectors generating the highest turnover in 2018 mostly dependent on foreign capital they own (out-going links from the UK) these are extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods.
-In terms of in-going links to the UK, activities of head offices and holding companies, sale of cars and manufacture of motor vehicles are sectors mostly dependent on foreign ownership coming from abroad.
-
-\FloatBarrier
-
-\begin{table}[!ht]
-\begin{adjustwidth}{-2.25in}{0in}
-\caption{Top 10 UK industries depending on foreign ownership expressed in total turnover (out-going links from the UK)\label{tab:uk_outgoing}}
-\centering
-\medskip
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Classification  & NACE Industries & Total turnover \\ 
-\hline
-1 &     Extraction of crude petroleum &   96,458,258,642  \\
-\hline
-2 & Manufacture of tobacco products &  5,477,272,464  \\
-\hline
-3 & Manufacture of basic pharmaceutical products &  4,825,881,930 \\
-\hline
-4 &  Retail sale in non-specialised stores with food, beverages or tobacco predominating  &   3,681,373,374  \\
-\hline
-5 &  Wireless telecommunications activities &   3,379,297,218  \\
-\hline
-6 & Event catering activities & 2,871,593,754  \\
-\hline
-7 &  Other mining and quarrying n.e.c. &  1,590,548,873  \\
-\hline
-8 & Business and other management consultancy activities &  1,394,373,014 \\
-\hline
-9 &  Activities of head offices &  1,322,692,365 \\
-\hline
-10 & Manufacture of air and spacecraft and related machinery &  1,292,065,497 \\
-\hline
-\end{tabular}
-\end{adjustwidth}
-\end{table}
-
-\begin{table}[!ht]
-\begin{adjustwidth}{-2.25in}{0in}
-\caption{Top 10 UK industries depending on foreign ownership expressed in total turnover (in-going links to the UK)\label{tab:uk_ingoing}}
-\centering
-\medskip
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Classification  & NACE Industries & Total turnover \\ 
-\hline
-1 &      Activities of head offices &  358,099,870  \\
-\hline
-2 & Activities of holding companies  & 259,008,123  \\
-\hline
-3 & Other business support service activities n.e.c. & 183,098,170 \\
-\hline
-4 &  Security and commodity contracts brokerage  &  72,482,167 \\
-\hline
-5 &  Manufacture of motor vehicles  &   65,162,757 \\
-\hline
-6 &  Sale of cars and light motor vehicles  &    61,019,026  \\
-\hline
-7 &  Other financial service activities, except insurance and pension funding & 47,003,937 \\
-\hline
-8 &  Other personal service activities n.e.c.  & 37,926,668  \\
-\hline
-9 &  Wholesale of solid, liquid and gaseous fuels and related products & 36,693,362 \\
-\hline
-10 &  Other information technology and computer service activities  &  35,668,779 \\
-\hline
-\end{tabular}
-\end{adjustwidth}
-\end{table}
-
-\newpage
-When expressed in total number of links, activities mostly dependent on out-going links from the UK are activities of holding companies, of head offices and fund management activities.
-In terms of in-going links these are business and consultancy activities and renting and operating of own and leased real estate.
-
-\begin{table}[!ht]
-\begin{adjustwidth}{-2.25in}{0in}
-\caption{Top 10 UK industries depending on foreign ownership expressed in total links (out-going links from the UK)\label{tab:uk_outgoinglinks}}
-\centering
-\medskip
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Classification  & NACE Industries & Total links \\ 
-\hline
-1 & Activities of holding companies &   686  \\
-\hline
-2 & Other business support service activities n.e.c.  &  567  \\
-\hline
-3 & Activities of head offices &  503 \\
-\hline
-4 &  Fund management activities &   407  \\
-\hline
-5 &  Business and other management consultancy activities &  398  \\
-\hline
-6 &  Extraction of crude petroleum &  366  \\
-\hline
-7 &  Other activities auxiliary to financial services, except insurance and pension funding & 240 \\
-\hline
-8 &  Development of building projects  & 221 \\
-\hline
-9 &  Manufacture of basic pharmaceutical products &  212 \\
-\hline
-10 &  Other information technology and computer service activities  &  181 \\
-\hline
-\end{tabular}
-\end{adjustwidth}
-\end{table}
-
-\begin{table}[!ht]
-\begin{adjustwidth}{-2.25in}{0in}
-\caption{Top 10 UK industries depending on foreign ownership expressed in total links (in-going links to the UK)\label{tab:uk_ingoinglinks}}
-\centering
-\medskip
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Classification  & NACE Industries & Total links \\ 
-\hline
-1 &      Other business support service activities n.e.c. &   16,140  \\
-\hline
-2 & Renting and operating of own or leased real estate  &  6,320  \\
-\hline
-3 & Business and other management consultancy activities &  6,094  \\
-\hline
-4 &  Activities of head offices &   5,628  \\
-\hline
-5 &  Computer consultancy activities &   5,047  \\
-\hline
-6 &  Activities of holding companies &  4,728  \\
-\hline
-7 &  Other personal service activities n.e.c. &  3,996 \\
-\hline
-8 &  Development of building projects &  3,946 \\
-\hline
-9 &  Other information technology and computer service activities &  3,757 \\
-\hline
-10 &  Other financial service activities, except insurance and pension funding  &  3,321 \\
-\hline
-\end{tabular}
-\end{adjustwidth}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%
-\newpage
-
-\section*{S2 Text: Socio-economic distance matrix}
-
-% \todo{more detail / descriptive stats of the matrix / function of the distance between countries}
-
-%%%%%%%%%%%%%%%%%%
-\newpage
-
-\section*{S3 Text: Synthetic model setup}
-
-The rationale for the synthetic sector setup is the following:
-\begin{itemize}
-    \item 
-\end{itemize}
-
-%solve for mu,sigma of the log-normal as a function of gdp
-%       we aim at having a log-normal mostly in [0,1] such that
-%       (mode,variance) =  1 / K for the smallest log(E_i) and = 1/2 for the largest log(E_i)
-%        writing
-%        e_i = (log(E_i) - log(E_imin))/(log(E_imax) - log(E_imin)) * (1/2 - 1/K) + 1/K
-%        this yields
-%         (1)  mu - sigma^2 = log(e_i)
-%         (2) -3 sigma^2 - 2 log(exp(sigma^2) - 1) = log(e_i)
-%        => sigma^2 is the unique positive root of f(X)=0 with f(X) = -3X - 2 ln(exp(X) - 1) - ln(e_i)
-
-%%%%%%%%%%%%%%%%%%
-\newpage
-
-\section*{S4 Text: Sensitivity analysis}
-
-%%%%%%%%%%%%%
-\begin{table}
-\begin{adjustwidth}{-2.25in}{0in}
-\caption{Saltelli sensitivity indices, for indicators in rows and parameters in columns. We give for each pair the first order index (F) and the total order index (T).\label{tab:saltelli}}
-\hspace{-1cm}\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|c|c|}
-\hline
- & \multicolumn{2}{|c|}{$\gamma_G$} & \multicolumn{2}{|c|}{$\gamma_D$} & \multicolumn{2}{|c|}{$\gamma_S$} & \multicolumn{2}{|c|}{$\gamma_W$} & \multicolumn{2}{|c|}{$\gamma_F$} & \multicolumn{2}{|c|}{$\gamma_T$} \\
- & F & T & F & T & F & T & F & T & F & T & F & T \\
- \hline
-Internationalisation & 0.2 & 0.3 & 0.7 & 0.7 & 0.001 & 0.009 & $4\cdot 10^{-4}$ & 0.007 & 0.03 & 0.04 & 0.02 & 0.04 \\
-Metropolisation & 0.02 & 0.1 & 0.02 & 0.2 & 0.002 & 0.1 & 0.001 & 0.09 & 0.2 & 0.6 & 0.3 & 0.6 \\
-Modularity & 0.3 & 0.4 & 0.6 & 0.6 & 0.004 & 0.02 & $3\cdot 10^{-4}$ & 0.01 & 0.005 & 0.03 & 0.002 & 0.03 \\
-Avg. com. size & 0.008 & 0.09 & 0.01 & 0.1 & 0.002 & 0.07 & 0.003 & 0.04 & 0.3 & 0.6 & 0.4 & 0.6 \\
-Degree entropy & 0.006 & 0.02 & 0.003 & 0.02 & 0.006 & 0.03 & 0.008 & 0.02 & 0.5 & 0.5 & 0.5 & 0.5 \\
-Weight entropy & 0.04 & 0.1 & 0.03 & 0.1 & 0.008 & 0.08 & 0.01 & 0.07 & 0.4 & 0.5 & 0.4 & 0.5 \\\hline
-\end{tabular}
-\end{adjustwidth}
-\end{table}
-%%%%%%%%%%%%%
-
 \end{document}

### 2020-09-01 main.tex
Update on Overleaf.

@@
 {\Large
-\textbf\newline{A generative model for urban firm networks} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{Modeling growth of urban firm networks} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
 % alternative titles?
-% Modeling the growth of urban firm networks
+% Old title Generative Urban Network Model
 \newline
@@
 The situation of a minimised modularity corresponds to dense connections between cities of different countries of Europe and sparse connections between cities of the same European country.
-In practice, directed modularity is computed by
+In practice, modularity is computed by
 \begin{equation}
-
+    I = \frac{1}{2 \left|E\right|} \cdot \sum_{i,j} \left(w_{ij} - \frac{d_i d_j}{2 \left|E\right|}\right) \mathbb{1}_{c_i = c_j}
 \end{equation}
-% put the actual modularity
+where $d_i$ is the weighted degree of cities and $c_i$ the countries.
 % add some reference values - some kind of null model
@@
 
-% comment upon Netherlands Poland cluster justifying the ownership from Netherlands of polish firms (Zdanowska, 2018+ Zdanowska 2019) + Germany Czech Republic Hungary Romania Bulgaria corridor (firms+ trade road to Turkey)
-
+The communities reflect characteristic socio-cultural structures and cross-border economic ties throughout Europe.
+Cities in France, Belgium and Luxembourg belong to the same cluster as they share French as a common business language, but also cross-border tax arrangements creating easier economic interactions \cite{DecovilleDurand2019}.
+Cities of the United Kingdom and Ireland are in the same cluster for the same socio-cultural proximity matters, so as Sweden, Finland, Estonia, Latvia and Lithuania.
+The cluster composed of Dutch and Polish cities is justified by the fact that firms from the Netherlands are among the first foreign owners of polish firms since 2010, ranked even before the German ones \cite{2019arXiv191014652Z}.
+The cluster including cities in Germany, Czech Republic, Austria, Slovakia, Hungary and partly in Romania, find its origin in the dependence of the Central and Eastern European economy on German and Austrian large companies mostly from the automotive sector (ie.
+Audi, Volkswagen).
+In addition all cities of the later cluster have in common to be at the intersection of a very important freight transport route in Europe from Germany to Turkey \cite{Zdanowska2017}.
 %%%%%%%%%%%%%%
@@
 We apply Saltelli's Global Sensitivity method introduced by \cite{saltelli2008global}.
-This produces values for each parameter and each output of a first order sensitivity index, defined as $Var\left[\mathbb{E}_{\mathbf{X}_{\sim i}}\left(Y_j | X_i\right)\right] / Var\left[Y\right]$, for parameter $X_i$ and output $Y_j$, $\mathbf{X}_{\sim i}$ being all other parameters.
+This produces values for each parameter and each output of a first order sensitivity index, defined as $Var\left[\mathbb{E}_{\mathbf{X}_{\sim i}}\left(Y_j | X_i\right)\right] / Var\left[Y_j\right]$, for parameter $X_i$ and output $Y_j$, $\mathbf{X}_{\sim i}$ being all other parameters.
 This captures the total influence of $X_i$ all other things being equal.
@@
 This means that the generative model captures some complex behavior of the system.
-% The very different important parameters depending on outputs studied confirms the relevance of a more exhaustive grid exploration.
 
@@
 In that particular context, understanding how this parameter can impact urban dynamics within artificial systems of cities in which it can be modified to extreme values not observed in real systems, should inform on underlying processes.
-% too abstract: clarify why hypothetical systems of cities bring info on real ones
+% too abstract: clarify why hypothetical systems of cities bring info on real ones \cite{raimbault2019space}
 We change the initial hierarchy $\alpha$ from 0.2 to 2.0 with a step of 0.1, and also vary interaction range $d_0$.
@@
 
+%%%%%%%%%%%%%%
+\begin{figure}
+	\begin{center}
+    %\includegraphics[width=\linewidth]{figures/Fig9.png}
+    \end{center}
+	\caption{\textbf{Application of the calibrated model to an economic shock scenario.}\label{fig:fig9}}
+\end{figure}
+%%%%%%%%%%%%%%
+
 %%%%%%%%%%%%%%%
@@
 This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMole platform and calibrated on real data on ownership linkages of firms for Europe from Amadeus dataset.
-The simulation on a synthetic system of cities provides an broad knowledge on model behavior in its parameter space, and its calibration suggests its potential application to policy issues.
-This kind of model can therefore have practical applications in the future in terms of testing the effect of exogenous shocks in the socio-economic structure as the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to COVID19 on the single market and on the strongly foreign dependent economic sectors.
-More precisely the UK sectors generating the highest turnover in 2018 mostly dependent on foreign capital they own (out-going links from the UK) are extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods.
-In terms of in-going links to the UK, activities of head offices and holding companies, sale of cars and manufacture of motor vehicles are sectors mostly dependent on foreign ownership coming from abroad.
-
+The simulation on a synthetic system of cities provides a broad knowledge on model behavior in its parameter space.
 Even in such a simple model (close to directly tractable stationary state) the behaviour is highly non-linear in many dimensions.
@@
 This paper emphasis on the fact that exploration of intrinsic dynamics on synthetic data is of great importance and should be driven before an application on real data so as to put aside the geographical effects from model dynamics.
+
+The calibration of model suggests its potential application to policy issues and can therefore have practical applications in the future.
+For example the effect of exogenous shocks in the socio-economic structure could be tested as the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to further pandemic outbreaks on the single market.
+In this sense the UK economic sectors mostly dependent on foreign capital and generating the highest turnovers in 2018 could be examined: extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods (in terms of out-going links from the UK) ; activities of head offices and holding companies, sale of cars and manufacture of motor vehicles (in terms of in-going links to the UK).
 

### 2020-09-02 figures.sh
fig_7

@@ -24,6 +24,7 @@ cp figuresraw/internationalization_countryGravityDecay200_gammaDestination0_face
 cp figuresraw/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png figures/Fig6.png
 
 # Fig 7
+montage figuresraw/internationalization-alpha_colorgravityDecay_errorBar.png figuresraw/rhoDegreeSize-alpha_colorgravityDecay_errorBar.png -tile 2x1 -geometry +"$HORIZONTALPADDING"+0 -resize "$WIDTH"x figures/Fig7.png
 
 # Fig 8
 cp figuresraw/pareto_colorgammaOrigin.png figures/Fig8.png

### 2020-09-02 main.tex

@@
 
-The rest of this paper is organized as follows: We first describe the generative model for urban networks and proceed to an empirical analysis of the network used as a case study.
+The rest of this paper is organized as follows: we first describe the generative model for urban networks and proceed to an empirical analysis of the network used as a case study.
 We then do a sensitivity analysis and exploration of the model on synthetic data.

### 2020-09-03 main.tex
Update on Overleaf.

@@
 
-\cite{derudder2018central}
-\cite{storme2019introducing} cities firms interactions
+%is it something you want to add?
+%\cite{derudder2018central}
+%\cite{storme2019introducing} cities firms interactions
 
@@
 Rank-size law in cities is a central stylized fact to urban theories, despite the discussions on how to estimate the power-law exponent, its actual values \cite{cottineau2017metazipf} and how its depends on the definition of cities considered \cite{corral2020truncated}.
-In that particular context, understanding how this parameter can impact urban dynamics within artificial systems of cities in which it can be modified to extreme values not observed in real systems, should inform on underlying processes.
-% too abstract: clarify why hypothetical systems of cities bring info on real ones \cite{raimbault2019space}
-We change the initial hierarchy $\alpha$ from 0.2 to 2.0 with a step of 0.1, and also vary interaction range $d_0$.
+In that particular context, understanding how this parameter can impact urban dynamics within artificial systems of cities in which it can be modified to extreme values not observed in real systems, should inform on underlying processes \cite{raimbault2019space}.
+We change the initial hierarchy $\alpha$ from 0.5 to 2.0 with a step of 0.1, and also vary the interaction range $d_0$.
 
@@
     \begin{center}
-        %\includegraphics[width=\linewidth]{figures/Fig7.png}
+        \includegraphics[width=\linewidth]{figures/Fig7.png}
     \end{center}

### 2020-09-10 CoverLetter.tex
Update on Overleaf.

\documentclass[11pt,a4paper,sans]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')

\usepackage[document]{ragged2e}

% moderncv themes
\moderncvstyle{banking}                            % style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
\moderncvcolor{red}                                % color options 'blue' (default), 'orange', 'green', 'red', 'purple', 'grey' and 'black'
\renewcommand{\familydefault}{\rmdefault}         % to set the default font; use '\sfdefault' for the default sans serif font, '\rmdefault' for the default roman one, or any tex font name
%\nopagenumbers{}                                  % uncomment to suppress automatic page numbering for CVs longer than one page

% character encoding
\usepackage[utf8]{inputenc}                       % if you are not using xelatex ou lualatex, replace by the encoding you are using
%\usepackage{CJKutf8}                              % if you need to use CJK to typeset your resume in Chinese, Japanese or Korean

% adjust the page margins
\usepackage[scale=0.75]{geometry}

\firstname{}
\lastname{}
\begin{document}
%-----       letter       ---------------------------------------------------------

% recipient data
\recipient{Editors PLoS ONE}{}
\date{September 11th, 2020}
\opening{Dear Editors,}
\closing{Yours faithfully,\\
Juste Raimbault\\
University College London
}
         % use an optional argument to use a string other than "Enclosure", or redefine \enclname
\makelettertitle

\justify
We are pleased, with my co-authors Natalia Zdanowska (University College London) and Elsa Arcaute (University College London), to submit an original research article entitled ``Modeling growth of urban firm networks'' for consideration for publication in PLoS ONE. This article introduces a generative model for urban firm networks, to understand underlying processes and with possible application to policies. It does not contain previously published work. Preliminary results were presented as a satellite abstract at the Conference on Complex Systems 2019 (September 2019) and will be presented as a main track abstract at the NetScience 2020 conference (September 2020). We did not have any previous interaction with PLOS regarding this manuscript.

This manuscript has not been published and is not under consideration for publication elsewhere. A preprint was deposited on arxiv simultaneously to the submission.

We have no conflicts of interest to disclose. We do not have any opposed reviewer. This paper is submitted to the PloS ONE collection ``Cities as Complex Systems'' edited by M. Gonzalez (University of California, Berkeley) and D. Rybski (Potsdam Institute for Climate Impact Research).

Thank you for your consideration.
\justify

\makeletterclosing

\end{document}

%% end of file `template.tex'.

### 2020-09-10 main.tex
Update on Overleaf.

@@
 
-% Title must be 250 characters or less.
 \begin{flushleft}
@@
 }
-% alternative titles?
-% Old title Generative Urban Network Model
 \newline
@@
 Understanding the drivers behind the growth of such networks -- in particular urban firm networks --, is essential for the economic resilience of urban systems.
-We introduce in this paper a generative network model for firm networks at urban area level including several complementary processes: the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distance.
+We introduce in this paper a generative network model for firm networks at the urban area level including several complementary processes: the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distance.
 An empirical network analysis on European firm ownership data confirms the relevance of each of these factors.
@@
 
-%is it something you want to add?
-%\cite{derudder2018central}
-%\cite{storme2019introducing} cities firms interactions
-
 Networks constitute the new social morphology of our societies, since the intensification of globalisation of the economy in the late 20th century \cite{castells2000networksociety}.
 A world-city system has since then emerged \cite{taylor2001specification}, characterized by highly interconnected urban centres at the global level \cite{sassen1991}.
-These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity implying spatial and non-spatial patterns \cite{martinus2018global}.
-Interactions of economic nature between transnational firms, are together with international trade patterns among the most representative of the current economic trends \cite{taylor2001specification}.
+These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity inducing spatial and non-spatial patterns \cite{martinus2018global}.
+Interactions of economic nature between transnational firms \cite{derudder2018central}, are together with international trade patterns among the most representative of the current economic trends \cite{taylor2001specification}.
 Understanding the drivers of growth and geographical properties of such an urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}.
-Straightening these city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}, by developing strategic industries for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
-The functioning of such a city-network and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.
+In addition, straightening these city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}, by developing strategic industries for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
+Exploring city/firm interactions \cite{storme2019introducing} and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.
 
-The use of network models to understand processes underlying the emergence and dynamics of such networks has already been tackled in the literature, but still not extensively at city level.
-Generative urban network models combining geographical factors with network topological properties exist at the regional scale {\cite{dai2016simulating}}.
-%Air transportation networks can also be understood as urban networks.
-\cite{zanin2013modelling} reviews several models for such networks.
+The use of network models to understand processes underlying the emergence and dynamics of such networks has already been tackled in the literature, but still not extensively and never explicitly at city level \cite{taylor2001specification}.
+In fact generative urban network models combining geographical factors with network topological properties exist at the regional scale {\cite{dai2016simulating}}.
 Trade networks have been extensively studied from the complex networks perspective, but these are generally driven at the country-level \cite{garlaschelli2005structure} because of a lack of city-level data.
@@
 
-More precisely, our contribution consists in the following: (i) we propose an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database \cite{AMADEUS2018}; (ii) we introduce a generative network model to simulate the growth of such linkages at urban areas scale, which combines multiple factors influencing link formation, as economic intensity of origin and destination of urban areas, industrial sector proximity, the strength of previous links, and geographical and socio-cultural distance; (iii) the model allows us to compare the effect of different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; and (iv) we calibrate the model on the empirical network.
+More precisely, our contribution consists in the following: (i) we propose an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database \cite{AMADEUS2018}; (ii) we introduce a generative network model to simulate the growth of such linkages at urban areas scale, which combines multiple factors influencing link formation, such as the economic intensity of the urban areas corresponding to the origin and destination nodes, the industrial sector proximity, the strength of previous links, and the geographical and socio-cultural distance; (iii) the model allows us to compare the effect of different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; and (iv) we calibrate the model on the empirical network.
 
-The rest of this paper is organized as follows: we first describe the generative model for urban networks and proceed to an empirical analysis of the network used as a case study.
-We then do a sensitivity analysis and exploration of the model on synthetic data.
+The rest of this paper is organized as follows: we firstly describe the generative model for urban networks and proceed to an empirical analysis of the network used as a case study.
+We then proceed to a sensitivity analysis and exploration of the model on synthetic data.
 The model is calibrated on real European firm ownership data.
@@
 Geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}.
-Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalized to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (confirmed relevant by the fixed effect regression described above); (iii) city economic size, which has furthermore scaling laws properties; (iv) industrial proximity in terms of firm sector composition; (v) previous linkages (influence of already existing links in the past).
-This last point is crucial to capture accumulated effects and include path-dependency, which fits well the use of a simulation model instead of a more simple statistical model.
+Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalized to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (confirmed relevant by the fixed effect regression described below); (iii) city economic size, which presents agglomeration effects; (iv) industrial proximity in terms of firm sector composition; and (v) previous linkages (influence of already existing links in the past).
+This last point is crucial to capture accumulated effects and include path-dependencies, justifying further the use of a simulation model instead of a more simple statistical model.
 
@@
 
-Let $1 \leq i \leq N$ cities defined at time $t$ by their economic structure such that $E_i(t)$ is the total economic volume (GDP) for city $i$ and $S_{ik}(t)$ is a matrix giving economic volume proportions within each economic sector $k$, assuming $K$ economic sectors.
+Let $1 \leq i \leq N$ cities be defined at time $t$ by their economic structure such that $E_i(t)$ is the total economic volume (GDP) for city $i$ and $S_{ik}(t)$ is a matrix giving economic volume proportions within each economic sector $k$, assuming there are $K$ economic sectors.
 Formally, the model creates links between cities which are characterized by their economic size $E_i$ (GDP) and economic structure $S_{ik}$ in terms of activity sectors (probability distribution of firms within $K$ sectors).
@@
 \end{equation}
-
-where $E  =  \sum_k E_k$, $W  = \sum_{i,j} w_{ij}$ weights of previous links, $s(S_{ik},S_{jk})$ is a similarity function between activity sectors $S_{ik}$ and $S_{jk}$ taken as a cosine similarity, $d_{ij}$ euclidian distance, and $g_{ij}$ a socio-cultural distance.
+where $E  =  \sum_k E_k$, $W  = \sum_{i,j} w_{ij}$ weights of previous links, $s(S_{ik},S_{jk})$ is a similarity function between activity sectors $S_{ik}$ and $S_{jk}$ taken as a cosine similarity, $d_{ij}$ is the Euclidian distance, and $c_{ij}$ is a socio-cultural distance.
 In the case of an already existing link between two areas, the weight of the latter is incremented by a unit volume $w_0$.
@@
 
-Note that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalise the exponents to 1 (which is different than renormalising the probability), to ensure to include convex functions.
+Note that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalise the exponents to 1
+%(which is different than renormalising the probability), 
+to ensure to include convex functions.
 The ``geopolitical/sociocultural'' distance remains abstract and should be parametrised (see setup below) and estimated in the data-driven setting.
-We assume no evolution of economic size in this simple version of the model, which means that adjusted parameter are valid on a restricted time period.
+In this simple version of the model we assume an absence of evolution of economic size, which means that the adjusted parameters are valid on a restricted time period.
 
@@
 
-Besides model dynamics, its initialisation (or parametrisation on real data) regarding the spatial distribution, sizes, countries and industrial sector compositions, is also important to understand model behavior.
-In addition  \cite{raimbault2019space} showed that the initial spatial configuration can have significant impacts on model outcomes.
+%Besides model dynamics, its initialisation (or parametrisation on real data) regarding the spatial distribution, sizes and industrial sector compositions, it remains important to understand the model behavior.
+When setting the model, it is important to consider that different initial conditions (or parametrisation on real data), given by the spatial distribution, the sizes and industrial sector compositions, are going to affect the behaviour of the model.
+One of the authors \cite{raimbault2019space} has already shown that the initial spatial configuration can have significant impacts on model outcomes.
 We therefore consider two versions: one with a synthetic system of cities which can be stochastically repeated, and one other based on real data.
@@
 
-The real setup is done using the empirical network analyzed above.
+The real setup is done using the empirical network of ownership links.
 Urban areas are distributed according to their positions in the actual geographical space, and economic size and sector composition attributes are used.
@@
 
-Initial network is empty, and the real links are taken as targets for the generated network.
+Initial the network is empty, and the real links are taken as targets for the generated network.
 We constrain the total volume of flows by setting $w_0 = \sum_{ij} w_{ij}^{(obs)} / t_{f}$ when the number of time steps $t_f$ is given as a parameter.
@@
 The model parameters are the Cobb-Douglas weights $\gamma_O,\gamma_D,\gamma_W,\gamma_S$ and distance decay parameters $d_0,c_0$.
-As the model also includes path-dependency by integrating the influence of previous links, it can not be reduced to a simple statistical model.
+As the model also includes path-dependencies by integrating the influence of previous links, it cannot be reduced to a simple statistical model.
 Finally, the role of space introduces even further complexity, yielding a generative model, which has to be simulated.
@@
 We consider two main macroscopic indicators to quantify the generated network.
-Firstly geographical structures are captured by internationalisation (modularity of countries in the network).
-The situation of a minimised modularity corresponds to dense connections between cities of different countries of Europe and sparse connections between cities of the same European country.
-In practice, modularity is computed by
+Firstly, geographical structures are captured by internationalisation (modularity of countries in the network).
+The situation of a minimised modularity corresponds to dense connections between cities of different countries in Europe and sparse connections between cities of the same European country.
+In practice, modularity is computed as follows
 \begin{equation}
@@
 \end{equation}
-where $d_i$ is the weighted degree of cities and $c_i$ the countries.
+where $d_i$ is the weighted degree of city $i$ in country $c_i$.
 % add some reference values - some kind of null model
 
-Secondly we consider the metropolisation indicator as the correlation between the weighted degree and the GDP of the city.
+Secondly, we consider the metropolisation indicator as the correlation between the weighted degree and the GDP of the city.
 With this indicator we capture the effect of large cities or metropolises of regional importance attracting the biggest amount of links and involved in high interactions with other cities.
@@
 
-%I think we need to keep this paragraph but shorten it, because to my understanding network average community size is tested in Figure 5, but not the Louvain modularity for ex.
-%We also consider several network indicators as Louvain modularity, sizes of maximal modularity communities, degree and link weight distributions (described by their average, hierarchy, and entropy), and correlations between degree and city size, and between link weight and distance.
-
-We take a geographical approach to study the behavior of the generative model.
+%We take a geographical approach to study the behavior of the generative model.
 Other related approaches include indicators of economic complexity, estimating for example the diversity of products produced by countries and how they depend on each other \cite{hidalgo2007product}.
 These are less relevant in our case as the sector structure of cities does not evolve in the model.
+In this work, we do not consider the relationship between countries given their economic complexity, such as the framework developed by  \cite{hidalgo2007product}, where the diversity of products produced plays an essential role in determining co-dependencies and resilience.
+What we investigate instead, is the role of linkages given by geography and history, and hence, the sector structure of cities does not evolve within the model.
 
@@
 Companies from \emph {AMADEUS} are georeferenced using the Geonames database (using postcode or address depending on the information available).
-We then join this data with the Global Human Settlement dataset GHS-FUA for Functional Urban Areas \cite{Florczyk2019ghs} -- the most relevant objects for applying our model.
+We then join this data with the Global Human Settlement dataset GHS-FUA for Functional Urban Areas \cite{Florczyk2019ghs} -- which delineate the commuting areas for urban centres.
 Starting from a firm-level dataset of 1,562,862 nodes and 1,866,936 links, we obtain a directed network of 573 urban areas and 9323 ownership links.
 The weight of links is obtained by computing the owned share of turnover at destination, i.e.
-$w_{ij} = \sum_{k \in i,l \in j} p_{kl} \cdot T_l$ where $p_{kl}$ of the ownership share of company $l$ by company $k$ and $T_l$ the turnover of company $l$.
-Node attributes include an economic size of urban areas, that we compute as the sum of turnovers of companies within the area.
+$w_{ij} = \sum_{k \in i,l \in j} p_{kl} \cdot T_l$ where $p_{kl}$ is the ownership share of company $l$ by company $k$ and $T_l$ is the turnover of company $l$.
+Node attributes include the economic size of urban areas, that we compute as the sum of turnovers of companies within the area.
 This quantity is highly correlated with the GDP of the area-level included in the GHS database (Spearman correlation $\rho = 0.71$) and also with population ($\rho = 0.60$).
 We then expect to capture size effects while keeping the consistence of a single main data source.
-An other attribute of the nodes is the industrial sector composition, expressed as a proportion of turnover in the area associated to a given industrial sector.
+Another attribute of the nodes is the industrial sector composition, expressed as a proportion of turnover in the area associated to a given industrial sector.
 Detailed sectors of firms are available up to the four digits of the NACE classification in the raw data \cite{EUROSTAT2008}.
@@
 We fit power law and log-normal distributions, including minimal value cutoff following \cite{clauset2009power}, using the \texttt{powerlaw} R package \cite{powerlawpackage}.
-For both, log-normal distributions appear to be a better fit ($\mu=18.8$, $\sigma=2.3$ for weighted degree; $\mu=13.8$, $\sigma=2.8$ for edge weight), and include a much broader range of the empirical distributions with lower estimated cut-offs.
+For both, log-normal distributions appear to be a better fit ($\mu=18.8$, $\sigma=2.3$ for weighted degree; $\mu=13.8$, $\sigma=2.8$ for edge weight), and include a much broader range of empirical distributions with lower estimated cut-offs.
 These heavy tail properties suggest the relevance of a generative model including self-reinforcing processes, which are known to produce such distributions.
@@
 
-Regarding the community structure of the network, different modularity maximisation algorithms (run on the undirected corresponding network) give the same undirected modularity of 0.38, and a directed modularity \cite{nicosia2009extending} of 0.34 for the greedy algorithm of \cite{clauset2004finding} and 0.36 for the Louvain algorithm \cite{blondel2008fast}, corresponding to 36 communities (resp.
+When focusing on the community structure of the network, different modularity maximisation algorithms (ran on the undirected corresponding network) result in the same undirected modularity of 0.38, a directed modularity \cite{nicosia2009extending} of 0.34 for the greedy algorithm of \cite{clauset2004finding} and of 0.36 for the Louvain algorithm \cite{blondel2008fast}, corresponding to 36 communities (resp.
 15).
-In comparison, with 29 countries, the classification of urban areas by country has a modularity of 0.32.
-This means that the structure of flows present certain regional patterns.
+In comparison, when taking into account 29 countries, the classification of urban areas by country has a modularity of 0.32.
+This indicates that the structure of flows present certain regional patterns.
 Indeed, a null model assigning random labels with 30 communities, bootstrap 1000 times, yields an average modularity of $0.049 \pm 0.002$, confirming the statistical significance of these communities.
@@
 Cities in France, Belgium and Luxembourg belong to the same cluster as they share French as a common business language, but also cross-border tax arrangements creating easier economic interactions \cite{DecovilleDurand2019}.
-Cities of the United Kingdom and Ireland are in the same cluster for the same socio-cultural proximity matters, so as Sweden, Finland, Estonia, Latvia and Lithuania.
-The cluster composed of Dutch and Polish cities is justified by the fact that firms from the Netherlands are among the first foreign owners of polish firms since 2010, ranked even before the German ones \cite{2019arXiv191014652Z}.
-The cluster including cities in Germany, Czech Republic, Austria, Slovakia, Hungary and partly in Romania, find its origin in the dependence of the Central and Eastern European economy on German and Austrian large companies mostly from the automotive sector (ie.
+The same socio-cultural proximity effect is seen in cities of the United Kingdom and Ireland, and those of Sweden, Finland, Estonia, Latvia and Lithuania.
+The cluster composed of Dutch and Polish cities is justified by the fact that firms from the Netherlands are among the first foreign owners of polish firms since 2010 \cite{2019arXiv191014652Z}.
+The cluster including cities in Germany, Czech Republic, Austria, Slovakia, Hungary and partly in Romania, find its origin in the dependence of the Central and Eastern European economies on German and Austrian large companies mostly from the automotive sector (i.e.
 Audi, Volkswagen).
-In addition all cities of the later cluster have in common to be at the intersection of a very important freight transport route in Europe from Germany to Turkey \cite{Zdanowska2017}.
+In addition, all cities of the later cluster have in common to be at the intersection of a very important freight transport route in Europe from Germany to Turkey \cite{Zdanowska2017}.
 %%%%%%%%%%%%%%
@@
     \end{center}
-    \caption{{\bf Map of network communities obtained by modularity maximisation.} Area of circles gives the population with the Functional Urban Area, and color the community it belongs to.
+    \caption{{\bf Map of network communities obtained by modularity maximisation.} The area of circles gives the population within the Functional Urban Area, and color the community it belongs to.
 Urban Areas in communities of size smaller than 3 were not colored for readability.}
@@
 
-In order to understand which are the most determining geo-economic factors influencing creation of links between cities, we run a statistical analysis as applied in the case of unconstrained spatial interaction models \cite{wilson1975some}.
-We test five different statistical models, including progressively the processes to be included in the network generative model.
-The most general statistical model is:
+In order to understand the most determining geo-economic factors influencing the creation of links between cities, we run a statistical analysis as applied in the case of unconstrained spatial interaction models \cite{wilson1975some}.
+We test five different statistical models, which progressively include the processes for the generative network model.
+The most general statistical one is:
 
@@
 \end{equation}
-
-where $\alpha_{c_i,c_j}$ is a fixed effect term for the coupling of countries $c_i,c_j$.
-A multiplicative form is applied by taking logarithms as usually done for spatial interaction models.
+where $\alpha_{c_i,c_j}$ is a fixed effect term for the coupling of countries $c_i,c_j$, and $T_i$, $T_j$ correspond to the turnovers of cities $i$ and $j$ respectively. 
+%A multiplicative form is applied by taking logarithms as usually done for spatial interaction models.
 Furthermore, origin and/or destination constrained models are not relevant in our case, as links include information about the turnover and a share of ownership at the same time.
+This multiplicative form linearly transformed through logarithms is common practice for spatial interaction models.
+Note that no constraints are considered, since links include information about the turnover and a share of ownership at the same time.
 Estimation results are given in Table~\ref{tab:reg}.
@@
 Across all models, consistent qualitative stylized facts regarding link creation between cities are revealed: (i) distance has a negative influence; (ii) economic sizes have a positive effect, with the size of the origin city being more important (consistent in the ownership links context, as strongly driven by decision at the headquarter/owner level); (iii) industrial proximity has a positive influence, i.e.
-similarity is more important than complementarity for this dimension; (iv) an large part of country fixed-effects are statistically significant; and (v) the order of magnitude of the influence of each factor is roughly the same, which means that no process totally dominates the others.
+similarity is more important than complementarity for this dimension; (iv) a large part of country fixed-effects are statistically significant; and (v) the order of magnitude of the influence of each factor is roughly the same, which means that no process totally dominates the others.
 This confirms the relevance of including all these different dimensions in the generative network model, since they have a comparable explanatory role in statistical models.
@@
 We compare model performance using $R^2$, Mean Square Error, and Akaike Information Criterion.
-For the Poisson model (model 5), AIC is not given as it is not comparable, and the pseudo-$R^2$ is given either.
+For the Poisson model (model 5), AIC is not given as it is not comparable.
 \label{tab:reg}}
@@
 Numerical experiments are done by integrating the model into the OpenMOLE software for model exploration \cite{reuillon2013openmole}, which provides sensitivity analysis and calibration methods and a transparent access to high-performance computing infrastructures.
-Model source code, aggregated network data (raw initial firm network can not be made available for data ownership reasons), and results are available on the open Git repository of the project at \url{https://github.com/JusteRaimbault/ABMCitiesFirms}.
+The model source code, the aggregated network data (the raw initial firm network cannot be made available for data ownership reasons), and the results are available on the open Git repository of the project at \url{https://github.com/JusteRaimbault/ABMCitiesFirms}.
 Large simulation result files are available on the dataverse repository at \url{https://doi.org/10.7910/DVN/UPX23S}.
@@
     \end{center}
-    \caption{Example of simulated networks at $t=1500$, for a synthetic system of cities (Top row) and the European urban network (Bottom row).
-, with a high (resp.
+    \caption{Example of simulated networks at $t=1500$, for a synthetic system of cities (Top row) and the European urban network (Bottom row), with a high (resp.
 low) gravity decay parameter for the left column (resp.
@@
 We use thus 20 repetitions in following larger experiments (in the case of a normal distribution, a 95\% confidence interval on the average is of size $2\sigma \cdot 1.96 / \sqrt{n}$, what gives $n=16$ repetitions for a CI of width of the standard deviation).
-Behavior of some indicators are shown in Fig.~\ref{fig:onefactor}.
-For example, internationalisation and metropolisation both show a transition as a function of $\gamma_G$, from a local to a global regime.
+Behavior of some indicators are shown in Fig.~\ref{fig:fig4}.
+For example, internationalisation and metropolisation both show a transition as a function of d$_o$, from a local to a global regime.
 The sector proximity parameter $\gamma_S$ influences the internationalization linearly, but induces a maximum for the correlation between size and degree, which can be interpreted as a setting where size and the total volume in a city are the most correlated.
@@
 
-The model behaviour is then studied with a grid sampling for parameters and 20 repetitions (the computations run on the European Grid Infrastructure, and are equivalent to 2.5 years of CPU time).
+The model behaviour is then studied with a grid sampling for parameters and 20 repetitions (the computations are run on the European Grid Infrastructure, and are equivalent to 2.5 years of CPU time).
 Distance decay and sector  parameters are varied with 10 steps, other parameters with three.
-The influence of gravity decay parameters is confirmed when plotting the internationalization index against the distance, which exhibits an exponential decay as shown in Fig.~\ref{fig:internat}.
-It interacts with the role of sectors and the size of the origin, with a more significant effect of sector proximity when size has the largest influence (third panel).
+The influence of gravity decay parameters is confirmed when plotting the internationalization index against the distance, which exhibits an exponential decay as shown in Fig.~\ref{fig:fig5}.
+It interacts with the role of sectors and the size of the origin, with a more significant effect of sector proximity, when size has the largest influence (third panel).
 
@@
 Other indicators exhibit non-trivial patterns.
-For example, when considering average community size of the final network as shown in Fig.~\ref{fig:comsize}, we obtain a maximal integration in terms of communities at an intermediate value of the gravity decay, what can be interpreted as the emergence of a regional regime.
+For example, when considering average community size of the final network as shown in Fig.~\ref{fig:fig6}, we obtain a maximal integration in terms of communities at an intermediate value of the gravity decay.
+This can be interpreted as the emergence of a regional regime.
 The size of communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.
 In particular, we observe a qualitative inversion of the role of $\gamma_S$ when introducing an effect of the origin (switching $\gamma_O$ from 0 to 1, first panel to second panel).
-The maximal community size disappears when $\gamma_O = 2$, implying a regime with small local communities where large cities mostly connect with there hinterland.
+The maximal community size disappears when $\gamma_O = 2$, implying a regime with small local communities where large cities mostly connect with their hinterland.
 These experiments reveal a complex interplay between processes and how the model produces diverse stylised facts.
@@
     \includegraphics[width=\textwidth]{figures/Fig6.png}
-	\caption{Maximal integration in terms of community size is achieved at an intermediate value of $d_G$: emergence of a regional regime; Maximal size depends on the role of sectors $\gamma_S$, in a decreasing way when origin size is deactivated, and increasing way when $\gamma_F=1$; This regime disappear when origin size influence is too large.\label{fig:fig6}}
+	\caption{Maximal integration in terms of community size is achieved at an intermediate value of d$_o$: emergence of a regional regime; maximal size depends on the role of sectors $\gamma_S$ in a decreasing way when the origin size is deactivated, and in an increasing way when $\gamma_o=1$; this regime disappears when the influence of the origin size becomes too large.\label{fig:fig6}}
 \end{figure}
@@
 We run a specific experiment to study the impact of urban hierarchy on firms dynamics.
-Rank-size law in cities is a central stylized fact to urban theories, despite the discussions on how to estimate the power-law exponent, its actual values \cite{cottineau2017metazipf} and how its depends on the definition of cities considered \cite{corral2020truncated}.
-In that particular context, understanding how this parameter can impact urban dynamics within artificial systems of cities in which it can be modified to extreme values not observed in real systems, should inform on underlying processes \cite{raimbault2019space}.
+Rank-size law in cities is a central stylized fact within urban theories, despite the discussions regarding the estimation of the power-law exponent, its actual values and its variation depending on the definition of cities considered \cite{cottineau2017metazipf,corral2020truncated}.
+In this particular context, understanding how this parameter can impact urban dynamics within artificial systems of cities -- in which it can be modified to extreme values non observable in real systems--, should bring evidence on underlying processes \cite{raimbault2019space}.
 We change the initial hierarchy $\alpha$ from 0.5 to 2.0 with a step of 0.1, and also vary the interaction range $d_0$.
@@
     \end{center}
-    \caption{\textbf{Impact of urban hierarchy.}\label{fig:fig7}}
+    \caption{\textbf{Impact of urban hierarchy.} (Left) Internationalisation \label{fig:fig7}}
 \end{figure}
@@
 The other objective would correspond to statistical models directly on variables: therefore, our model covers in a sense a set of intermediate models which make a compromise between the two objectives, and is therefore much more flexible.
-The values for $\varepsilon_L$ are for a large proportion of points lower than for statistical models, our model outperforming these in this predictive sense.
-As shown in Fig.~\ref{fig:fig7}, increasing values for $\gamma_O$ lead to better performance regarding $\varepsilon_L$: increasing the importance of the origin will provide a better fit on all order of magnitudes of link weights (while $\varepsilon_M$ favors mostly the largest links given the fat-tail distribution).
+The values for $\varepsilon_L$ are for a large proportion of points lower than for statistical models, our model outperforms these in this predictive sense.
+As shown in Fig.~\ref{fig:fig8}, increasing values for $\gamma_O$ lead to better performance regarding $\varepsilon_L$: increasing the importance of the origin will provide a better fit on all order of magnitudes of link weights (while $\varepsilon_M$ favors mostly the largest links given the fat-tail distribution).
 
@@
 Based on calibrated parameter values, we design scenarios to evaluate the impact of economic shocks.
-These can be the consequence of diverse factors, including socio-economic, geopolitical factors (example of Brexit), and natural disasters such as COVID-19.
+These can be the consequence of diverse factors, including socio-economic, geopolitical factors (example of Brexit), or other  unpredicted human disasters such as the COVID-19 pandemic.
 We simulate a full border closure by setting $\gamma_D = 0$, and intermediate scenarios.
@@
 
-This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMole platform and calibrated on real data on ownership linkages of firms for Europe from Amadeus dataset.
+This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMole platform and calibrated on real data on ownership linkages of firms for Europe from the Amadeus dataset.
 The simulation on a synthetic system of cities provides a broad knowledge on model behavior in its parameter space.
@@
 Other formulations of the model might be taken into consideration as other formulation of the combination of factors or multi-objective optimisation depending on sectors using Pareto fronts.
+
+% I believe than something on the results of the COVID/lockdown scenario should be commented in the discussion as well
 
@@
 We thank the European Grid Infrastructure and its supporting National Grid Initiatives (France-Grilles in particular) for providing the technical support and infrastructure.
+We also acknowledge the funding of the EPSRC grant EP/M023583/1.
 

### 2020-09-10 S2_Text.tex

@@
-\input{si.tex}
-\begin{document}
-\vspace*{0.35in}
-\justify
+%\input{si.tex}
+%\begin{document}
+%\vspace*{0.35in}
+%\justify
 
@@
 
-% \todo{more detail / descriptive stats of the matrix / function of the distance between countries}
+In the real model setup, socio-economic distance between countries $c_{ij}$ is constructed using fixed effects coefficients from the statistical model.
 
@@
 
-The rationale for the synthetic sector setup is the following:
+The procedure for the synthetic sector setup is the following:
 \begin{itemize}
     \item Sectors distributions follow log-normal densities with most mass in $\left[0;1\right]$
+    \item Large cities are more innovative and more diverse.
+Assuming that sectors are ordered by innovativity (the larger the sector index the larger the innovativity), this assumption is translated by taking a mode and variance of 0.5 for the largest city and a mode and variance of $1/K$ for the smallest, and a linear interpolation between the two.
+For each city, we define
+    \[
+    \log(e_i) = \frac{(\log(E_i) - \log(E_{imin}))}{(\log(E_{imax}) - \log(E_{imin}))} * (1/2 - 1/K) + 1/K
+    \]
+    \item Log-normal parameters for each city are then fixed by $mu_i - sigma_i^2 = log(e_i)$ and $-3 sigma_i^2 - 2 log(exp(sigma_i^2) - 1) = log(e_i)$
+    \item $\sigma_i^2$ is the unique positive root of $f(X)=0$ with $f(X) = -3X - 2 ln(exp(X) - 1) - ln(e_i)$
 \end{itemize}
@@
 
+The table~\ref{tab:saltelli} gives the full results for the Global Sensitivity Analysis, for all model indicators and free parameters. We give the first order indices and the total indices.
+
 %%%%%%%%%%%%%
-\begin{table}
+\begin{table}[h!]
 \begin{adjustwidth}{-2.25in}{0in}
@@
 
-\end{document}
+%\end{document}

### 2020-09-11 figures.sh
fig_9

@@ -30,3 +30,4 @@ montage figuresraw/internationalization-alpha_colorgravityDecay_errorBar.png fig
 cp figuresraw/pareto_colorgammaOrigin.png figures/Fig8.png
 
 # Fig 9
+montage figuresraw/internationalization_Time_colorgravityScaleFactor_facetcountryScaleFactor.png figuresraw/networkAvgCommunitySize_Time_colorgravityScaleFactor_facetcountryScaleFactor.png -tile 1x2 -geometry +0+"$VERTICALPADDING" -resize "$WIDTH"x figures/Fig9.png

### 2020-09-11 S2_Text.tex
Update on Overleaf.

@@
-%\input{si.tex}
-%\begin{document}
-%\vspace*{0.35in}
-%\justify
+\input{si.tex}
+
+\renewcommand{\thetable}{S2-\Alph{table}}
+
+\begin{document}
+\vspace*{0.35in}
+\justify
 
@@
 In the real model setup, socio-economic distance between countries $c_{ij}$ is constructed using fixed effects coefficients from the statistical model.
+In Table~1 in main text, we estimated several statistical models, including some with fixed effects by pairs of countries.
+The socio-cultural distance between two countries is then taken as the opposite of the fixed effect coefficient for this pair, for the full statistical model.
+Summary statistics for the 29x29 matrix (28 EU countries and Jersey which is in the database a distinct country from UK but has an important total turnover since many British companies are located there for fiscal advantage reasons) are shown below in Table~\ref{tab:fixedeff}.
+We observe that distances are rather localized but with large outliers, and an important proportion are not defined (339 out of 841), corresponding to couple of countries having no exchange at all in the dataset.
+
+\begin{table}[h!]
+\caption{\textbf{Summary statistics of the socio-cultural distance estimated with a fixed effects model.}\label{tab:fixedeff}}
+\begin{tabular}{ccccccc}
+   Min. & 1st Qu. &  Median &  Mean &  3rd Qu. &  Max. &  NA's \\
+ -2.327 &  1.440 &  2.451 &  2.538 &  3.447 & 11.797 &  339 
+\end{tabular}
+\end{table}
 
@@
 
-%\end{document}
+\end{document}

### 2020-09-11 S1_Text.tex
Update on Overleaf.
SuppMat

@@
 \input{si.tex}
+
+\renewcommand{\thetable}{S1-\Alph{table}}
+
 \begin{document}
@@
 
-When considering the UK sectors generating the highest turnover in 2018 mostly dependent on foreign capital they own (out-going links from the UK) these are extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods.
-In terms of in-going links to the UK, activities of head offices and holding companies, sale of cars and manufacture of motor vehicles are sectors mostly dependent on foreign ownership coming from abroad.
+When considering the UK sectors generating the highest turnover in 2018 mostly dependent on foreign capital they own (out-going links from the UK - see below Table~\ref{tab:uk_outgoing}) these are extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods.
+In terms of in-going links to the UK (see below Table~\ref{tab:uk_ingoing}), activities of head offices and holding companies, sale of cars and manufacture of motor vehicles are sectors mostly dependent on foreign ownership coming from abroad.
 
@@
 \newpage
-When expressed in total number of links, activities mostly dependent on out-going links from the UK are activities of holding companies, of head offices and fund management activities.
+When expressed in total number of links (Table~\ref{tab:uk_outgoinglinks} and Table~\ref{tab:uk_ingoinglinks}), activities mostly dependent on out-going links from the UK are activities of holding companies, of head offices and fund management activities.
 In terms of in-going links these are business and consultancy activities and renting and operating of own and leased real estate.

### 2020-09-11 S2_Text.tex
SuppMat

@@
     \]
-    \item Log-normal parameters for each city are then fixed by $mu_i - sigma_i^2 = log(e_i)$ and $-3 sigma_i^2 - 2 log(exp(sigma_i^2) - 1) = log(e_i)$
-    \item $\sigma_i^2$ is the unique positive root of $f(X)=0$ with $f(X) = -3X - 2 ln(exp(X) - 1) - ln(e_i)$
+    \item Log-normal parameters $(\mu_i,\sigma_i)$ for each city are then fixed by $\mu_i - \sigma_i^2 = \log(e_i)$ and $-3 \sigma_i^2 - 2 \log(\exp(\sigma_i^2) - 1) = log(e_i)$
+    \item $\sigma_i^2$ is the unique positive root of $f(X)=0$ with $f(X) = -3X - 2 \log(\exp(X) - 1) - \log(e_i)$ and $\mu_i = \log(e_i) + \sigma_i^2$.
 \end{itemize}

### 2020-09-11 main.tex
Update on Overleaf.

@@
 
-In synthetic experiments, the network is generated from an empty initial network until reaching $t_f=1500$, with a unit link volume of $w_0 = 1$.
+In synthetic experiments, the network is generated from an initial complete network with dummy links of weight 1 until reaching $t_f=1500$, with a unit link volume of $w_0 = 1$.
 
@@
 
-Initial the network is empty, and the real links are taken as targets for the generated network.
+The initial network is the same than in the synthetic setup, and the real links are taken as targets for the generated network.
 We constrain the total volume of flows by setting $w_0 = \sum_{ij} w_{ij}^{(obs)} / t_{f}$ when the number of time steps $t_f$ is given as a parameter.
@@
 where $d_i$ is the weighted degree of city $i$ in country $c_i$.
-% add some reference values - some kind of null model
 
@@
 
-%We take a geographical approach to study the behavior of the generative model.
-Other related approaches include indicators of economic complexity, estimating for example the diversity of products produced by countries and how they depend on each other \cite{hidalgo2007product}.
-These are less relevant in our case as the sector structure of cities does not evolve in the model.
 In this work, we do not consider the relationship between countries given their economic complexity, such as the framework developed by  \cite{hidalgo2007product}, where the diversity of products produced plays an essential role in determining co-dependencies and resilience.
@@
 where $\alpha_{c_i,c_j}$ is a fixed effect term for the coupling of countries $c_i,c_j$, and $T_i$, $T_j$ correspond to the turnovers of cities $i$ and $j$ respectively.
-%A multiplicative form is applied by taking logarithms as usually done for spatial interaction models.
-Furthermore, origin and/or destination constrained models are not relevant in our case, as links include information about the turnover and a share of ownership at the same time.
+
 This multiplicative form linearly transformed through logarithms is common practice for spatial interaction models.
@@
 We change the initial hierarchy $\alpha$ from 0.5 to 2.0 with a step of 0.1, and also vary the interaction range $d_0$.
+The behavior of internationalisation and metropolisation indicators are shown in Fig.~\ref{fig:fig7}.
+As expected, the hierarchy of the urban system has an important impact on model outcomes.
+We find that more hierarchical urban systems tend to produce more international networks (internationalisation indicator is a modularity which is minimized when networks span more accross different countries), what is consistent with the existence of globalized metropolis \cite{sassen1991global}.
+This effect is mitigated by lower interaction ranges.
+Regarding metropolisation, we observe a maximum values as a function of initial hierarchy consistent across different interaction ranges $d_0$.
+This means that a too high hierarchy is detrimental to the global metropolisation when considering all cities, what may be due to the fact that in such a case very large cities are capturing all flows, leaving nothing to intermediate and medium-sized cities.
+Interestingly, the optimal value $\alpha \simeq 1.3$ is not far from observed values (which usually range between 0.9 and 1.1 but can take larger values such as 1.5 when extending the urban system \cite{Raimbault_2020}), what may imply that this optimality property is endogenous to the emergence of urban systems and linked to their intrinsic hierarchy.
 
@@
     \end{center}
-    \caption{\textbf{Impact of urban hierarchy.} (Left) Internationalisation \label{fig:fig7}}
+    \caption{\textbf{Impact of urban hierarchy.} (Left) Internationalisation as a function of the initial synthetic hierarchy $\alpha$, for different values of interaction range $d_0$ (color); (Right) Metropolisation as a function of $\alpha$, for different values of interaction range $d_0$ (color).\label{fig:fig7}}
 \end{figure}
@@
 
-Based on calibrated parameter values, we design scenarios to evaluate the impact of economic shocks.
+We finally design scenarios to evaluate the impact of economic shocks.
 These can be the consequence of diverse factors, including socio-economic, geopolitical factors (example of Brexit), or other unpredicted human disasters such as the COVID-19 pandemic.
-We simulate a full border closure by setting $\gamma_D = 0$, and intermediate scenarios.
-Setting also $\gamma_G$ to 0 allows simulating inter-country lock-downs.
-% run simulations
-% note that time scales are in fact smaller - or assume that crisis keeping border closed is rather long
+We simulate border closures by rescaling $c_0$, and intra-country lock-downs by rescaling $d_0$ during the simulation, with other parameters being fixed to their default values.
+More precisely, with $t_f = 5000$, the model is run for $t_f / 2$ steps when parameters are changed as $\kappa_G \cdot d_0$ and $\kappa_C \cdot c_0$, and then run for the remaining time steps.
+We study the temporal trajectory of indicators to see the impact of the shock in time.
+The experiment is run for $(\kappa_G,\kappa_C)$ ranging between 0.2 (strong restrictions) to 1 (reference trajectory), with 100 model repetitions.
 
@@
 	\begin{center}
-    %\includegraphics[width=\linewidth]{figures/Fig9.png}
+    \includegraphics[width=\linewidth]{figures/Fig9.png}
     \end{center}
-	\caption{\textbf{Application of the calibrated model to an economic shock scenario.}\label{fig:fig9}}
+	\caption{\textbf{Application of the calibrated model to an economic shock scenario.} All plots show the trajectory of indicators (internationalisation for the top row and average community size for the bottom row) in time, for different values of interaction decay rescaling $\kappa_G$ (color) and socio-economic distance rescaling $\kappa_C$ (panels).
+The vertical dotted red line shows the moment the economic shock is introduced and $d_0$ (respectively $c_0$) is set to $\kappa_G \cdot d_0$ (resp.
+$\kappa_C \cdot c_0$).\label{fig:fig9}}
 \end{figure}

### 2020-09-11 main.tex
Update on Overleaf.

@@
 
-% Plos one requirements https://journals.plos.org/plosone/s/submission-guidelines
-% Deadline for the Cities as Complex Systems Call for Papers: 4th September https://collections.plos.org/s/cities
-% Manuscripts can be any length. There are no restrictions on word count, number of figures, or amount of supporting information. 
-% PLOS uses “Vancouver” references style
-% We need to reorganise the paper to have more or less the following sections (which can be reordered or renamed):
-% 1. Title page: List title, authors, and affiliations as first page of manuscript 2. Abstract 3. Introduction 4. Materials and Methods 5. Results 6. Discussion 7. Conclusions (optional)
-
 \section*{Abstract}
@@
 We then simulate network growth for synthetic systems of cities, unveiling stylized facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range.
-We finally calibrate the model on the European network, outperforming statistical models and showing a strong role of path-dependency.
-Potential applications of the model include the study of mitigation policies to deal with exogenous shocks such as economic crisis or potential lockdowns of countries.
+We calibrate the model on the European network, outperforming statistical models and showing a strong role of path-dependency.
+Potential applications of the model include the study of mitigation policies to deal with exogenous shocks such as economic crisis or potential lockdowns of countries, which we illustrate with an application on stylized scenarios.
 
@@
 
-Note that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalise the exponents to 1
-%(which is different than renormalising the probability), 
-to ensure to include convex functions.
+Note that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalise the exponents to 1 to ensure to include convex functions.
 The ``geopolitical/sociocultural'' distance remains abstract and should be parametrised (see setup below) and estimated in the data-driven setting.
@@
 
-%Besides model dynamics, its initialisation (or parametrisation on real data) regarding the spatial distribution, sizes and industrial sector compositions, it remains important to understand the model behavior.
 When setting the model, it is important to consider that different initial conditions (or parametrisation on real data), given by the spatial distribution, the sizes and industrial sector compositions, are going to affect the behaviour of the model.
@@
 Parameters of the log-normal law are obtained by solving the corresponding two equations system for each city, and each distribution is discretised and renormalised to be a vector of probabilities.
-See \ref{S2_Text} for all details of this synthetic setup.
+See S2 Text for all details of this synthetic setup.
 
@@
 The socio-economic distance is constructed using information from the statistical modeling, by considering fixed-effect coefficients estimated for each pair of countries.
-See \ref{S2_Text} for more details on the construction of the socio-economic distance matrix.
+See S2 Text for more details on the construction of the socio-economic distance matrix.
 
@@
 \begin{equation}
-    I = \frac{1}{2 \left|E\right|} \cdot \sum_{i,j} \left(w_{ij} - \frac{d_i d_j}{2 \left|E\right|}\right) \mathbb{1}_{c_i = c_j}
+    I = \frac{1}{2 \left|E\right|} \cdot \sum_{i,j} \left(w_{ij} - \frac{d_i d_j}{2 \left|E\right|}\right) \mathbbm{1}_{c_i = c_j}
 \end{equation}
@@
     \end{center}
-    \caption{{\bf Map of network communities obtained by modularity maximisation.} The area of circles gives the population within the Functional Urban Area, and color the community it belongs to.
+    \caption{{\bf Map of network communities obtained by modularity maximisation.} The area of circles gives the turnover within the Functional Urban Area, and color the community it belongs to.
 Urban Areas in communities of size smaller than 3 were not colored for readability.}
@@
 These can be the consequence of diverse factors, including socio-economic, geopolitical factors (example of Brexit), or other unpredicted human disasters such as the COVID-19 pandemic.
-We simulate border closures by rescaling $c_0$, and intra-country lock-downs by rescaling $d_0$ during the simulation, with other parameters being fixed to their default values.
+We simulate border closures by rescaling $c_0$, and intra-country lock-downs by rescaling $d_0$ during the simulation, with other parameters being fixed to their default values and the model setup in the real configuration.
 More precisely, with $t_f = 5000$, the model is run for $t_f / 2$ steps when parameters are changed as $\kappa_G \cdot d_0$ and $\kappa_C \cdot c_0$, and then run for the remaining time steps.
@@
 The experiment is run for $(\kappa_G,\kappa_C)$ ranging between 0.2 (strong restrictions) to 1 (reference trajectory), with 100 model repetitions.
+
+Results are shown in Fig.~\ref{fig:fig9}.
+We find an important impact of the shock on internationalisation, but less on the average community size.
+The impact of changing $d_0$ is more important than $c_0$, but a combination of both yield the stronger effect.
+For minimal values of $\kappa_G$ and $\kappa_C$, internationalisation is more than doubled, confirming a reconfiguration of urban networks within countries.
+We confirm the statistical significance of effects observed in Fig.~\ref{fig:fig9} by comparing distributions with Kolmogorov-Smirnov (KS) tests.
+More precisely, we proceed for each indicator and each value of $(\kappa_G,\kappa_C)$ to a two-sided KS test between the statistical distribution of the indicator over repetitions at final time with the distribution in the reference configuration.
+For internationalisation, p-values of the test are all smaller than 0.002 for $\kappa_D < 0.6$ but all larger than $0.15$ for $\kappa_D \geq 0.6$, what means that the shock has a statistically significant impact only when interaction range is more than halved.
+For community size, significance (p-value smaller than 0.01) is reached only for the smallest value $\kappa_D = 0.2$.
+Thus, the trajectory of the urban network is rather resilient to moderate shocks.
+This application shows how the model could be applied to practical policy issues.
+
+% computePvalMat(res$internationalizationTS.98)
+%             [,1]         [,2]        [,3]         [,4]        [,5]
+%[1,] 4.884981e-15 4.440892e-16 0.000000000 0.0000000000 0.000000000
+%[2,] 1.364656e-03 9.570235e-06 0.001364656 0.0002468196 0.002318458
+%[3,] 8.127483e-01 6.993742e-01 0.999633292 0.4675585921 0.154538055
+%[4,] 5.806178e-01 8.127483e-01 0.366726444 0.6993741999 0.906206390
+%[5,] 8.127483e-01 2.809600e-01 0.210551633 0.4675585921 1.000000000
+% computePvalMat(res$networkAvgCommunitySizeTS.98)
+%             [,1]         [,2]         [,3]         [,4]         [,5]
+%[1,] 4.365916e-08 4.705150e-06 3.728748e-05 2.267454e-06 1.071070e-06
+%[2,] 5.410262e-02 6.993742e-01 6.302223e-03 1.545381e-01 4.675586e-01
+%[3,] 9.670685e-01 1.545381e-01 9.996333e-01 9.996333e-01 4.675586e-01
+%[4,] 4.675586e-01 5.806178e-01 4.675586e-01 9.670685e-01 6.993742e-01
+%[5,] 9.670685e-01 5.806178e-01 4.675586e-01 6.993742e-01 1.000000e+00
 
@@
 The calibration of model suggests its potential application to policy issues and can therefore have practical applications in the future.
-For example the effect of exogenous shocks in the socio-economic structure could be tested as the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to further pandemic outbreaks on the single market.
+For example the effect of exogenous shocks in the socio-economic structure can be tested as we did in the stylized application example above.
+Such shocks can for example be the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to further pandemic outbreaks on the single market.
 In this sense the UK economic sectors mostly dependent on foreign capital and generating the highest turnovers in 2018 could be examined: extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods (in terms of out-going links from the UK) ; activities of head offices and holding companies, sale of cars and manufacture of motor vehicles (in terms of in-going links to the UK).
+We showed in our experiment that although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalization and thus can be expected to have a detrimental effect on UK economy due to these foreign dependency links (see also S1 Text).
 
@@
 
-% I believe than something on the results of the COVID/lockdown scenario should be commented in the discussion as well
-
 \section*{Supporting information}
@@
 \paragraph*{S1 Text.}
-\label{S1_Text}
+\label{S1Text}
 {\bf UK's economic sectors dependent on international links.}
 
+Summary of links between UK and the rest of the EU, by economic sector.
+
 \paragraph*{S2 Text.}
-\label{S2_Text}
+\label{S2Text}
 {\bf Model setup and sensitivity analysis.}
 
-Socio-economic distance matrix, Synthetic model setup and Sensitivity analysis.
+Details on the simulation model: summary statistics for the socio-economic distance matrix, procedure for the synthetic model setup and detailed Global Sensitivity Analysis.
 

### 2020-09-11 CoverLetter.tex
Update on Overleaf.

@@
 We are pleased, with my co-authors Natalia Zdanowska (University College London) and Elsa Arcaute (University College London), to submit an original research article entitled ``Modeling growth of urban firm networks'' for consideration for publication in PLoS ONE.
-This article introduces a generative model for urban firm networks, to understand underlying processes and with possible application to policies.
-It does not contain previously published work.
+This article introduces a generative model for urban firm networks, to understand underlying processes, with possible applications to public policies.
+%It does not contain previously published work.
 Preliminary results were presented as a satellite abstract at the Conference on Complex Systems 2019 (September 2019) and will be presented as a main track abstract at the NetScience 2020 conference (September 2020).
-We did not have any previous interaction with PLOS regarding this manuscript.
+We did not have any previous interaction with PLoS regarding this manuscript.
 
-This manuscript has not been published and is not under consideration for publication elsewhere. A preprint was deposited on arxiv simultaneously to the submission.
+This manuscript has not been published and is not under consideration for publication elsewhere, nor contains previously published work.
+A preprint is being deposited on arxiv simultaneously to the submission.
 
@@
 Rybski (Potsdam Institute for Climate Impact Research).
+We know that the official deadline has passed, nevertheless, we hope that it could still be considered for this special issue.
 

### 2020-09-11 main.tex
Update on Overleaf.

@@
 \begin{figure}
-    \begin{center}
-        \includegraphics[width=\linewidth]{figures/Fig1.png}
-    \end{center}
+    %\begin{center}
+    %    \includegraphics[width=\linewidth]{figures/Fig1.png}
+    %\end{center}
+    \vspace{2cm}
     \caption{Distributions of network properties.
@@
 \begin{figure}
-    \begin{center}
-    \includegraphics[width=\linewidth]{figures/Fig2.png}
-    \end{center}
+    %\begin{center}
+    %\includegraphics[width=\linewidth]{figures/Fig2.png}
+    %\end{center}
+    \vspace{2cm}
     \caption{{\bf Map of network communities obtained by modularity maximisation.} The area of circles gives the turnover within the Functional Urban Area, and color the community it belongs to.
@@
 \begin{table}[!ht]
-\begin{adjustwidth}{-2.25in}{0in}
+\begin{adjustwidth}{-1in}{0in}
 \caption{{\bf Statistical models.} Ordinary Least Square estimations of statistical models explaining $\log(w_{ij})$ (standard errors in parenthesis).
@@
 \begin{figure}
-    \begin{center}
-        \includegraphics[width=\linewidth]{figures/Fig3.png}
-    \end{center}
+    %\begin{center}
+    %    \includegraphics[width=\linewidth]{figures/Fig3.png}
+    %\end{center}
+    \vspace{2cm}
     \caption{Example of simulated networks at $t=1500$, for a synthetic system of cities (Top row) and the European urban network (Bottom row), with a high (resp.
@@
 Indices were estimated with a Sobol sequence sampling of 20,000 model runs \cite{saltelli2010variance}.
-We give in \ref{S4_Text} the full estimated results.
+We give in S2 Text the full estimated results for sensitivity indices.
 To summarize, we find that indicators related to modularity (internationalisation and optimal modularity) are mostly influenced by geographical parameters.
@@
 \begin{figure}
-	\begin{center}
-	    \includegraphics[width=\linewidth]{figures/Fig4.png}
-	\end{center}
+	%\begin{center}
+	%    \includegraphics[width=\linewidth]{figures/Fig4.png}
+	%\end{center}
+	\vspace{2cm}
     \caption{(Top Left) Internationalisation index decreases exponentially with gravity decay; (Top Right) Correlation between city weighted degree and size.
@@
 \begin{figure}
-    \includegraphics[width=\linewidth]{figures/Fig5.png}
+    %\includegraphics[width=\linewidth]{figures/Fig5.png}
+    \vspace{2cm}
     \caption{The transition as a function of interaction range depends on the influence of origin size $\gamma_O$; sector proximity $\gamma_S$ plays a role only for a large influence of the origin.\label{fig:fig5}}
@@
 \begin{figure}
-    \includegraphics[width=\textwidth]{figures/Fig6.png}
+    %\includegraphics[width=\textwidth]{figures/Fig6.png}
+    \vspace{2cm}
 	\caption{Maximal integration in terms of community size is achieved at an intermediate value of d$_o$: emergence of a regional regime; maximal size depends on the role of sectors $\gamma_S$ in a decreasing way when the origin size is deactivated, and in an increasing way when $\gamma_o=1$; this regime disappears when the influence of the origin size becomes too large.\label{fig:fig6}}
@@
 \begin{figure}
-    \begin{center}
-        \includegraphics[width=\linewidth]{figures/Fig7.png}
-    \end{center}
+    %\begin{center}
+    %    \includegraphics[width=\linewidth]{figures/Fig7.png}
+    %\end{center}
+    \vspace{2cm}
     \caption{\textbf{Impact of urban hierarchy.} (Left) Internationalisation as a function of the initial synthetic hierarchy $\alpha$, for different values of interaction range $d_0$ (color); (Right) Metropolisation as a function of $\alpha$, for different values of interaction range $d_0$ (color).\label{fig:fig7}}
@@
 \begin{figure}
-	\begin{center}
-    \includegraphics[width=0.75\linewidth]{figures/Fig8.png}
-    \end{center}
+	%\begin{center}
+    %\includegraphics[width=0.75\linewidth]{figures/Fig8.png}
+    %\end{center}
+    \vspace{2cm}
 	\caption{\textbf{Model calibration results.} Pareto front obtained for the bi-objective model calibration for $\varepsilon_L$ and $\varepsilon_M$ objectives.
@@
 
-\subsection*{Application: impact of economic shock}
+\section*{Application: impact of economic shock}
 
@@
 These can be the consequence of diverse factors, including socio-economic, geopolitical factors (example of Brexit), or other unpredicted human disasters such as the COVID-19 pandemic.
-We simulate border closures by rescaling $c_0$, and intra-country lock-downs by rescaling $d_0$ during the simulation, with other parameters being fixed to their default values and the model setup in the real configuration.
-More precisely, with $t_f = 5000$, the model is run for $t_f / 2$ steps when parameters are changed as $\kappa_G \cdot d_0$ and $\kappa_C \cdot c_0$, and then run for the remaining time steps.
+We simulate border closures by rescaling $c_0$ -- the range parameter for socio-economic interactions --, and intra-country lock-downs by rescaling $d_0$ -- the range parameter for direct interactions -- during the simulation, with other parameters being fixed to their default values and the model setup in the real configuration.
+More precisely, with $t_f = 5000$, the model is run for $t_f / 2$ steps.
+Then, following two new rate parameters $\kappa_C \in \left[0;1\right]$ and $\kappa_G \in \left[0;1\right]$ defining the shock, model parameters are updated as $\kappa_G \cdot d_0$ and $\kappa_C \cdot c_0$.
+The model is then run for the remaining time steps until $t_f$.
 We study the temporal trajectory of indicators to see the impact of the shock in time.
@@
 For minimal values of $\kappa_G$ and $\kappa_C$, internationalisation is more than doubled, confirming a reconfiguration of urban networks within countries.
-We confirm the statistical significance of effects observed in Fig.~\ref{fig:fig9} by comparing distributions with Kolmogorov-Smirnov (KS) tests.
-More precisely, we proceed for each indicator and each value of $(\kappa_G,\kappa_C)$ to a two-sided KS test between the statistical distribution of the indicator over repetitions at final time with the distribution in the reference configuration.
-For internationalisation, p-values of the test are all smaller than 0.002 for $\kappa_D < 0.6$ but all larger than $0.15$ for $\kappa_D \geq 0.6$, what means that the shock has a statistically significant impact only when interaction range is more than halved.
+We validate the statistical significance of effects observed in Fig.~\ref{fig:fig9} by comparing distributions with Kolmogorov-Smirnov (KS) tests.
+More precisely, we proceed for each indicator and value of $(\kappa_G,\kappa_C)$ to a two-sided KS test between the statistical distribution of the indicator over repetitions at final time with the distribution in the reference configuration.
+For internationalisation, p-values of the test are all smaller than 0.002 for $\kappa_D < 0.6$ but larger than $0.15$ for $\kappa_D \geq 0.6$.
+This means that the shock has a statistically significant impact only when interaction range is more than halved.
 For community size, significance (p-value smaller than 0.01) is reached only for the smallest value $\kappa_D = 0.2$.
@@
 \begin{figure}
-	\begin{center}
-    \includegraphics[width=\linewidth]{figures/Fig9.png}
-    \end{center}
+	%\begin{center}
+    %\includegraphics[width=\linewidth]{figures/Fig9.png}
+    %\end{center}
+    \vspace{2cm}
 	\caption{\textbf{Application of the calibrated model to an economic shock scenario.} All plots show the trajectory of indicators (internationalisation for the top row and average community size for the bottom row) in time, for different values of interaction decay rescaling $\kappa_G$ (color) and socio-economic distance rescaling $\kappa_C$ (panels).
@@
 Such shocks can for example be the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to further pandemic outbreaks on the single market.
-In this sense the UK economic sectors mostly dependent on foreign capital and generating the highest turnovers in 2018 could be examined: extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods (in terms of out-going links from the UK) ; activities of head offices and holding companies, sale of cars and manufacture of motor vehicles (in terms of in-going links to the UK).
-We showed in our experiment that although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalization and thus can be expected to have a detrimental effect on UK economy due to these foreign dependency links (see also S1 Text).
+In this sense the UK economic sectors mostly dependent on foreign capital and generating the highest turnovers in 2018, that would be the most affected and worth examining in details are: extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods (in terms of out-going links from the UK) ; activities of head offices and holding companies, sale of cars and manufacture of motor vehicles (in terms of in-going links to the UK).
+We showed in our experiment that although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalization and thus can be expected to have a detrimental effect on UK economy due to these foreign ownership links (see also S1 Text).
 

### 2020-09-11 main.tex
Update on Overleaf.

@@
 More precisely, with $t_f = 5000$, the model is run for $t_f / 2$ steps.
-Then, following two new rate parameters $\kappa_C \in \left[0;1\right]$ and $\kappa_G \in \left[0;1\right]$ defining the shock, model parameters are updated as $\kappa_G \cdot d_0$ and $\kappa_C \cdot c_0$.
+Then, following two new rate parameters $\kappa_C \in \left[0;1\right]$ and $\kappa_G \in \left[0;1\right]$ defining the shock, model parameters are updated as $d_0 = \kappa_G \cdot d_0$ and $c_0 = \kappa_C \cdot c_0$.
 The model is then run for the remaining time steps until $t_f$.

### 2020-09-16 communications/NetSci2020/slidesNetSci.tex

--- 2019-10-02 slides.tex
@@
 
-\title[Agent-based modeling city networks]{An agent-based model for city networks based on interactions between firms}
-\author[Raimbault \and Zdanowska]{J.~Raimbault$^{1,2,3\ast}$ \and N. Zdanowska$^{1,3}$\\\medskip
+\title
+[A generative model for urban firm networks]{A generative model for urban firm networks}
+\author[Raimbault \and Zdanowska \and Arcaute]{J.~Raimbault$^{1,2,3\ast}$ \and N. Zdanowska$^{1,3}$ \and E. Arcaute$^{1}$\\\medskip
 $^{\ast}$\texttt{j.raimbault@ucl.ac.uk}
@@
 
-\date[2nd October 2019]{CCS 2019\\
-Satellite Simulation models\\
-October 2nd 2019
+\date[24 September 2020]{NetSci\\
+September 24 2020
 }
@@
 \begin{itemize}
-    \item Cities as cross-overs of socio-economic interactions 
-    \item Network economies \cite{sassen1991global} \cite{castells1996information}
-    \item World city network driven by interactions between firms \cite{taylor2001specification} \cite{martinus2018global}
-    \item Asymmetrical spatial interactions -- different relative position of cities
+\item Cities as cross-overs of socio-economic interactions           \cite{castells1996information}
+    \item Networks as the new social morphology of societies \cite{castells2000networksociety}, giving rise to network economies \cite{sassen1991global} 
+    \item Emergence of an interconnected world city network as a crucial feature of globalisation 
+    \item Interurban interactions between firms as one of the most representative of economic trends \cite{taylor2001specification} \cite{martinus2018global}
+
 \end{itemize}
@@
     \item Evolutionary Theory of Urban Systems \cite{pumain1997pour} \cite{pumain2006villes}
-    \item Adaptive cycles and diffusion on innovation \cite{Hagerstrand1968Innovation}
+    \item Adaptive cycles and diffusion of innovation \cite{Hagerstrand1968Innovation}
     \item Path dependence \cite{martin2006path} \cite{pumain2012urban}
-    \item Selection and emerging structures of systems % pas clair
+    \item Selection and emerging structures of systems 
     \item Evolutionary models for urban systems dynamics: \cite{favaro2011gibrat} \cite{cottineau2015growing} \cite{schmitt2015half} \cite{raimbault2018indirect} \cite{raimbault2018calibration}
@@
 }
-
-\sframe{Firm linkages structuring urban systems}{
-
- \begin{center}
- \includegraphics[width=0.7\textwidth]{figures/Figure1.png}
- \end{center}
-
-  \textit{Central and Eastern European cities within ownership links of firms in 2013}
-
-}
-
-\sframe{Firm linkages structuring urban systems}{
-
- \begin{center}
- \includegraphics[width=0.7\textwidth]{figures/Figure2.jpg}
- \end{center}
-
-  \textit{UK cities within transnational ownership links of firms in 2016}
-
-}
-
-% bon je ne sais pas si c'est utile mais au moins ca fait un exempple en plus (bon ok on ne voit rien)
 
@@
 
-% j'ai mis ca plus pour moi,enfin pour moi c'est plus clair pour la presentation
-%  => oui c'est exactement ce quil fallait mettre !
-% question a la con, mais est-ce que ce n'est pas tautolgique de dire que les villes et les liens sont definis par le GDP, ou le lien est bien defini par le GDP des noeuds.
 }
@@
 
-% excuses-moi, mais je vais commencer a partir de la a te demander de m'expliquer, ca va te paraitre evident je suis desolee, mais c'est pour que je ne trompe pas en presentant
-% a quoi correspond le i et j? En gros c'est ville i, puis la ville j? Le k c'est bien le nombre de secteurs? 
+% Le k c'est bien le nombre de secteurs? 
 % -> oui exactement
@@
 
-\sframe{Simulation of urban networks}{
+\sframe{Simulation of urban networks}
 
+{
     % examples
@@
     \begin{center}
-    \includegraphics[width=0.48\textwidth]{figures/ex_alleq-highgravity_seed-12102_t1500.png}\hspace{0.1cm}
-    \includegraphics[width=0.48\textwidth]{figures/ex_alleq-lowgravity_seed-12102_t1500.png}
+    \includegraphics[width=0.48\textwidth]{figuresraw/ex_alleq-highgravity_seed-12102_t1500.png}\hspace{0.1cm}
+    \includegraphics[width=0.48\textwidth]{figuresraw/ex_alleq-lowgravity_seed-12102_t1500.png}
 	\end{center}
@@
 
-    \includegraphics[width=0.48\textwidth]{figures/internationalization-gravityDecay_errorbars.png}
-    \includegraphics[width=0.48\textwidth]{figures/rhoDegreeSize-gravityDecay_errorbars.png}
+    \includegraphics[width=0.48\textwidth]{figuresraw/internationalisation-gravityDecay_errorbars.png}
+    \includegraphics[width=0.48\textwidth]{figuresraw/rhoDegreeSize-gravityDecay_errorbars.png}
 
@@
 
-     \includegraphics[width=0.48\textwidth]{figures/internationalization-gammaSectors_errorbars.png}
-    \includegraphics[width=0.48\textwidth]{figures/rhoDegreeSize-gammaSectors_errorbars.png}
+     \includegraphics[width=0.48\textwidth]{figuresraw/internationalisation-gammaSectors_errorbars.png}
+    \includegraphics[width=0.48\textwidth]{figuresraw/rhoDegreeSize-gammaSectors_errorbars.png}
 
@@
     \begin{center}
-    \includegraphics[width=\textwidth]{figures/internationalization_countryGravityDecay200_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+    \includegraphics[width=\textwidth]{figuresraw/internationalization_countryGravityDecay200_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
     \end{center}
@@
     \begin{center}
-    	\includegraphics[height=0.6\textheight]{figures/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+    	\includegraphics[height=0.6\textheight]{figuresraw/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+
     \end{center}

### 2020-09-16 communications/NetSci2020/slidesNetSci.tex
fix_compil

@@
 
-\sframe{Simulation of urban networks}
-
-{
+\sframe{Simulation of urban networks}{
     % examples
@@
     \begin{center}
-    \includegraphics[width=0.48\textwidth]{figuresraw/ex_alleq-highgravity_seed-12102_t1500.png}\hspace{0.1cm}
-    \includegraphics[width=0.48\textwidth]{figuresraw/ex_alleq-lowgravity_seed-12102_t1500.png}
+    \includegraphics[width=0.48\textwidth]{../../figuresraw/ex_alleq-highgravity_seed-12102_t1500.png}\hspace{0.1cm}
+    \includegraphics[width=0.48\textwidth]{../../figuresraw/ex_alleq-lowgravity_seed-12102_t1500.png}
 	\end{center}
@@
 
-    \includegraphics[width=0.48\textwidth]{figuresraw/internationalisation-gravityDecay_errorbars.png}
-    \includegraphics[width=0.48\textwidth]{figuresraw/rhoDegreeSize-gravityDecay_errorbars.png}
+    \includegraphics[width=0.48\textwidth]{../../figuresraw/internationalisation-gravityDecay_errorbars.png}
+    \includegraphics[width=0.48\textwidth]{../../figuresraw/metropolisation-gravityDecay_errorbars.png}
 
@@
 
-     \includegraphics[width=0.48\textwidth]{figuresraw/internationalisation-gammaSectors_errorbars.png}
-    \includegraphics[width=0.48\textwidth]{figuresraw/rhoDegreeSize-gammaSectors_errorbars.png}
+     \includegraphics[width=0.48\textwidth]{../../figuresraw/internationalisation-gammaSectors_errorbars.png}
+    \includegraphics[width=0.48\textwidth]{../../figuresraw/metropolisation-gammaSectors_errorbars.png}
 
@@
     \begin{center}
-    \includegraphics[width=\textwidth]{figuresraw/internationalization_countryGravityDecay200_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+    \includegraphics[width=\textwidth]{../../figuresraw/internationalization_countryGravityDecay200_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
     \end{center}
@@
     \begin{center}
-    	\includegraphics[height=0.6\textheight]{figuresraw/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+    	\includegraphics[height=0.6\textheight]{../../figuresraw/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
 

### 2020-09-18 communications/NetSci2020/slidesNetSci.tex
Update on Overleaf.

@@
 [A generative model for urban firm networks]{A generative model for urban firm networks}
-\author[Raimbault \and Zdanowska \and Arcaute]{J.~Raimbault$^{1,2,3\ast}$ \and N. Zdanowska$^{1,3}$ \and E. Arcaute$^{1}$\\\medskip
-$^{\ast}$\texttt{j.raimbault@ucl.ac.uk}
+\author[Raimbault \and Zdanowska]{J.~Raimbault$^{1,2,3\ast}$ \and N. Zdanowska$^{1,3}$\\\medskip
+$^{\ast}$\texttt{n.zdanowska@ucl.ac.uk}
 }
 
-\institute[UCL]{$^{1}$CASA, UCL\\
+\institute[UCL]{$^{1}$Center for Advanced Spatial Analysis, University College London\\
 $^{2}$UPS CNRS 3611 Complex Systems Institute Paris\\
@@
 
-\date[24 September 2020]{NetSci\\
-September 24 2020
+\date[24/09/2020]{NetSci 2020\\
+Session 12F: Diffusion\\
+September 24th, 2020
 }
@@
 
+\medskip
+
 \begin{itemize}
@@
     \item Networks as the new social morphology of societies \cite{castells2000networksociety}, giving rise to network economies \cite{sassen1991global} 
-    \item Emergence of an interconnected world city network as a crucial feature of globalisation 
-    \item Interurban interactions between firms as one of the most representative of economic trends \cite{taylor2001specification} \cite{martinus2018global}
+    \item Emergence of an interconnected world city network as a crucial feature of globalisation \cite{taylor2001specification}  
+    \item Interurban interactions between firms as one of the most representative of economic trends \cite{martinus2018global}
 
@@
     \item Selection and emerging structures of systems 
-    \item Evolutionary models for urban systems dynamics: \cite{favaro2011gibrat} \cite{cottineau2015growing} \cite{schmitt2015half} \cite{raimbault2018indirect} \cite{raimbault2018calibration}
+    \item Evolutionary models for urban systems dynamics: \cite{favaro2011gibrat} \cite{cottineau2015growing} \cite{schmitt2015half} \cite{raimbault2018calibration} \cite{raimbault2018indirect} \cite{Raimbault_2020}
 \end{itemize}
@@
 
-Assumptions
+Assumptions:
+
+\medskip
 
 \begin{itemize}
-    \item Metropolisation vs. regionalisation effects
-    \item Relationship local/global of multinational firm linkages 
+    \item Metropolisation \& internationalisation effects
     \item Specialisation-driven factors and drivers of innovation
@@
 	\item Includes path-dependency
+\end{itemize}
+
+}
+
+\sframe{Contribution}{
+
+We examine the interactions of European cities within firm linkages defined by
+corporate ownership links and introduce
+
+\medskip
+
+\begin{itemize}
+    \item an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's AMADEUS database;
+    \item a generative network model to simulate the growth of such linkages at urban areas scale;
+    \item the model allows us to compare the effect of different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration;
+   \item we calibrate the model on the empirical network and apply it to stylized economic shock scenarios.
 \end{itemize}
@@
 
-% Le k c'est bien le nombre de secteurs? 
-% -> oui exactement
+\medskip
 
-\bigskip
+Starting from an initial network, at each time step, add a fixed number of links randomly, following a probability as a generalized Cobb-Douglas function
 
-Starting from an initial network, at each time step:
-\begin{enumerate}
-    \item Evolve city sizes $E_i \left(t \textrm{ + } 1 \right) \textrm{ = } f\left(E_j\left(t\right)\right)$ with an interaction model (\cite{raimbault2018indirect} or \cite{cottineau2015growing})
-    % ici le = n'apparait pas dans la formule. le t c'est bien le temps?
-    %  => oui t est le temps ; les = et + et ( ) qui passent pas c'est la template ucl, j'arrive pas a fixer alors il faut passer par le reswitch en mode texte
-    \item Add a fixed number of links randomly, following a probability function of sizes, sector proximity, and geographical and socio-cultural proximity
-\end{enumerate}
+\[
+     p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_O} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_D} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s\left(S_{ik},S_{jk}\right)^{\gamma_S} \cdot \exp \left(- \frac{d_{ij}}{ d_0}\right) \cdot \exp \left(- \frac{c_{ij}}{c_0}\right)
+\]
+
+where $E \textrm{ = } \sum_k E_k$, $W \textrm{ = } \sum_{i,j} w_{ij}$, $s$ is a proximity measure given by cosine similarity, $d_{ij}$ euclidian distance, and $c_{ij}$ a socio-cultural distance
+
+\medskip
+
+\textbf{Model parameters: } $\gamma_O, \gamma_D, \gamma_W, \gamma_S, d_0, c_0$
+
+\medskip
+
+\textbf{Model indicators: } internationalisation (modularity of countries in the network), metropolisation (correlation between weighted degree and city size), optimal modularity, average community size
 
@@
 
-\sframe{Formalization}{
+\sframe{Empirical data}{
 
-Probability for a new link follows a generalized Cobb-Douglas function
+ \begin{center}
+ \includegraphics[width=0.7\textwidth]{figures/Fig2.png}
+ \end{center}
 
-\medskip
+ \footnotesize
 
-\[
-     p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_F} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_T} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s\left(S_{ik},S_{jk}\right)^{\gamma_S} \cdot \exp \left(- \gamma_G \cdot d_{ij}\right) \cdot \exp \left(- \gamma_D \cdot g_{ij}\right)
-\]
+  \textit{Map of network communities obtained by modularity maximisation (European ownership network between Functional Urban Areas, constructed from the AMADEUS database).}
 
-\medskip
+}
 
-where $E \textrm{ = } \sum_k E_k$, $W \textrm{ = } \sum_{i,j} w_{ij}$, $s$ is a proximity measure given by cosine similarity, $d_{ij}$ euclidian distance, and $g_{ij}$ a socio-cultural distance
+\sframe{Statistical models}{
+
+Most general spatial interaction model with fixed effects:
+
+\begin{equation}
+\log w_{ij} \textrm{ = } \log d_{ij} \textrm{ + } \log T_i \textrm{ + } \log T_j \textrm{ + } \log s_{ij} \textrm{ + } \alpha_{c_i,c_j} \textrm{ + } \varepsilon_{ij}
+\end{equation}
 
@@
 
-\textbf{Model parameters: } $\gamma_F, \gamma_T, \gamma_W, \gamma_S, \gamma_D \textrm{ = } \frac{1}{d_G}, \gamma_G \textrm{ = } \frac{1}{d_G}$
+\resizebox{\linewidth}{!}{
+\begin{tabular}{|l|c|c|c|c|c|}
+\hline
+Model  & (1) & (2) & (3) & (4) & (5 - Poisson GLM) \\ 
+\hline
+$\log(d_{ij})$ &      -0.06** (0.03) &   -0.11** (0.05)  & -0.41*** (0.02)  & -0.35*** (0.04)  &  -0.26*** ($5\cdot 10^{-6}$) \\
+$\log(T_i)$ &   &   & 0.56*** (0.01) &  0.56*** (0.01) & 0.79*** ($2\cdot 10^{-6}$) \\
+$\log(T_j)$ &     &   & 0.39*** (0.01) &  0.39*** (0.01) & 0.66***  ($1.5\cdot 10^{-6}$) \\
+$\log(s_{ij})$ &     &   &  &  0.19*** (0.07) & 0.53*** ($9\cdot 10^{-6}$)  \\
+Countries &    &  28.3\% &   &  26.5\% & 100\% \\
+\hline
+$R^2$ &       0.00059   &  0.17 & 0.21 &  0.31  &  0.56 \\
+MSE & 7.75 & 6.84 & 6.10 & 5.33 & 8.72 \\
+AIC &        44304   &  43578  &  42131  & 41917 &   \\
+\hline
+\end{tabular}
+}
 
@@
 
-\sframe{Model indicators}{
+\begin{frame}[label=setupmain]
 
-\textbf{Geographical indicators:}
-
-\begin{itemize}
-    \item Internationalisation (modularity of countries in the network)
-    \item Metropolisation (correlation between weighted degree and city size)
-    \item Regionalisation (correlation between length and flow of links, stratified by size of extremities)
-    \item Specialisation (correlation between sector proximity and flow of links, stratified by size of extremities)
-\end{itemize}
-
-\bigskip
-
-\textbf{Network and flows indicators:}
-
-%  networkModularity, networkAvgCommunitySize, networkDegreeHierarchy, networkAvgDegree, networkDegreeEntropy, flowsHierarchy, flowsEntropy, rhoDegreeSize, rhoFlowDistancePos, rhoFlowDistance
-
-\begin{itemize}
-    \item Louvain modularity, community sizes
-    \item Degree and flows distribution (average, hierarchy, entropy)
-    \item Correlations (degree-size, flow-distance)
-\end{itemize}
-
-}
-
-\sframe{Simulation on synthetic systems of cities}{
+\frametitle{Model setup}
 
@@
 
-\textit{Following \cite{raimbault2018space}, geosimulation models must be studied within synthetic controllable urban contexts in order to (i) understand intrinsic behavior of the model and robust qualitative stylized facts; (ii) study the sensitivity to the spatial configuration}
-
+\begin{enumerate} 
+    \item Simulation on synthetic systems of cities \cite{raimbault2019space}: generation of a continent-scale urban system with stylized order of magnitude corresponding to Europe (urban hierarchy, countries, industrial sector distributions - \hyperlink{syntheticsetup}{\beamerbutton{Details}})
     \bigskip
-
-Generation of a continent-scale urban system with stylized order of magnitude corresponding to Europe:
-
-\medskip
-
-\footnotesize
-
-\begin{enumerate}
-    \item Generate $N \textrm{ = } 700$ cities with size following a power law $E_i \textrm{ = } E_0 \cdot i^{-\alpha}$ with $E_0 \textrm{ = } 10^{11}$ and $\alpha \textrm{ = } 1.1$ (computed on Europe for GDP with cities larger than 50.000 inhabitants)
-    \item Distribute them randomly in space (\cite{simini2019testing} vs \cite{banos2011christaller})
-    \item Create countries with k-means clustering ($C \textrm{ = } 30$)
-    \item Distribute sectors such that (i) smaller cities are more specialized and (ii) larger cities are more knowledge-based, with a one dimensional axis to position sectors $1/K \ldots 1$ where the density $f\left(k\right)$ follows a log-normal with $\left(\mu,\sigma\right)$ such that $\sqrt{\Var f} \textrm{ = } K/2$ for the largest, $\sqrt{\Var f} \textrm{ = } 1/K$ for the smallest
+    \item Simulation on the European system of cities (AMADEUS data at FUA level)
 \end{enumerate}
 
-}
+\end{frame}
 
@@
 
+\footnotesize
+
 \textbf{Implementation}
@@
 	\item Model implemented in NetLogo (good compromise interactivity / ergonomy), with fast data structures (matrix/table extensions)
-	\item Integrated seamlessly into OpenMOLE \cite{reuillon2013openmole} for model exploration (NetLogoTask)
+	\item Integrated seamlessly into OpenMOLE \cite{reuillon2013openmole} for model exploration \url{https://openmole.org/}
 \end{itemize}
+
+\begin{center}
+\includegraphics[width=0.1\linewidth]{../../figuresraw/iconOM.png}
+\includegraphics[width=0.4\linewidth]{../../figuresraw/openmole.png}
+\end{center}
 
@@
 
-% for now no dynamics of cities
-
 \begin{itemize}
-	\item Current experiment: only network dynamics (short time scale)
 	\item One-factor sampling with 100 repetitions to assess statistical properties (good convergence, average sharpe ratios for indicators all larger than 5)
-	\item Grid sampling with 20 repetitions for model behavior %(2years 5months of computation)
+	\item Grid sampling with 20 repetitions for model behavior
+	\item Targeted experiment on impact of synthetic hierarchy
+	\item Global Sensitivity Analysis \cite{saltelli2008global}
+	\item Calibration on real setup with a Genetic Algorithm
+	\item Application to economic shocks scenarios
 \end{itemize}
@@
 \sframe{Simulation of urban networks}{
-    % examples
 
  \begin{center}
-    \includegraphics[width=0.48\textwidth]{../../figuresraw/ex_alleq-highgravity_seed-12102_t1500.png}\hspace{0.1cm}
-    \includegraphics[width=0.48\textwidth]{../../figuresraw/ex_alleq-lowgravity_seed-12102_t1500.png}
+ \includegraphics[width=0.67\textwidth]{figures/Fig3.png}
  \end{center}
 
-	\medskip
+\footnotesize
 
-	\textit{Networks at $t \textrm{ = } 1500$, for default parameters values and high gravity (left) and low gravity (right)}
+	\textit{Simulated networks in a synthetic setup (top) and real setup (bottom), for a large (resp. low) interaction range $d_0$ (left, resp. right).}
 
 }
-
-%\sframe{Statistical convergence}{
-
-%$\rightarrow$ Good convergence for most indicators: average sharpes ratios on a one-factor sampling above 5 (except for metropolization index $\simeq 1.5$)
-
-% + histograms
-
-%\bigskip
-%}
 
@@
 }
-
-% alors la je ne comprend pas desolee l'interpretation de l'evolution des courbes, l'internationalisation (qui est defini par la modularite diminue avec le temp?
-% et le p(di,E) proba entre distance et poids economique? comment tu interpretes ca? il y a un seuil a la fin
-% du coup je ne comprend pas bien comment interpreter
-%  => c'est pas dans le temps, c'est en fonction du parametre de gravity ; internationalization comme c'est la modularite elle est minimisee plus c'est internataional
 
@@
     \begin{center}
-    \includegraphics[width=\textwidth]{../../figuresraw/internationalization_countryGravityDecay200_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+    \includegraphics[width=\textwidth]{../../figures/Fig5.png}
     \end{center}
@@
 
-% alors tu as explique pendant la presentqtion pour Mike, mais je n'ai pas compris pourquoi il a trois graphiques 0 1 2 et pourquoi le 3 est si different:?
-% je pense qu'il faut reconmmer en un truc plus clair l'axe des ordonnees et les 0, 1, 2
-% dG c'est quoi, c'est toujours le pqs de temps? donc la meme chose que le gravity decay?
-
 \sframe{Grid exploration: community size}{
@@
     \begin{center}
-    	\includegraphics[height=0.6\textheight]{../../figuresraw/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png}
+    	\includegraphics[height=0.6\textheight]{../../figures/Fig6.png}
 
@@
 
-% alors ici il faut reconnomer l'axe des ordonnees ca peut etre community size?
-qui est defini comment d'ailleurs ici?
-C'est la ou tu as dit que le pic aussi rapide est interessant, donc en si peu de temps la taille du reseaux grossi c'est ca :'(?
-et du coup comment expliquer le 2 a droite?
+\sframe{Calibration on the European urban network}{
 
-%\sframe{Grid exploration}{
-%    \centering
-%
-%    \includegraphics[width=\textwidth]{figures/rhoFlowDistance_countryGravityDecay2100_gammaDestination2_facetwrapgammaOrigin_colorgammaSectors.png}
-%}
+    \begin{center}
+    	\includegraphics[height=0.75\textheight]{figures/Fig8.png}
 
-% idem la c'est pas facile a lire pour quelqu'un qui ne connait pas. c'est quoi rhoFlowDistance?
-% le gamma sector c'est le parametre de la variable secteurs donc s? 
-% en fait quand j'y pense pourquoi il yq u gradien du gamma et comment comprendre ces varations pour chqaue pqs de temps
+    \end{center}
 
-% other experiments ? PSE on correlations ?
+    \footnotesize
+
+\textit{Pareto front obtained for the bi-objective model calibration using a NSGA2 algorithm (outperforms statistical models that have a MSE of 5.33).}
+
+}
+
+\sframe{Economic shock scenarios}{
+
+    \begin{center}
+    	\includegraphics[height=0.72\textheight]{figures/Fig9.png}
+
+    \end{center}
+
+    %\vspace{-0.5cm}
+
+    \footnotesize
+
+\textit{Although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalization and thus can be
+expected to have a detrimental effect on UK economy due to these foreign ownership links.}
+
+}
 
 \section{Discussion}
-
-% - future work etc (classical discussion)
-% - on the role of model exploration and simulation to gain knowledge from the model
-
-% further comparing with real data?
 
@@
 \justify
-\footnotesize
 
@@
 
-$\rightarrow$ effect of exogenous shocks in the socio-economic structure (``subsystem-xit'')
+\medskip
 
-\medskip
+$\rightarrow$ Potential application to public policy issues;
+
+$\rightarrow$ Effect of exogenous shocks on specific economic sectors mostly dependent on foreign ownership links (extraction of crude petroleum, manufacture of tobacco \& basic pharmaceutical goods, activities of head offices).
+
+\bigskip
 
@@
 
-$\rightarrow$ evolution of city sizes (co-evolution model)
+$\rightarrow$ co-evolution model (evolution of city sizes) \cite{raimbault2018modeling};
 
-$\rightarrow$ parametrization/calibration on real data
+$\rightarrow$ a model with firm agents (multi-scale ABM) \cite{raimbault:halshs-02351722};
 
-$\rightarrow$ role of path-dependency
+$\rightarrow$ targeted study of path dependency;
 
-$\rightarrow$ towards a model with firm agents? (multi-scale ABM)
-
-\medskip
-
-\textbf{On the role of model exploration}
-
-$\rightarrow$ even with such a ``simple'' model (close to directly tractable stationary state), behavior is highly non-linear in many dimensions
-
-$\rightarrow$ model exploration allows to overcome hidden parameters (deactivated mechanisms or default parameter values)
-
-$\rightarrow$ exploration of intrinsic dynamics on synthetic data is a crucial step before an application on real data (disentangle effects of geography from model dynamics)
+$\rightarrow$ other formulation of the combination of factors or multi-objective optimisation depending on sectors using Pareto fronts.
 
@@
 
+\textbf{Preprint at} \texttt{https://arxiv.org/abs/2009.05528}
+
+\bigskip
+
 \textbf{Open repository for model and results at}
@@
 
-\textbf{Acknowledgments}: thanks to the \textit{European Grid Infrastructure} for access to the infrastructure.
+\textbf{Acknowledgments}: thanks to the \textit{European Grid Infrastructure} for access to the infrastructure. We also acknowledge the funding of the EPSRC grant EP/M023583/1
 
@@
 
+\sframe{Reserve slides}{
+
+\Huge
+
+\centering
+
+Reserve slides
+
+}
+
+\sframe{Empirical network properties}{
+
+\begin{center}
+    \includegraphics[width=\linewidth]{../../figures/Fig1.png}
+\end{center}
+
+\textit{Empirical distribution of network properties, with log-normal and power law fits}
+
+}
+
+\begin{frame}[label=syntheticsetup]
+
+\frametitle{Synthetic setup}
+
+\begin{enumerate}
+    \item Generate $N \textrm{ = } 700$ cities with size following a power law $E_i \textrm{ = } E_0 \cdot i^{-\alpha}$ with $E_0 \textrm{ = } 10^{11}$ and $\alpha \textrm{ = } 1.1$ (computed on Europe for GDP with cities larger than 50.000 inhabitants)
+    \item Distribute them randomly in space (\cite{simini2019testing} vs \cite{banos2011christaller})
+    \item Create countries with k-means clustering ($C \textrm{ = } 30$)
+    \item Distribute sectors such that (i) smaller cities are more specialized and (ii) larger cities are more knowledge-based, with a one dimensional axis to position sectors $1/K \ldots 1$ where the density $f\left(k\right)$ follows a log-normal with $\left(\mu,\sigma\right)$ such that $\sqrt{\Var f} \textrm{ = } K/2$ for the largest, $\sqrt{\Var f} \textrm{ = } 1/K$ for the smallest
+\end{enumerate}
+
+\hyperlink{setupmain}{\beamerbutton{Back}}
+
+\end{frame}
+
+\sframe{Global sensitivity analysis}{
+
+\begin{center}
+\resizebox{\linewidth}{!}{
+    \begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|c|c|}
+\hline
+ & \multicolumn{2}{|c|}{$d_0$} & \multicolumn{2}{|c|}{$c_0$} & \multicolumn{2}{|c|}{$\gamma_S$} & \multicolumn{2}{|c|}{$\gamma_W$} & \multicolumn{2}{|c|}{$\gamma_O$} & \multicolumn{2}{|c|}{$\gamma_D$} \\
+ & F & T & F & T & F & T & F & T & F & T & F & T \\
+ \hline
+Internationalisation & 0.2 & 0.3 & 0.7 & 0.7 & 0.001 & 0.009 & $4\cdot 10^{-4}$ & 0.007 & 0.03 & 0.04 & 0.02 & 0.04 \\
+Metropolisation & 0.02 & 0.1 & 0.02 & 0.2 & 0.002 & 0.1 & 0.001 & 0.09 & 0.2 & 0.6 & 0.3 & 0.6 \\
+Modularity & 0.3 & 0.4 & 0.6 & 0.6 & 0.004 & 0.02 & $3\cdot 10^{-4}$ & 0.01 & 0.005 & 0.03 & 0.002 & 0.03 \\
+Avg. com. size & 0.008 & 0.09 & 0.01 & 0.1 & 0.002 & 0.07 & 0.003 & 0.04 & 0.3 & 0.6 & 0.4 & 0.6 \\
+Degree entropy & 0.006 & 0.02 & 0.003 & 0.02 & 0.006 & 0.03 & 0.008 & 0.02 & 0.5 & 0.5 & 0.5 & 0.5 \\
+Weight entropy & 0.04 & 0.1 & 0.03 & 0.1 & 0.008 & 0.08 & 0.01 & 0.07 & 0.4 & 0.5 & 0.4 & 0.5 \\\hline
+\end{tabular}
+}
+\end{center}
+
+\medskip
+
+\textit{First order and total sensitivity indices \cite{saltelli2010variance}, for all parameters and indicators.}
+
+}
+
+\sframe{Role of urban hierarchy}{
+
+\begin{center}
+    \includegraphics[width=\linewidth]{../../figures/Fig7.png}
+\end{center}
+
+\medskip
+
+\textit{Impact of urban hierarchy on model indicators in synthetic systems of cities.}
+
+}
+
 \end{document}

### 2020-09-24 communications/NetSci2020/slidesNetSci.tex
Update on Overleaf.

@@
 [A generative model for urban firm networks]{A generative model for urban firm networks}
-\author[Raimbault \and Zdanowska]{J.~Raimbault$^{1,2,3\ast}$ \and N. Zdanowska$^{1,3}$\\\medskip
+\author[Raimbault \and Zdanowska \and Arcaute] {J.~Raimbault$^{1,2,3}$ \and N. Zdanowska$^{1,3}\ast$ \and E. Arcaute$^{1}$\\\medskip
 $^{\ast}$\texttt{n.zdanowska@ucl.ac.uk}
@@
  \begin{center}
- \includegraphics[width=0.7\textwidth]{figures/Fig2.png}
+ \includegraphics[width=0.7\textwidth]{../../figures/Fig2.png}
  \end{center}
@@
  \begin{center}
- \includegraphics[width=0.67\textwidth]{figures/Fig3.png}
+ \includegraphics[width=0.67\textwidth]{../../figures/Fig3.png}
  \end{center}
@@
 
-\sframe{Effect of sector proximity}{
+% \sframe{Effect of sector proximity}{
 
@@
 
-     \includegraphics[width=0.48\textwidth]{../../figuresraw/internationalisation-gammaSectors_errorbars.png}
-    \includegraphics[width=0.48\textwidth]{../../figuresraw/metropolisation-gammaSectors_errorbars.png}
+     % \includegraphics[width=0.48\textwidth]{../../figuresraw/internationalisation-gammaSectors_errorbars.png}
+    % \includegraphics[width=0.48\textwidth]{../../figuresraw/metropolisation-gammaSectors_errorbars.png}
 
-    \medskip
-    \footnotesize
+    % \medskip
+    % \footnotesize
 
-    \textit{(Left) Internationalization varies linearly with sector proximity $\gamma_S$; (Right) Correlation between degree and size exhibits a maximum, witnessing an intermediate regime where size is the most important}
+    % \textit{(Left) Internationalization varies linearly with sector proximity $\gamma_S$; (Right) Correlation between degree and size exhibits a maximum, witnessing an intermediate regime where size is the most important}
 
-}
+% }
 
-\sframe{Grid exploration: internationalization}{
+% \sframe{Grid exploration: internationalization}{
 
-    \begin{center}
-    \includegraphics[width=\textwidth]{../../figures/Fig5.png}
-    \end{center}
+    % \begin{center}
+    % \includegraphics[width=\textwidth]{../../figures/Fig5.png}
+    % \end{center}
 
-    \medskip
+    % \medskip
 
-    \footnotesize
+    % \footnotesize
 
-    \textit{The transition as a function of interaction range depends on the influence of origin size $\gamma_F$; sector proximity $\gamma_S$ plays a role only for a large influence of the origin.}
+    % \textit{The transition as a function of interaction range % % depends on the influence of origin size $\gamma_F$; sector % proximity $\gamma_S$ plays a role only for a large influence of the origin.}
 
-}
+% }
 
@@
     \begin{center}
-    	\includegraphics[height=0.75\textheight]{figures/Fig8.png}
+    	\includegraphics[height=0.75\textheight]{../../figures/Fig8.png}
 
@@
     \begin{center}
-    	\includegraphics[height=0.72\textheight]{figures/Fig9.png}
+    	\includegraphics[height=0.72\textheight]{../../figures/Fig9.png}
 

### 2020-11-18 Response_to_reviewers.tex
add_resp_to_reviewers

%% start of file `template.tex'.
%% Copyright 2006-2013 Xavier Danaux (xdanaux@gmail.com).
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License version 1.3c,
% available at http://www.latex-project.org/lppl/.

\documentclass[10pt,a4paper,sans]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')

\usepackage[document]{ragged2e}
% pour justifier

% moderncv themes
\moderncvstyle{banking}                            % style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
\moderncvcolor{red}                                % color options 'blue' (default), 'orange', 'green', 'red', 'purple', 'grey' and 'black'
\renewcommand{\familydefault}{\rmdefault} 
%\nopagenumbers{} 
\usepackage[utf8]{inputenc} 
\usepackage[scale=0.9]{geometry}

\def \draft {1}
\usepackage{xparse}
\usepackage{ifthen}
\DeclareDocumentCommand{\comment}{o m o o o o}
{\ifthenelse{\draft=1}{
  \IfValueT{#1}{
      \textcolor{red}{\textbf{C (#1) : }#2}
      \IfValueT{#3}{\textcolor{blue}{\textbf{A1 : }#3}}
      \IfValueT{#4}{\textcolor{green}{\textbf{A2 : }#4}}
      \IfValueT{#5}{\textcolor{red!50!blue}{\textbf{A3 : }#5}}
      \IfValueT{#6}{\textcolor{blue}{\textbf{A4 : }#6}}
    }
    \IfNoValueT{#1}{
      \textcolor{red}{\textbf{C : }#2}
      \IfValueT{#3}{\textcolor{blue}{\textbf{A1 : }#3}}
      \IfValueT{#4}{\textcolor{green}{\textbf{A2 : }#4}}
      \IfValueT{#5}{\textcolor{red!50!blue}{\textbf{A3 : }#5}}
      \IfValueT{#6}{\textcolor{blue}{\textbf{A4 : }#6}}
    }
 }{}
}

\firstname{}
\lastname{}
\begin{document}

% recipient data
\recipient{Editor PONE}{}
\date{\today}
\opening{Dear Editor,}
\closing{Yours faithfully,\\
Juste Raimbault and co-authors\\
University College London
}
         % use an optional argument to use a string other than "Enclosure", or redefine \enclname

%\makelettertitle

\justify

\textbf{Reply to the editor / Journal requirements}

\textit{}

\medskip

$\rightarrow$

\bigskip

\textbf{Reply to reviewers}

\medskip

\textbf{First referee:}

\medskip

%### Summary of the research and your overall impression

%The authors develop a model aiming at capturing the determinants of inter-city firms ownership links formation. The model can be seen as an interesting combination of gravity models and preferential attachment that allows to account for the effect of path-dependence (or more accurately, the effect of circular causation/feedback).
%The theoretical implications of the model are rigorously explored by uncovering the link between the model parameters and some of the city network structural features.
%The model is eventually fitted to empirical data to evaluate the impact of the model variables on the formation of firms ownership links. The results are consistent with the literature where the gravity model has been applied in similar contexts (mainly trade and migration), although this is not explicitly mentioned by the authors. The effect of path dependence is found to be strongly significant.
%The authors then use the calibrated model to evaluate the effect of movement restrictions and change in institutional proximity on the city network structure. While the observed effects seem qualitatively relevant, it is difficult to assess the practical usefulness of this exercise. Because: 1) the authors don't discuss the ability of their model to reproduce the structural features of the empirical network that it was calibrated on, and, 2) the global predictive performance of the model (through the Mean squared error on log and the log of mean squared error) seems insufficient for such an exercise (roughly almost an order of magnitude of difference) seems too large at a first glance.
%In conclusion, the idea of the model is interesting and novel. The developed model appears to be a useful tool to analyze the effects of the chosen variables in the case of inter-city firm ownership links formation. Using it as a predictive tool seems however less convincing.

%### Major issues

\begin{enumerate}
%1.1 -

	\item \textit{The work is at the intersection of three strands of literature: a) spatial interaction models, b) network analysis, and c) preferential attachment. Yet this does not appear in the literature review. The work could gain more depth from a better discussion of the literature it is grounded in.}

	$\rightarrow$

	\medskip

%1.2 -

	\item \textit{The empirical network is sparse (density ~ 0.03). Do the authors consider the absent links in the regression and/or the generative model they propose ? (from the fact that the log is taken in the statistical model and the computation of the mean square error on logs, it appears that it is not the case). This could be made clear and discussed, especially that the absence of links is the norm rather than the exception, meaning that the model only treats only a minor proportion of cases.}

	$\rightarrow$

	\medskip

%1.3 -

	\item \textit{Given that the model is generative, the temporal aspect of the model can be an issue. Indeed, if the number of iterations (around 5000) needed to generate the empirical network correspond to a long time period (for example decades), the economic size and sectoral proximity are expected to change, probably implying a substantially different network, if these changes were accounted for in the model. Can the authors mention (and may be discuss) this issue?}

	$\rightarrow$

	\medskip

%1.4 -

	\item \textit{Application: Two potentially problematic aspects: 1) the MSE on logs and the log MSE appear to be quite large to perform an accurate prediction (for an MSE on logs $\sim 5$, the average multiplicative error can roughly be expected to be around $\exp(5^0.5) \sim 9$, which is almost an order of magnitude of difference; the same quick calculation can be applied to the log MSE). Can the authors discuss this point? (May be using a mean absolute error instead of mean squared error is a more tangible estimate of the error?)}

	$\rightarrow$

	\medskip

	\item \textit{2) The authors do not show how well the model reproduces the metropolization and the communities of the empirical network. If the model does not perform well at reproducing these properties of the network, it can mean that the predictors they used are not relevant to such features. This would imply that the model is of limited relevancy to study the impact of crises on these structural features.}

	$\rightarrow$

	\medskip

%### Minor issues
%2.1 -

	\item \textit{Do you consider self-connections in the network ? These have been shown to represent an important proportion of firm ownership links.}

	$\rightarrow$

	\medskip

% 2.2 - 
	\item \textit{Could you please better justify the choice of indicators (Internationalizationand Metropolization).}

	$\rightarrow$

	\medskip

%2.3 -

	\item \textit{The model is ambiguously described. The link probability update $p_{ij}$ is between 0 and 1, does it mean that 1) at each time step each link ij has a $p_{ij}$ probability to be updated or that 2) like in the preferential attachment model, only one link is updated at each time step (and thus the real probability is $p_ij/sum_{lk}(p_{lk})$)?}

	$\rightarrow$

	\medskip

%2.4

	\item \textit{Statistical analysis: A negative binomial or Gamma regression can also be tried. In the case the absence of links is taken into account in regression, a zero-inflated or hurdle regression can be done.}

	$\rightarrow$

	\medskip

%### Misc:

	\item \textit{For a better clarity, please precise that the model generates a directed network.}

	$\rightarrow$

	\medskip

	\item \textit{Link creation probability formula - industrial proximity: oversight to remove k (should not depend on k).}

	$\rightarrow$

	\medskip

	\item \textit{"Model 4 also indicates the absence of overfitting": Can the authors be more precise?}

	$\rightarrow$

	\medskip

	\item \textit{Model exploration: The variation of community size was explored but the indicator has not been defined.}

\end{enumerate}

\bigskip
\bigskip

\textbf{Second referee:}

\medskip

%This article presents a very interesting and novel approach to identifying corporate city networks, and for testing their effects in terms of economic resilience in cities. The approach is well-supported by previous literature and techniques, and it is clear that the authors have a solid background in statistical analysis.
%My overall assessment is that the paper is high-quality and along the lines of what should be published in PLOS One. Thus rather than focusing on what the paper's merits are, I will point to a number of areas for improvement. If I am not mistaken, PLOS One does not have a word count limit, meaning that -- unlike many other journals -- there are effectively no limitations on the amount of detail the authors can provide.
%My main comment is that this paper would be almost unintelligible to the 'average' reader, even an economic geographer. It is full of jargon (not in a bad way) that is clearly of more use to a network scientist, statistician, or even a physicist with an understanding of basic geography. However, as an economic geographer myself, I must say that I would have liked to see a much greater focus on the actual geography, and theory, behind the analysis. To rectify this, I suggest the following:

\begin{enumerate}

%1.

	\item \textit{Clarify the paper's aims. The part about border closures is interesting but seems almost tacked-on. Is the paper about 'what if' borders closed (e.g. COVID, Brexit), or is it about the communities that exist and the relative hierarchies within the network. To be honest, I don't know if I could pick one after reading it. The conclusion claims that "this paper [is] aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMole platform and calibrated on real data on ownership linkages of firms for Europe from the Amadeus dataset". This should be reoriented toward a more theoretical framing, and speak to the network geographic ties behind it.}

	$\rightarrow$

	\medskip

%2.
	\item \textit{Clarify the jargon. What is a 'generative model'? What is 'synthetic' data?}

	$\rightarrow$

	\medskip

%3.
	\item \textit{Clarfiy the methods and data. Explain what Amadeus is, how it is gathered etc. Explain the tools that are used to analyse it. Why those statistical packages and not others? It would also be useful to describe the data. How many cities and how many firms?}

	$\rightarrow$

	\medskip

%4.
	\item \textit{And most importantly, clarify the theoretical framing. There is plenty of literature cited, but which of these provided inspiration for this study? For example, is this speaking to the 'global city networks' literature (e.g. Neal, Derudder) or more the economic connectivity literature (e.g. Acemoglu)?}

	$\rightarrow$

	\medskip

%Once these issues are addressed I feel the paper will be in much better shape. There also a few awkward phrases (e.g. 'crow fly' distance) throughout, so I recommend a carefully re-reading prior to resubmission.

\end{enumerate}

\bigskip

\textbf{Third referee:}

\medskip

%It is with great interest that I read this paper, as I feel it has much to contribute to current debates on cities - in particular how to measure specific processes. The authors have done well to think about how the complexity of city processes might be measured, but I feel there are issues that need to be addressed before it is ready to be published in this journal.

\begin{enumerate}
%1)

	\item \textit{The main literature discussion appears to be in the introduction - but this is relatively high level. The research gap that the authors appear to be filling is the operationalisation of the processes (internationalisation, metropolisation, regionalisation and specialisation) within a city. However, it is unclear how this links to the processes you idetnfy as driving the network (listed 1-4 on pg.2) and then the model itself. The processes mentioned appear to be various ways of seeing proximity (ie agglomeration is spatial proximity, and previous linkages could be related to social or institutional proximity) - and whilst link back to previous work mentioned, there needs to be more discussion of past research.}

	$\rightarrow$

	\medskip

%2)
	\item \textit{The processes are mentioned in the methods section - but again exactly what is being measured and what each means is not well related to the process. For example, why exactly are geographic structures captured by internationalisation? Why have you chosen metropolisation to be represented in this way?}

	$\rightarrow$

	\medskip

%3)
	\item \textit{More information on why the synthetic and real world models have been set up are needed. What do they bring to our understandings? You mention how internationalisation, metropolisation, fit within the model .... but what about regionalisation and specialisation? Are these latter just an outcome of the analysis, whilst the former are model parameters? Then why are they treated as the same in the introduction?}

	$\rightarrow$

	\medskip

%4)
	\item \textit{Why in the statistical analysis do you then look at the 'most determining geo-economic factors influencing the creation of links between cities' - there was not real literature review to identify these - how do you then identify them? What is the relationship to the research question? Also pg 6 again describe the processes driving the networks (from pg 2 as noted above) - but these measures where not really discussed in when the model was being explained. How do they fit?}

	$\rightarrow$

	\medskip

%5)
	\item \textit{The application of economic shocks appears to be an additional part of the model that is not properly discussed again in the literature review. I wonder if this is actually a different paper? I felt there were lots of interesting aspects to the research - and thinking about how to divide out the components to allow better discussion for each would benefit the research in that you could dive alot deeper into specific discussions.}

	$\rightarrow$

	\medskip

%I hope these suggestions help the authors develop this work, I feel it is very interesting and worthwhile. Deeper discussion on the different elements would really help to add to the contribution they are making to the literature. Given word limits - this is likely done by rethinking the research into several papers. Good luck with your revisions.

\end{enumerate}

\end{document}

%% end of file `template.tex'.

### 2021-01-05 Response_to_reviewers.tex
Update on Overleaf.

@@
 
+\textbf{Action points}
+
+\begin{itemize}
+    \item[All]
+    \begin{itemize}
+        \item Rewrite literature review
+        \item discuss possible extension with self-connex
+        \item better model description
+        \item paper positioning
+        \item paper writing (ex jargon, clarity)
+    \end{itemize}
+    \item[JR]
+    \begin{itemize}
+        \item Discuss choice/possibly test other stat models - Hurdle, etc.
+        \item total time is linked to $w_0$ - discuss this
+        \item this is not a prediction model - possibly give values of abs error - discuss more model function
+        \item compute more empirical indicators and compare  with generated networks
+        \item technical clarifications
+    \end{itemize}
+    \item[NZ]
+    \begin{itemize}
+        \item geo lit choice indics
+        \item more precise data description
+    \end{itemize}
+\end{itemize}
+
 \textbf{Reply to the editor / Journal requirements}
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(All) Rewrite lit review}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(JR) Discuss choice/possibly test other stat models}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(JR) total time is linked to $w_0$ - discuss this}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(JR) this is not a prediction model - possibly give values of abs error - discuss more model function}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(JR) compute more empirical indicators and compare  with generated networks}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(All) discuss possible extension with self-connex}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(NZ) geo lit choice indics}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(All) model description}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(JR) other stat models (see above)}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(JR) clarify}
 
@@
 	\item \textit{Model exploration: The variation of community size was explored but the indicator has not been defined.}
+
+	$\rightarrow$ \todo{(All) model description}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(All) paper positioning}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(All) paper writing (ex jargon, clarity)}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(NZ) more precise data description}
 

### 2021-02-23 Response_to_reviewers.tex
Update on Overleaf.

--- 2020-11-18 Response_to_reviewers.tex
@@
 
+\textbf{Action points}
+
+\begin{itemize}
+    \item[All]
+    \begin{itemize}
+        \item Paper writing (jargon, clarity, nonsense in Fig 4 legend on gravity (?))
+        \item Improve model description
+        \item Rewrite literature review, theoretical framing, more literature on included processes
+        \item Paper positioning: rewrite
+        \item Discuss possible extension with self-connection e.g.
+        \item Justify application to economic shocks is a proof-of-concept - add literature
+    \end{itemize}
+    \item[JR]
+    \begin{itemize}
+        \item Discuss choices of stat models (link with literature); possibly test other stat models - Hurdle, Gamma etc. (discuss missing data treated differently than actual zeros)
+        \item This is not a prediction model: discuss more model function; possibly give values of abs error
+        \item Compute more indicators on empirical data and compare with generated networks (the comparison is only on fitting indicators for now)
+        \item Technical clarifications (total time is linked to $w_0$; over-fitting)
+    \end{itemize}
+    \item[NZ]
+    \begin{itemize}
+        \item More geographical literature for the choice of indicators, theoretical framing of processes
+        \item More precise data description, descriptive statistics
+    \end{itemize}
+\end{itemize}
+
 \textbf{Reply to the editor / Journal requirements}
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(All) Rewrite lit review. Include applications of gravity models to trade and migration.}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(NZ) The choice of indicators has been better justified and explained based on citations of previous literature, already using these indicators to capture important globalisation processes in a city level network}
 
@@
 	\item \textit{"Model 4 also indicates the absence of overfitting": Can the authors be more precise?}
+	% \todo{(JR) clarify}
 
-	$\rightarrow$
+	$\rightarrow$ This sentence was unclear, we meant that fit improvement in model 4 was not due to overfitting compared to other models. This was reformulated.
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(NZ) The ADAMDEUS dataset has been better explained in terms of sources and representativity}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(All) Literature review/theoretical framing has been reviewed in order to justify the inspiration by the global city networks litterature}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(NZ) The choice and the purpose of choosing thes indicators to identify economic processes has been better described in the introduction and hopefully it links better with the model}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(NZ) This has been better explained in the literature review and introduction}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(All) improve model description/contextualization}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(All) Literature on geo-economic factors has been enhanced - processes included}
 
@@
 
-	$\rightarrow$
+	$\rightarrow$ \todo{(All) justify proof of concept - add literature}
 

### 2021-02-23 main.tex
Update on Overleaf.

@@
 
+% Keywords
+% Urban firm networks; Simulation model; Model calibration
+
 \section*{Introduction}
@@
 Understanding the drivers of growth and geographical properties of such an urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}.
-In addition, straightening these city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}, by developing strategic industries for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
+In addition, strengthening these city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}.
+This can be achieved through strategic development of certain industrial sectors for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
 Exploring city/firm interactions \cite{storme2019introducing} and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.
@@
 
-In this paper we choose to focus on urban networks of firms to question and capture geographical and economic processes as internationalisation, metropolisation, regionalisation and specialisation of cities within a generative model.
+In this paper we choose to focus on interurban interactions between firms.
+Firm networks have already been investigated in complexity science at city level to unveil agglomeration economies and competitiveness of cities \cite{Rozenblat2010}, regionalisation patterns \cite{Rozenblatetal2017} or to model hierarchy and specialisation \cite{BidaRozenblat2020} \cite{BidaRozenblat2020}.
+Moreover economic complexity indicators have been developed in this sense to characterise the production structure and specialisation of different countries due to economic activity, including those of trade firms \cite{HaussmanHidalgo2009} \cite{OCleary2016}.
+Following this literature the present paper positions itself within the global networks theoretical framework and contributes to question and capture geographical and economic processes as internationalisation, metropolisation, regionalisation and specialisation of cities in terms of firm networks within a generative model.
 The specific question we tackle is the complementarity of different processes to generate such urban networks.
-The originality of the present paper is to examine these issues within the system of cities framework \cite{berry1964cities}.
+The originality of the present paper is to examine these issues within the system of cities \cite{berry1964cities} by following previous literature the field of urban firm networks \cite{Rozenblatetal2017} .
 In this approach the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}.
 We examine the interactions of European cities within firm linkages defined by corporate ownership links.
-Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2019arXiv191014652Z}.
+Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2020Zdanowska}.
 
@@
 We then proceed to a sensitivity analysis and exploration of the model on synthetic data.
-The model is calibrated on real European firm ownership data.
+The model is calibrated on empirical European firm ownership data.
 We finally discuss implications of our results and possible extensions.
@@
 Geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}.
-Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalized to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (confirmed relevant by the fixed effect regression described below); (iii) city economic size, which presents agglomeration effects; (iv) industrial proximity in terms of firm sector composition; and (v) previous linkages (influence of already existing links in the past).
+Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalised to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (confirmed relevant by the fixed effect regression described below); (iii) city economic size, which presents agglomeration effects; (iv) industrial proximity in terms of firm sector composition; and (v) previous linkages (influence of already existing links in the past).
 This last point is crucial to capture accumulated effects and include path-dependencies, justifying further the use of a simulation model instead of a more simple statistical model.
@@
 
-We consider two main macroscopic indicators to quantify the generated network.
-Firstly, geographical structures are captured by internationalisation (modularity of countries in the network).
+We consider two main geographical phenomena and therefore two macroscopic indicators to quantify the generated network.
+Firstly, geographical structures are captured by internationalisation of city-interactions in the context of intensification of globalisation and transnational city networking at the worldwide level \cite{taylor2001specification}.
+Hence we define the internationalisation indicator as the modularity of countries in the network.
 The situation of a minimised modularity corresponds to dense connections between cities of different countries in Europe and sparse connections between cities of the same European country.
@@
 
-Secondly, we consider the metropolisation indicator as the correlation between the weighted degree and the GDP of the city.
-With this indicator we capture the effect of large cities or metropolises of regional importance attracting the biggest amount of links and involved in high interactions with other cities.
-It is computed with a Pearson correlation estimator as
+Secondly, we consider the geographical process of metropolisation.
+The later can be defined as a concentration of population and major economic decision functions in metropolises, in particular at the world-system level \cite{sassen1991global}.
+Metropolises playing the role of gateways of globalisation, are competing between each other to attract the most qualified and dynamic activities \cite{RozenblatPumain2007}.
+Metropolisation gradients have been already explored in literature in city-systems considering several economic and political aspects \cite{RozenblatPumain2018} \cite{PumainRozenblat2019} \cite{Zdanowskaetal2020}.
+We introduce in this paper a metropolisation indicator defined as the correlation between the weighted degree and the GDP of the city, in order to capture the effect of large cities or metropolises of regional importance attracting the biggest amount of links and involved in high interactions with other cities.
+
+The later is computed with a Pearson correlation estimator as
 \begin{equation}
@@
 \subsection*{Empirical data}
+
+We base our study on empirical data from \emph {ADAMDEUS} dataset produced by the Bureau Van Dijk (BVD) on transnational firms in Europe.
+AMADEUS indicates the exact geolocation, turnover and activity codes for owners and owned companies in all sectors in Europe.
+The Bureau Van Dick gathers, homogenizes and completes the information on these firms from national sources, publications of consultancies and audit firms such as KPMG, Deloitte, PricewaterhouseCoopers (PwC), Ernst & Young (EY ), Mazars, and those of other financial organisations such as Moody's, Fitch, Standard & Poor's, Thomson Reuters or Duns and Bradstreet.
+A representativity study shows that \emph {ORBIS} --the Bureau Van Dijk's dataset for firms at the worldwide scale--, captures around 10 percent of reality in terms of amount of foreign direct investments in 2011 in Poland \cite{Zdanowska2018}.
 
@@
 \begin{figure}
-    %\begin{center}
-    %    \includegraphics[width=\linewidth]{figures/Fig1.png}
-    %\end{center}
+    \begin{center}
+        \includegraphics[width=\linewidth]{figures/Fig1.png}
+    \end{center}
     \vspace{2cm}
@@
 The same socio-cultural proximity effect is seen in cities of the United Kingdom and Ireland, and those of Sweden, Finland, Estonia, Latvia and Lithuania.
-The cluster composed of Dutch and Polish cities is justified by the fact that firms from the Netherlands are among the first foreign owners of polish firms since 2010 \cite{2019arXiv191014652Z}.
+The cluster composed of Dutch and Polish cities is justified by the fact that firms from the Netherlands are among the first foreign owners of polish firms since 2010 \cite{2020Zdanowska}.
 The cluster including cities in Germany, Czech Republic, Austria, Slovakia, Hungary and partly in Romania, find its origin in the dependence of the Central and Eastern European economies on German and Austrian large companies mostly from the automotive sector (i.e.
@@
 \begin{figure}
-    %\begin{center}
-    %\includegraphics[width=\linewidth]{figures/Fig2.png}
-    %\end{center}
+    \begin{center}
+    \includegraphics[width=\linewidth]{figures/Fig2.png}
+    \end{center}
     \vspace{2cm}
@@
     \label{fig:fig2}
+    % TODO redo map with open resources (see editor request) https://www.naturalearthdata.com/downloads/110m-cultural-vectors/
 \end{figure}
@@
 Adding the economic size at origin and destination (turnovers $T_i$ and $T$, model 3) slightly improves the explanatory power (model 3).
-The best model among the linear ones is the model 4, with all factors and fixed effects, both in terms of explained variance ($R^2 = 0.31$) and Akaike information criterion.
-Model 4 also indicates the absence of overfitting.
+The best model in terms of explained variance ($R^2 = 0.31$) among the linear ones is the model 4, which includes all factors and fixed effects.
+This model also have the lowest Akaike information criterion among the comparable linear models, meaning that the fit improvement is not due to overfitting.
 Model 5, which is a Poisson model with all variables is much a better fit regarding pseudo-$R^2$ and provides much more significant estimates (all fixed effects significant and much smaller standard deviations).
+
 Across all models, consistent qualitative stylized facts regarding link creation between cities are revealed: (i) distance has a negative influence; (ii) economic sizes have a positive effect, with the size of the origin city being more important (consistent in the ownership links context, as strongly driven by decision at the headquarter/owner level); (iii) industrial proximity has a positive influence, i.e.
@@
 \begin{figure}
-    %\begin{center}
-    %    \includegraphics[width=\linewidth]{figures/Fig3.png}
-    %\end{center}
+    \begin{center}
+        \includegraphics[width=\linewidth]{figures/Fig3.png}
+    \end{center}
     \vspace{2cm}
@@
 \begin{figure}
-	%\begin{center}
-	%    \includegraphics[width=\linewidth]{figures/Fig4.png}
-	%\end{center}
+	\begin{center}
+	    \includegraphics[width=\linewidth]{figures/Fig4.png}
+	\end{center}
 	\vspace{2cm}
@@
 \begin{figure}
-    %\includegraphics[width=\linewidth]{figures/Fig5.png}
+    \includegraphics[width=\linewidth]{figures/Fig5.png}
     \vspace{2cm}
@@
 \begin{figure}
-    %\includegraphics[width=\textwidth]{figures/Fig6.png}
+    \includegraphics[width=\textwidth]{figures/Fig6.png}
     \vspace{2cm}
@@
 As expected, the hierarchy of the urban system has an important impact on model outcomes.
-We find that more hierarchical urban systems tend to produce more international networks (internationalisation indicator is a modularity which is minimized when networks span more accross different countries), what is consistent with the existence of globalized metropolis \cite{sassen1991global}.
+We find that more hierarchical urban systems tend to produce more international networks (internationalisation indicator corresponds to a modularity which is minimized when networks span more across different countries), what is consistent with the existence of globalised metropolis \cite{sassen1991global}.
 This effect is mitigated by lower interaction ranges.
@@
 \begin{figure}
-    %\begin{center}
-    %    \includegraphics[width=\linewidth]{figures/Fig7.png}
-    %\end{center}
+    \begin{center}
+        \includegraphics[width=\linewidth]{figures/Fig7.png}
+    \end{center}
     \vspace{2cm}
@@
 \begin{figure}
-	%\begin{center}
-    %\includegraphics[width=0.75\linewidth]{figures/Fig8.png}
-    %\end{center}
+	\begin{center}
+    \includegraphics[width=0.75\linewidth]{figures/Fig8.png}
+    \end{center}
     \vspace{2cm}
@@
 \begin{figure}
-	%\begin{center}
-    %\includegraphics[width=\linewidth]{figures/Fig9.png}
-    %\end{center}
+	\begin{center}
+    \includegraphics[width=\linewidth]{figures/Fig9.png}
+    \end{center}
     \vspace{2cm}

### 2021-03-02 Response_to_reviewers.tex
.

@@
 
-	$\rightarrow$ \todo{(All) Rewrite lit review. Include applications of gravity models to trade and migration.}
+	% TODO (All) Rewrite lit review
+	$\rightarrow$ \comment{(All) Rewrite lit review. Include applications of gravity models to trade and migration.}
 
@@
 
-	$\rightarrow$ \todo{(JR) Discuss choice/possibly test other stat models}
+	% TODO (JR) Hurdle and zero inflated models
+	$\rightarrow$ \comment{(JR) Discuss choice/possibly test other stat models}
 
@@
 
-	$\rightarrow$ \todo{(JR) total time is linked to $w_0$ - discuss this}
+	% TODO (JR) total time is linked to $w_0$ - discuss this
+	$\rightarrow$ The way the model captures time steps and deals with time granularity was indeed not discussed well in the previous version of the paper.
+Effective time simulated by the model can be set by modifying both $t_f$ (number of steps) and $w_0$ parameters: % develop
 
@@
 
-	\item \textit{2) The authors do not show how well the model reproduces the metropolization and the communities of the empirical network.
+	\item \textit{2) The authors do not show how well the model reproduces the metropolisation and the communities of the empirical network.
 If the model does not perform well at reproducing these properties of the network, it can mean that the predictors they used are not relevant to such features.
@@
 
-	$\rightarrow$ \todo{(All) discuss possible extension with self-connex}
+	% TODO (All) discuss possible extension with self-connex
+	$\rightarrow$ Self-connections within urban areas are not considered in the current version of the model. We added in the discussion this possibility and issues related to such an extension.
 
@@
 % 2.2 - 
-	\item \textit{Could you please better justify the choice of indicators (Internationalizationand Metropolization).}
+	\item \textit{Could you please better justify the choice of indicators (Internationalization and Metropolization).}
 
-	$\rightarrow$ \todo{(NZ) The choice of indicators has been better justified and explained based on citations of previous literature, already using these indicators to capture important globalisation processes in a city level network}
+	% TODO check lit - which lit stream
+	$\rightarrow$ The choice of indicators has been better justified and explained based on citations of previous literature, already using these indicators to capture important globalisation processes in a city level network.
 
@@
 
-	$\rightarrow$ \todo{(All) model description}
+	% TODO (All) model description
+	$\rightarrow$
 
@@
 
-	$\rightarrow$ \todo{(JR) other stat models (see above)}
+	% TODO (JR) other stat models (see above)
+	$\rightarrow$ This remark converges with one point raised by the second referee (see ).
 
@@
 
-	$\rightarrow$ \todo{(All) model description}
+	% TODO model description: directed network
+	$\rightarrow$
 

### 2021-03-02 main.tex
.

@@
 
-We base our study on empirical data from \emph {ADAMDEUS} dataset produced by the Bureau Van Dijk (BVD) on transnational firms in Europe.
+We base our study on empirical data from \emph{ADAMDEUS} dataset produced by the Bureau Van Dijk (BVD) on transnational firms in Europe.
 AMADEUS indicates the exact geolocation, turnover and activity codes for owners and owned companies in all sectors in Europe.
-The Bureau Van Dick gathers, homogenizes and completes the information on these firms from national sources, publications of consultancies and audit firms such as KPMG, Deloitte, PricewaterhouseCoopers (PwC), Ernst & Young (EY ), Mazars, and those of other financial organisations such as Moody's, Fitch, Standard & Poor's, Thomson Reuters or Duns and Bradstreet.
-A representativity study shows that \emph {ORBIS} --the Bureau Van Dijk's dataset for firms at the worldwide scale--, captures around 10 percent of reality in terms of amount of foreign direct investments in 2011 in Poland \cite{Zdanowska2018}.
+The Bureau Van Dick gathers, homogenizes and completes the information on these firms from national sources, publications of consultancies and audit firms such as KPMG, Deloitte, PricewaterhouseCoopers (PwC), Ernst \& Young (EY), Mazars, and those of other financial organisations such as Moody's, Fitch, Standard \& Poor's, Thomson Reuters or Duns and Bradstreet.
+A representativity study shows that \emph{ORBIS} --the Bureau Van Dijk's dataset for firms at the worldwide scale--, captures around 10 percent of existing foreign direct investments in 2011 in Poland \cite{Zdanowska2018}.
+\comment[JR]{To what database is Orbis compared to in this study - what is the confidence on its exhaustivity}
+\cite{bajgar2020coverage} find a representative coverage of the \emph{ORBIS} database on OECD countries, but with a better performance on larger firms.
+As we aggregate company turnovers on urban areas in our analysis, we thus do not expect much representativity bias in the inter-urban network.
 
-Companies from \emph {AMADEUS} are georeferenced using the Geonames database (using postcode or address depending on the information available).
+Companies from \emph{AMADEUS} are georeferenced using the Geonames database (using postcode or address depending on the information available).
 We then join this data with the Global Human Settlement dataset GHS-FUA for Functional Urban Areas \cite{Florczyk2019ghs} -- which delineate the commuting areas for urban centres.
@@
 Estimation results are given in Table~\ref{tab:reg}.
-We also fit a Poisson generalized linear model with a logarithmic link on the rounded weights \cite{flowerdew1988fitting}.
+We also fit a Poisson generalised linear model with a logarithmic link on the rounded weights \cite{flowerdew1988fitting}.
 The findings show that when distance only is taken into account (model 1), it has a significant negative effect on link creation, but with an explained variance close to zero.
@@
 The best model in terms of explained variance ($R^2 = 0.31$) among the linear ones is the model 4, which includes all factors and fixed effects.
-This model also have the lowest Akaike information criterion among the comparable linear models, meaning that the fit improvement is not due to overfitting.
-Model 5, which is a Poisson model with all variables is much a better fit regarding pseudo-$R^2$ and provides much more significant estimates (all fixed effects significant and much smaller standard deviations).
+This model also have the lowest Akaike Information Criterion (AIC) among the comparable linear models, meaning that the fit improvement is not due to overfitting.
+Model 5, which is a Poisson model with all variables is much a better fit regarding pseudo-$R^2$ and provides more significant estimates (all fixed effects significant and smaller standard deviations).
 
-Across all models, consistent qualitative stylized facts regarding link creation between cities are revealed: (i) distance has a negative influence; (ii) economic sizes have a positive effect, with the size of the origin city being more important (consistent in the ownership links context, as strongly driven by decision at the headquarter/owner level); (iii) industrial proximity has a positive influence, i.e.
+To investigate the role of missing links in these results (these have been removed since previous models are fitted on logarithms), we furthermore fit Zero-inflated Poisson \cite{lambert1992zero}.
+%and Hurdle models.
+% TODO fit zeroinfl and Hurdle
+% TODO in discussion: spatial models such as GWR
+
+Across all models, consistent qualitative stylised facts regarding link creation between cities are revealed: (i) distance has a negative influence; (ii) economic sizes have a positive effect, with the size of the origin city being more important (consistent in the ownership links context, as strongly driven by decision at the headquarter/owner level); (iii) industrial proximity has a positive influence, i.e.
 similarity is more important than complementarity for this dimension; (iv) a large part of country fixed-effects are statistically significant; and (v) the order of magnitude of the influence of each factor is roughly the same, which means that no process totally dominates the others.
-This confirms the relevance of including all these different dimensions in the generative network model, since they have a comparable explanatory role in statistical models.
+
+This preliminary analysis confirms the relevance of including all these different dimensions in the generative network model, since they have a comparable explanatory role in statistical models.
 
@@
 The model is implemented in NetLogo, which provides a good compromise between performance and interactive prototyping and exploration \cite{railsback2017improving}.
-Numerical experiments are done by integrating the model into the OpenMOLE software for model exploration \cite{reuillon2013openmole}, which provides sensitivity analysis and calibration methods and a transparent access to high-performance computing infrastructures.
+Numerical experiments are done by integrating the model into the OpenMOLE software for model exploration and validation \cite{reuillon2013openmole}.
+This platform implements sensitivity analysis and calibration methods and offers a transparent access to high-performance computing infrastructures, from local servers to clusters and computing grids.
 The model source code, the aggregated network data (the raw initial firm network cannot be made available for data ownership reasons), and the results are available on the open Git repository of the project at \url{https://github.com/JusteRaimbault/ABMCitiesFirms}.
@@
 We also show networks simulated in the real setting.
+% TODO work on data readme on the repo / do a separate dataverse?
 
@@
 For example, internationalisation and metropolisation both show a transition as a function of d$_o$, from a local to a global regime.
-The sector proximity parameter $\gamma_S$ influences the internationalization linearly, but induces a maximum for the correlation between size and degree, which can be interpreted as a setting where size and the total volume in a city are the most correlated.
+The sector proximity parameter $\gamma_S$ influences the internationalisation linearly, but induces a maximum for the correlation between size and degree, which can be interpreted as a setting where size and the total volume in a city are the most correlated.
 
@@
 Distance decay and sector  parameters are varied with 10 steps, other parameters with three.
-The influence of gravity decay parameters is confirmed when plotting the internationalization index against the distance, which exhibits an exponential decay as shown in Fig.~\ref{fig:fig5}.
+The influence of gravity decay parameters is confirmed when plotting the internationalisation index against the distance, which exhibits an exponential decay as shown in Fig.~\ref{fig:fig5}.
 It interacts with the role of sectors and the size of the origin, with a more significant effect of sector proximity, when size has the largest influence (third panel).

### 2021-05-28 main.tex
Update on Overleaf.

@@
 Juste Raimbault\textsuperscript{1,2,3*},
-Natalia Zdanowska\textsuperscript{1,3},
-Elsa Arcaute\textsuperscript{3}
+Natalia Zdanowska\textsuperscript{4,1,3},
+Elsa Arcaute\textsuperscript{1}
 \\
@@
 \\
+\textbf{4} Luxembourg Institute of Socio-Economic Research, Luxembourg
+
 \bigskip
@@
 We calibrate the model on the European network, outperforming statistical models and showing a strong role of path-dependency.
-Potential applications of the model include the study of mitigation policies to deal with exogenous shocks such as economic crisis or potential lockdowns of countries, which we illustrate with an application on stylized scenarios.
+Potential applications of the model include the study of mitigation policies to deal with exogenous shocks such as economic crisis or potential lockdowns of countries, which we illustrate with an application on stylised scenarios.
 
@@
 % Urban firm networks; Simulation model; Model calibration
+
+\cite{rozenblat2021intra}
+
+% https://zenodo.org/record/4668663#.YHmorxNKi3I GISRUK paper on urban nw - typology
 
@@
 
+% TODO towards dynamical graph generative model, to allow for dynamical calibration (discuss possibilities of dyn calib? ; no edge deletion in the model: also discuss?)
+
 \section*{Supporting information}

### 2021-05-28 main.tex
.

--- 2021-03-02 main.tex
@@
     \vspace{2cm}
-    \caption{{\bf Map of network communities obtained by modularity maximisation.} The area of circles gives the turnover within the Functional Urban Area, and color the community it belongs to.
-Urban Areas in communities of size smaller than 3 were not colored for readability.}
+    \caption{{\bf Map of network communities obtained by modularity maximisation.} The area of circles gives the turnover within the Functional Urban Area, and colour the community it belongs to.
+Urban Areas in communities of size smaller than 3 were not coloured for readability.}
     \label{fig:fig2}
@@
 For the fixed effect by origin-destination pairs of countries, the proportion of significant coefficients is attributed (where $p<0.1$).
-Models (1) to (4) are ordinary least square models, while model (5) is a Poisson generalized linear model with logarithmic link.
+Models (1) to (4) are ordinary least square models, while model (5) is a Poisson generalised linear model with logarithmic link.
 We compare model performance using $R^2$, Mean Square Error, and Akaike Information Criterion.

### 2022-03-10 Response_to_reviewers.tex
Update on Overleaf.

@@
 
-\textbf{Action points}
+%\textbf{Action points}
 
-\begin{itemize}
-    \item[All]
-    \begin{itemize}
-        \item Paper writing (jargon, clarity, nonsense in Fig 4 legend on gravity (?))
-        \item Improve model description
-        \item Rewrite literature review, theoretical framing, more literature on included processes
-        \item Paper positioning: rewrite
-        \item Discuss possible extension with self-connection e.g.
-        \item Justify application to economic shocks is a proof-of-concept - add literature
-    \end{itemize}
-    \item[JR]
-    \begin{itemize}
-        \item Discuss choices of stat models (link with literature); possibly test other stat models - Hurdle, Gamma etc. (discuss missing data treated differently than actual zeros)
-        \item This is not a prediction model: discuss more model function; possibly give values of abs error
-        \item Compute more indicators on empirical data and compare with generated networks (the comparison is only on fitting indicators for now)
-        \item Technical clarifications (total time is linked to $w_0$; over-fitting)
-    \end{itemize}
-    \item[NZ]
-    \begin{itemize}
-        \item More geographical literature for the choice of indicators, theoretical framing of processes
-        \item More precise data description, descriptive statistics
-    \end{itemize}
-\end{itemize}
+%\begin{itemize}
+%    \item[All]
+%    \begin{itemize}
+%        \item Paper writing (jargon, clarity, nonsense in Fig 4 legend on gravity (?))
+%        \item Improve model description
+%        \item Rewrite literature review, theoretical framing, more literature on included processes
+%        \item Paper positioning: rewrite
+%        \item Discuss possible extension with self-connection e.g.
+%        \item Justify application to economic shocks is a proof-of-concept - add literature
+%    \end{itemize}
+%    \item[JR]
+%    \begin{itemize}
+%        OK \item Discuss choices of stat models (link with literature); possibly test other stat models - Hurdle, Gamma etc. (discuss missing data treated differently than actual zeros)
+%        OK\item This is not a prediction model: discuss more model function; possibly give values of abs error
+%        OK \item Compute more indicators on empirical data and compare with generated networks (the comparison is only on fitting indicators for now)
+%        OK \item Technical clarifications (total time is linked to $w_0$; over-fitting)
+%    \end{itemize}
+%    \item[NZ]
+%    \begin{itemize}
+%        OK \item More geographical literature for the choice of indicators, theoretical framing of processes
+%        OK \item More precise data description, descriptive statistics
+%    \end{itemize}
+%\end{itemize}
 
@@
 
-\textit{}
+% redo map with open resources (see editor request) https://www.naturalearthdata.com/downloads/110m-cultural-vectors/
+
+\textit{Maps should use only open resources.}
 
@@
 
-$\rightarrow$
+$\rightarrow$ The map in Figure 2 was redrawn using countries from Natural Earth Data.
 
@@
 
-	% TODO (All) Rewrite lit review
-	$\rightarrow$ \comment{(All) Rewrite lit review. Include applications of gravity models to trade and migration.}
+	$\rightarrow$ Literature review and paper positioning was reworked and extended to reflect this intersection of disciplines and currents.
+	%\comment{(All) Rewrite lit review. Include applications of gravity models to trade and migration.}
 
@@
 
-	% TODO (JR) Hurdle and zero inflated models
-	$\rightarrow$ \comment{(JR) Discuss choice/possibly test other stat models}
+	$\rightarrow$ Additional statistical models extended to account for zeros were fitted and commented. This aspect was better explained.
+	%(JR) Hurdle and zero inflated models\comment{(JR) Discuss choice/possibly test other stat models}
 
@@
 
-	% TODO (JR) total time is linked to $w_0$ - discuss this
 	$\rightarrow$ The way the model captures time steps and deals with time granularity was indeed not discussed well in the previous version of the paper.
-Effective time simulated by the model can be set by modifying both $t_f$ (number of steps) and $w_0$ parameters: % develop
+Effective time simulated by the model can be set by modifying both $t_f$ (number of steps) and $w_0$ parameters.
+This was better explained in the model description.
+	% develop - (JR) total time is linked to $w_0$ - discuss this
 
@@
 
-	$\rightarrow$ \todo{(JR) this is not a prediction model - possibly give values of abs error - discuss more model function}
+	$\rightarrow$ The purpose of both the statistical analysis and the generative model is not prediction.
+Statistical analysis is a preliminary exploration to determine potential processes to be included in the generative model, while the latest is an explanatory model, i.e.
+aimed at explaining the role of different factors in the growth of the urban network (such as the role of urban hierarchy) and test hypothesis in a virtual laboratory.
+It is also used to prototype policy applications, and in that sense is potentially a decision-making model.
+A discussion on model function and the precise positioning of this model was added.
+	%\todo{(JR) this is not a prediction model - possibly give values of abs error - discuss more model function}
 
@@
 
-	% TODO (All) discuss possible extension with self-connex
+	% (All) discuss possible extension with self-connex
 	$\rightarrow$ Self-connections within urban areas are not considered in the current version of the model. We added in the discussion this possibility and issues related to such an extension.
@@
 
-	% TODO check lit - which lit stream
+	% check lit - which lit stream
 	$\rightarrow$ The choice of indicators has been better justified and explained based on citations of previous literature, already using these indicators to capture important globalisation processes in a city level network.
@@
 
-	% TODO (JR) other stat models (see above)
-	$\rightarrow$ This remark converges with one point raised by the second referee (see ).
+	% (JR) other stat models (see above)
+	$\rightarrow$ This remark converges with one point raised by the second referee (see above) - more statistical models were tested.
 
@@
 %3.
-	\item \textit{Clarfiy the methods and data.
+	\item \textit{Clarify the methods and data.
 Explain what Amadeus is, how it is gathered etc.
@@
 
-	$\rightarrow$ \todo{(NZ) The ADAMDEUS dataset has been better explained in terms of sources and representativity}
+	$\rightarrow$ The AMADEUS dataset has been better explained in terms of sources and exhaustiveness.
 
@@
 
-	$\rightarrow$ \todo{(All) Literature review/theoretical framing has been reviewed in order to justify the inspiration by the global city networks litterature}
+	$\rightarrow$ Literature review/theoretical framing has been reviewed in order to justify the inspiration by the global city networks literature.
 
@@
 The research gap that the authors appear to be filling is the operationalisation of the processes (internationalisation, metropolisation, regionalisation and specialisation) within a city.
-However, it is unclear how this links to the processes you idetnfy as driving the network (listed 1-4 on pg.2) and then the model itself.
+However, it is unclear how this links to the processes you identify as driving the network (listed 1-4 on pg.2) and then the model itself.
 The processes mentioned appear to be various ways of seeing proximity (ie agglomeration is spatial proximity, and previous linkages could be related to social or institutional proximity) - and whilst link back to previous work mentioned, there needs to be more discussion of past research.}
 
-	$\rightarrow$ \todo{(NZ) The choice and the purpose of choosing thes indicators to identify economic processes has been better described in the introduction and hopefully it links better with the model}
+	$\rightarrow$ The choice and the purpose of choosing these indicators to identify economic processes has been better described in the introduction and hopefully it links better with the model.
 
@@
 
-	$\rightarrow$ \todo{(NZ) This has been better explained in the literature review and introduction}
+	$\rightarrow$ This has been better explained in the literature review and introduction.
 

### 2022-03-10 main.tex
Update on Overleaf.

--- 2021-05-28 main.tex
@@
 \\
-Juste Raimbault\textsuperscript{1,2,3*},
-Natalia Zdanowska\textsuperscript{1,3},
-Elsa Arcaute\textsuperscript{3}
+Juste Raimbault\textsuperscript{1,2,3,4 *},
+Natalia Zdanowska\textsuperscript{5,2,4},
+Elsa Arcaute\textsuperscript{2}
 \\
 \bigskip
-\textbf{1} Centre for Advanced Spatial Analysis, UCL, London, United Kingdom
+\textbf{1} LASTIG, Univ Gustave Eiffel, IGN-ENSG, Saint-Mand{\'e}, France
 \\
-\textbf{2} UPS CNRS 3611 ISC-PIF, Paris, France
+\textbf{2} Centre for Advanced Spatial Analysis, UCL, London, United Kingdom
 \\
-\textbf{3} UMR CNRS 8504 G{\'e}ographie-cit{\'e}s, Paris, France
+\textbf{3} UPS CNRS 3611 ISC-PIF, Paris, France
 \\
+\textbf{4} UMR CNRS 8504 G{\'e}ographie-cit{\'e}s, Paris, France
+\\
+\textbf{5} Luxembourg Institute of Socio-Economic Research, Luxembourg
+
 \bigskip
@@
 Understanding the drivers behind the growth of such networks -- in particular urban firm networks --, is essential for the economic resilience of urban systems.
-We introduce in this paper a generative network model for firm networks at the urban area level including several complementary processes: the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distance.
-An empirical network analysis on European firm ownership data confirms the relevance of each of these factors.
-We then simulate network growth for synthetic systems of cities, unveiling stylized facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range.
-We calibrate the model on the European network, outperforming statistical models and showing a strong role of path-dependency.
-Potential applications of the model include the study of mitigation policies to deal with exogenous shocks such as economic crisis or potential lockdowns of countries, which we illustrate with an application on stylized scenarios.
+We introduce in this paper a generative model for firm networks at the urban area level including several complementary processes: the economic size of urban areas at origin and destination of links, industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distance between urban areas.
+An empirical network analysis on European urban firm ownership data confirms the relevance of each of these factors on new link creation.
+We then simulate network growth on synthetic systems of cities, unveiling stylized facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range.
+We calibrate the model on the European network, outperforming statistical models showing a strong role of path-dependency.
+The model opens up potential applications to formulate public policies recommendations to deal with exogenous shocks such as economic crisis or potential lockdowns of countries.
+This is illustrated with an application on stylised scenarios.
 
@@
 
-In this paper we choose to focus on interurban interactions between firms.
-Firm networks have already been investigated in complexity science at city level to unveil agglomeration economies and competitiveness of cities \cite{Rozenblat2010}, regionalisation patterns \cite{Rozenblatetal2017} or to model hierarchy and specialisation \cite{BidaRozenblat2020} \cite{BidaRozenblat2020}.
-Moreover economic complexity indicators have been developed in this sense to characterise the production structure and specialisation of different countries due to economic activity, including those of trade firms \cite{HaussmanHidalgo2009} \cite{OCleary2016}.
-Following this literature the present paper positions itself within the global networks theoretical framework and contributes to question and capture geographical and economic processes as internationalisation, metropolisation, regionalisation and specialisation of cities in terms of firm networks within a generative model.
+In this paper we choose to focus on interurban interactions between firms considering the preferential attachment paradigm in a network \cite{BarabasiAlbert1999} and spatial interaction model properties % add a paragraph with the most important literature on spatial interaction models and preferential attachment in network and how it can be linked to firm networks.
+Firm networks remains an important aspect of network and system of cities theory \cite{RozenblatNeal2021}.
+They have already been investigated in complexity science at city level to unveil agglomeration economies and competitiveness of cities \cite{Rozenblat2010}, regionalisation patterns \cite{Rozenblatetal2017} or to model hierarchy and specialisation \cite{BidaRozenblat2020}.
+Moreover economic complexity indicators have been developed in this sense to characterise the production structure and specialisation of different countries due to economic activity, including those of trade firms \cite{HaussmanHidalgo2009,OCleary2016}.
+
+Following this literature the present paper positions itself at the intersection of global networks theoretical framework, spatial interaction models and preferential attachment.
+It contributes to question and capture geographical and economic processes as internationalisation, metropolisation, regionalisation and specialisation of cities in terms of firm networks within a generative model.
 The specific question we tackle is the complementarity of different processes to generate such urban networks.
-The originality of the present paper is to examine these issues within the system of cities \cite{berry1964cities} by following previous literature the field of urban firm networks \cite{Rozenblatetal2017} .
+The originality of the present paper is to examine these issues within the system of cities \cite{berry1964cities} by following previous literature the field of urban firm networks \cite{Rozenblatetal2017}.
 In this approach the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}.
@@
 This last point is crucial to capture accumulated effects and include path-dependencies, justifying further the use of a simulation model instead of a more simple statistical model.
+
+It is worth mentioning that our model is not aimed to be predictive, but rather designed as an explanatory model in order to disentangle the influence of diverse processes in the growth of the urban network and to test diverse hypothesis \cite{conte2012manifesto}.
+It also tends to be a decision-making model as we explore potential policy applications.
+Our focus on explanation rather than prediction can be understood in a context of a broad diversity of possible model functions \cite{varenne2018models}.
 
@@
 Parameters of the log-normal law are obtained by solving the corresponding two equations system for each city, and each distribution is discretised and renormalised to be a vector of probabilities.
-See S2 Text for all details of this synthetic setup.
+See S2 Text\ref{S2Text} for all details of this synthetic setup.
 
@@
 A representativity study shows that \emph{ORBIS} --the Bureau Van Dijk's dataset for firms at the worldwide scale--, captures around 10 percent of existing foreign direct investments in 2011 in Poland \cite{Zdanowska2018}.
-\comment[JR]{To what database is Orbis compared to in this study - what is the confidence on its exhaustivity}
-\cite{bajgar2020coverage} find a representative coverage of the \emph{ORBIS} database on OECD countries, but with a better performance on larger firms.
+\cite{bajgar2020coverage} find that the coverage of the \emph{ORBIS} database on OECD countries is a representative sample, but with a better performance on larger firms.
 As we aggregate company turnovers on urban areas in our analysis, we thus do not expect much representativity bias in the inter-urban network.
+% \comment[JR]{To what database is Orbis compared to in this study - what is the confidence on its exhaustivity}
 
@@
 
-The urban network has heavy tail weighted degree and edge weight distributions, as shown in Fig.~\ref{fig:nwdist}.
+% network density: 9323/(573*572) = 0.02844494
+
+The urban network is sparse with a density around $0.028$.
+It has heavy tail weighted degree and edge weight distributions, as shown in Fig.~\ref{fig:nwdist}.
 We fit power law and log-normal distributions, including minimal value cutoff following \cite{clauset2009power}, using the \texttt{powerlaw} R package \cite{powerlawpackage}.
@@
     \caption{Distributions of network properties.
-\textit{(Left)} Cumulated distribution for edge weight, with power law fit (red) and log-normal fit (green); \textit{(Right)} Cumulated distribution for weighted degree.}
+\textit{(Left)} Cumulative distribution for edge weight, with power law fit (red) and log-normal fit (green); \textit{(Right)} Cumulative distribution for weighted degree.}
     \label{fig:nwdist}
@@
 In addition, all cities of the later cluster have in common to be at the intersection of a very important freight transport route in Europe from Germany to Turkey \cite{Zdanowska2017}.
+
 %%%%%%%%%%%%%%
@@
     \vspace{2cm}
-    \caption{{\bf Map of network communities obtained by modularity maximisation.} The area of circles gives the turnover within the Functional Urban Area, and colour the community it belongs to.
+    \caption{{\bf Map of network communities obtained by modularity maximisation.} We show communities with the higher modularity, obtained with the Louvain algorithm.
+The area of circles gives the turnover within the Functional Urban Area, and colour the community it belongs to.
 Urban Areas in communities of size smaller than 3 were not coloured for readability.}
     \label{fig:fig2}
-    % TODO redo map with open resources (see editor request) https://www.naturalearthdata.com/downloads/110m-cultural-vectors/
 \end{figure}
 %%%%%%%%%%%%%%
+
+% TODO redo map with non licensed data for countries: https://www.naturalearthdata.com/downloads/110m-cultural-vectors/
 
@@
 
-To investigate the role of missing links in these results (these have been removed since previous models are fitted on logarithms), we furthermore fit Zero-inflated Poisson \cite{lambert1992zero}.
-%and Hurdle models.
-% TODO fit zeroinfl and Hurdle
-% TODO in discussion: spatial models such as GWR
+To investigate the role of missing links in these results (these have been removed since previous models are fitted on logarithms), we furthermore fit extended statistical models for count data focused on the role of zeros.
+Two different flavors of such models are the hurdle model \cite{mullahy1986specification} which corresponds to a conditional count model, and a zero-inflated Poisson model \cite{lambert1992zero} which mixes a count model with an additional component to generate zeros.
+Full estimation results, done using \cite{zeileis2008regression}, are shown in S2 Text.
+As the network is rather sparse, these more detailed statistical models do not perform better than more simple models, as missing links seem rather hard to predict.
+The best zero-inflated model has a $R^2 = 0.16$, while the best hurdle is just above with $R^2 = 0.17$.
 
@@
 
+% TODO recall not a predictive model
+
 \section*{Application: impact of economic shock}
@@
 We showed in our experiment that although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalization and thus can be expected to have a detrimental effect on UK economy due to these foreign ownership links (see also S1 Text).
+% ~\ref{S1Text} ref not working
 
-Several possible extensions of the current work can be undertaken in the future as a co-evolution model with an evolution of city sizes or a model with firm agents leading to a multi-scale agent-based model.
-A more precise study of the role of path dependency can also be undertaken in order to understand the creation of the future links.
-Other formulations of the model might be taken into consideration as other formulation of the combination of factors or multi-objective optimisation depending on sectors using Pareto fronts.
+Several possible extensions of the current work can be undertaken in the future.
+A co-evolution model with an evolution of city sizes in parallels to the firm network would be more realistic for an application on longer time scales, but would also be more difficult to parametrise and calibrate \cite{raimbault2018characterising}.
+A model with additional firm agents would lead to a multi-scale agent-based model, with also more flexibility for policy applications.
+A precise study of the role of path dependency should also be undertaken in order to anticipate the creation of future links.
+Other formulations of the generative model might be taken into consideration as other formulation of the combination of factors or multi-objective optimisation depending on sectors using Pareto fronts.
+Regarding the statistical analysis, no spatial or neighborhood effects were taken into account, and spatial non-stationarity may be accounted for using Geographically Weighted Regression models for example \cite{comber2021route}.
+Finally, an adaptation at the world scale using for example urban regions and data explored by \cite{rozenblat2021intra}, would be an extension with important practical and theoretical potentialities.
+
+%one can also think about multi-level impact of economic unforeseen shock following \cite{Dorretal2022} study on firm firm-level impact of coronavirus.
+% TODO towards dynamical graph generative model, to allow for dynamical calibration (discuss possibilities of dyn calib? ; no edge deletion in the model: also discuss?)
+
+% Discuss possible extension with self-connection
 
@@
 
-\paragraph*{S1 Text.}
-\label{S1Text}
-{\bf UK's economic sectors dependent on international links.}
+\subsection*{S1 Text: UK's economic sectors dependent on international links.\label{S1Text}}
 
@@
 
-\paragraph*{S2 Text.}
-\label{S2Text}
-{\bf Model setup and sensitivity analysis.}
+\subsection*{S2 Text: Statistical models, generative model setup and sensitivity analysis.\label{S2Text}}
 
-Details on the simulation model: summary statistics for the socio-economic distance matrix, procedure for the synthetic model setup and detailed Global Sensitivity Analysis.
+Additional zero-inflated and hurdle statistical models; details on the simulation model: summary statistics for the socio-economic distance matrix, procedure for the synthetic model setup and detailed Global Sensitivity Analysis.
 

### 2022-03-28 Response_to_reviewers.tex
Update on Overleaf.

@@
 
-$\rightarrow$ The map in Figure 2 was redrawn using countries from Natural Earth Data.
+$\rightarrow$ The map in Figure 2 was redrawn using countries from Natural Earth Data (public domain).
 
@@
 	\item \textit{Given that the model is generative, the temporal aspect of the model can be an issue.
-Indeed, if the number of iterations (around 5000) needed to generate the empirical network correspond to a long time period (for example decades), the economic size and sectoral proximity are expected to change, probably implying a substantially different network, if these changes were accounted for in the model.
+Indeed, if the number of iterations (around 5000) needed to generate the empirical network correspond to a long time period (for example decades), the economic size and sectorial proximity are expected to change, probably implying a substantially different network, if these changes were accounted for in the model.
 Can the authors mention (and may be discuss) this issue?}
@@
 Effective time simulated by the model can be set by modifying both $t_f$ (number of steps) and $w_0$ parameters.
+The model assumes that context remains constant, so that medium-length periods can be captured.
 This was better explained in the model description.
@@
 
-	$\rightarrow$ \todo{(JR) compute more empirical indicators and compare  with generated networks}
+	$\rightarrow$ We added the computation of empirical macroscopic indicators (internationalisation and metropolisation), and how the model gets close to these through the exploration.
+The calibration at the link level (mean square errors) remains the most relevant for discussion - this was further discussed.
+	%\todo{(JR) compute more empirical indicators and compare  with generated networks}
 
@@
 
-	$\rightarrow$ \todo{(All) paper positioning}
+	$\rightarrow$ The paper positioning and goal were reframed.
+The purpose is to introduce a generative model to understand how different geographical and economic factors influence the growth of urban firm networks, with both potentialities for theoretical explanations (through exploration and calibration) and practical applications (the latest being a proof-of-concept in this work).
+	%\todo{(All) paper positioning}
 
@@
 
-	$\rightarrow$ \todo{(All) justify proof of concept - add literature}
+	$\rightarrow$ Echoing a remark by the second referee, the model application is not the main focus of the paper, but a proof-of-concept illustration for possible developments.
+This was clarified in paper positioning.
+	%\todo{(All) justify proof of concept - add literature}
 

### 2022-03-28 main.tex
Update on Overleaf.

@@
 Understanding the drivers behind the growth of such networks -- in particular urban firm networks --, is essential for the economic resilience of urban systems.
-We introduce in this paper a generative model for firm networks at the urban area level including several complementary processes: the economic size of urban areas at origin and destination of links, industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distance between urban areas.
+We introduce in this paper a generative model for firm networks at the urban area level, which considers the following features: the economic size of the areas representing the origin and destination nodes, the industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distance between urban areas.
 An empirical network analysis on European urban firm ownership data confirms the relevance of each of these factors on new link creation.
@@
 This can be achieved through strategic development of certain industrial sectors for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
-Exploring city/firm interactions \cite{storme2019introducing} and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.
+Exploring city/firm interactions \cite{storme2019introducing} and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future unforeseen exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.
 
-The use of network models to understand processes underlying the emergence and dynamics of such networks has already been tackled in the literature, but still not extensively and never explicitly at city level \cite{taylor2001specification}.
+Network models to understand processes underlying the emergence and dynamics of such networks have been proposed in the literature, nevertheless, these are not extensive and do not explicitly model the city level \cite{taylor2001specification}.
+At regional scale, generative urban network models combining geographical factors with network topological properties have been proposed \cite{dai2016simulating}.
+At country level, trade networks have been extensively studied from the complex networks perspective \cite{garlaschelli2005structure}, but these are generally not explored at a more granular level due to the lack of city-level data.
+Within regional science, input-output models are used as generative urban models of flows between geographical areas \cite{jin1993generation}.
+In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for different types of flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2018indirect} at different geographical scales.
+
+%aca: change
+
+% add 3 mains contributions to literature of this paper 1.
+approach by cities and not regions, approach by cities and not firm to firm links only 2.
+manière dont on modélise le poids avec le cobb-douglas pas avec le GDP, mais avec les % owernship links - vérifier ce que les économiustes utilisent pour modéliser les flux 3.
+modèle génératif et dynamique
+In this paper we choose to focus on interurban interactions between firms.
+Firm networks remains an central aspect of network and system of cities theory \cite{RozenblatNeal2021} from the economic perspective.
+They have already been investigated in complexity science at city level to unveil agglomeration economies and competitiveness of cities \cite{Rozenblat2010}, regionalisation patterns \cite{Rozenblatetal2017} or to model hierarchy and specialisation \cite{BidaRozenblat2020}.
+Moreover economic complexity indicators have been developed in this sense to characterise the production structure and specialisation of different countries due to economic activity, including those of trade firms \cite{HaussmanHidalgo2009,OCleary2016}.
+
+Following this literature the present paper contributes to question and capture geographical and economic processes as internationalisation and metropolisation in terms of firm networks within a generative model.
+The specific question we tackle is the complementarity of different processes to generate such urban networks.
+The originality of the present paper is to examine these issues within the system of cities \cite{berry1964cities} by following previous literature the field of urban firm networks \cite{Rozenblatetal2017}.
+In this approach the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}.
+We examine the interactions of European cities within firm linkages defined by corporate ownership links.
+Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2020Zdanowska}.
+% regionalisation and specialisation of cities
+% aca: to here
+
+%The use of network models to understand processes underlying the emergence and dynamics of such networks has already been tackled in the literature, but still not extensively and never explicitly at city level \cite{taylor2001specification}.
 In fact generative urban network models combining geographical factors with network topological properties exist at the regional scale {\cite{dai2016simulating}}.
@@
 
-In this paper we choose to focus on interurban interactions between firms considering the preferential attachment paradigm in a network \cite{BarabasiAlbert1999} and spatial interaction model properties % add a paragraph with the most important literature on spatial interaction models and preferential attachment in network and how it can be linked to firm networks.
-Firm networks remains an important aspect of network and system of cities theory \cite{RozenblatNeal2021}.
-They have already been investigated in complexity science at city level to unveil agglomeration economies and competitiveness of cities \cite{Rozenblat2010}, regionalisation patterns \cite{Rozenblatetal2017} or to model hierarchy and specialisation \cite{BidaRozenblat2020}.
-Moreover economic complexity indicators have been developed in this sense to characterise the production structure and specialisation of different countries due to economic activity, including those of trade firms \cite{HaussmanHidalgo2009,OCleary2016}.
-
-Following this literature the present paper positions itself at the intersection of global networks theoretical framework, spatial interaction models and preferential attachment.
-It contributes to question and capture geographical and economic processes as internationalisation, metropolisation, regionalisation and specialisation of cities in terms of firm networks within a generative model.
-The specific question we tackle is the complementarity of different processes to generate such urban networks.
-The originality of the present paper is to examine these issues within the system of cities \cite{berry1964cities} by following previous literature the field of urban firm networks \cite{Rozenblatetal2017}.
-In this approach the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}.
-We examine the interactions of European cities within firm linkages defined by corporate ownership links.
-Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2020Zdanowska}.
+% Positioning a preciser
 
@@
 
-The rest of this paper is organized as follows: we firstly describe the generative model for urban networks and proceed to an empirical analysis of the network used as a case study.
-We then proceed to a sensitivity analysis and exploration of the model on synthetic data.
-The model is calibrated on empirical European firm ownership data.
+%aca: review and extend this
+The rest of this paper is organized as follows: we firstly describe the generative model for urban networks and then present an empirical analysis of the network used as a case study.
+We then perform a sensitivity analysis and exploration of the model parameters using synthetic data.
+The model is then calibrated on empirical European firm ownership data.
 We finally discuss implications of our results and possible extensions.
@@
 \subsubsection*{Rationale}
-
+%aca: change
 The model is expected to capture a single macroscopic urban scale, i.e.
@@
 As the authors of \cite{martinus2018global} point out, a combination of several distances may play a role in establishing linkages.
-Geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}.
-Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalised to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (confirmed relevant by the fixed effect regression described below); (iii) city economic size, which presents agglomeration effects; (iv) industrial proximity in terms of firm sector composition; and (v) previous linkages (influence of already existing links in the past).
+%aca: to here
+Geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}. Therefore, we combine several features %processes 
+driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalised to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (this is confirmed to be  relevant by the fixed effect regression described below); (iii) city economic size, which is known to give rise to agglomeration effects; (iv) industrial proximity in terms of firm sector composition; and (v) previous linkages, which capture the influence of existing linkages created in the past.
 This last point is crucial to capture accumulated effects and include path-dependencies, justifying further the use of a simulation model instead of a more simple statistical model.
 
-It is worth mentioning that our model is not aimed to be predictive, but rather designed as an explanatory model in order to disentangle the influence of diverse processes in the growth of the urban network and to test diverse hypothesis \cite{conte2012manifesto}.
-It also tends to be a decision-making model as we explore potential policy applications.
+It is worth mentioning that our model is not aimed to be predictive, but rather designed as an explanatory model in order to disentangle the influence of diverse processes and drivers in the growth of the urban network, and to test different hypotheses \cite{conte2012manifesto,varenne2018models}.
+The model can also be used a decision-making model, where different scenarios for potential policy applications can be explored. 
+%It also tends to be a decision-making model as we explore potential policy applications.
 Our focus on explanation rather than prediction can be understood in a context of a broad diversity of possible model functions \cite{varenne2018models}.
@@
 \end{equation}
-where $E  =  \sum_k E_k$, $W  = \sum_{i,j} w_{ij}$ weights of previous links, $s(S_{ik},S_{jk})$ is a similarity function between activity sectors $S_{ik}$ and $S_{jk}$ taken as a cosine similarity, $d_{ij}$ is the Euclidian distance, and $c_{ij}$ is a socio-cultural distance.
+where $E  =  \sum_k E_k$; $w_{ij}$ are the weights of previous links, and $W  = \sum_{i,j} w_{ij}$; $s(S_{ik},S_{jk})$ is a similarity function between activity sectors $S_{ik}$ and $S_{jk}$, taken to be the cosine similarity; $d_{ij}$ is the Euclidian distance; and $c_{ij}$ is a socio-cultural distance.
 In the case of an already existing link between two areas, the weight of the latter is incremented by a unit volume $w_0$.
-The model is stopped either when a final time is reached or when the cumulated volume of links reaches a target quantity.
-This formulation can be considered both as an economic utility model and a generalisation of spatial interaction models.
+The model is stopped either when a final time $t_f$ is reached or when the cumulative volume of links reaches a target quantity.
+Therefore, the duration of the time period simulated can be adjusted by changing either $t_f$ or $w_0$ parameter, given a total volume.
 
-Note that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalise the exponents to 1 to ensure to include convex functions.
+This model formulation can be considered both as an economic utility model and a generalisation of spatial interaction models.
+We must furthermore precise that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalise the exponents to 1 to ensure to include convex functions.
 The ``geopolitical/sociocultural'' distance remains abstract and should be parametrised (see setup below) and estimated in the data-driven setting.
-In this simple version of the model we assume an absence of evolution of economic size, which means that the adjusted parameters are valid on a restricted time period.
+In this simple version of the model we assume an absence of evolution of economic size or sectorial proximity, which means that the adjusted parameters are valid on a restricted time period.
 
@@
 
-In synthetic experiments, the network is generated from an initial complete network with dummy links of weight 1 until reaching $t_f=1500$, with a unit link volume of $w_0 = 1$.
+In synthetic experiments, the network is generated from an initial complete network with dummy links of weight $1$ until reaching $t_f=1500$, with a unit link volume of $w_0 = 1$.
+% missing links are not constrained
 
@@
 Metropolises playing the role of gateways of globalisation, are competing between each other to attract the most qualified and dynamic activities \cite{RozenblatPumain2007}.
-Metropolisation gradients have been already explored in literature in city-systems considering several economic and political aspects \cite{RozenblatPumain2018} \cite{PumainRozenblat2019} \cite{Zdanowskaetal2020}.
+Metropolisation gradients have been linked in the literature on systems of cities with several economic and political aspects \cite{RozenblatPumain2018} \cite{PumainRozenblat2019} \cite{Zdanowskaetal2020}.
 We introduce in this paper a metropolisation indicator defined as the correlation between the weighted degree and the GDP of the city, in order to capture the effect of large cities or metropolises of regional importance attracting the biggest amount of links and involved in high interactions with other cities.
@@
 
-We base our study on empirical data from \emph{ADAMDEUS} dataset produced by the Bureau Van Dijk (BVD) on transnational firms in Europe.
-AMADEUS indicates the exact geolocation, turnover and activity codes for owners and owned companies in all sectors in Europe.
+We base our study on empirical data from the \emph{AMADEUS} dataset produced by the Bureau Van Dijk (BVD) on transnational firms in Europe.
+\emph{AMADEUS} provides a precise location, turnover and activity codes for owners and owned companies in all sectors in Europe.
 The Bureau Van Dick gathers, homogenizes and completes the information on these firms from national sources, publications of consultancies and audit firms such as KPMG, Deloitte, PricewaterhouseCoopers (PwC), Ernst \& Young (EY), Mazars, and those of other financial organisations such as Moody's, Fitch, Standard \& Poor's, Thomson Reuters or Duns and Bradstreet.
@@
 As we aggregate company turnovers on urban areas in our analysis, we thus do not expect much representativity bias in the inter-urban network.
-% \comment[JR]{To what database is Orbis compared to in this study - what is the confidence on its exhaustivity}
+% To what database is Orbis compared to in this study - what is the confidence on its exhaustivity
 
@@
 The area of circles gives the turnover within the Functional Urban Area, and colour the community it belongs to.
-Urban Areas in communities of size smaller than 3 were not coloured for readability.}
+Urban Areas in communities of size smaller than 3 were not coloured for readability.
+Made with Natural Earth Data \cite{naturalearth}.}
     \label{fig:fig2}
@@
 
-% TODO redo map with non licensed data for countries: https://www.naturalearthdata.com/downloads/110m-cultural-vectors/
+Model macroscopic indicators of internationalisation and metropolisation can also be computed on the empirical network.
+Internationalisation corresponds to the modularity of countries in the network, which was computed above as $I = 0.32$.
+Metropolisation computed with turnover has a very high empirical value $M = 0.961$ (95 \% confidence interval $\left[ 0.954 ; 0.967\right]$).
+A variation of this indicator computed with GDP from the GHSL database is not as high but still a relatively high correlation ($0.776 \left[ 0.736 ; 0.811\right]$).
+In practice, this European urban network is thus highly internationalised and metropolised, following what can be expected from the empirical literature on urban networks.
 
@@
 The number of time steps $t_f$ is left as an additional parameter, which in a sense determines the granularity of the cumulative network generation process.
-The objective functions for calibration are the average mean squared error on logarithms of weights given by $\varepsilon_L = \frac{1}{N^2} \sum_{i,j} \left(\log w_{ij} - \log \hat{w}_{ij} \right)^2$ if $\hat{w}_{ij}$ are the simulated weights, and the logarithm of the mean squared error $\varepsilon_M = \log\left(\frac{1}{N^2} \sum_{i,j} \left(w_{ij} - \hat{w}_{ij}\right)^2 \right)$.
-\cite{raimbault2018indirect} showed that these two are complementary to calibrate urban systems models.
+The model is fitted at the microscopic level on link structure: the objective functions for calibration are the average mean squared error on logarithms of weights given by
+
+\begin{equation}
+\varepsilon_L = \frac{1}{N^2} \sum_{i,j} \left(\log w_{ij} - \log \hat{w}_{ij} \right)^2
+\end{equation}
+
+where $\hat{w}_{ij}$ are the simulated weights, and the logarithm of the mean squared error
+
+\begin{equation}
+\varepsilon_M = \log\left(\frac{1}{N^2} \sum_{i,j} \left(w_{ij} - \hat{w}_{ij}\right)^2 \right)
+\end{equation}
+
+\cite{raimbault2018indirect} showed that these two are complementary to calibrate urban systems models, capturing the overall performance but also how the model behaves on medium-sized and small links.
+A calibration on macroscopic indicators of metropolisation and internationalisation would not bring much knowledge on model processes, since several points from the grid experiment above - not designed for this - already provide a good performance (19 parameter points out of 6534 are simultaneously below a relative error of 10\% on internationalisation and below 10\% on metropolisation).
+% TODO in progress: test macro calib ; rerun explo metropolisation? -> ok
+
 Calibration is done using a genetic NSGA2 algorithm \cite{deb2002fast}, which is particularly suited for such a bi-objective calibration of a simulation model.
@@
 	\caption{\textbf{Model calibration results.} Pareto front obtained for the bi-objective model calibration for $\varepsilon_L$ and $\varepsilon_M$ objectives.
-$\varepsilon_L$ can be compared to the MSE of statistical models fitted above, and the generative model outperform these for several parametrizations (lowest MSE 5.33 for statistical models).
+$\varepsilon_L$ can be compared to the MSE of statistical models fitted above, and the generative model outperform these for several parametrisations (lowest MSE 5.33 for statistical models).
 Point size gives the number of stochastic samples and color the value of $\gamma_O$.\label{fig:fig8}}
@@
 Besides these stylized facts, these calibration results altogether demonstrate the feasibility of model application to real data.
-
-% TODO recall not a predictive model
+As stated previously, the model is not predictive and the relatively high error values would not make it satisfactory in that manner, but are reasonable for its explanatory and application functions.
 
@@
 
-This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMole platform and calibrated on real data on ownership linkages of firms for Europe from the Amadeus dataset.
+This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMOLE platform and calibrated on real data on ownership linkages of firms for Europe from the Amadeus dataset.
 The simulation on a synthetic system of cities provides a broad knowledge on model behavior in its parameter space.
@@
 In this sense the UK economic sectors mostly dependent on foreign capital and generating the highest turnovers in 2018, that would be the most affected and worth examining in details are: extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods (in terms of out-going links from the UK) ; activities of head offices and holding companies, sale of cars and manufacture of motor vehicles (in terms of in-going links to the UK).
-We showed in our experiment that although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalization and thus can be expected to have a detrimental effect on UK economy due to these foreign ownership links (see also S1 Text).
-% ~\ref{S1Text} ref not working
+We showed in our experiment that although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalization and thus can be expected to have a detrimental effect on UK economy due to these foreign ownership links (see also S1 Text\ref{S1Text}).
+The impact of such unforeseen economic shocks could also be considered from a multi-level perspective, coupling our work with other approaches such as \cite{Dorretal2022} studying the firm-level impact of the Covid crisis.
 
 Several possible extensions of the current work can be undertaken in the future.
+A simple extension with additional processes could include for example: (i) self-connections to account for links within FUAs; (ii) edge deletion which would capture the sell of subsidies to other companies; (iii) a precise study of the role of path dependency should also be undertaken in order to anticipate the creation of future links; (iv) other formulations of the generative model might be taken into consideration as other approaches to combine factors such as multi-objective optimisation depending on sectors using Pareto fronts; (v) allow for a dynamical calibration with a more detailed realistic representation of temporal dynamics, accounting for temporal non-stationarity for example.
+
+We also suggest that our simple agent-based model can be a basis for or a component of more complicated economic agent-based models.
 A co-evolution model with an evolution of city sizes in parallels to the firm network would be more realistic for an application on longer time scales, but would also be more difficult to parametrise and calibrate \cite{raimbault2018characterising}.
 A model with additional firm agents would lead to a multi-scale agent-based model, with also more flexibility for policy applications.
-A precise study of the role of path dependency should also be undertaken in order to anticipate the creation of future links.
-Other formulations of the generative model might be taken into consideration as other formulation of the combination of factors or multi-objective optimisation depending on sectors using Pareto fronts.
-Regarding the statistical analysis, no spatial or neighborhood effects were taken into account, and spatial non-stationarity may be accounted for using Geographically Weighted Regression models for example \cite{comber2021route}.
+The link with data could also be more elaborated: indeed the statistical analysis does not include spatial or neighborhood effects, and spatial non-stationarity may be accounted for using Geographically Weighted Regression models for example \cite{comber2021route}, to then be coupled with the ABM in an hybrid approach.
 Finally, an adaptation at the world scale using for example urban regions and data explored by \cite{rozenblat2021intra}, would be an extension with important practical and theoretical potentialities.
 
-%one can also think about multi-level impact of economic unforeseen shock following \cite{Dorretal2022} study on firm firm-level impact of coronavirus.
-% TODO towards dynamical graph generative model, to allow for dynamical calibration (discuss possibilities of dyn calib? ; no edge deletion in the model: also discuss?)
-
+% towards dynamical graph generative model, to allow for dynamical calibration (discuss possibilities of dyn calib?
+%no edge deletion in the model: also discuss?)
 % Discuss possible extension with self-connection
@@
 
-Additional zero-inflated and hurdle statistical models; details on the simulation model: summary statistics for the socio-economic distance matrix, procedure for the synthetic model setup and detailed Global Sensitivity Analysis.
+Details on the empirical analysis and the simulation model:
+
+\begin{itemize}
+    \item additional zero-inflated and hurdle statistical models;
+    \item summary statistics for the socio-economic distance matrix;
+    \item procedure for the synthetic model setup;
+    \item Global Sensitivity Analysis
+\end{itemize}
 

### 2022-03-30 main.tex
Update on Overleaf.

@@
 
-A first numerical experiment of one-factor sampling on all Cobb-Douglas parameters and 100 stochastic repetitions confirms a good statistical convergence (average Sharpe ratios for indicators all larger than 5).
+A first numerical experiment of one-factor sampling on all Cobb-Douglas parameters and 100 stochastic repetitions confirms a good statistical convergence (average Sharpe ratios for indicators are all larger than 5).
 We use thus 20 repetitions in following larger experiments (in the case of a normal distribution, a 95\% confidence interval on the average is of size $2\sigma \cdot 1.96 / \sqrt{n}$, what gives $n=16$ repetitions for a CI of width of the standard deviation).
@@
 The sector proximity parameter $\gamma_S$ influences the internationalisation linearly, but induces a maximum for the correlation between size and degree, which can be interpreted as a setting where size and the total volume in a city are the most correlated.
+% TODO comment max metropolisation
 
@@
 Both plots show a transition from a local to a global regime.
-(Bottom Left) Internationalisation varies linearly with sector proximity $\gamma_S$; (Bottom Right) Correlation between degree and size exhibits a maximum, witnessing an intermediate regime where size is the most important.
+(Bottom Left) Internationalisation varies linearly with sector proximity $\gamma_S$; (Bottom Right) Metropolisation exhibits a maximum, witnessing an intermediate regime where size is the most important.
 \label{fig:fig4}}
@@
 
+% TODO JR redo comment
 Other indicators exhibit non-trivial patterns.

### 2022-03-31 CoverLetter.tex

@@
 We do not have any opposed reviewer.
-This paper is submitted to the PloS ONE collection ``Cities as Complex Systems'' edited by M.
+This paper was initially submitted to the PloS ONE collection ``Cities as Complex Systems'' edited by M.
 Gonzalez (University of California, Berkeley) and D.
 Rybski (Potsdam Institute for Climate Impact Research).
-We know that the official deadline has passed, nevertheless, we hope that it could still be considered for this special issue.
+We know that the official deadline has passed, nevertheless, we hope that it could still be considered for a non-special issue.
 

### 2022-03-31 figures.sh
new_fig_alpha

@@ -24,7 +24,7 @@ cp figuresraw/internationalization_countryGravityDecay200_gammaDestination0_face
 cp figuresraw/networkAvgCommunitySize_countryGravityDecay2100_gammaDestination0_facetwrapgammaOrigin_colorgammaSectors.png figures/Fig6.png
 
 # Fig 7
-montage figuresraw/internationalization-alpha_colorgravityDecay_errorBar.png figuresraw/rhoDegreeSize-alpha_colorgravityDecay_errorBar.png -tile 2x1 -geometry +"$HORIZONTALPADDING"+0 -resize "$WIDTH"x figures/Fig7.png
+montage figuresraw/internationalization-alpha_colorgravityDecay_errorBar.png figuresraw/metropolization-alpha_colorgravityDecay_errorBar.png -tile 2x1 -geometry +"$HORIZONTALPADDING"+0 -resize "$WIDTH"x figures/Fig7.png
 
 # Fig 8
 cp figuresraw/pareto_colorgammaOrigin.png figures/Fig8.png

### 2022-05-12 Response_to_reviewers.tex
.
Update on Overleaf.

@@
-%% start of file `template.tex'.
-%% Copyright 2006-2013 Xavier Danaux (xdanaux@gmail.com).
-%
-% This work may be distributed and/or modified under the
-% conditions of the LaTeX Project Public License version 1.3c,
-% available at http://www.latex-project.org/lppl/.
-
 \documentclass[10pt,a4paper,sans]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')
@@
 
-	% (All) discuss possible extension with self-connex
 	$\rightarrow$ Self-connections within urban areas are not considered in the current version of the model. We added in the discussion this possibility and issues related to such an extension.
@@
 
-	% TODO (All) model description
-	$\rightarrow$
+	$\rightarrow$ There was a lack of clarity in the description of the link addition process: the provided function is an utility and the probability is computed accordingly by renormalising it.
+A single link is updated at each time step.
+This was reformulated.
 
@@
 
-	% (JR) other stat models (see above)
 	$\rightarrow$ This remark converges with one point raised by the second referee (see above) - more statistical models were tested.
@@
 
-	% TODO model description: directed network
-	$\rightarrow$
+	$\rightarrow$ This was added in model description.
 
@@
 
-	$\rightarrow$ \todo{(All) model description}
+	$\rightarrow$ This was corrected using a matrix notation for industrial structure.
 
@@
 	\item \textit{"Model 4 also indicates the absence of overfitting": Can the authors be more precise?}
-	% \todo{(JR) clarify}
 
@@
 
-	$\rightarrow$ \todo{(All) model description}
+	$\rightarrow$ This additional indicator was added in model description.
 
@@
 The purpose is to introduce a generative model to understand how different geographical and economic factors influence the growth of urban firm networks, with both potentialities for theoretical explanations (through exploration and calibration) and practical applications (the latest being a proof-of-concept in this work).
-	%\todo{(All) paper positioning}
 
@@
 
-	$\rightarrow$ \todo{(All) paper writing (ex jargon, clarity)}
+	$\rightarrow$ Specific terms were better defined and the paper was screened for clarity.
 
@@
 
-	$\rightarrow$ \todo{(All) improve model description/contextualization}
+	$\rightarrow$ Regionalisation and specialisation indicators where explored in a previous version of the model, but did not seem to bear significant information and were difficult to interpret, so these were removed from model description as they are not used.
+Regarding model setup, an extensive description of the synthetic setup is given in supplementary material.
 
@@
 
-	$\rightarrow$ \todo{(All) Literature on geo-economic factors has been enhanced - processes included}
+	$\rightarrow$ The literature on geo-economic factors has been enhanced, and some references suggesting candidate processes are included in model rationale.
 
@@
 This was clarified in paper positioning.
-	%\todo{(All) justify proof of concept - add literature}
 

### 2022-05-12 S2_Text.tex
Update on Overleaf.

@@
 
-\section*{S2 Text: Model setup and sensitivity analysis}
+\section*{S2 Text: Statistical models, generative model setup and sensitivity analysis}
 
@@
 \end{table}
+
+\subsection*{Zero-inflated models}
+
+We show in Table~\ref{tab:zeroinfl} estimation results for zero-inflated and hurdle statistical models.
+We only show models with all variables included, which had the best AIC values.
+Main model coefficients are very close in both approaches, and do not deviate largely from values obtained with the simpler model.
+All coefficients have a high significance in this case.
+Regarding the additional components, both give the same qualitative outcome (with opposite signs, the hurdle interpretation being the opposite of the zero-inflated).
+Distance increase the likelihood of having a zero flow, while other variables decrease it, with the same order of magnitude of effect size than the main component.
+R-squared values are much lower than simpler model, since a large proportion of the dataset are zeros and the models are thus harder to fit when including them.
+
+%%%%%%%%%%%%%
+\begin{table}[!ht]
+\begin{adjustwidth}{-1in}{0in}
+\caption{{\bf Zero-inflated and Hurdle statistical models.}\label{tab:zeroinfl}}
+\medskip
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+Model  & \multicolumn{2}{|c|}{Zero-inflated} & \multicolumn{2}{|c|}{Hurdle} \\ 
+ & Count & Zero-infl. & Count & Zero-hurdle \\
+\hline
+$\log(d_{ij})$ &    -0.28*** (7e-4) &   0.88*** (0.01)      &   -0.29*** (8e-6)   &  -0.75*** (0.01)    \\
+$\log(T_i)$ & 0.74*** (4e-6)  & -0.74*** (7e-3)       &   0.74***  (5e-6)     &  0.67*** (5e-3)   \\
+$\log(T_j)$ & 0.64*** (4e-6)  & -0.52*** (7e-3)      &    0.64***  (4e-6)     &  0.48*** (5e-3)    \\
+$\log(s_{ij})$ & 0.46*** (2e-5)   &  -0.33*** (3e-2)  &  0.45*** (2e-5)       &  0.27*** (0.02)    \\
+\hline
+$R^2$ &   \multicolumn{2}{|c|}{0.1607}     &  \multicolumn{2}{|c|}{0.1661}  \\
+%MSE & & \\
+AIC &     \multicolumn{2}{|c|}{5.85e10}    &   \multicolumn{2}{|c|}{5.85e10}  \\
+\hline
+\end{tabular}
+\end{adjustwidth}
+\end{table}
+%%%%%%%%%%%%%
+
+%pscl::hurdle(formula = intweight ~ log(distance) + log(from_turnover) + log(to_turnover) + 
+%    log(sim), data = dz, dist = "poisson", zero.dist = "poisson")
+%
+%Pearson residuals:
+%       Min         1Q     Median         3Q        Max 
+%-9.195e+03 -1.373e-01 -7.627e-02 -4.204e-02  1.727e+04 
+%
+%Count model coefficients (truncated poisson with log link):
+%                     Estimate Std. Error z value Pr(>|z|)    
+%(Intercept)        -9.853e+00  1.345e-04  -73252   <2e-16 ***
+%log(distance)      -2.875e-01  7.603e-06  -37812   <2e-16 ***
+%log(from_turnover)  7.379e-01  4.708e-06  156720   <2e-16 ***
+%log(to_turnover)    6.379e-01  4.060e-06  157119   <2e-16 ***
+%log(sim)            4.550e-01  2.137e-05   21292   <2e-16 ***
+%Zero hurdle model coefficients (censored poisson with log link):
+%                     Estimate Std. Error z value Pr(>|z|)    
+%(Intercept)        -20.525823   0.155298 -132.17   <2e-16 ***
+%log(distance)       -0.747788   0.010991  -68.04   <2e-16 ***
+%log(from_turnover)   0.665678   0.005630  118.23   <2e-16 ***
+%log(to_turnover)     0.479214   0.005707   83.96   <2e-16 ***
+%log(sim)             0.278290   0.022439   12.40   <2e-16 ***
+%---
+%Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1 
+%
+%Number of iterations in BFGS optimization: 46 
+%Log-likelihood: -2.927e+10 on 10 Df
+
+% 0.1661997
+
+%mean((dz$intweight-hurdle2$fitted)^2)
+%[1] 6.119694e+12
+
+% > AIC(hurdle2)
+%[1] 58545461912
+
+%pscl::zeroinfl(formula = intweight ~ log(distance) + log(from_turnover) + log(to_turnover) + 
+%    log(sim), data = dz, dist = "poisson")
+%
+%Pearson residuals:
+%       Min         1Q     Median         3Q        Max 
+%-7.693e+00 -1.335e-01 -6.957e-02 -3.591e-02  1.760e+04 
+%
+%Count model coefficients (poisson with log link):
+%                     Estimate Std. Error z value Pr(>|z|)    
+%(Intercept)        -9.853e+00  1.345e-04  -73251   <2e-16 ***
+%log(distance)      -2.875e-01  7.603e-06  -37810   <2e-16 ***
+%log(from_turnover)  7.379e-01  4.708e-06  156719   <2e-16 ***
+%log(to_turnover)    6.379e-01  4.060e-06  157120   <2e-16 ***
+%log(sim)            4.550e-01  2.137e-05   21294   <2e-16 ***
+%
+%Zero-inflation model coefficients (binomial with logit link):
+%                    Estimate Std. Error z value Pr(>|z|)    
+%(Intercept)        22.046685   0.180605  122.07   <2e-16 ***
+%log(distance)       0.881724   0.013134   67.13   <2e-16 ***
+%log(from_turnover) -0.737441   0.006776 -108.83   <2e-16 ***
+%log(to_turnover)   -0.521642   0.006540  -79.76   <2e-16 ***
+%log(sim)           -0.330340   0.025360  -13.03   <2e-16 ***
+%---
+%Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1 
+%
+%Number of iterations in BFGS optimization: 1 
+%Log-likelihood: -2.927e+10 on 10 Df
+
+% 0.1607349
+
+%mean((dz$intweight-zinflpoisson2$fitted)^2)
+%[1] 6.159803e+12
+
+% AIC(zinflpoisson2)
+%[1] 58545461691
 
@@
 
-\section*{Sensitivity analysis}
+\subsection*{Sensitivity analysis}
 
@@
 \hline
- & \multicolumn{2}{|c|}{$\gamma_G$} & \multicolumn{2}{|c|}{$\gamma_D$} & \multicolumn{2}{|c|}{$\gamma_S$} & \multicolumn{2}{|c|}{$\gamma_W$} & \multicolumn{2}{|c|}{$\gamma_O$} & \multicolumn{2}{|c|}{$\gamma_D$} \\
+ & \multicolumn{2}{|c|}{$d_0$} & \multicolumn{2}{|c|}{$c_0$} & \multicolumn{2}{|c|}{$\gamma_S$} & \multicolumn{2}{|c|}{$\gamma_W$} & \multicolumn{2}{|c|}{$\gamma_O$} & \multicolumn{2}{|c|}{$\gamma_D$} \\
  & F & T & F & T & F & T & F & T & F & T & F & T \\

### 2022-05-12 main.tex
Update on Overleaf.

@@
 \\
-Juste Raimbault\textsuperscript{1,2,3,4 *},
-Natalia Zdanowska\textsuperscript{5,2,4},
-Elsa Arcaute\textsuperscript{2}
+Juste Raimbault\textsuperscript{1,2,3 *},
+Natalia Zdanowska\textsuperscript{3,4,5},
+Elsa Arcaute\textsuperscript{4}
 \\
@@
 \\
-\textbf{2} Centre for Advanced Spatial Analysis, UCL, London, United Kingdom
+\textbf{2} UPS CNRS 3611 ISC-PIF, Paris, France
 \\
-\textbf{3} UPS CNRS 3611 ISC-PIF, Paris, France
+\textbf{3} UMR CNRS 8504 G{\'e}ographie-cit{\'e}s, Paris, France
 \\
-\textbf{4} UMR CNRS 8504 G{\'e}ographie-cit{\'e}s, Paris, France
+\textbf{4} Centre for Advanced Spatial Analysis, UCL, London, United Kingdom
 \\
-\textbf{5} Luxembourg Institute of Socio-Economic Research, Luxembourg
+\textbf{5} Luxembourg Institute of Socio-Economic Research, Esch-sur-Alzette, Luxembourg
 
@@
 We calibrate the model on the European network, outperforming statistical models showing a strong role of path-dependency.
-The model opens up potential applications to formulate public policies recommendations to deal with exogenous shocks such as economic crisis or potential lockdowns of countries.
+The model opens up potential applications to formulate public policies recommendations to deal with exogenous shocks such as economic crisis or potential lock-downs of countries.
 This is illustrated with an application on stylised scenarios.
@@
 Understanding the drivers of growth and geographical properties of such an urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}.
-In addition, strengthening these city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}.
+In addition, strengthening city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}.
 This can be achieved through strategic development of certain industrial sectors for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
@@
 
-%aca: change
-
-% add 3 mains contributions to literature of this paper 1.
-approach by cities and not regions, approach by cities and not firm to firm links only 2.
-manière dont on modélise le poids avec le cobb-douglas pas avec le GDP, mais avec les % owernship links - vérifier ce que les économiustes utilisent pour modéliser les flux 3.
-modèle génératif et dynamique
-In this paper we choose to focus on interurban interactions between firms.
-Firm networks remains an central aspect of network and system of cities theory \cite{RozenblatNeal2021} from the economic perspective.
-They have already been investigated in complexity science at city level to unveil agglomeration economies and competitiveness of cities \cite{Rozenblat2010}, regionalisation patterns \cite{Rozenblatetal2017} or to model hierarchy and specialisation \cite{BidaRozenblat2020}.
+The first main contribution of this paper is to focus on interurban interactions between firms.
+Firm networks remains a central aspect of network and system of cities theory \cite{RozenblatNeal2021} from the economic perspective.
+In this approach the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary} \cite{berry1964cities}.
+An extensive literature exists to explore firm-to-firm level networks in economic geography withing the input-output models framework in order to study inter-sectors and inter-region transactions \cite{Rodriguesetal2016} or firm level productivity to understand macroeconomic level outcomes \cite{Grassi2016} and economic resilience \cite{Ialongo2021}.
+Studies of these firm networks at city level remain still scarce.
+Urban firm networks have been investigated in complexity science to unveil agglomeration economies and competitiveness of cities \cite{Rozenblat2010}, regionalisation patterns \cite{Rozenblatetal2017} or to model hierarchy and specialisation \cite{BidaRozenblat2020}.
 Moreover economic complexity indicators have been developed in this sense to characterise the production structure and specialisation of different countries due to economic activity, including those of trade firms \cite{HaussmanHidalgo2009,OCleary2016}.
 
-Following this literature the present paper contributes to question and capture geographical and economic processes as internationalisation and metropolisation in terms of firm networks within a generative model.
-The specific question we tackle is the complementarity of different processes to generate such urban networks.
-The originality of the present paper is to examine these issues within the system of cities \cite{berry1964cities} by following previous literature the field of urban firm networks \cite{Rozenblatetal2017}.
-In this approach the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}.
-We examine the interactions of European cities within firm linkages defined by corporate ownership links.
-Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2020Zdanowska}.
-% regionalisation and specialisation of cities
-% aca: to here
-
-%The use of network models to understand processes underlying the emergence and dynamics of such networks has already been tackled in the literature, but still not extensively and never explicitly at city level \cite{taylor2001specification}.
+The second contribution of the present study is to question and capture geographical and economic processes as internationalisation and metropolisation of cities in terms of firm networks within a generative model, where weights of cities are given by shares of turnovers generated by ownership links between firms.
+Previous literature considers mostly weights as GDP of countries for example in gravity models of export-import firms networks \cite{Kepaptsogloual2010} \cite{MinondoRequena2013}.
+The use of network models to understand processes underlying the emergence and dynamics of such networks has already been tackled in the literature, but still not extensively and never explicitly at city level \cite{taylor2001specification}.
 In fact generative urban network models combining geographical factors with network topological properties exist at the regional scale {\cite{dai2016simulating}}.
@@
 
-% Positioning a preciser
+The specific question we tackle is the complementarity of different processes to generate urban firm networks.
+We examine the interactions of European cities within firm linkages defined by corporate ownership links.
+Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2020Zdanowska}.
+More precisely : (i) we propose an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database \cite{AMADEUS2018}; (ii) we introduce a generative network model to simulate the growth of such linkages at urban areas scale, which combines multiple factors influencing link formation, such as the economic intensity of the urban areas corresponding to the origin and destination nodes, the industrial sector proximity, the strength of previous links, and the geographical and socio-cultural distance; (iii) the model allows us to compare the effect of different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; and (iv) we calibrate the model on the empirical network.
 
-More precisely, our contribution consists in the following: (i) we propose an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database \cite{AMADEUS2018}; (ii) we introduce a generative network model to simulate the growth of such linkages at urban areas scale, which combines multiple factors influencing link formation, such as the economic intensity of the urban areas corresponding to the origin and destination nodes, the industrial sector proximity, the strength of previous links, and the geographical and socio-cultural distance; (iii) the model allows us to compare the effect of different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; and (iv) we calibrate the model on the empirical network.
-
-%aca: review and extend this
 The rest of this paper is organized as follows: we firstly describe the generative model for urban networks and then present an empirical analysis of the network used as a case study.
@@
 \subsubsection*{Rationale}
-%aca: change
-The model is expected to capture a single macroscopic urban scale, i.e.
+%The model is expected to capture a single macroscopic urban scale, i.e.
 the level of the urban system where basic entities are cities.
@@
 As the authors of \cite{martinus2018global} point out, a combination of several distances may play a role in establishing linkages.
-%aca: to here
-Geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}. Therefore, we combine several features %processes 
-driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalised to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (this is confirmed to be  relevant by the fixed effect regression described below); (iii) city economic size, which is known to give rise to agglomeration effects; (iv) industrial proximity in terms of firm sector composition; and (v) previous linkages, which capture the influence of existing linkages created in the past.
+The model is expected to capture a single macroscopic urban scale, which corresponds to the level at which nodes are cities.
+These are characterised by their economic size and their industrial sector profile.
+%The links in the network are constructed from the relationship between these characteristics and hence induced from their underlying firms.
+Different distances, or proximity measures, play a role in establishing linkages, as pointed out by \cite{martinus2018global}.
+Links are then created considering these features and different distance, or proximity measures \cite{martinus2018global}. 
+Note that, geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}. 
+We therefore combine several features driving network growth: (i) geographical proximity (in our case the crow-fly distance between cities, but it can be generalised to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (this is confirmed to be  relevant by the fixed effect regression described below); (iii) city economic size, which is known to give rise to agglomeration effects; (iv) industrial proximity in terms of firm sector composition; and (v) previous linkages, which capture the influence of existing linkages created in the past.
 This last point is crucial to capture accumulated effects and include path-dependencies, justifying further the use of a simulation model instead of a more simple statistical model.
@@
 It is worth mentioning that our model is not aimed to be predictive, but rather designed as an explanatory model in order to disentangle the influence of diverse processes and drivers in the growth of the urban network, and to test different hypotheses \cite{conte2012manifesto,varenne2018models}.
-The model can also be used a decision-making model, where different scenarios for potential policy applications can be explored. 
+The model can also be used as a decision-making model, where different scenarios for potential policy applications can be explored. 
 %It also tends to be a decision-making model as we explore potential policy applications.
@@
 
+In this work, we do not consider the relationship between countries given their economic complexity, such as the framework developed by  \cite{hidalgo2007product}, where the diversity of products produced plays an essential role in determining co-dependencies and resilience.
+What we investigate instead, is the role of linkages given by geography and history, and hence, the sector structure of cities does not evolve within the model.
+
 \subsubsection*{Model description}
 
-Let $1 \leq i \leq N$ cities be defined at time $t$ by their economic structure such that $E_i(t)$ is the total economic volume (GDP) for city $i$ and $S_{ik}(t)$ is a matrix giving economic volume proportions within each economic sector $k$, assuming there are $K$ economic sectors.
-Formally, the model creates links between cities which are characterized by their economic size $E_i$ (GDP) and economic structure $S_{ik}$ in terms of activity sectors (probability distribution of firms within $K$ sectors).
-Starting with an existing network with no links, the model iteratively adds links, following a probability given by a generalized Cobb-Douglas function \cite{vilcu2011geometric} as
+Let $1 \leq i \leq N$ cities be defined at time $t$ by their economic structure such that $E_i(t)$ is the total economic volume (GDP) for city $i$ and $\mathbf{S_i} = S_{ik}(t)$ is a matrix giving economic volume proportions within each economic sector $k$, assuming that there are $K$ economic sectors.
+Given these characteristics, the model evolves firm links between cities.
+Starting with an existing initial network, the model iteratively adds links or updates their weights to generate a final directed weighted network.
+
+At each time step, one directed link is randomly drawn. We define a generalized Cobb-Douglas utility function \cite{vilcu2011geometric} as
 
 \begin{equation}
-p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_O} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_D} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s\left(S_{ik},S_{jk}\right)^{\gamma_S} \cdot \exp \left(- d_{ij} / d_0\right) \cdot \exp \left(- c_{ij} / c_0\right)
+u_{ij} = \left(\frac{E_{i}}{E}\right)^{\gamma_O} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_D} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s\left(\mathbf{S_{i}},\mathbf{S_{j}}\right)^{\gamma_S} \cdot \exp \left(- d_{ij} / d_0\right) \cdot \exp \left(- c_{ij} / c_0\right)
 \end{equation}
-where $E  =  \sum_k E_k$; $w_{ij}$ are the weights of previous links, and $W  = \sum_{i,j} w_{ij}$; $s(S_{ik},S_{jk})$ is a similarity function between activity sectors $S_{ik}$ and $S_{jk}$, taken to be the cosine similarity; $d_{ij}$ is the Euclidian distance; and $c_{ij}$ is a socio-cultural distance.
-In the case of an already existing link between two areas, the weight of the latter is incremented by a unit volume $w_0$.
-The model is stopped either when a final time $t_f$ is reached or when the cumulative volume of links reaches a target quantity.
-Therefore, the duration of the time period simulated can be adjusted by changing either $t_f$ or $w_0$ parameter, given a total volume.
+where $E  =  \sum_k E_k$; $w_{ij}$ are the weights of previous links, and $W  = \sum_{i,j} w_{ij}$; $s(\mathbf{S_{i}},\mathbf{S_{j}})$ is a similarity function between the distributions of activity sectors, taken to be the cosine similarity; $d_{ij}$ is the Euclidian distance; and $c_{ij}$ is a socio-cultural distance.
 
-This model formulation can be considered both as an economic utility model and a generalisation of spatial interaction models.
-We must furthermore precise that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalise the exponents to 1 to ensure to include convex functions.
+This allows us to define the probability to select one link as $p_{ij} = u_{ij} / \sum_{k,l} u_{kl}$. The weight of this link is incremented by a unit volume $w_0$.
+
+The simulation is stopped either when a final time $t_f$ is reached or when the cumulative volume of links reaches a target quantity.
+Therefore, the duration of the time period simulated can be adjusted by changing either $t_f$ or the $w_0$ parameter, given a total volume.
+
+This model formulation can be considered both as an economic utility model and as a generalisation of a spatial interaction model.
+We must furthermore precise that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (equivalently the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalise the exponents to 1 to ensure the inclusion of convex functions.
 The ``geopolitical/sociocultural'' distance remains abstract and should be parametrised (see setup below) and estimated in the data-driven setting.
-In this simple version of the model we assume an absence of evolution of economic size or sectorial proximity, which means that the adjusted parameters are valid on a restricted time period.
+In this simple version of the model, we do not consider the evolution of either the economic size nor of the sectorial proximity, which means that the adjusted parameters are valid on a limited time period.
 
@@
 
-The synthetic system of cities is generated to be at the scale of a continent with properties similar to Europe: (i) $N=700$ cities with an empirical power-law from the Global Human Settlements database for economic sizes ($\alpha = 1.1$), distributed randomly into a uniform and isotropic square space of width $d_{max}=3000km$;
-(ii) clustered into $C = 20$ countries using a k-means algorithm (this remains consistent with non-correlated city sizes at the scale of the countries as our cities are randomly distributed in space \cite{simini2019testing});
-(iii) with synthetic distribution of sectors following log-normal distributions adjusted in a way that large cities have more high-value activities and are more diverse.
+The synthetic system of cities is generated to be at the scale of a continent with properties similar to Europe: (i) $N=700$ cities with an empirical power-law from the Global Human Settlements database for economic sizes ($\alpha = 1.1$), distributed randomly into a uniform and isotropic square space of width $d_{max}=3000$km;
+(ii) cities are clustered into $C = 20$ countries using a k-means algorithm (this remains consistent with non-correlated city sizes at the scale of the countries as our cities are randomly distributed in space \cite{simini2019testing});
+(iii) their synthetic distribution of sectors follows a log-normal distribution adjusted in a way that large cities have more high-value activities and are more diverse.
 
-More precisely for this last point, assuming sectors are distributed continuously along a one-dimensional axis, cities are assumed to have a sector distribution with a log-normal density function.
-The largest city is constrained to have a standard deviation and a mode of $1/2$ and the smallest $1/K$, with a linear interpolation between the two for other cities.
+More precisely, assuming that sectors are distributed continuously along a one-dimensional axis, cities are assumed to have a sector distribution with a log-normal density function.
+The largest city is constrained to have a standard deviation and a mode of $1/2$ while for the smallest city it will be $1/K$, with a linear interpolation between the two for other cities.
 Parameters of the log-normal law are obtained by solving the corresponding two equations system for each city, and each distribution is discretised and renormalised to be a vector of probabilities.
@@
 
-In synthetic experiments, the network is generated from an initial complete network with dummy links of weight $1$ until reaching $t_f=1500$, with a unit link volume of $w_0 = 1$.
+In the synthetic experiments, the network is generated from an initial complete network with dummy links of weight $1$ until reaching $t_f=1500$, and with a unit link volume of $w_0 = 1$.
 % missing links are not constrained
@@
 As the model also includes path-dependencies by integrating the influence of previous links, it cannot be reduced to a simple statistical model.
-Finally, the role of space introduces even further complexity, yielding a generative model, which has to be simulated.
+Finally, the role of space introduces even further complexity.
+These aspects yield a generative model, in the sense that the successive states have to be simulated \cite{epstein1999agent}.
 
@@
 Metropolises playing the role of gateways of globalisation, are competing between each other to attract the most qualified and dynamic activities \cite{RozenblatPumain2007}.
-Metropolisation gradients have been linked in the literature on systems of cities with several economic and political aspects \cite{RozenblatPumain2018} \cite{PumainRozenblat2019} \cite{Zdanowskaetal2020}.
-We introduce in this paper a metropolisation indicator defined as the correlation between the weighted degree and the GDP of the city, in order to capture the effect of large cities or metropolises of regional importance attracting the biggest amount of links and involved in high interactions with other cities.
-
-The later is computed with a Pearson correlation estimator as
+Metropolisation gradients have been linked in the literature on systems of cities to several economic and political aspects \cite{RozenblatPumain2018} \cite{PumainRozenblat2019} \cite{Zdanowskaetal2020}.
+We introduce in this paper a metropolisation indicator defined as the correlation between the weighted degree and the GDP of the city, in order to capture the effect of large cities, or metropolises of regional importance, attracting the biggest amount of links and involved in a high number of interactions with other cities.
+The later is computed using a Pearson correlation estimator on logarithms as follows:
 \begin{equation}
-    M = \hat{\rho} \left[d_i , E_i \right]
+    M = \hat{\rho} \left[\log (d_i) , \log (E_i) \right]
 \end{equation}
 
-In this work, we do not consider the relationship between countries given their economic complexity, such as the framework developed by  \cite{hidalgo2007product}, where the diversity of products produced plays an essential role in determining co-dependencies and resilience.
-What we investigate instead, is the role of linkages given by geography and history, and hence, the sector structure of cities does not evolve within the model.
+In addition, we also consider an indicator giving insight of the structure of the network: this is the average size of communities obtained with a modularity optimisation algorithm (weighted Louvain algorithm \cite{blondel2008fast}).
+This is complementary to internationalisation, as it captures the endogenous structure of the network and the average size of international business clusters.
 
@@
 
-The urban network is sparse with a density around $0.028$.
+The urban network is sparse with a density of around $0.028$.
 It has heavy tail weighted degree and edge weight distributions, as shown in Fig.~\ref{fig:nwdist}.
@@
 
-Model macroscopic indicators of internationalisation and metropolisation can also be computed on the empirical network.
+Macroscopic indicators of internationalisation and metropolisation can also be computed on the empirical network.
 Internationalisation corresponds to the modularity of countries in the network, which was computed above as $I = 0.32$.
 Metropolisation computed with turnover has a very high empirical value $M = 0.961$ (95 \% confidence interval $\left[ 0.954 ; 0.967\right]$).
-A variation of this indicator computed with GDP from the GHSL database is not as high but still a relatively high correlation ($0.776 \left[ 0.736 ; 0.811\right]$).
+A variation of this indicator computed with GDP from the GHSL database still shows a relatively high correlation ($0.776 \left[ 0.736 ; 0.811\right]$).
 In practice, this European urban network is thus highly internationalised and metropolised, following what can be expected from the empirical literature on urban networks.
@@
 We also fit a Poisson generalised linear model with a logarithmic link on the rounded weights \cite{flowerdew1988fitting}.
-The findings show that when distance only is taken into account (model 1), it has a significant negative effect on link creation, but with an explained variance close to zero.
+The findings show that when only distance is taken into account (model 1), it has a significant negative effect on link creation, but with an explained variance close to zero.
 Model 2 adds fixed effects by pair of countries, for which around a third are significant, and with a better $R^2$ of 0.17.
@@
 
-To investigate the role of missing links in these results (these have been removed since previous models are fitted on logarithms), we furthermore fit extended statistical models for count data focused on the role of zeros.
-Two different flavors of such models are the hurdle model \cite{mullahy1986specification} which corresponds to a conditional count model, and a zero-inflated Poisson model \cite{lambert1992zero} which mixes a count model with an additional component to generate zeros.
-Full estimation results, done using \cite{zeileis2008regression}, are shown in S2 Text.
-As the network is rather sparse, these more detailed statistical models do not perform better than more simple models, as missing links seem rather hard to predict.
-The best zero-inflated model has a $R^2 = 0.16$, while the best hurdle is just above with $R^2 = 0.17$.
+To investigate the role of missing links in these results (these have been removed since previous models are fitted on logarithms), we fit extended statistical models for count data focused on the role of zeros.
+Two different flavors of such models are: the hurdle model \cite{mullahy1986specification}, which corresponds to a conditional count model; and a zero-inflated Poisson model \cite{lambert1992zero}, which mixes a count model with an additional component to generate zeros.
+Full estimation results using \cite{zeileis2008regression}, are shown in S2 Text.
+As the network is rather sparse, these more detailed statistical models do not perform better than more simple models, as missing links are difficult to predict.
+The best zero-inflated model has a $R^2 = 0.16$, and it is similar to the performance of the best hurdle moel $R^2 = 0.17$.
 
-Across all models, consistent qualitative stylised facts regarding link creation between cities are revealed: (i) distance has a negative influence; (ii) economic sizes have a positive effect, with the size of the origin city being more important (consistent in the ownership links context, as strongly driven by decision at the headquarter/owner level); (iii) industrial proximity has a positive influence, i.e.
+Across all models, consistent qualitative stylised facts regarding link creation between cities are revealed: (i) distance has a negative influence; (ii) economic sizes have a positive effect, with the size of the origin city being more important (this is consistent within the ownership links context, as these are strongly driven by decisions at the headquarter/owner level); (iii) industrial proximity has a positive influence, i.e.
 similarity is more important than complementarity for this dimension; (iv) a large part of country fixed-effects are statistically significant; and (v) the order of magnitude of the influence of each factor is roughly the same, which means that no process totally dominates the others.
@@
 Low values of the gravity decay parameter yield local networks around main cities, while long interaction ranges yield more integrated networks.
-We also show networks simulated in the real setting.
+The same results can be observed for the networks generated using the real data in the lower panels.
 % TODO work on data readme on the repo / do a separate dataverse?
@@
 
-We first study extensively the behavior of the model on synthetic systems of cities, to isolate its intrinsic behavior independently of a contingent geographical situation.
+We first study extensively the behaviour of the model on synthetic systems of cities, to isolate its intrinsic behaviour independently of a contingent geographical situation.
 
@@
 This captures the total influence of $X_i$ all other things being equal.
-The total order index is given by $\mathbb{E}_{\mathbf{X}_{\sim i}} \left[Var(Y_j | \mathbf{X}_{\sim i})\right] / Var\left[Y_j\right]$ and captures non-homogeneity in behavior and interaction between parameters.
+The total order index is given by $\mathbb{E}_{\mathbf{X}_{\sim i}} \left[Var(Y_j | \mathbf{X}_{\sim i})\right] / Var\left[Y_j\right]$ and captures non-homogeneity in behaviour and interaction between parameters.
 Indices were estimated with a Sobol sequence sampling of 20,000 model runs \cite{saltelli2010variance}.
@@
 The influence of sectors and historical links has a secondary but not negligible role.
-Altogether, the analysis witnesses of strong interaction effects between parameters, since for several indicators the total order index is one order of magnitude larger than the first order index.
-This means that the generative model captures some complex behavior of the system.
+Altogether, the analysis shows strong interaction effects between parameters, since for several indicators the total order index is one order of magnitude larger than the first order index.
+This means that the generative model captures some aspects of the complex behaviour of the system.
 
@@
 
-A first numerical experiment of one-factor sampling on all Cobb-Douglas parameters and 100 stochastic repetitions confirms a good statistical convergence (average Sharpe ratios for indicators are all larger than 5).
-We use thus 20 repetitions in following larger experiments (in the case of a normal distribution, a 95\% confidence interval on the average is of size $2\sigma \cdot 1.96 / \sqrt{n}$, what gives $n=16$ repetitions for a CI of width of the standard deviation).
-Behavior of some indicators are shown in Fig.~\ref{fig:fig4}.
+A first numerical experiment of one-factor sampling on all Cobb-Douglas parameters and 100 stochastic repetitions confirms a good statistical convergence (the average Sharpe ratios for the indicators are all larger than 5).
+We thus use 20 repetitions in the following larger experiments (in the case of a normal distribution, a 95\% confidence interval on the average is of size $2\sigma \cdot 1.96 / \sqrt{n}$, what gives $n=16$ repetitions for a CI of width of the standard deviation).
+The behaviour of some indicators are shown in Fig.~\ref{fig:fig4}.
 For example, internationalisation and metropolisation both show a transition as a function of d$_o$, from a local to a global regime.
@@
 The model behaviour is then studied with a grid sampling for parameters and 20 repetitions (the computations are run on the European Grid Infrastructure, and are equivalent to 2.5 years of CPU time).
-Distance decay and sector  parameters are varied with 10 steps, other parameters with three.
-The influence of gravity decay parameters is confirmed when plotting the internationalisation index against the distance, which exhibits an exponential decay as shown in Fig.~\ref{fig:fig5}.
-It interacts with the role of sectors and the size of the origin, with a more significant effect of sector proximity, when size has the largest influence (third panel).
+The distance decay and sector parameters are varied with 10 steps, other parameters with three.
+The influence of the gravity decay parameters is confirmed when plotting the internationalisation index against the distance, which exhibits an exponential decay as shown in Fig.~\ref{fig:fig5}.
+We show that the size of the origin and the sector proximity also have a significant effect, although the latter only becomes relevant when the size of the origin has a strong influence (third panel).
+%It interacts with the role of sectors and the size of the origin, with a more significant effect of sector proximity, when size has the largest influence (third panel).
 
@@
     \vspace{2cm}
-    \caption{The transition as a function of interaction range depends on the influence of origin size $\gamma_O$; sector proximity $\gamma_S$ plays a role only for a large influence of the origin.\label{fig:fig5}}
+    \caption{The transition as a function of the interaction range depends on the influence of the origin size $\gamma_O$; the sector proximity $\gamma_S$ plays a role only for a large influence of the origin.\label{fig:fig5}}
 \end{figure}
@@
 Other indicators exhibit non-trivial patterns.
-For example, when considering average community size of the final network as shown in Fig.~\ref{fig:fig6}, we obtain a maximal integration in terms of communities at an intermediate value of the gravity decay.
+For example, when considering the average community size of the final network as shown in Fig.~\ref{fig:fig6}, we obtain a maximal integration in terms of the communities at an intermediate value of the gravity decay parameter.
 This can be interpreted as the emergence of a regional regime.
-The size of communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.
+The size of the communities is largely influenced by the value of the elasticities for the similarity function and the ratio of the economic output of the area.
 In particular, we observe a qualitative inversion of the role of $\gamma_S$ when introducing an effect of the origin (switching $\gamma_O$ from 0 to 1, first panel to second panel).
 The maximal community size disappears when $\gamma_O = 2$, implying a regime with small local communities where large cities mostly connect with their hinterland.
-These experiments reveal a complex interplay between processes and how the model produces diverse stylised facts.
+These experiments show the diverse stylised facts that the model is capable of producing, while at the same time they reveal the complex interplay between the different processes involved.
 
@@
     \vspace{2cm}
-	\caption{Maximal integration in terms of community size is achieved at an intermediate value of d$_o$: emergence of a regional regime; maximal size depends on the role of sectors $\gamma_S$ in a decreasing way when the origin size is deactivated, and in an increasing way when $\gamma_o=1$; this regime disappears when the influence of the origin size becomes too large.\label{fig:fig6}}
+	\caption{Maximal integration in terms of the community size is achieved at an intermediate value of d$_o$: emergence of a regional regime; the maximal size depends on the role of the sectors $\gamma_S$ in a decreasing way when the origin size is deactivated, and in an increasing way when $\gamma_o=1$; this regime disappears when the influence of the origin size becomes too large.\label{fig:fig6}}
 \end{figure}
@@
 
-We run a specific experiment to study the impact of urban hierarchy on firms dynamics.
+% TODO new fig (range?)
+We run a specific experiment to study the impact of the urban hierarchy on firms dynamics.
 Rank-size law in cities is a central stylized fact within urban theories, despite the discussions regarding the estimation of the power-law exponent, its actual values and its variation depending on the definition of cities considered \cite{cottineau2017metazipf,corral2020truncated}.
 In this particular context, understanding how this parameter can impact urban dynamics within artificial systems of cities -- in which it can be modified to extreme values non observable in real systems--, should bring evidence on underlying processes \cite{raimbault2019space}.
-We change the initial hierarchy $\alpha$ from 0.5 to 2.0 with a step of 0.1, and also vary the interaction range $d_0$.
-The behavior of internationalisation and metropolisation indicators are shown in Fig.~\ref{fig:fig7}.
+We change the initial hierarchical exponent $\alpha$ from 0.5 to 2.0 with a step of 0.1, and also vary the interaction range $d_0$.
+The behaviour of the internationalisation and metropolisation indicators are shown in Fig.~\ref{fig:fig7}.
 As expected, the hierarchy of the urban system has an important impact on model outcomes.
-We find that more hierarchical urban systems tend to produce more international networks (internationalisation indicator corresponds to a modularity which is minimized when networks span more across different countries), what is consistent with the existence of globalised metropolis \cite{sassen1991global}.
-This effect is mitigated by lower interaction ranges.
-Regarding metropolisation, we observe a maximum values as a function of initial hierarchy consistent across different interaction ranges $d_0$.
+We find that more hierarchical urban systems tend to produce more international networks (recall that the internationalisation indicator corresponds to the modularity, the latter is minimized when the networks span more across different countries).
+This is consistent with the existence of globalised metropolis \cite{sassen1991global}.
+This effect is mitigated by reducing the interaction ranges.
+Regarding metropolisation, we observe consistent maximum values as a function of the initial hierarchy across different interaction ranges $d_0$.
 This means that a too high hierarchy is detrimental to the global metropolisation when considering all cities, what may be due to the fact that in such a case very large cities are capturing all flows, leaving nothing to intermediate and medium-sized cities.
-Interestingly, the optimal value $\alpha \simeq 1.3$ is not far from observed values (which usually range between 0.9 and 1.1 but can take larger values such as 1.5 when extending the urban system \cite{Raimbault_2020}), what may imply that this optimality property is endogenous to the emergence of urban systems and linked to their intrinsic hierarchy.
+Interestingly, the optimal value $\alpha \simeq 1.3$ is not far from the observed values (which usually range between 0.9 and 1.1 but can take larger values such as 1.5 when extending the urban system \cite{Raimbault_2020}), implying that this optimality property is endogenous to the emergence of urban systems and linked to their intrinsic hierarchy.
 
@@
     \vspace{2cm}
-    \caption{\textbf{Impact of urban hierarchy.} (Left) Internationalisation as a function of the initial synthetic hierarchy $\alpha$, for different values of interaction range $d_0$ (color); (Right) Metropolisation as a function of $\alpha$, for different values of interaction range $d_0$ (color).\label{fig:fig7}}
+    \caption{\textbf{Impact of the urban hierarchy.} (Left) Internationalisation as a function of the initial synthetic hierarchical exponent $\alpha$, for different values of the interaction range $d_0$ (color); (Right) Metropolisation as a function of $\alpha$, for different values of the interaction range $d_0$ (color).\label{fig:fig7}}
 \end{figure}
@@
 
+% TODO precise calib setting + details on rescaling
 We then apply the model to a real system of cities by calibrating it on the European ownership network.
-Model setup is done following the real setup described above.
+The model setup is done following the real setup described above.
 The number of time steps $t_f$ is left as an additional parameter, which in a sense determines the granularity of the cumulative network generation process.
-The model is fitted at the microscopic level on link structure: the objective functions for calibration are the average mean squared error on logarithms of weights given by
+The model is fitted at the microscopic level on the link structure: the objective functions for calibration are the average mean squared error on logarithms of weights given by
 
@@
 \end{equation}
-
 where $\hat{w}_{ij}$ are the simulated weights, and the logarithm of the mean squared error
@@
 The values for $\varepsilon_L$ are for a large proportion of points lower than for statistical models, our model outperforms these in this predictive sense.
-As shown in Fig.~\ref{fig:fig8}, increasing values for $\gamma_O$ lead to better performance regarding $\varepsilon_L$: increasing the importance of the origin will provide a better fit on all order of magnitudes of link weights (while $\varepsilon_M$ favors mostly the largest links given the fat-tail distribution).
+As shown in Fig.~\ref{fig:fig8}, increasing values for $\gamma_O$ lead to better performance regarding $\varepsilon_L$: increasing the importance of the origin will provide a better fit on all orders of magnitude of the link weights (while $\varepsilon_M$ favours mostly the largest links given the fat-tail distribution).
 
-When considering optimal points with $\varepsilon_L < 5.0$, the adjusted parameters are: long range gravity interactions and socio-economic interactions given by $d_0 = 5972 \pm 2148$ and $c_0 = 113 \pm 33$ (the maximal $c_{ij}$ is 11.8); a stronger influence of destination than origin as $\gamma_O = 1.46 \pm 0.10$ and $\gamma_D = 2.0 \pm 0.05$; a comparable influence of sectors as $\gamma_S = 1.76 \pm 0.21$; a fine time granularity as $t_f = 4951 \pm 124$; and a very strong influence of previous links as $\gamma_W = 6.43 \pm 4.11$.
-This last point furthermore shows in particular a role of the path-dependency parameter, confirming the relevance of this complex generative model including path-dependency.
-The interaction of this process with others implies an inversion of the relative importance of origin and destination (on the contrary to statistical models).
+When considering optimal points with $\varepsilon_L < 5.0$, the adjusted parameters are: long range gravity interactions and socio-economic interactions given by $d_0 = 5972 \pm 2148$ and $c_0 = 113 \pm 33$ (the maximal $c_{ij}$ is 11.8); a stronger influence of the destination than of the origin as $\gamma_O = 1.46 \pm 0.10$ and $\gamma_D = 2.0 \pm 0.05$; a comparable influence of sectors as $\gamma_S = 1.76 \pm 0.21$; a fine time granularity as $t_f = 4951 \pm 124$; and a very strong influence of previous links as $\gamma_W = 6.43 \pm 4.11$.
+This last point furthermore shows in particular the importance of including path-dependency in this complex generative model.
+The interaction of this process with others implies an inversion of the relative importance of origin and destination (in opposition to the statistical models).
 The fact that calibrated interaction distances are large is consistent with long range economic interactions within Europe.
-Besides these stylized facts, these calibration results altogether demonstrate the feasibility of model application to real data.
+Besides these stylized facts, these calibration results altogether demonstrate the feasibility of the application of the model to real data.
 As stated previously, the model is not predictive and the relatively high error values would not make it satisfactory in that manner, but are reasonable for its explanatory and application functions.
@@
 
-We finally design scenarios to evaluate the impact of economic shocks.
-These can be the consequence of diverse factors, including socio-economic, geopolitical factors (example of Brexit), or other unpredicted human disasters such as the COVID-19 pandemic.
-We simulate border closures by rescaling $c_0$ -- the range parameter for socio-economic interactions --, and intra-country lock-downs by rescaling $d_0$ -- the range parameter for direct interactions -- during the simulation, with other parameters being fixed to their default values and the model setup in the real configuration.
+Let us design scenarios to evaluate the impact of economic shocks.
+These can be the consequence of diverse factors, including socio-economic, geopolitical factors (the example of Brexit), or other unpredicted human disasters such as the COVID-19 pandemic.
+We introduce the shock of the latter by simulating border closures by rescaling $c_0$ -- the range parameter for socio-economic interactions --, and intra-country lock-downs by rescaling $d_0$ -- the range parameter for direct interactions -- leaving other parameters fixed to their default values, and the model setup in the real configuration.
 More precisely, with $t_f = 5000$, the model is run for $t_f / 2$ steps.
-Then, following two new rate parameters $\kappa_C \in \left[0;1\right]$ and $\kappa_G \in \left[0;1\right]$ defining the shock, model parameters are updated as $d_0 = \kappa_G \cdot d_0$ and $c_0 = \kappa_C \cdot c_0$.
+Then, we introduce the effect of the shock through two new rate parameters $\kappa_C \in \left[0;1\right]$ and $\kappa_G \in \left[0;1\right]$, such that the model parameters are updated as $d_0 = \kappa_G \cdot d_0$ and $c_0 = \kappa_C \cdot c_0$.
 The model is then run for the remaining time steps until $t_f$.
@@
 We find an important impact of the shock on internationalisation, but less on the average community size.
-The impact of changing $d_0$ is more important than $c_0$, but a combination of both yield the stronger effect.
+The impact of changing $d_0$ is more important than $c_0$, but a combination of both yields the stronger effect.
 For minimal values of $\kappa_G$ and $\kappa_C$, internationalisation is more than doubled, confirming a reconfiguration of urban networks within countries.
 We validate the statistical significance of effects observed in Fig.~\ref{fig:fig9} by comparing distributions with Kolmogorov-Smirnov (KS) tests.
-More precisely, we proceed for each indicator and value of $(\kappa_G,\kappa_C)$ to a two-sided KS test between the statistical distribution of the indicator over repetitions at final time with the distribution in the reference configuration.
+More precisely, we proceed for each indicator and value of $(\kappa_G,\kappa_C)$ to a two-sided KS test between the statistical distribution of the indicator over repetitions at the final time with the distribution in the reference configuration.
 For internationalisation, p-values of the test are all smaller than 0.002 for $\kappa_D < 0.6$ but larger than $0.15$ for $\kappa_D \geq 0.6$.
-This means that the shock has a statistically significant impact only when interaction range is more than halved.
+This means that the shock has a statistically significant impact only when the interaction range is more than halved.
 For community size, significance (p-value smaller than 0.01) is reached only for the smallest value $\kappa_D = 0.2$.
@@
 This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMOLE platform and calibrated on real data on ownership linkages of firms for Europe from the Amadeus dataset.
-The simulation on a synthetic system of cities provides a broad knowledge on model behavior in its parameter space.
+The simulation on a synthetic system of cities provides a broad knowledge on model behaviour in its parameter space.
 Even in such a simple model (close to directly tractable stationary state) the behaviour is highly non-linear in many dimensions.
 Our study shows how crucial model exploration is to overcome hidden parameters (deactivated mechanisms or default parameter values).
-This paper emphasis on the fact that exploration of intrinsic dynamics on synthetic data is of great importance and should be driven before an application on real data so as to put aside the geographical effects from model dynamics.
+This paper highlights the fact that the exploration of the intrinsic dynamics on synthetic data is of great importance and should be driven before an application on real data, in order to set aside the geographical effects in the model dynamics.
 
-The calibration of model suggests its potential application to policy issues and can therefore have practical applications in the future.
+The calibration of the model suggests its potential application to policy issues and can therefore have practical applications in the future.
 For example the effect of exogenous shocks in the socio-economic structure can be tested as we did in the stylized application example above.
 Such shocks can for example be the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to further pandemic outbreaks on the single market.
-In this sense the UK economic sectors mostly dependent on foreign capital and generating the highest turnovers in 2018, that would be the most affected and worth examining in details are: extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods (in terms of out-going links from the UK) ; activities of head offices and holding companies, sale of cars and manufacture of motor vehicles (in terms of in-going links to the UK).
-We showed in our experiment that although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalization and thus can be expected to have a detrimental effect on UK economy due to these foreign ownership links (see also S1 Text\ref{S1Text}).
+In this sense the UK economic sectors mostly dependent on foreign capital and generating the highest turnovers in 2018, that would be the most affected and worth examining in detail are: the extraction of crude petroleum, the manufacture of tobacco products and of basic pharmaceutical goods (in terms of out-going links from the UK) ; the activities of head offices and holding companies, the sale of cars and manufacture of motor vehicles (in terms of in-going links to the UK).
+We showed in our experiment that although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalisation and thus can be expected to have a detrimental effect on the UK economy due to these foreign ownership links (see also S1 Text\ref{S1Text}).
 The impact of such unforeseen economic shocks could also be considered from a multi-level perspective, coupling our work with other approaches such as \cite{Dorretal2022} studying the firm-level impact of the Covid crisis.
@@
 We also suggest that our simple agent-based model can be a basis for or a component of more complicated economic agent-based models.
-A co-evolution model with an evolution of city sizes in parallels to the firm network would be more realistic for an application on longer time scales, but would also be more difficult to parametrise and calibrate \cite{raimbault2018characterising}.
+A co-evolution model with an evolution of city sizes in parallel to the firm network would be more realistic for an application on longer time scales, but would also be more difficult to parametrise and calibrate \cite{raimbault2018characterising}.
 A model with additional firm agents would lead to a multi-scale agent-based model, with also more flexibility for policy applications.

### 2022-05-12 Response_to_reviewers.tex
Update on Overleaf.

@@
-\documentclass[10pt,a4paper,sans]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')
+\documentclass[10pt,a4paper,sans]{moderncv}
 
 \usepackage[document]{ragged2e}
-% pour justifier
 
-% moderncv themes
-\moderncvstyle{banking}                            % style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
-\moderncvcolor{red}                                % color options 'blue' (default), 'orange', 'green', 'red', 'purple', 'grey' and 'black'
+\moderncvstyle{banking}
+\moderncvcolor{red}
 \renewcommand{\familydefault}{\rmdefault} 
@@
 \usepackage[utf8]{inputenc} 
-\usepackage[scale=0.9]{geometry}
+\usepackage[margin=3cm]{geometry}
 
@@
 \begin{document}
-
-% recipient data
-\recipient{Editor PONE}{}
-\date{\today}
-\opening{Dear Editor,}
-\closing{Yours faithfully,\\
-Juste Raimbault and co-authors\\
-University College London
-}
-         % use an optional argument to use a string other than "Enclosure", or redefine \enclname
-
-%\makelettertitle
 

### 2022-05-12 CoverLetter.tex
Update on Overleaf.

--- 2020-09-11 CoverLetter.tex
@@
 \recipient{Editors PLoS ONE}{}
-\date{September 11th, 2020}
+\date{May 12th 2022}
 \opening{Dear Editors,}
@@
 Juste Raimbault\\
-University College London
+LASTIG, Univ Gustave Eiffel, IGN-ENSG
 }
@@
 \justify
-We are pleased, with my co-authors Natalia Zdanowska (University College London) and Elsa Arcaute (University College London), to submit an original research article entitled ``Modeling growth of urban firm networks'' for consideration for publication in PLoS ONE.
+We are pleased, with my co-authors Natalia Zdanowska (Luxembourg Institute of Socio-economic Research) and Elsa Arcaute (University College London), to submit an original research article entitled ``Modeling growth of urban firm networks'' for consideration for publication in PLoS ONE.
 This article introduces a generative model for urban firm networks, to understand underlying processes, with possible applications to public policies.
@@
 
+A previous version of this manuscript was submitted to PLoS ONE in September 2020, with the submission identifier PONE-S-20-35765.
+Due to diverse unforeseen circumstances including job uncertainty and instability through the pandemic, we were not able to revise the manuscript in time after receiving a first round of reviews suggesting a revision.
+This current submission thus corresponds to the revision of this previous submission, and we included the response to reviewers and manuscript with track changes in submitted files.
+
 This manuscript has not been published and is not under consideration for publication elsewhere, nor contains previously published work.
-A preprint is being deposited on arxiv simultaneously to the submission.
+A preprint of a previous version was deposited on arxiv with the identifier 2009.05528.
 
-We have no conflicts of interest to disclose.
-We do not have any opposed reviewer.
-This paper is submitted to the PloS ONE collection ``Cities as Complex Systems'' edited by M.
+We have no conflicts of interest to disclose. We do not have any opposed reviewer.
+
+%This paper was initially submitted to the PloS ONE collection ``Cities as Complex Systems'' edited by M.
 Gonzalez (University of California, Berkeley) and D.
 Rybski (Potsdam Institute for Climate Impact Research).
-We know that the official deadline has passed, nevertheless, we hope that it could still be considered for this special issue.
+We know that the official deadline has passed, nevertheless, we hope that it could still be considered for a non-special issue.
 

### 2022-05-12 CoverLetter.tex
Update on Overleaf.

@@
-\documentclass[11pt,a4paper,sans]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')
+\documentclass[11pt,a4paper,sans]{moderncv}
 
@@
 
-% moderncv themes
-\moderncvstyle{banking}                            % style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
-\moderncvcolor{red}                                % color options 'blue' (default), 'orange', 'green', 'red', 'purple', 'grey' and 'black'
-\renewcommand{\familydefault}{\rmdefault}         % to set the default font; use '\sfdefault' for the default sans serif font, '\rmdefault' for the default roman one, or any tex font name
-%\nopagenumbers{}                                  % uncomment to suppress automatic page numbering for CVs longer than one page
-
-% character encoding
-\usepackage[utf8]{inputenc}                       % if you are not using xelatex ou lualatex, replace by the encoding you are using
-%\usepackage{CJKutf8}                              % if you need to use CJK to typeset your resume in Chinese, Japanese or Korean
-
-% adjust the page margins
+\moderncvstyle{banking}   
+\moderncvcolor{red} 
+\renewcommand{\familydefault}{\rmdefault}
+\usepackage[utf8]{inputenc}
 \usepackage[scale=0.75]{geometry}
@@
 \begin{document}
-%-----       letter       ---------------------------------------------------------
 
-% recipient data
 \recipient{Editors PLoS ONE}{}
@@
 }
-         % use an optional argument to use a string other than "Enclosure", or redefine \enclname
+
 \makelettertitle
@@
 This article introduces a generative model for urban firm networks, to understand underlying processes, with possible applications to public policies.
-%It does not contain previously published work.
-Preliminary results were presented as a satellite abstract at the Conference on Complex Systems 2019 (September 2019) and will be presented as a main track abstract at the NetScience 2020 conference (September 2020).
-We did not have any previous interaction with PLoS regarding this manuscript.
 
@@
 
-%This paper was initially submitted to the PloS ONE collection ``Cities as Complex Systems'' edited by M.
-Gonzalez (University of California, Berkeley) and D.
-Rybski (Potsdam Institute for Climate Impact Research).
-We know that the official deadline has passed, nevertheless, we hope that it could still be considered for a non-special issue.
+This paper was initially submitted to the PloS ONE collection ``Cities as Complex Systems''.
+We therefore suggest as editor D.
+Rybski (ORCID 0000-0001-6125-7705, Potsdam Institute for Climate Impact Research) who is on the editorial board and was editor of the collection.
+We also suggest as editor F.
+Biljecki (ORCID 000-0002-6229-7749, National University of Singapore).
 
@@
 \end{document}
-
-%% end of file `template.tex'.

### 2022-05-12 main.tex
Update on Overleaf.

@@
 {\Large
-\textbf\newline{Modeling growth of urban firm networks} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{Modelling growth of urban firm networks} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
@@
 Understanding the drivers behind the growth of such networks -- in particular urban firm networks --, is essential for the economic resilience of urban systems.
-We introduce in this paper a generative model for firm networks at the urban area level, which considers the following features: the economic size of the areas representing the origin and destination nodes, the industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distance between urban areas.
+We introduce in this paper a generative model for firm networks at the urban area level, which considers the following features: the economic size of the areas representing the source and target nodes, the industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distances between urban areas.
 An empirical network analysis on European urban firm ownership data confirms the relevance of each of these factors on new link creation.
 We then simulate network growth on synthetic systems of cities, unveiling stylized facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range.
-We calibrate the model on the European network, outperforming statistical models showing a strong role of path-dependency.
-The model opens up potential applications to formulate public policies recommendations to deal with exogenous shocks such as economic crisis or potential lock-downs of countries.
-This is illustrated with an application on stylised scenarios.
+We calibrate the model on the European network, outperforming statistical models, and showing a strong role of path-dependency.
+The model is applied to stylised scenarios, opening up to the formulation of public policies recommendations to deal with exogenous shocks such as economic crises or potential lock-downs of countries.
 
@@
 This can be achieved through strategic development of certain industrial sectors for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}.
-Exploring city/firm interactions \cite{storme2019introducing} and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future unforeseen exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.
+Exploring city/firm interactions \cite{storme2019introducing} and the position of cities within these networks \cite{gluckler2016relational} can also allow us to infer the consequences of unforeseen exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.
 
-Network models to understand processes underlying the emergence and dynamics of such networks have been proposed in the literature, nevertheless, these are not extensive and do not explicitly model the city level \cite{taylor2001specification}.
+%Network models to understand processes underlying the emergence and dynamics of such networks have been proposed in the literature, nevertheless, these are not extensive and do not explicitly model the city level \cite{taylor2001specification}.
 At regional scale, generative urban network models combining geographical factors with network topological properties have been proposed \cite{dai2016simulating}.
@@
 
-The first main contribution of this paper is to focus on interurban interactions between firms.
+Network models to understand processes underlying the emergence and dynamics of such networks have been proposed in the literature, nevertheless, these are not extensive and do not explicitly model the city level \cite{taylor2001specification}.
+At regional scale, generative urban network models combining geographical factors with network topological properties have been proposed \cite{dai2016simulating}.
+At country level, trade networks have been extensively studied from the complex networks perspective \cite{garlaschelli2005structure}, but these are generally not explored at a more granular level due to the lack of city-level data.
+Within this work, we contribute to the research at an interurban level, exploring the position and dynamics of cities in the socio-economic world-wide space \cite{pumain2018evolutionary} \cite{berry1964cities}, by considering the interactions between the firms in the cities.
+An extensive literature exists exploring firm-to-firm level networks in economic geography.
+For example, within the input-output models framework, these are used to study inter-sectors and inter-region transactions \cite{Rodriguesetal2016}, or firm level productivity, to understand macroeconomic level outcomes \cite{Grassi2016} and economic resilience \cite{Ialongo2021}.
+%Studies of these firm networks at city level remain still scarce.
+At a more granular level, urban firm networks have been investigated in complexity science \cite{RozenblatNeal2021} to unveil agglomeration economies and competitiveness of cities \cite{Rozenblat2010}, regionalisation patterns \cite{Rozenblatetal2017} or to model hierarchy and specialisation \cite{BidaRozenblat2020}.
+Moreover economic complexity indicators have been developed in this sense to characterise the production structure and specialisation of different countries due to economic activity, including those of trade firms \cite{HaussmanHidalgo2009,OCleary2016}.
+This work complements this effort by proposing a generative mechanism of such networks.
+
+Another contribution of the present study is the derivation of geographical and economic processes such as internationalisation and metropolisation of cities, through the proposed generative model of firm networks, where weights of cities are given by shares of turnovers generated by ownership links between firms.
+Previous literature considers mostly weights as GDP of countries, for example in gravity models of export-import firms networks \cite{Kepaptsogloual2010} \cite{MinondoRequena2013}.
+Input-output models, mostly used in regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}.
+In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for different types of flows \cite{dai2016generative} or for the physical infrastructure \cite{raimbault2018indirect} at different geographical scales.
+
+%The first main contribution of this paper is to focus on interurban interactions between firms.
 Firm networks remains a central aspect of network and system of cities theory \cite{RozenblatNeal2021} from the economic perspective.
@@
 
-The second contribution of the present study is to question and capture geographical and economic processes as internationalisation and metropolisation of cities in terms of firm networks within a generative model, where weights of cities are given by shares of turnovers generated by ownership links between firms.
+%The second contribution of the present study is to question and capture geographical and economic processes as internationalisation and metropolisation of cities in terms of firm networks within a generative model, where weights of cities are given by shares of turnovers generated by ownership links between firms.
 Previous literature considers mostly weights as GDP of countries for example in gravity models of export-import firms networks \cite{Kepaptsogloual2010} \cite{MinondoRequena2013}.
@@
 
-The specific question we tackle is the complementarity of different processes to generate urban firm networks.
+The specific question we tackle is the complementarity of the different processes to generate urban firm networks.
 We examine the interactions of European cities within firm linkages defined by corporate ownership links.
-Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2020Zdanowska}.
+Since transnational firms structure is one of the determinants of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2020Zdanowska}.
+More precisely : (i) we firstly introduce a generative network model to simulate the growth of such linkages at urban areas scale, combining multiple factors influencing link formation, such as the economic intensity of the urban areas corresponding to the source and target nodes, the industrial sector proximity, the strength of previous links, and the geographical and socio-cultural distances; (ii) the model allows us to then compare the effect of the different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; (iii) we develop an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database \cite{AMADEUS2018};  (iv) we use this data to calibrate the model on the empirical network; and (v) we explore an application of the model to an economic shock scenario, with pandemic-like stylised facts.
+
+%The specific question we tackle is the complementarity of different processes to generate urban firm networks.
+We examine the interactions of European cities within firm linkages defined by corporate ownership links.
+Since transnational firms structure is one of the determinants of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2020Zdanowska}.
 More precisely : (i) we propose an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database \cite{AMADEUS2018}; (ii) we introduce a generative network model to simulate the growth of such linkages at urban areas scale, which combines multiple factors influencing link formation, such as the economic intensity of the urban areas corresponding to the origin and destination nodes, the industrial sector proximity, the strength of previous links, and the geographical and socio-cultural distance; (iii) the model allows us to compare the effect of different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; and (iv) we calibrate the model on the empirical network.
 
-The rest of this paper is organized as follows: we firstly describe the generative model for urban networks and then present an empirical analysis of the network used as a case study.
+%The rest of this paper is organized as follows: we firstly describe the generative model for urban networks and then present an empirical analysis of the network used as a case study.
 We then perform a sensitivity analysis and exploration of the model parameters using synthetic data.
@@
 The model can also be used as a decision-making model, where different scenarios for potential policy applications can be explored.
-%It also tends to be a decision-making model as we explore potential policy applications.
-Our focus on explanation rather than prediction can be understood in a context of a broad diversity of possible model functions \cite{varenne2018models}.
 
@@
 Parameters of the log-normal law are obtained by solving the corresponding two equations system for each city, and each distribution is discretised and renormalised to be a vector of probabilities.
-See S2 Text\ref{S2Text} for all details of this synthetic setup.
+See S2 Text for all details of this synthetic setup.
 
 In the synthetic experiments, the network is generated from an initial complete network with dummy links of weight $1$ until reaching $t_f=1500$, and with a unit link volume of $w_0 = 1$.
-% missing links are not constrained
 
@@
 Urban areas are distributed according to their positions in the actual geographical space, and economic size and sector composition attributes are used.
-The socio-economic distance is constructed using information from the statistical modeling, by considering fixed-effect coefficients estimated for each pair of countries.
+The socio-economic distance is constructed using information from the statistical modelling, by considering fixed-effect coefficients estimated for each pair of countries.
 See S2 Text for more details on the construction of the socio-economic distance matrix.
@@
 The later can be defined as a concentration of population and major economic decision functions in metropolises, in particular at the world-system level \cite{sassen1991global}.
-Metropolises playing the role of gateways of globalisation, are competing between each other to attract the most qualified and dynamic activities \cite{RozenblatPumain2007}.
+Metropolises playing the role of gateways of globalisation, are competing between each other to attract the most skilled workers and innovative industries \cite{RozenblatPumain2007}.
 Metropolisation gradients have been linked in the literature on systems of cities to several economic and political aspects \cite{RozenblatPumain2018} \cite{PumainRozenblat2019} \cite{Zdanowskaetal2020}.
@@
 As we aggregate company turnovers on urban areas in our analysis, we thus do not expect much representativity bias in the inter-urban network.
-% To what database is Orbis compared to in this study - what is the confidence on its exhaustivity
 
@@
 This quantity is highly correlated with the GDP of the area-level included in the GHS database (Spearman correlation $\rho = 0.71$) and also with population ($\rho = 0.60$).
-We then expect to capture size effects while keeping the consistence of a single main data source.
 Another attribute of the nodes is the industrial sector composition, expressed as a proportion of turnover in the area associated to a given industrial sector.
@@
 This allows us to define an industrial proximity matrix between urban areas, taking a cosine similarity between vectors as $s_{ij} = \sum_{k=1}^{K} S_{ik}\cdot S_{jk}$ where $S_{ik}$ are sector proportions.
-
-% network density: 9323/(573*572) = 0.02844494
 
@@
 15).
-In comparison, when taking into account 29 countries, the classification of urban areas by country has a modularity of 0.32.
+In comparison, when taking as communities 29 existing countries, the (directed?)  modularity is 0.32.
 This indicates that the structure of flows present certain regional patterns.
@@
     \end{center}
-    \vspace{2cm}
     \caption{{\bf Map of network communities obtained by modularity maximisation.} We show communities with the higher modularity, obtained with the Louvain algorithm.
@@
 The same results can be observed for the networks generated using the real data in the lower panels.
-% TODO work on data readme on the repo / do a separate dataverse?
 
@@
     \end{center}
-    \vspace{2cm}
     \caption{Example of simulated networks at $t=1500$, for a synthetic system of cities (Top row) and the European urban network (Bottom row), with a high (resp.
@@
 The behaviour of some indicators are shown in Fig.~\ref{fig:fig4}.
-For example, internationalisation and metropolisation both show a transition as a function of d$_o$, from a local to a global regime.
-The sector proximity parameter $\gamma_S$ influences the internationalisation linearly, but induces a maximum for the correlation between size and degree, which can be interpreted as a setting where size and the total volume in a city are the most correlated.
-% TODO comment max metropolisation
+For example, internationalisation and metropolisation both show a transition as a function of $d_0$, from a local to a global regime.
+The sector proximity parameter $\gamma_S$ has a linearly increasing influence on internationalisation.
+On the contrary, it induces a maximum for metropolisation, what can be interpreted as the fact that too few or too much specialisation leading to less integrated cities in terms of metropolisation.
 
@@
 	\end{center}
-	\vspace{2cm}
     \caption{(Top Left) Internationalisation index decreases exponentially with gravity decay; (Top Right) Correlation between city weighted degree and size.
 Both plots show a transition from a local to a global regime.
-(Bottom Left) Internationalisation varies linearly with sector proximity $\gamma_S$; (Bottom Right) Metropolisation exhibits a maximum, witnessing an intermediate regime where size is the most important.
+(Bottom Left) Internationalisation varies linearly with sector proximity $\gamma_S$; (Bottom Right) Metropolisation exhibits a maximum at $\gamma_S \simeq 1.2$, suggesting an intermediate regime where specialisation has more influence.
 \label{fig:fig4}}
@@
 We show that the size of the origin and the sector proximity also have a significant effect, although the latter only becomes relevant when the size of the origin has a strong influence (third panel).
-%It interacts with the role of sectors and the size of the origin, with a more significant effect of sector proximity, when size has the largest influence (third panel).
 
@@
     \includegraphics[width=\linewidth]{figures/Fig5.png}
-    \vspace{2cm}
     \caption{The transition as a function of the interaction range depends on the influence of the origin size $\gamma_O$; the sector proximity $\gamma_S$ plays a role only for a large influence of the origin.\label{fig:fig5}}
@@
 
-% TODO JR redo comment
 Other indicators exhibit non-trivial patterns.
@@
     \includegraphics[width=\textwidth]{figures/Fig6.png}
-    \vspace{2cm}
 	\caption{Maximal integration in terms of the community size is achieved at an intermediate value of d$_o$: emergence of a regional regime; the maximal size depends on the role of the sectors $\gamma_S$ in a decreasing way when the origin size is deactivated, and in an increasing way when $\gamma_o=1$; this regime disappears when the influence of the origin size becomes too large.\label{fig:fig6}}
@@
 
-% TODO new fig (range?)
 We run a specific experiment to study the impact of the urban hierarchy on firms dynamics.
 Rank-size law in cities is a central stylized fact within urban theories, despite the discussions regarding the estimation of the power-law exponent, its actual values and its variation depending on the definition of cities considered \cite{cottineau2017metazipf,corral2020truncated}.
-In this particular context, understanding how this parameter can impact urban dynamics within artificial systems of cities -- in which it can be modified to extreme values non observable in real systems--, should bring evidence on underlying processes \cite{raimbault2019space}.
-We change the initial hierarchical exponent $\alpha$ from 0.5 to 2.0 with a step of 0.1, and also vary the interaction range $d_0$.
+I this particular context, understanding how this parameter can impact urban dynamics within artificial systems of cities -- in which it can be modified to extreme values non observable in real systems--, should help disentangling the variability due to model parameters from the variability due to the specific configuration of the urban system \cite{raimbault2019space}.
+We change the initial hierarchical exponent $\alpha$ from 0.1 to 2.0 with a step of 0.1, and also vary the interaction range $d_0$.
 The behaviour of the internationalisation and metropolisation indicators are shown in Fig.~\ref{fig:fig7}.
@@
 This effect is mitigated by reducing the interaction ranges.
-Regarding metropolisation, we observe consistent maximum values as a function of the initial hierarchy across different interaction ranges $d_0$.
+Regarding metropolisation, we observe the existence of maxima as a function of the initial hierarchy across different interaction ranges $d_0$.
 This means that a too high hierarchy is detrimental to the global metropolisation when considering all cities, what may be due to the fact that in such a case very large cities are capturing all flows, leaving nothing to intermediate and medium-sized cities.
-Interestingly, the optimal value $\alpha \simeq 1.3$ is not far from the observed values (which usually range between 0.9 and 1.1 but can take larger values such as 1.5 when extending the urban system \cite{Raimbault_2020}), implying that this optimality property is endogenous to the emergence of urban systems and linked to their intrinsic hierarchy.
+Interestingly, the optimal value $\alpha \simeq 1.1$ is not far from the observed values (which usually range between 0.9 and 1.1 but can take larger values such as 1.5 when extending the urban system \cite{Raimbault_2020}), implying that this optimality property is endogenous to the emergence of urban systems and linked to their intrinsic hierarchy.
 
@@
     \end{center}
-    \vspace{2cm}
     \caption{\textbf{Impact of the urban hierarchy.} (Left) Internationalisation as a function of the initial synthetic hierarchical exponent $\alpha$, for different values of the interaction range $d_0$ (color); (Right) Metropolisation as a function of $\alpha$, for different values of the interaction range $d_0$ (color).\label{fig:fig7}}
@@
 The number of time steps $t_f$ is left as an additional parameter, which in a sense determines the granularity of the cumulative network generation process.
+
 The model is fitted at the microscopic level on the link structure: the objective functions for calibration are the average mean squared error on logarithms of weights given by
@@
 \end{equation}
-where $\hat{w}_{ij}$ are the simulated weights, and the logarithm of the mean squared error
+where $\hat{w}_{ij}$ are the simulated weights, and the logarithm of the mean squared error (MSE)
 
@@
 
-\cite{raimbault2018indirect} showed that these two are complementary to calibrate urban systems models, capturing the overall performance but also how the model behaves on medium-sized and small links.
+These two indicators capture the overall model performance with the logarithm of MSE, but also performance on medium-sized and small links, and are therefore complementary as shown in the case of cities population dynamics by \cite{raimbault2018indirect}.
 A calibration on macroscopic indicators of metropolisation and internationalisation would not bring much knowledge on model processes, since several points from the grid experiment above - not designed for this - already provide a good performance (19 parameter points out of 6534 are simultaneously below a relative error of 10\% on internationalisation and below 10\% on metropolisation).
-% TODO in progress: test macro calib ; rerun explo metropolisation? -> ok
 
@@
     \end{center}
-    \vspace{2cm}
 	\caption{\textbf{Model calibration results.} Pareto front obtained for the bi-objective model calibration for $\varepsilon_L$ and $\varepsilon_M$ objectives.
@@
 
-% computePvalMat(res$internationalizationTS.98)
-%             [,1]         [,2]        [,3]         [,4]        [,5]
-%[1,] 4.884981e-15 4.440892e-16 0.000000000 0.0000000000 0.000000000
-%[2,] 1.364656e-03 9.570235e-06 0.001364656 0.0002468196 0.002318458
-%[3,] 8.127483e-01 6.993742e-01 0.999633292 0.4675585921 0.154538055
-%[4,] 5.806178e-01 8.127483e-01 0.366726444 0.6993741999 0.906206390
-%[5,] 8.127483e-01 2.809600e-01 0.210551633 0.4675585921 1.000000000
-% computePvalMat(res$networkAvgCommunitySizeTS.98)
-%             [,1]         [,2]         [,3]         [,4]         [,5]
-%[1,] 4.365916e-08 4.705150e-06 3.728748e-05 2.267454e-06 1.071070e-06
-%[2,] 5.410262e-02 6.993742e-01 6.302223e-03 1.545381e-01 4.675586e-01
-%[3,] 9.670685e-01 1.545381e-01 9.996333e-01 9.996333e-01 4.675586e-01
-%[4,] 4.675586e-01 5.806178e-01 4.675586e-01 9.670685e-01 6.993742e-01
-%[5,] 9.670685e-01 5.806178e-01 4.675586e-01 6.993742e-01 1.000000e+00
-
 %%%%%%%%%%%%%%
@@
     \end{center}
-    \vspace{2cm}
 	\caption{\textbf{Application of the calibrated model to an economic shock scenario.} All plots show the trajectory of indicators (internationalisation for the top row and average community size for the bottom row) in time, for different values of interaction decay rescaling $\kappa_G$ (color) and socio-economic distance rescaling $\kappa_C$ (panels).
@@
 
-% towards dynamical graph generative model, to allow for dynamical calibration (discuss possibilities of dyn calib?
-%no edge deletion in the model: also discuss?)
-% Discuss possible extension with self-connection
-
 \section*{Supporting information}
 
-\subsection*{S1 Text: UK's economic sectors dependent on international links.\label{S1Text}}
+\subsection*{S1 Text: UK's economic sectors dependent on international links.}
+\label{S1Text}
 
@@
 
-\subsection*{S2 Text: Statistical models, generative model setup and sensitivity analysis.\label{S2Text}}
+\subsection*{S2 Text: Statistical models, generative model setup and sensitivity analysis.}
+\label{S2Text}
 

### 2022-05-16 main.tex
Update on Overleaf.

@@
 {\Large
-\textbf\newline{Modelling growth of urban firm networks} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{Modelling growth of urban firm networks}
 }
 \newline
-% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+
 \\
@@
 
-%Network models to understand processes underlying the emergence and dynamics of such networks have been proposed in the literature, nevertheless, these are not extensive and do not explicitly model the city level \cite{taylor2001specification}.
-At regional scale, generative urban network models combining geographical factors with network topological properties have been proposed \cite{dai2016simulating}.
-At country level, trade networks have been extensively studied from the complex networks perspective \cite{garlaschelli2005structure}, but these are generally not explored at a more granular level due to the lack of city-level data.
-Within regional science, input-output models are used as generative urban models of flows between geographical areas \cite{jin1993generation}.
-In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for different types of flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2018indirect} at different geographical scales.
-
 Network models to understand processes underlying the emergence and dynamics of such networks have been proposed in the literature, nevertheless, these are not extensive and do not explicitly model the city level \cite{taylor2001specification}.
@@
 For example, within the input-output models framework, these are used to study inter-sectors and inter-region transactions \cite{Rodriguesetal2016}, or firm level productivity, to understand macroeconomic level outcomes \cite{Grassi2016} and economic resilience \cite{Ialongo2021}.
-%Studies of these firm networks at city level remain still scarce.
 At a more granular level, urban firm networks have been investigated in complexity science \cite{RozenblatNeal2021} to unveil agglomeration economies and competitiveness of cities \cite{Rozenblat2010}, regionalisation patterns \cite{Rozenblatetal2017} or to model hierarchy and specialisation \cite{BidaRozenblat2020}.
@@
 Since transnational firms structure is one of the determinants of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2020Zdanowska}.
-More precisely : (i) we firstly introduce a generative network model to simulate the growth of such linkages at urban areas scale, combining multiple factors influencing link formation, such as the economic intensity of the urban areas corresponding to the source and target nodes, the industrial sector proximity, the strength of previous links, and the geographical and socio-cultural distances; (ii) the model allows us to then compare the effect of the different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; (iii) we develop an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database \cite{AMADEUS2018};  (iv) we use this data to calibrate the model on the empirical network; and (v) we explore an application of the model to an economic shock scenario, with pandemic-like stylised facts.
-
-%The specific question we tackle is the complementarity of different processes to generate urban firm networks.
-We examine the interactions of European cities within firm linkages defined by corporate ownership links.
-Since transnational firms structure is one of the determinants of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2020Zdanowska}.
-More precisely : (i) we propose an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database \cite{AMADEUS2018}; (ii) we introduce a generative network model to simulate the growth of such linkages at urban areas scale, which combines multiple factors influencing link formation, such as the economic intensity of the urban areas corresponding to the origin and destination nodes, the industrial sector proximity, the strength of previous links, and the geographical and socio-cultural distance; (iii) the model allows us to compare the effect of different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; and (iv) we calibrate the model on the empirical network.
+More precisely: (i) we firstly introduce a generative network model to simulate the growth of such linkages at urban areas scale, combining multiple factors influencing link formation, such as the economic intensity of the urban areas corresponding to the source and target nodes, the industrial sector proximity, the strength of previous links, and the geographical and socio-cultural distances; (ii) the model allows us to then compare the effect of the different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; (iii) we develop an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database \cite{AMADEUS2018};  (iv) we use this data to calibrate the model on the empirical network; and (v) we explore an application of the model to an economic shock scenario, with pandemic-like stylised facts.
 
@@
 \subsubsection*{Rationale}
-%The model is expected to capture a single macroscopic urban scale, i.e.
-the level of the urban system where basic entities are cities.
-Links are induced by underlying firms but these are not explicit.
-Cities are defined by their economic size, but also by an industrial sector profile.
-As the authors of \cite{martinus2018global} point out, a combination of several distances may play a role in establishing linkages.
+
 The model is expected to capture a single macroscopic urban scale, which corresponds to the level at which nodes are cities.
 These are characterised by their economic size and their industrial sector profile.
-%The links in the network are constructed from the relationship between these characteristics and hence induced from their underlying firms.
-Different distances, or proximity measures, play a role in establishing linkages, as pointed out by \cite{martinus2018global}.
 Links are then created considering these features and different distance, or proximity measures \cite{martinus2018global}.
@@
 
-When focusing on the community structure of the network, different modularity maximisation algorithms (ran on the undirected corresponding network) result in the same undirected modularity of 0.38, a directed modularity \cite{nicosia2009extending} of 0.34 for the greedy algorithm of \cite{clauset2004finding} and of 0.36 for the Louvain algorithm \cite{blondel2008fast}, corresponding to 36 communities (resp.
-15).
-In comparison, when taking as communities 29 existing countries, the (directed?)  modularity is 0.32.
+To focus on the community structure of the network, we test different modularity maximisation algorithms ran on the undirected corresponding network.
+The Louvain algorithm \cite{blondel2008fast} provides an undirected modularity of $0.388$, with 19 communities.
+The corresponding directed modularity computed following \cite{nicosia2009extending} is $0.362$.
+The greedy algorithm of \cite{clauset2004finding} results in 35 communities for a undirected modularity of $0.376$ and an directed modularity of $0.352$.
+In comparison, when taking as communities 29 existing countries, the corresponding directed modularity is $0.323$.
 This indicates that the structure of flows present certain regional patterns.
-Indeed, a null model assigning random labels with 30 communities, bootstrap 1000 times, yields an average modularity of $0.049 \pm 0.002$, confirming the statistical significance of these communities.
+Indeed, a null model assigning random labels with 30 communities, bootstrap 1000 times, yields an average modularity of $0.050 \pm 0.0017$, confirming the statistical significance of these communities.
 We show in Fig.~\ref{fig:fig2} a map of the communities obtained.

### 2022-05-18 Response_to_reviewers.tex
Update on Overleaf.

@@
 
-	$\rightarrow$ Literature review and paper positioning was reworked and extended to reflect this intersection of disciplines and currents.
+	$\rightarrow$ Literature review and paper positioning was reworked and extended to reflect this intersection of disciplines (although preferential attachment is not as central in our study).
 	%\comment{(All) Rewrite lit review. Include applications of gravity models to trade and migration.}
@@
 	\item \textit{Given that the model is generative, the temporal aspect of the model can be an issue.
-Indeed, if the number of iterations (around 5000) needed to generate the empirical network correspond to a long time period (for example decades), the economic size and sectorial proximity are expected to change, probably implying a substantially different network, if these changes were accounted for in the model.
+Indeed, if the number of iterations (around 5000) needed to generate the empirical network correspond to a long time period (for example decades), the economic size and sectoral proximity are expected to change, probably implying a substantially different network, if these changes were accounted for in the model.
 Can the authors mention (and may be discuss) this issue?}
@@
 Statistical analysis is a preliminary exploration to determine potential processes to be included in the generative model, while the latest is an explanatory model, i.e.
-aimed at explaining the role of different factors in the growth of the urban network (such as the role of urban hierarchy) and test hypothesis in a virtual laboratory.
+aimed at explaining the role of different factors in the growth of the urban network (such as the one of urban hierarchy) and test hypothesis in a virtual laboratory.
 It is also used to prototype policy applications, and in that sense is potentially a decision-making model.
@@
 	% check lit - which lit stream
-	$\rightarrow$ The choice of indicators has been better justified and explained based on citations of previous literature, already using these indicators to capture important globalisation processes in a city level network.
+	$\rightarrow$ The choice of indicators has been better justified and explained based on citations of previous literature, already using these indicators to capture important globalisation processes in a city-level network.
 

### 2022-05-18 S1_Text.tex
Update on Overleaf.

--- 2020-09-01 S1_Text.tex
@@
 
-Based on \emph {AMADEUS} dataset, the most vulnerable economic sectors in the United Kingdom have been highlighted.
-The later are the ones due to be mostly affected by eventual future lock-downs of the international economy (international ownership links).
+This section provides details on the selection of economic sectors in the United Kingdom that contain considerable international ownership links, and could hence become vulnerable to shocks related to those links being compromised, such as protectionist strategies, new trading agreements, reshuffling of the international trading organization, etc.
 
-When considering the UK sectors generating the highest turnover in 2018 mostly dependent on foreign capital they own (out-going links from the UK) these are extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods.
-In terms of in-going links to the UK, activities of head offices and holding companies, sale of cars and manufacture of motor vehicles are sectors mostly dependent on foreign ownership coming from abroad.
+Looking specifically at sectors in the UK whose turnover is mostly dependent on their ownership of foreign capital in 2018, these correspond to extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods.
+Ownership relationships correspond to out-going links, see Table~\ref{tab:uk_outgoing} for details.
+
+On the other hand, the sectors in the UK that mostly dependent on foreign ownership coming from abroad correspond to activities of head offices and holding companies, sale of cars and manufacture of motor vehicles, and these are represented by in-going links to the UK, see Table~\ref{tab:uk_ingoing}.
 
@@
 \newpage
-When expressed in total number of links, activities mostly dependent on out-going links from the UK are activities of holding companies, of head offices and fund management activities.
-In terms of in-going links these are business and consultancy activities and renting and operating of own and leased real estate.
+When expressed in total number of links (Table~\ref{tab:uk_outgoinglinks} and Table~\ref{tab:uk_ingoinglinks}), the activities that mostly depend on out-going links from the UK are activities of holding companies, of head offices and fund management activities.
+In terms of in-going links, these are business and consultancy activities and renting and operating of own and leased real estate.
 

### 2022-05-18 S2_Text.tex
Update on Overleaf.

@@
 
-In the real model setup, socio-economic distance between countries $c_{ij}$ is constructed using fixed effects coefficients from the statistical model.
-In Table~1 in main text, we estimated several statistical models, including some with fixed effects by pairs of countries.
+In the real model setup, the socio-economic distance between countries $c_{ij}$ is constructed using the fixed effects coefficients from the statistical model.
+In Table~1 in the main text, we estimated several statistical models, including some with fixed effects by pairs of countries.
 The socio-cultural distance between two countries is then taken as the opposite of the fixed effect coefficient for this pair, for the full statistical model.
-Summary statistics for the 29x29 matrix (28 EU countries and Jersey which is in the database a distinct country from UK but has an important total turnover since many British companies are located there for fiscal advantage reasons) are shown below in Table~\ref{tab:fixedeff}.
-We observe that distances are rather localized but with large outliers, and an important proportion are not defined (339 out of 841), corresponding to couple of countries having no exchange at all in the dataset.
+The summary statistics for the 29x29 matrix (28 EU countries and Jersey, which is in the database corresponds to a distinct country from the UK, although it has an important total turnover since many British companies are located there for fiscal advantage reasons) are shown below in the Table~\ref{tab:fixedeff}.
+We observe that distances are rather localised but with large outliers, and an important proportion are not defined (339 out of 841), corresponding to a couple of countries having no exchange at all in the dataset.
 
@@
 
-%pscl::hurdle(formula = intweight ~ log(distance) + log(from_turnover) + log(to_turnover) + 
-%    log(sim), data = dz, dist = "poisson", zero.dist = "poisson")
-%
-%Pearson residuals:
-%       Min         1Q     Median         3Q        Max 
-%-9.195e+03 -1.373e-01 -7.627e-02 -4.204e-02  1.727e+04 
-%
-%Count model coefficients (truncated poisson with log link):
-%                     Estimate Std. Error z value Pr(>|z|)    
-%(Intercept)        -9.853e+00  1.345e-04  -73252   <2e-16 ***
-%log(distance)      -2.875e-01  7.603e-06  -37812   <2e-16 ***
-%log(from_turnover)  7.379e-01  4.708e-06  156720   <2e-16 ***
-%log(to_turnover)    6.379e-01  4.060e-06  157119   <2e-16 ***
-%log(sim)            4.550e-01  2.137e-05   21292   <2e-16 ***
-%Zero hurdle model coefficients (censored poisson with log link):
-%                     Estimate Std. Error z value Pr(>|z|)    
-%(Intercept)        -20.525823   0.155298 -132.17   <2e-16 ***
-%log(distance)       -0.747788   0.010991  -68.04   <2e-16 ***
-%log(from_turnover)   0.665678   0.005630  118.23   <2e-16 ***
-%log(to_turnover)     0.479214   0.005707   83.96   <2e-16 ***
-%log(sim)             0.278290   0.022439   12.40   <2e-16 ***
-%---
-%Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1 
-%
-%Number of iterations in BFGS optimization: 46 
-%Log-likelihood: -2.927e+10 on 10 Df
-
-% 0.1661997
-
-%mean((dz$intweight-hurdle2$fitted)^2)
-%[1] 6.119694e+12
-
-% > AIC(hurdle2)
-%[1] 58545461912
-
-%pscl::zeroinfl(formula = intweight ~ log(distance) + log(from_turnover) + log(to_turnover) + 
-%    log(sim), data = dz, dist = "poisson")
-%
-%Pearson residuals:
-%       Min         1Q     Median         3Q        Max 
-%-7.693e+00 -1.335e-01 -6.957e-02 -3.591e-02  1.760e+04 
-%
-%Count model coefficients (poisson with log link):
-%                     Estimate Std. Error z value Pr(>|z|)    
-%(Intercept)        -9.853e+00  1.345e-04  -73251   <2e-16 ***
-%log(distance)      -2.875e-01  7.603e-06  -37810   <2e-16 ***
-%log(from_turnover)  7.379e-01  4.708e-06  156719   <2e-16 ***
-%log(to_turnover)    6.379e-01  4.060e-06  157120   <2e-16 ***
-%log(sim)            4.550e-01  2.137e-05   21294   <2e-16 ***
-%
-%Zero-inflation model coefficients (binomial with logit link):
-%                    Estimate Std. Error z value Pr(>|z|)    
-%(Intercept)        22.046685   0.180605  122.07   <2e-16 ***
-%log(distance)       0.881724   0.013134   67.13   <2e-16 ***
-%log(from_turnover) -0.737441   0.006776 -108.83   <2e-16 ***
-%log(to_turnover)   -0.521642   0.006540  -79.76   <2e-16 ***
-%log(sim)           -0.330340   0.025360  -13.03   <2e-16 ***
-%---
-%Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1 
-%
-%Number of iterations in BFGS optimization: 1 
-%Log-likelihood: -2.927e+10 on 10 Df
-
-% 0.1607349
-
-%mean((dz$intweight-zinflpoisson2$fitted)^2)
-%[1] 6.159803e+12
-
-% AIC(zinflpoisson2)
-%[1] 58545461691
-
 \subsection*{Synthetic model setup}
@@
     \item Large cities are more innovative and more diverse.
-Assuming that sectors are ordered by innovativity (the larger the sector index the larger the innovativity), this assumption is translated by taking a mode and variance of 0.5 for the largest city and a mode and variance of $1/K$ for the smallest, and a linear interpolation between the two.
-For each city, we define
+Assuming that sectors are ordered by innovativity (the larger the sector index the larger the innovativity), this assumption is translated by taking a mode and variance of $1/2$ for the sector distribution of the largest city and a mode and variance of $1/K$ for the sector distribution of the smallest city.
+Intermediate cities have their sector distribution mode and variance following a linear interpolation between the two.
+For each city, we therefore define this value for mode and variance as $e_i$ by taking
     \[
-    \log(e_i) = \frac{(\log(E_i) - \log(E_{imin}))}{(\log(E_{imax}) - \log(E_{imin}))} * (1/2 - 1/K) + 1/K
+    e_i = \frac{(\log(E_i) - \log(E_{imin}))}{(\log(E_{imax}) - \log(E_{imin}))} \cdot (1/2 - 1/K) + 1/K
     \]
-    \item Log-normal parameters $(\mu_i,\sigma_i)$ for each city are then fixed by $\mu_i - \sigma_i^2 = \log(e_i)$ and $-3 \sigma_i^2 - 2 \log(\exp(\sigma_i^2) - 1) = log(e_i)$
-    \item $\sigma_i^2$ is the unique positive root of $f(X)=0$ with $f(X) = -3X - 2 \log(\exp(X) - 1) - \log(e_i)$ and $\mu_i = \log(e_i) + \sigma_i^2$.
+    where $E_i$ is the economic size of city $i$, $E_{imin}$ size of smallest city and $E_{imax}$ size of the largest city;
+    \item Log-normal parameters $(\mu_i,\sigma_i)$ for each city are then determined by using the formula for the mode and variance of a log-normal law; the mode is directly given by $\mu_i - \sigma_i^2 = \log(e_i)$, while the variance verifies $\log(\exp(\sigma_i^2) - 1) + 2\mu + \sigma^2  = log(e_i)$; using the first equation to have an expression for $\mu$ and substituting into the second, we obtain
+    \[
+    -3 \sigma_i^2 - 2 \log(\exp(\sigma_i^2) - 1) = log(e_i)
+    \]
+    \item $\sigma_i^2$ is thus obtained as the unique positive root of $f(X)=0$ with $f(X) = -3X - 2 \log(\exp(X) - 1) - \log(e_i)$ (this function has a derivative always negative on $\mathbb{R}+$, tends to $\infty$ when $X \rightarrow 0+$ and to $-\infty$ when $X \rightarrow \infty$); we finally compute $\mu_i$ as $\mu_i = \log(e_i) + \sigma_i^2$.
 \end{itemize}
-
-%solve for mu,sigma of the log-normal as a function of gdp
-%       we aim at having a log-normal mostly in [0,1] such that
-%       (mode,variance) =  1 / K for the smallest log(E_i) and = 1/2 for the largest log(E_i)
-%        writing
-%        e_i = (log(E_i) - log(E_imin))/(log(E_imax) - log(E_imin)) * (1/2 - 1/K) + 1/K
-%        this yields
-%         (1)  mu - sigma^2 = log(e_i)
-%         (2) -3 sigma^2 - 2 log(exp(sigma^2) - 1) = log(e_i)
-%        => sigma^2 is the unique positive root of f(X)=0 with f(X) = -3X - 2 ln(exp(X) - 1) - ln(e_i)
 
@@
  \hline
-Internationalisation & 0.2 & 0.3 & 0.7 & 0.7 & 0.001 & 0.009 & $4\cdot 10^{-4}$ & 0.007 & 0.03 & 0.04 & 0.02 & 0.04 \\
-Metropolisation & 0.02 & 0.1 & 0.02 & 0.2 & 0.002 & 0.1 & 0.001 & 0.09 & 0.2 & 0.6 & 0.3 & 0.6 \\
-Modularity & 0.3 & 0.4 & 0.6 & 0.6 & 0.004 & 0.02 & $3\cdot 10^{-4}$ & 0.01 & 0.005 & 0.03 & 0.002 & 0.03 \\
-Avg. com. size & 0.008 & 0.09 & 0.01 & 0.1 & 0.002 & 0.07 & 0.003 & 0.04 & 0.3 & 0.6 & 0.4 & 0.6 \\
-Degree entropy & 0.006 & 0.02 & 0.003 & 0.02 & 0.006 & 0.03 & 0.008 & 0.02 & 0.5 & 0.5 & 0.5 & 0.5 \\
-Weight entropy & 0.04 & 0.1 & 0.03 & 0.1 & 0.008 & 0.08 & 0.01 & 0.07 & 0.4 & 0.5 & 0.4 & 0.5 \\\hline
+Internationalisation & 0.2 & 0.3 & 0.7 & 0.7 & 0.001 & 0.008 & $5\cdot 10^{-4}$ & 0.007 & 0.03 & 0.04 & 0.02 & 0.04 \\
+Metropolisation & 0.02 & 0.04 & 0.02 & 0.04 & 0.02 & 0.3 & 0.002 & 0.01 & 0.2 & 0.7 & 0.2 & 0.7 \\
+Modularity & 0.3 & 0.4 & 0.6 & 0.6 & $5\cdot 10^{-4}$ & 0.02 & 0.002 & 0.02 & 0.007 & 0.03 & 0.003 & 0.03 \\
+Avg. com. size & 0.004 & 0.1 & 0.01 & 0.1 & 0.003 & 0.07 & 0.001 & 0.04 & 0.3 & 0.6 & 0.3 & 0.6 \\
+Degree entropy & 0.003 & 0.03 & 0.002 & 0.03 & 0.008 & 0.04 & 0.003 & 0.03 & 0.5 & 0.6 & 0.5 & 0.6 \\
+Weight entropy & 0.04 & 0.2 & 0.03 & 0.2 & 0.002 & 0.1 & 0.001 & 0.1 & 0.5 & 0.6 & 0.5 & 0.6 \\\hline
 \end{tabular}

### 2022-05-18 main.tex
Update on Overleaf.

@@
 \\
-Juste Raimbault\textsuperscript{1,2,3 *},
-Natalia Zdanowska\textsuperscript{3,4,5},
-Elsa Arcaute\textsuperscript{4}
+Juste Raimbault\textsuperscript{1,2,3,4 *},
+Natalia Zdanowska\textsuperscript{2,4,5},
+Elsa Arcaute\textsuperscript{2}
 \\
@@
 \\
-\textbf{2} UPS CNRS 3611 ISC-PIF, Paris, France
+\textbf{2} Centre for Advanced Spatial Analysis, UCL, London, United Kingdom
 \\
-\textbf{3} UMR CNRS 8504 G{\'e}ographie-cit{\'e}s, Paris, France
+\textbf{3} UPS CNRS 3611 ISC-PIF, Paris, France
 \\
-\textbf{4} Centre for Advanced Spatial Analysis, UCL, London, United Kingdom
+\textbf{4} UMR CNRS 8504 G{\'e}ographie-cit{\'e}s, Paris, France
 \\
@@
 
-%The first main contribution of this paper is to focus on interurban interactions between firms.
-Firm networks remains a central aspect of network and system of cities theory \cite{RozenblatNeal2021} from the economic perspective.
-In this approach the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary} \cite{berry1964cities}.
-An extensive literature exists to explore firm-to-firm level networks in economic geography withing the input-output models framework in order to study inter-sectors and inter-region transactions \cite{Rodriguesetal2016} or firm level productivity to understand macroeconomic level outcomes \cite{Grassi2016} and economic resilience \cite{Ialongo2021}.
-Studies of these firm networks at city level remain still scarce.
-Urban firm networks have been investigated in complexity science to unveil agglomeration economies and competitiveness of cities \cite{Rozenblat2010}, regionalisation patterns \cite{Rozenblatetal2017} or to model hierarchy and specialisation \cite{BidaRozenblat2020}.
-Moreover economic complexity indicators have been developed in this sense to characterise the production structure and specialisation of different countries due to economic activity, including those of trade firms \cite{HaussmanHidalgo2009,OCleary2016}.
-
-%The second contribution of the present study is to question and capture geographical and economic processes as internationalisation and metropolisation of cities in terms of firm networks within a generative model, where weights of cities are given by shares of turnovers generated by ownership links between firms.
-Previous literature considers mostly weights as GDP of countries for example in gravity models of export-import firms networks \cite{Kepaptsogloual2010} \cite{MinondoRequena2013}.
-The use of network models to understand processes underlying the emergence and dynamics of such networks has already been tackled in the literature, but still not extensively and never explicitly at city level \cite{taylor2001specification}.
-In fact generative urban network models combining geographical factors with network topological properties exist at the regional scale {\cite{dai2016simulating}}.
-Trade networks have been extensively studied from the complex networks perspective, but these are generally driven at the country-level \cite{garlaschelli2005structure} because of a lack of city-level data.
-Input-output models, mostly used in regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}.
-In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for different types of flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2018indirect}.
-
 The specific question we tackle is the complementarity of the different processes to generate urban firm networks.
@@
 More precisely: (i) we firstly introduce a generative network model to simulate the growth of such linkages at urban areas scale, combining multiple factors influencing link formation, such as the economic intensity of the urban areas corresponding to the source and target nodes, the industrial sector proximity, the strength of previous links, and the geographical and socio-cultural distances; (ii) the model allows us to then compare the effect of the different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; (iii) we develop an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database \cite{AMADEUS2018};  (iv) we use this data to calibrate the model on the empirical network; and (v) we explore an application of the model to an economic shock scenario, with pandemic-like stylised facts.
-
-%The rest of this paper is organized as follows: we firstly describe the generative model for urban networks and then present an empirical analysis of the network used as a case study.
-We then perform a sensitivity analysis and exploration of the model parameters using synthetic data.
-The model is then calibrated on empirical European firm ownership data.
-We finally discuss implications of our results and possible extensions.
 
@@
 As the network is rather sparse, these more detailed statistical models do not perform better than more simple models, as missing links are difficult to predict.
-The best zero-inflated model has a $R^2 = 0.16$, and it is similar to the performance of the best hurdle moel $R^2 = 0.17$.
+The best zero-inflated model has a $R^2 = 0.16$, and it is similar to the performance of the best hurdle model $R^2 = 0.17$.
 
@@
 The total order index is given by $\mathbb{E}_{\mathbf{X}_{\sim i}} \left[Var(Y_j | \mathbf{X}_{\sim i})\right] / Var\left[Y_j\right]$ and captures non-homogeneity in behaviour and interaction between parameters.
-Indices were estimated with a Sobol sequence sampling of 20,000 model runs \cite{saltelli2010variance}.
+Indices were estimated with a Sobol sequence sampling of 20,000 parameter points (corresponding to 160,000 model runs) \cite{saltelli2010variance}.
 We give in S2 Text the full estimated results for sensitivity indices.
@@
 
-The calibration of the model suggests its potential application to policy issues and can therefore have practical applications in the future.
+The model calibration results provide model parametrisations reasonably approaching the observed network.
+This validates its potential application to stylised policy issues with practical applications in the future.
 For example the effect of exogenous shocks in the socio-economic structure can be tested as we did in the stylized application example above.

### 2022-05-23 main.tex
Update on Overleaf.

@@
 
-\\
 Juste Raimbault\textsuperscript{1,2,3,4 *},
@@
 These two indicators capture the overall model performance with the logarithm of MSE, but also performance on medium-sized and small links, and are therefore complementary as shown in the case of cities population dynamics by \cite{raimbault2018indirect}.
-A calibration on macroscopic indicators of metropolisation and internationalisation would not bring much knowledge on model processes, since several points from the grid experiment above - not designed for this - already provide a good performance (19 parameter points out of 6534 are simultaneously below a relative error of 10\% on internationalisation and below 10\% on metropolisation).
+A calibration on macroscopic indicators of metropolisation and internationalisation would not bring much knowledge on model processes, since several points from the grid experiment above - not designed for this - already provide a good performance: 19 parameter points out of 6534 are simultaneously below a relative error of 10\% on internationalisation and below 10\% on metropolisation.
+These points correspond to diverse parametrisation, suggesting the presence of equifinality when trying to calibrate at the macroscopic level.
 

### 2022-05-23 S2_Text.tex
Update on Overleaf.

@@
 
-\section*{S2 Text: Statistical models, generative model setup and sensitivity analysis}
+\section*{S2 Text: Statistical models, generative model setup, sensitivity analysis, and model calibration}
 
@@
 
+\subsection*{Model calibration}
+
+When all parameters except $w_0$ are fixed, an external constraint can be enforced, hence fixing its value.
+A simple example is to set the total volume of simulated links to that it equals the total volume of observed links.
+Since by construction we have at final time $\sum_{i,j} w_{ij} = w_0 \cdot t_f$, this constraint yields $w_0 = \frac{\sum_{ij} w_{ij}^{(obs)}}{t_f}$ with $w_{ij}^{(obs)}$ observed weights.
+
+For the model calibration, we choose a constraint minimising one of the two objectives.
+Rescaling the final simulated flows by some constant $k_0$ while having $w_0 = 1$ is equivalent to set $w_0 = k_0$.
+We can thus determine the optimal value of $k_0$ that will minimise
+
+\[
+\varepsilon_L (k_0) = \frac{1}{N^2} \sum_{i,j} \left(\log w_{ij} - \log \left(k_0 \cdot \hat{w}_{ij} \right) \right)^2
+\]
+
+The derivative along $k_0$ must thus be zero, what yields
+
+\[
+0 =  \frac{1}{N^2} \sum_{i,j} \log w_{ij} - \frac{1}{N^2} \sum_{i,j} \log \hat{w}_{ij} - \frac{1}{N^2} \sum_{i,j} \log k_0
+\]
+
+and therefore
+
+\[
+k_0 = \exp \left[\frac{1}{N^2} \sum_{i,j} \log w_{ij} - \frac{1}{N^2} \sum_{i,j} \log \hat{w}_{ij} \right]
+\]
+
+In practice, we rescale the final simulated flows by this value in the calibration procedure.
+Note that only one constraint can be satisfied, and that this rescaling which improves the $\varepsilon_L$ objective probably deteriorates the $\varepsilon_M$ objective.
+We choose to rescale this dimension as it is directly comparable with the statistical models.
+
 \end{document}

### 2022-05-23 main.tex
Update on Overleaf.

@@
 An empirical network analysis on European urban firm ownership data confirms the relevance of each of these factors on new link creation.
-We then simulate network growth on synthetic systems of cities, unveiling stylized facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range.
+We then simulate network growth on synthetic systems of cities, unveiling stylised facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range.
 We calibrate the model on the European network, outperforming statistical models, and showing a strong role of path-dependency.
@@
 Networks constitute the new social morphology of our societies, since the intensification of globalisation of the economy in the late 20th century \cite{castells2000networksociety}.
-A world-city system has since then emerged \cite{taylor2001specification}, characterized by highly interconnected urban centres at the global level \cite{sassen1991}.
+A world-city system has since then emerged \cite{taylor2001specification}, characterised by highly interconnected urban centres at the global level \cite{sassen1991}.
 These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity inducing spatial and non-spatial patterns \cite{martinus2018global}.
@@
 
-At each time step, one directed link is randomly drawn. We define a generalized Cobb-Douglas utility function \cite{vilcu2011geometric} as
+At each time step, one directed link is randomly drawn. We define a generalised Cobb-Douglas utility function \cite{vilcu2011geometric} as
 
@@
 The initial network is the same than in the synthetic setup, and the real links are taken as targets for the generated network.
-We constrain the total volume of flows by setting $w_0 = \sum_{ij} w_{ij}^{(obs)} / t_{f}$ when the number of time steps $t_f$ is given as a parameter.
+We constrain the total volume of flows by setting $w_0 = \sum_{ij} w_{ij}^{(obs)} / t_{f}$ when the number of time steps $t_f$ is given as a parameter and $w_{ij}^{(obs)}$ are observed flows.
 
@@
 \emph{AMADEUS} provides a precise location, turnover and activity codes for owners and owned companies in all sectors in Europe.
-The Bureau Van Dick gathers, homogenizes and completes the information on these firms from national sources, publications of consultancies and audit firms such as KPMG, Deloitte, PricewaterhouseCoopers (PwC), Ernst \& Young (EY), Mazars, and those of other financial organisations such as Moody's, Fitch, Standard \& Poor's, Thomson Reuters or Duns and Bradstreet.
+The Bureau Van Dick gathers, homogenises and completes the information on these firms from national sources, publications of consultancies and audit firms such as KPMG, Deloitte, PricewaterhouseCoopers (PwC), Ernst \& Young (EY), Mazars, and those of other financial organisations such as Moody's, Fitch, Standard \& Poor's, Thomson Reuters or Duns and Bradstreet.
 A representativity study shows that \emph{ORBIS} --the Bureau Van Dijk's dataset for firms at the worldwide scale--, captures around 10 percent of existing foreign direct investments in 2011 in Poland \cite{Zdanowska2018}.
@@
 We give in S2 Text the full estimated results for sensitivity indices.
-To summarize, we find that indicators related to modularity (internationalisation and optimal modularity) are mostly influenced by geographical parameters.
+To summarise, we find that indicators related to modularity (internationalisation and optimal modularity) are mostly influenced by geographical parameters.
 On the contrary, indicators of network structure are driven by origin and destination sizes.
@@
 We run a specific experiment to study the impact of the urban hierarchy on firms dynamics.
-Rank-size law in cities is a central stylized fact within urban theories, despite the discussions regarding the estimation of the power-law exponent, its actual values and its variation depending on the definition of cities considered \cite{cottineau2017metazipf,corral2020truncated}.
+Rank-size law in cities is a central stylised fact within urban theories, despite the discussions regarding the estimation of the power-law exponent, its actual values and its variation depending on the definition of cities considered \cite{cottineau2017metazipf,corral2020truncated}.
 I this particular context, understanding how this parameter can impact urban dynamics within artificial systems of cities -- in which it can be modified to extreme values non observable in real systems--, should help disentangling the variability due to model parameters from the variability due to the specific configuration of the urban system \cite{raimbault2019space}.
@@
 As expected, the hierarchy of the urban system has an important impact on model outcomes.
-We find that more hierarchical urban systems tend to produce more international networks (recall that the internationalisation indicator corresponds to the modularity, the latter is minimized when the networks span more across different countries).
+We find that more hierarchical urban systems tend to produce more international networks (recall that the internationalisation indicator corresponds to the modularity, the latter is minimised when the networks span more across different countries).
 This is consistent with the existence of globalised metropolis \cite{sassen1991global}.
@@
     \end{center}
-    \caption{\textbf{Impact of the urban hierarchy.} (Left) Internationalisation as a function of the initial synthetic hierarchical exponent $\alpha$, for different values of the interaction range $d_0$ (color); (Right) Metropolisation as a function of $\alpha$, for different values of the interaction range $d_0$ (color).\label{fig:fig7}}
+    \caption{\textbf{Impact of the urban hierarchy.} (Left) Internationalisation as a function of the initial synthetic hierarchical exponent $\alpha$, for different values of the interaction range $d_0$ (colour); (Right) Metropolisation as a function of $\alpha$, for different values of the interaction range $d_0$ (colour).\label{fig:fig7}}
 \end{figure}
@@
 
-% TODO precise calib setting + details on rescaling
 We then apply the model to a real system of cities by calibrating it on the European ownership network.
@@
 The number of time steps $t_f$ is left as an additional parameter, which in a sense determines the granularity of the cumulative network generation process.
+To avoid overfitting and including information from statistical models, we deactivate the socio-economic gravity interactions by setting $c_0 = 0$ (distances are obtained as fixed effects from a statistical model).
+Finally, $w_0$ can be set so that a given constraint is satisfied in a way similar to spatial interaction models \cite{wilson1975some}: in this case, we take an optimal value for the $\varepsilon_L$ objective given below as detailed in S2 Text.
 
@@
 $\varepsilon_L$ can be compared to the MSE of statistical models fitted above, and the generative model outperform these for several parametrisations (lowest MSE 5.33 for statistical models).
-Point size gives the number of stochastic samples and color the value of $\gamma_O$.\label{fig:fig8}}
+Point size gives the number of stochastic samples and colour the value of $\gamma_O$.\label{fig:fig8}}
 \end{figure}
@@
 
-When considering optimal points with $\varepsilon_L < 5.0$, the adjusted parameters are: long range gravity interactions and socio-economic interactions given by $d_0 = 5972 \pm 2148$ and $c_0 = 113 \pm 33$ (the maximal $c_{ij}$ is 11.8); a stronger influence of the destination than of the origin as $\gamma_O = 1.46 \pm 0.10$ and $\gamma_D = 2.0 \pm 0.05$; a comparable influence of sectors as $\gamma_S = 1.76 \pm 0.21$; a fine time granularity as $t_f = 4951 \pm 124$; and a very strong influence of previous links as $\gamma_W = 6.43 \pm 4.11$.
+When considering optimal points with $\varepsilon_L < 5.0$, the adjusted parameters are: long range gravity interactions given by $d_0 = 7432 \pm 2350$; a stronger influence of the destination than of the origin as $\gamma_O = 1.41 \pm 0.11$ and $\gamma_D = 1.95 \pm 0.06$; a comparable influence of sectors as $\gamma_S = 1.20 \pm 0.14$; a fine time granularity as $t_f = 4927 \pm 119$; and a very strong influence of previous links as $\gamma_W = 7.72 \pm 5.15$.
 This last point furthermore shows in particular the importance of including path-dependency in this complex generative model.
 The interaction of this process with others implies an inversion of the relative importance of origin and destination (in opposition to the statistical models).
-The fact that calibrated interaction distances are large is consistent with long range economic interactions within Europe.
-Besides these stylized facts, these calibration results altogether demonstrate the feasibility of the application of the model to real data.
+The fact that calibrated interaction distance is large is consistent with long range economic interactions within Europe.
+Besides these stylised facts, these calibration results altogether demonstrate the feasibility of the application of the model to real data.
 As stated previously, the model is not predictive and the relatively high error values would not make it satisfactory in that manner, but are reasonable for its explanatory and application functions.
@@
     \end{center}
-	\caption{\textbf{Application of the calibrated model to an economic shock scenario.} All plots show the trajectory of indicators (internationalisation for the top row and average community size for the bottom row) in time, for different values of interaction decay rescaling $\kappa_G$ (color) and socio-economic distance rescaling $\kappa_C$ (panels).
+	\caption{\textbf{Application of the calibrated model to an economic shock scenario.} All plots show the trajectory of indicators (internationalisation for the top row and average community size for the bottom row) in time, for different values of interaction decay rescaling $\kappa_G$ (colour) and socio-economic distance rescaling $\kappa_C$ (panels).
 The vertical dotted red line shows the moment the economic shock is introduced and $d_0$ (respectively $c_0$) is set to $\kappa_G \cdot d_0$ (resp.
@@
 This validates its potential application to stylised policy issues with practical applications in the future.
-For example the effect of exogenous shocks in the socio-economic structure can be tested as we did in the stylized application example above.
+For example the effect of exogenous shocks in the socio-economic structure can be tested as we did in the stylised application example above.
 Such shocks can for example be the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to further pandemic outbreaks on the single market.
 In this sense the UK economic sectors mostly dependent on foreign capital and generating the highest turnovers in 2018, that would be the most affected and worth examining in detail are: the extraction of crude petroleum, the manufacture of tobacco products and of basic pharmaceutical goods (in terms of out-going links from the UK) ; the activities of head offices and holding companies, the sale of cars and manufacture of motor vehicles (in terms of in-going links to the UK).
-We showed in our experiment that although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalisation and thus can be expected to have a detrimental effect on the UK economy due to these foreign ownership links (see also S1 Text\ref{S1Text}).
+We showed in our experiment that although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalisation and thus can be expected to have a detrimental effect on the UK economy due to these foreign ownership links (see also S1 Text).
 The impact of such unforeseen economic shocks could also be considered from a multi-level perspective, coupling our work with other approaches such as \cite{Dorretal2022} studying the firm-level impact of the Covid crisis.
@@
 A model with additional firm agents would lead to a multi-scale agent-based model, with also more flexibility for policy applications.
-The link with data could also be more elaborated: indeed the statistical analysis does not include spatial or neighborhood effects, and spatial non-stationarity may be accounted for using Geographically Weighted Regression models for example \cite{comber2021route}, to then be coupled with the ABM in an hybrid approach.
+The link with data could also be more elaborated: indeed the statistical analysis does not include spatial or neighbourhood effects, and spatial non-stationarity may be accounted for using Geographically Weighted Regression models for example \cite{comber2021route}, to then be coupled with the ABM in an hybrid approach.
 Finally, an adaptation at the world scale using for example urban regions and data explored by \cite{rozenblat2021intra}, would be an extension with important practical and theoretical potentialities.
@@
 
-\subsection*{S1 Text: UK's economic sectors dependent on international links.}
-\label{S1Text}
+\subsection*{S1 Text: UK's economic sectors dependent on international links.\label{S1Text}}
 
@@
     \item procedure for the synthetic model setup;
-    \item Global Sensitivity Analysis
+    \item Global Sensitivity Analysis;
+    \item model calibration.
 \end{itemize}

### 2022-05-23 CoverLetter.tex
submission_ok

@@
 \recipient{Editors PLoS ONE}{}
-\date{May 12th 2022}
+\date{May 23th 2022}
 \opening{Dear Editors,}
@@
 \justify
-We are pleased, with my co-authors Natalia Zdanowska (Luxembourg Institute of Socio-economic Research) and Elsa Arcaute (University College London), to submit an original research article entitled ``Modeling growth of urban firm networks'' for consideration for publication in PLoS ONE.
+We are pleased, with my co-authors Natalia Zdanowska (Luxembourg Institute of Socio-economic Research) and Elsa Arcaute (University College London), to submit an original research article entitled ``Modelling growth of urban firm networks'' for consideration for publication in PLoS ONE.
 This article introduces a generative model for urban firm networks, to understand underlying processes, with possible applications to public policies.

### 2022-05-23 main.tex
submission_ok

@@
 Networks constitute the new social morphology of our societies, since the intensification of globalisation of the economy in the late 20th century \cite{castells2000networksociety}.
-A world-city system has since then emerged \cite{taylor2001specification}, characterised by highly interconnected urban centres at the global level \cite{sassen1991}.
+A world-city system has since then emerged \cite{taylor2001specification}, characterised by highly interconnected urban centres at the global level \cite{Sassen1991}.
 These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity inducing spatial and non-spatial patterns \cite{martinus2018global}.
@@
 Secondly, we consider the geographical process of metropolisation.
-The later can be defined as a concentration of population and major economic decision functions in metropolises, in particular at the world-system level \cite{sassen1991global}.
+The later can be defined as a concentration of population and major economic decision functions in metropolises, in particular at the world-system level \cite{Sassen1991}.
 Metropolises playing the role of gateways of globalisation, are competing between each other to attract the most skilled workers and innovative industries \cite{RozenblatPumain2007}.
@@
 We find that more hierarchical urban systems tend to produce more international networks (recall that the internationalisation indicator corresponds to the modularity, the latter is minimised when the networks span more across different countries).
-This is consistent with the existence of globalised metropolis \cite{sassen1991global}.
+This is consistent with the existence of globalised metropolis \cite{Sassen1991}.
 This effect is mitigated by reducing the interaction ranges.

### 2022-07-19 misc/extended_abstract_urbansys.tex
Update on Overleaf.

\documentclass[10pt]{article}

\usepackage{graphicx}
\usepackage[margin=2cm]{geometry}

\title{\vspace{-2cm}Modelling growth of urban firm networks}
\author{J. Raimbault$^{1,\ast}$, N. Zdanowska$^2$ and E. Arcaute$^3$\smallskip\\
$^1$ LASTIG, IGN-ENSG; $^2$ LISER; $^3$ CASA, UCL\\\smallskip
$\ast$ Presenting author: \texttt{juste.raimbault@ign.fr}
}
\date{}

\begin{document}

\maketitle

\pagenumbering{gobble}

The emergence of interconnected urban networks is a crucial feature of globalisation processes. Understanding the drivers behind the growth of such networks -- in particular urban firm networks --, is essential for the economic resilience of urban systems. We introduce in this paper a generative model for firm networks at the urban area level, which considers the following features: the economic size of the areas representing the source and target nodes, the industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distances between urban areas. These are combined into a generalised spatial interaction model giving the probability to add links between urban areas, used in an iterative manner to generate the network from an initial configuration.

An empirical network analysis on European urban firm ownership data confirms the relevance of each of these factors on new link creation. We construct a directed network between functional urban areas in Europe, based on the AMADEUS data which is a well covering dataset describing ownership links between companies and their subsidiaries. The links between firms are aggregated at the urban area level and economic size of areas are computed by aggregating turnovers of firms. A community detection analysis provides a community structure which is partly geographical but also captures some known economic collaboration areas (see Fig.~1). Statistical models (various specifications of spatial interaction models) assess the role of various factors to be included in the simulation model.

We then simulate network growth on synthetic systems of cities with the generative model. Examples of generated networks are shown in Fig.~1. Global sensitivity analysis provides the influence of parameters on model outputs characterising network structure. A grid sampling of the parameter space unveils stylised facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range.

We calibrate the model on the European network, with a real geographical setup and economic sector composition obtained from the empirical network data. Calibration is done on two objectives, in order to reproduce the real network in terms of link weights (mean squared error on weights and mean squared error on logarithms of weights), with a genetic algorithm for bi-objective optimisation (NSGA2). Considering the common fitting objective, the simulation model outperforms statistical models. This furthermore shows a strong role of path-dependency for model performance.

The model is finally applied to stylised scenarios, where economic shocks modify spatial interaction parameters after a fixed number of time steps. This opens up to the formulation of public policies recommendations to deal with exogenous shocks such as economic crises or potential lock-downs of countries.

\begin{figure}[h!]
    \centering
    \vspace{-0.2cm}
    \includegraphics[width=0.45\linewidth]{../figures/Fig2.png}
    \includegraphics[width=0.42\linewidth]{../figures/Fig3.png}
    \caption{(Left) Empirical analysis of the AMADEUS data: urban area size corresponds to economic size, while color gives the community obtained with a Louvain modularity optimisation algorithm; (Right) Example of simulated networks: top row in a synthetic system with long-range interactions (left) and short range interactions (right), bottom row with same parameter values but in the real setting.}
\end{figure}

\end{document}

### 2022-10-19 communications/CCS2022_UrbanSys/slides.tex
slides_urbansysCCS22

--- 2020-09-18 communications/NetSci2020/slidesNetSci.tex
@@
 
-\title
-[A generative model for urban firm networks]{A generative model for urban firm networks}
-\author[Raimbault \and Zdanowska]{J.~Raimbault$^{1,2,3\ast}$ \and N. Zdanowska$^{1,3}$\\\medskip
-$^{\ast}$\texttt{n.zdanowska@ucl.ac.uk}
+\title{Modelling growth of urban firm networks}
+\author{J.~Raimbault$^{1,2,3,4,\ast}$ \and N. Zdanowska$^{5,2,4}$ \and E. Arcaute$^{2}$\\
+$\ast$\texttt{juste.raimbault@ign.fr}
 }
 
-\institute[UCL]{$^{1}$Center for Advanced Spatial Analysis, University College London\\
-$^{2}$UPS CNRS 3611 Complex Systems Institute Paris\\
-$^{3}$UMR CNRS 8504 G{\'e}ographie-cit{\'e}s
+\institute{$^{1}$LASTIG, Univ Gustave Eiffel, IGN-ENSG\\
+$^{2}$CASA, UCL\\
+$^{3}$UPS CNRS 3611 ISC-PIF\\
+$^{4}$UMR CNRS 8504 G{\'e}ographie-cit{\'e}s\\
+$^{5}$Luxembourg Institute of Socio-Economic Research
 }
 
-\date[24/09/2020]{NetSci 2020\\
-Session 12F: Diffusion\\
-September 24th, 2020
+\date{CCS 2022\\
+UrbanSys Satellite\\
+19/10/2022
 }
@@
 \frame{\maketitle}
+
+% extended abstract
+% The emergence of interconnected urban networks is a crucial feature of globalisation processes.
+Understanding the drivers behind the growth of such networks – in particular urban firm networks –, is essential for the economic resilience of urban systems.
+We introduce in this paper a generative model for firm networks at the urban area level, which considers the following features: the economic size of the areas representing the source and target nodes, the industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distances between urban areas.
+These are combined into a generalised spatial interaction model giving the probability to add links between urban areas, used in an iterative manner to generate the network from an initial configuration.
+%An empirical network analysis on European urban firm ownership data confirms the relevance of each of these factors on new link creation.
+We construct a directed network between functional urban areas in Europe, based on the AMADEUS data which is a well covering dataset describing ownership links between companies and their subsidiaries.
+The links between firms are aggregated at the urban area level and economic size of areas are computed by aggregating turnovers of firms.
+A community detection analysis provides a community structure which is partly geographical but also captures some known economic collaboration areas (see Fig.
+1).
+Statistical models (various specifications of spatial interaction models) assess the role of various factors to be included in the simulation model.
+%We then simulate network growth on synthetic systems of cities with the generative model.
+Examples of generated networks are shown in Fig.
+1.
+Global sensitivity analysis provides the influence of parameters on model outputs characterising network structure.
+A grid sampling of the parameter space unveils stylised facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range.
+%We calibrate the model on the European network, with a real geographical setup and economic sector composition obtained from the empirical network data.
+Calibration is done on two objectives, in order to reproduce the real network in terms of link weights (mean squared error on weights and mean squared error on logarithms of weights), with a genetic algorithm for bi-objective optimisation (NSGA2).
+Considering the common fitting objective, the simulation model outperforms statistical models.
+This furthermore shows a strong role of path-dependency for model performance.
+%The model is finally applied to stylised scenarios, where economic shocks modify spatial interaction parameters after a fixed number of time steps.
+This opens up to the formulation of public policies recommendations to deal with exogenous shocks such as economic crises or potential lock-downs of countries.
+% Fig 2, 3
+% Figure 1: (Left) Empirical analysis of the AMADEUS data: urban area size corresponds to economic size, while color gives the community obtained with a Louvain modularity optimisation algorithm; (Right) Example of simulated networks: top row in a synthetic system with long-range interactions (left) and short range interactions (right), bottom row with same parameter values but in the real setting.
 
@@
 
-General framework:
+%General framework:
+\textit{Role of firm networks in urban systems:}
 
@@
 
+\textit{Relevance of dynamics urban network evolution models:}
+
+\medskip
+
 \begin{itemize}
@@
 
-Assumptions:
+Main geographical processes considered:
 
@@
 
+Advantages of the model introduced:
+
+\medskip
+
 \begin{itemize}
 	\item Indirect inference on processes
+	\item Multiple complementary factors driving inter-urban links
 	\item Includes path-dependency
@@
 
-Cities characterized by economic size $E_i$ (GDP) and economic structure $S_{ik}$ (probability distribution of firms within $K$ sectors)
+Cities characterised by economic size $E_i$ (GDP) and economic structure $S_{ik}$ (probability distribution of firms within $K$ sectors)
 
@@
 
-Starting from an initial network, at each time step, add a fixed number of links randomly, following a probability as a generalized Cobb-Douglas function
+Starting from an initial network, at each time step, add a fixed number of links randomly, following a probability as a generalised Cobb-Douglas function
 
@@
 
-\sframe{Statistical models}{
+\sframe{Empirical network properties}{
+
+\begin{center}
+    \includegraphics[width=\linewidth]{../../figures/Fig1.png}
+\end{center}
+
+\textit{Empirical distribution of network properties, with log-normal and power law fits}
+
+}
+
+\begin{frame}[label=statmain]
+
+\frametitle{Statistical models}
 
@@
 \resizebox{\linewidth}{!}{
-\begin{tabular}{|l|c|c|c|c|c|}
+\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|c|c|}
 \hline
-Model  & (1) & (2) & (3) & (4) & (5 - Poisson GLM) \\ 
+ & \multicolumn{2}{|c|}{$d_0$} & \multicolumn{2}{|c|}{$c_0$} & \multicolumn{2}{|c|}{$\gamma_S$} & \multicolumn{2}{|c|}{$\gamma_W$} & \multicolumn{2}{|c|}{$\gamma_O$} & \multicolumn{2}{|c|}{$\gamma_D$} \\
+ & F & T & F & T & F & T & F & T & F & T & F & T \\
  \hline
-$\log(d_{ij})$ &      -0.06** (0.03) &   -0.11** (0.05)  & -0.41*** (0.02)  & -0.35*** (0.04)  &  -0.26*** ($5\cdot 10^{-6}$) \\
-$\log(T_i)$ &   &   & 0.56*** (0.01) &  0.56*** (0.01) & 0.79*** ($2\cdot 10^{-6}$) \\
-$\log(T_j)$ &     &   & 0.39*** (0.01) &  0.39*** (0.01) & 0.66***  ($1.5\cdot 10^{-6}$) \\
-$\log(s_{ij})$ &     &   &  &  0.19*** (0.07) & 0.53*** ($9\cdot 10^{-6}$)  \\
-Countries &    &  28.3\% &   &  26.5\% & 100\% \\
-\hline
-$R^2$ &       0.00059   &  0.17 & 0.21 &  0.31  &  0.56 \\
-MSE & 7.75 & 6.84 & 6.10 & 5.33 & 8.72 \\
-AIC &        44304   &  43578  &  42131  & 41917 &   \\
-\hline
+Internationalisation & 0.2 & 0.3 & 0.7 & 0.7 & 0.001 & 0.008 & $5\cdot 10^{-4}$ & 0.007 & 0.03 & 0.04 & 0.02 & 0.04 \\
+Metropolisation & 0.02 & 0.04 & 0.02 & 0.04 & 0.02 & 0.3 & 0.002 & 0.01 & 0.2 & 0.7 & 0.2 & 0.7 \\
+Modularity & 0.3 & 0.4 & 0.6 & 0.6 & $5\cdot 10^{-4}$ & 0.02 & 0.002 & 0.02 & 0.007 & 0.03 & 0.003 & 0.03 \\
+Avg. com. size & 0.004 & 0.1 & 0.01 & 0.1 & 0.003 & 0.07 & 0.001 & 0.04 & 0.3 & 0.6 & 0.3 & 0.6 \\
+Degree entropy & 0.003 & 0.03 & 0.002 & 0.03 & 0.008 & 0.04 & 0.003 & 0.03 & 0.5 & 0.6 & 0.5 & 0.6 \\
+Weight entropy & 0.04 & 0.2 & 0.03 & 0.2 & 0.002 & 0.1 & 0.001 & 0.1 & 0.5 & 0.6 & 0.5 & 0.6 \\\hline
 \end{tabular}
@@
 
-}
+\bigskip
+
+Zero-inflated and Hurdle models: \hyperlink{zeroinfl}{\beamerbutton{Details}}
+
+\end{frame}
 
@@
     \bigskip
-    \item Simulation on the European system of cities (AMADEUS data at FUA level)
+    \item Simulation on the European system of cities, setup with AMADEUS data at FUA level (\hyperlink{realsetup}{\beamerbutton{Details}})
 \end{enumerate}
@@
 
-    % One factor sampling results
-    %  20190924_162740_ONEFACTOR_REPLICATIONS_SYNTHETIC_GRID
-
     \includegraphics[width=0.48\textwidth]{../../figuresraw/internationalisation-gravityDecay_errorbars.png}
@@
 
-    \textit{(Left) Internationalization index decreases exponentially with gravity decay; (Right) Correlation between city weighted degree and size.
-Both plots show a transition from a local to a global regime.}
+    \textit{(Left) Internationalisation index decreases exponentially with gravity decay; (Right) metropolisation also shows a progressive transition from a local to a global regime.}
 
@@
 \sframe{Effect of sector proximity}{
-
-    % One factor sampling results
-    %  20190924_162740_ONEFACTOR_REPLICATIONS_SYNTHETIC_GRID
 
@@
 
-    \textit{(Left) Internationalization varies linearly with sector proximity $\gamma_S$; (Right) Correlation between degree and size exhibits a maximum, witnessing an intermediate regime where size is the most important}
+    \textit{(Left) Internationalisation varies linearly with sector proximity $\gamma_S$; (Right) metropolisation witnesses an intermediate regime where size is the most important}
 
@@
 
-\sframe{Grid exploration: internationalization}{
+\sframe{Grid exploration: internationalisation}{
 
@@
 
-\sframe{Calibration on the European urban network}{
+\sframe{Role of urban hierarchy}{
 
 \begin{center}
-    	\includegraphics[height=0.75\textheight]{figures/Fig8.png}
+    \includegraphics[width=\linewidth]{../../figures/Fig7.png}
+\end{center}
+
+\medskip
+
+\textit{Impact of urban hierarchy on model indicators in synthetic systems of cities: existence of a maximum for metropolisation at intermediate hierarchies.}
+
+}
+
+\begin{frame}[label=calibmain]
+
+\frametitle{Calibration on the European urban network}
+
+    \begin{center}
+    	\includegraphics[height=0.7\textheight]{../../figures/Fig8.png}
 
@@
 
-}
+    % weird point out of pareto front: ?
+
+   \hyperlink{calibration}{\beamerbutton{Details}} on the calibration procedure
+
+\end{frame}
 
@@
 
-\textit{Although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalization and thus can be
+\textit{Although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalisation and thus can be
 expected to have a detrimental effect on UK economy due to these foreign ownership links.}
@@
 
-\textbf{Practical application}
+\textbf{Practical applications}
 
@@
 
+\medskip
+
 $\rightarrow$ Effect of exogenous shocks on specific economic sectors mostly dependent on foreign ownership links (extraction of crude petroleum, manufacture of tobacco \& basic pharmaceutical goods, activities of head offices).
 
+\bigskip
 \bigskip
@@
 
-$\rightarrow$ co-evolution model (evolution of city sizes) \cite{raimbault2018modeling};
+\medskip
+
+$\rightarrow$ co-evolution model (evolution of city sizes) \cite{raimbault2021modeling};
+
+\medskip
 
@@
 
+\medskip
+
 $\rightarrow$ targeted study of path dependency;
+
+\medskip
 
@@
 
-\sframe{Empirical network properties}{
-
-\begin{center}
-    \includegraphics[width=\linewidth]{../../figures/Fig1.png}
-\end{center}
-
-\textit{Empirical distribution of network properties, with log-normal and power law fits}
-
-}
-
 \begin{frame}[label=syntheticsetup]
@@
     \item Create countries with k-means clustering ($C \textrm{ = } 30$)
-    \item Distribute sectors such that (i) smaller cities are more specialized and (ii) larger cities are more knowledge-based, with a one dimensional axis to position sectors $1/K \ldots 1$ where the density $f\left(k\right)$ follows a log-normal with $\left(\mu,\sigma\right)$ such that $\sqrt{\Var f} \textrm{ = } K/2$ for the largest, $\sqrt{\Var f} \textrm{ = } 1/K$ for the smallest
+    \item Distribute sectors such that (i) smaller cities are more specialised and (ii) larger cities are more knowledge-based, with a one dimensional axis to position sectors $1/K \ldots 1$ where the density $f\left(k\right)$ follows a log-normal with $\left(\mu,\sigma\right)$ such that $\sqrt{\Var f} \textrm{ = } K/2$ for the largest, $\sqrt{\Var f} \textrm{ = } 1/K$ for the smallest % equations?
+too much -> refer to paper
 \end{enumerate}
+
+\hyperlink{setupmain}{\beamerbutton{Back}}
+
+\end{frame}
+
+\begin{frame}[label=realsetup]
+
+\frametitle{Real configuration setup}
+
+% sociocultural distance
+
+Urban areas (FUAs) positions, economic size initialised from the GHSL database, sector composition from the Amadeus database using classification of firms and aggregating
+
+\bigskip
+
+Initial dummy network ($w_0 = 1$)
+
+\bigskip
+
+\textbf{Socio-cultural distance} $c_{ij}$: fixed effects coefficients
+
+% Q: how are NAs handled?: infinite: ~ - anyway not used in calibration
+
+\begin{tabular}{ccccccc}
+   Min. & 1st Qu. &  Median &  Mean &  3rd Qu. &  Max. &  NA's \\
+ -2.327 &  1.440 &  2.451 &  2.538 &  3.447 & 11.797 &  339 
+\end{tabular}
 
@@
 
-\sframe{Role of urban hierarchy}{
+\begin{frame}[label=zeroinfl]
+
+\frametitle{Zero-inflated and Hurdle statistical models}
+
+Accounting for zeros with specific models: same qualitative results
+
+\bigskip
 
 \begin{center}
-    \includegraphics[width=\linewidth]{../../figures/Fig7.png}
+\resizebox{\linewidth}{!}{
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+Model  & \multicolumn{2}{|c|}{Zero-inflated} & \multicolumn{2}{|c|}{Hurdle} \\ 
+ & Count & Zero-infl. & Count & Zero-hurdle \\
+\hline
+$\log(d_{ij})$ &    -0.28*** (7e-4) &   0.88*** (0.01)      &   -0.29*** (8e-6)   &  -0.75*** (0.01)    \\
+$\log(T_i)$ & 0.74*** (4e-6)  & -0.74*** (7e-3)       &   0.74***  (5e-6)     &  0.67*** (5e-3)   \\
+$\log(T_j)$ & 0.64*** (4e-6)  & -0.52*** (7e-3)      &    0.64***  (4e-6)     &  0.48*** (5e-3)    \\
+$\log(s_{ij})$ & 0.46*** (2e-5)   &  -0.33*** (3e-2)  &  0.45*** (2e-5)       &  0.27*** (0.02)    \\
+\hline
+$R^2$ &   \multicolumn{2}{|c|}{0.1607}     &  \multicolumn{2}{|c|}{0.1661}  \\
+%MSE & & \\
+AIC &     \multicolumn{2}{|c|}{5.85e10}    &   \multicolumn{2}{|c|}{5.85e10}  \\
+\hline
+\end{tabular}
+}
 \end{center}
 
-\medskip
+\hyperlink{statmain}{\beamerbutton{Back}}
 
-\textit{Impact of urban hierarchy on model indicators in synthetic systems of cities.}
+\end{frame}
 
-}
+\begin{frame}[label=calibration]
+
+\frametitle{Model calibration}
+
+\footnotesize
+
+\textbf{To avoid overfitting:} $c_0 = 0$ to deactivate the mechanism
+
+\bigskip
+
+\textbf{Calibration objectives:}
+
+\[
+\varepsilon_L = \frac{1}{N^2} \sum_{i,j} \left(\log w_{ij} - \log \hat{w}_{ij} \right)^2
+\]
+
+\[
+\varepsilon_M = \log\left(\frac{1}{N^2} \sum_{i,j} \left(w_{ij} - \hat{w}_{ij}\right)^2 \right)
+\]
+
+\bigskip
+
+\textbf{On macro indicators?} Too much solutions, equifinality
+
+\bigskip
+
+\textbf{Rescale flows:} $w_0$ is arbitrary, so is optimised with rescaling:
+
+\[
+\varepsilon_L (k_0) = \frac{1}{N^2} \sum_{i,j} \left(\log w_{ij} - \log \left(k_0 \cdot \hat{w}_{ij} \right) \right)^2
+\]
+
+what yields
+
+\[
+k_0 = \exp \left[\frac{1}{N^2} \sum_{i,j} \log w_{ij} - \frac{1}{N^2} \sum_{i,j} \log \hat{w}_{ij} \right]
+\]
+
+\hyperlink{calibmain}{\beamerbutton{Back}}
+
+\end{frame}
 
